### 2022-06-03 reports/ACIC (2022)/poster/resources/rr.tex

% Created by tikzDevice version 0.12.3.1 on 2022-05-20 15:39:29
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,126.47);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,126.47);
\end{scope}
\begin{scope}
\path[clip] ( 35.37, 28.21) rectangle (283.58,120.97);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37, 28.21) rectangle (283.58,120.97);
\definecolor{drawColor}{RGB}{16,36,81}

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.11,101.41) --
	(213.11,105.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.11,103.58) --
	(118.54,103.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (118.54,101.41) --
	(118.54,105.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (197.44, 43.43) --
	(197.44, 47.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (197.44, 45.61) --
	( 86.05, 45.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 86.05, 43.43) --
	( 86.05, 47.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.09, 72.42) --
	(212.09, 76.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.09, 74.59) --
	(104.01, 74.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.01, 72.42) --
	(104.01, 76.77);
\definecolor{fillColor}{RGB}{16,36,81}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (165.35,103.58) circle (  1.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (139.20, 45.61) circle (  1.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (157.25, 74.59) circle (  1.32);
\definecolor{drawColor}{RGB}{16,36,81}

\path[draw=drawColor,draw opacity=0.70,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (207.83, 28.21) -- (207.83,120.97);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37, 28.21) rectangle (283.58,120.97);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 42.85) {0.05};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 71.84) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,100.83) {0.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 45.61) --
	( 35.37, 45.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 74.59) --
	( 35.37, 74.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,103.58) --
	( 35.37,103.58);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.88, 25.46) --
	( 78.88, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.36, 25.46) --
	(143.36, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.83, 25.46) --
	(207.83, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (272.30, 25.46) --
	(272.30, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 78.88, 17.75) {0.6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (143.36, 17.75) {0.8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (207.83, 17.75) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (272.30, 17.75) {1.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (159.47,  7.25) {Risk ratio};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70, 74.59) {Hypothetical limit (mg/m$^3$)};
\end{scope}
\end{tikzpicture}

\end{document}