### 2022-06-03 reports/paper/01-introduction.rmd

---
output:
  word_document:
    reference_docx: "~/HeadRs/rmd_templates/style-guide-arial-11.docx"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/asa.csl')`"
---

```{r, include=FALSE, eval=F}
knitr::opts_chunk$set(echo=F, results = 'asis')
library(bibtex)
biblio <- bibtex::read.bib('~/Box Sync/papers/EnvEpi.bib')
```

Metalworking fluids (MWF) are complex chemical mixtures used to lubricate and cool machinery during manufacturing processes. Straight MWF are the oldest class of MWF, comprising of simple oils or combinations of oils. By the mid 20th century, straight MWF waned in popularity and ceded market share to soluble MWF. Soluble MWF are water-oil mixtures with diverse additives for improving performance and durability. By the 1970s, semisynthetic and synthetic MWF, which contain little to no oil, overtook straight oils in market share. Nonetheless, at the turn of the millennium, soluble MWFs still retained the largest share of the MWF market [@Childers_2005]. Carcinogenicity of MWFs has been of concern since as early as the 1970s, when the International Agency for Research on Cancer (IARC) classified mineral oils as carcinogenic, citing studies of occupational exposures among workers in oil, textile, and metal industries [@IARC_1973]. Concerns over the health effects of straight MWF hastened the adoption of soluble MWF, which contain less oil per volume, but a richer cocktail of additives including chromates, cycloalkanes, phenols, organochlorines, nitrosamines, and triazines: classes of chemicals with known or plausible carcinogenicity [@IARC_1987;@IARC_2014;@IARC_2017]. Health and environmental considerations led to substantial variation in MWF composition over time, but the elimination of particular etiologic agents often leads to their replacement with other potentially hazardous chemical compounds. Exposure limits for MWF as a mixture encourage limiting occupational exposure to multiple agents simultaneously.

Here, we assess the effect of hypothetical exposure limits on non-Hodgkin lymphoma (NHL) cumulative incidence (risk) from 1985 to 2005 in the United Auto Workers-General Motors (UAW-GM) Occupational Cohort Study. The large size of the study population and rich time-varying, quantitative MWF exposure data provide an opportunity to study this relatively rare cancer while contrasting realistic interventions on MWF exposure in a longitudinal cohort setting. Though NHL accounts for a small proportion of cancer incidence of all sites, it has garnered intense research attention because its incidence nearly quadrupled since 1960 [@Ekstrom-Smedby_2006]. Immune modulation due to immunosuppressive therapy and infection by human immunodeficiency virus (HIV) are the most prominent risk factors of NHL, but they cannot fully explain this rise in NHL incidence [@Shiels_2013;@Chiu_2015]. Epidemiologic investigations of environmental and occupational exposures have yielded varied results [@Ekstrom-Smedby_2006;@Alexander_2007]. This variation in epidemiologic findings may reflect the substantial etiologic heterogeneity of NHL, but studies of specific NHL subtypes are limited by the lack of quantitative exposure information and the lack of control for the healthy worker survivor effect (HWSE) or other forms of time-varying confounding affected by prior exposure [@Evens_2008;@Morton_2014]. The HWSE, a key concern in occupational epidemiology, is the dynamic selection process by which healthier individuals remain at work, where they accumulate more occupational exposure, and less healthy individuals leave work [@Arrighi_1994]. Standard analyses of occupational cohorts affected by the HWSE result in a downward bias on the exposure-outcome associations of interest.

In the present analysis of the UAW-GM Cohort Study, we estimated NHL risk from 1985 to 2005 under hypothetical limits on average annual soluble MWF by applying the novel hazard-extended iterative conditional expectation (ICE) parametric g-formula estimator [@Wen_2020]. We leveraged time-varying quantitative MWF exposure data in tandem with employment data to adjust for the HWSE. In particular, we estimated the expected number of NHL cases we would observe if the hypothetical limits of 0.5, 0.25, and 0.05 mg/m^3^ were enforced for soluble MWF over workers' entire working lifetimes. Since workers were exposed to MWF of other types as well, we also evaluated the effect of dynamically reducing soluble MWF exposure with the goal of limiting total MWF exposure to 0.5 and 0.25 mg/m^3^.