### 2022-06-03 resources/exposure.tex

% Created by tikzDevice version 0.12.3.1 on 2022-04-22 17:18:12
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,180.67);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (433.62,180.68);
\end{scope}
\begin{scope}
\path[clip] ( 25.15, 28.21) rectangle (155.80,162.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 25.15, 28.21) rectangle (155.80,162.23);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15, 56.10) --
	(155.80, 56.10);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,101.47) --
	(155.80,101.47);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,146.83) --
	(155.80,146.83);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.63, 28.21) --
	( 48.63,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 75.63, 28.21) --
	( 75.63,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (102.62, 28.21) --
	(102.62,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.62, 28.21) --
	(129.62,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15, 33.42) --
	(155.80, 33.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15, 78.78) --
	(155.80, 78.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15,124.15) --
	(155.80,124.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.14, 28.21) --
	( 35.14,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.13, 28.21) --
	( 62.13,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.13, 28.21) --
	( 89.13,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.12, 28.21) --
	(116.12,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.12, 28.21) --
	(143.12,162.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.14,116.20) --
	( 36.49,154.33) --
	( 37.83,144.98) --
	( 39.18,118.42) --
	( 40.53,121.74) --
	( 41.88,121.74) --
	( 43.23,122.07) --
	( 44.58,133.91) --
	( 45.93,153.22) --
	( 47.28,154.33) --
	( 48.63,154.33) --
	( 49.98,156.14) --
	( 51.33,154.33) --
	( 52.68,133.91) --
	( 54.03,154.33) --
	( 55.38,156.14) --
	( 56.73,156.14) --
	( 58.08,156.14) --
	( 59.43,133.91) --
	( 60.78,154.33) --
	( 62.13,145.58) --
	( 63.48,141.88) --
	( 64.83,120.02) --
	( 66.18,133.91) --
	( 67.53,134.46) --
	( 68.88,122.01) --
	( 70.23, 94.47) --
	( 71.58,103.66) --
	( 72.93, 95.01) --
	( 74.28, 94.47) --
	( 75.63, 92.26) --
	( 76.98, 92.10) --
	( 78.33, 83.33) --
	( 79.68, 83.33) --
	( 81.03, 85.43) --
	( 82.38, 85.43) --
	( 83.73, 83.33) --
	( 85.08, 85.43) --
	( 86.43, 80.26) --
	( 87.78, 83.77) --
	( 89.13, 87.97) --
	( 90.48, 84.75) --
	( 91.83, 90.01) --
	( 93.18, 90.64) --
	( 94.53, 87.88) --
	( 95.87, 86.46) --
	( 97.22, 85.35) --
	( 98.57, 83.33) --
	( 99.92, 83.33) --
	(101.27, 85.19) --
	(102.62, 52.02) --
	(103.97, 52.30) --
	(105.32, 50.60) --
	(106.67, 50.94) --
	(108.02, 47.63) --
	(109.37, 48.18) --
	(110.72, 49.42) --
	(112.07, 49.79) --
	(113.42, 49.46) --
	(114.77, 49.79) --
	(116.12, 44.84) --
	(117.47, 46.82) --
	(118.82, 43.96) --
	(120.17, 44.33) --
	(121.52, 48.46) --
	(122.87, 46.75) --
	(124.22, 44.89) --
	(125.57, 44.67) --
	(126.92, 43.52) --
	(128.27, 43.38) --
	(129.62, 43.35) --
	(130.97, 42.97) --
	(132.32, 42.70) --
	(133.67, 41.95) --
	(135.02, 39.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 25.15, 28.21) rectangle (155.80,162.23);
\end{scope}
\begin{scope}
\path[clip] (161.30, 28.21) rectangle (291.96,162.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (161.30, 28.21) rectangle (291.96,162.23);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (161.30, 56.10) --
	(291.96, 56.10);

\path[draw=drawColor,line width= 0.3pt,line join=round] (161.30,101.47) --
	(291.96,101.47);

\path[draw=drawColor,line width= 0.3pt,line join=round] (161.30,146.83) --
	(291.96,146.83);

\path[draw=drawColor,line width= 0.3pt,line join=round] (184.79, 28.21) --
	(184.79,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (211.79, 28.21) --
	(211.79,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (238.78, 28.21) --
	(238.78,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (265.78, 28.21) --
	(265.78,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.30, 33.42) --
	(291.96, 33.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.30, 78.78) --
	(291.96, 78.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.30,124.15) --
	(291.96,124.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] (171.29, 28.21) --
	(171.29,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.29, 28.21) --
	(198.29,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.28, 28.21) --
	(225.28,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (252.28, 28.21) --
	(252.28,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.27, 28.21) --
	(279.27,162.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (171.29, 38.43) --
	(172.64, 41.00) --
	(173.99, 44.55) --
	(175.34, 41.44) --
	(176.69, 41.41) --
	(178.04, 43.59) --
	(179.39, 39.58) --
	(180.74, 45.48) --
	(182.09, 46.31) --
	(183.44, 48.44) --
	(184.79, 44.69) --
	(186.14, 48.44) --
	(187.49, 48.44) --
	(188.84, 40.74) --
	(190.19, 43.24) --
	(191.54, 44.57) --
	(192.89, 44.87) --
	(194.24, 48.44) --
	(195.59, 39.48) --
	(196.94, 48.44) --
	(198.29, 42.04) --
	(199.64, 51.78) --
	(200.99, 48.44) --
	(202.34, 51.09) --
	(203.69, 51.78) --
	(205.04, 47.41) --
	(206.39, 54.88) --
	(207.74, 57.99) --
	(209.09, 52.89) --
	(210.44, 53.09) --
	(211.79, 51.22) --
	(213.14, 43.75) --
	(214.49, 39.34) --
	(215.84, 38.74) --
	(217.19, 39.51) --
	(218.53, 39.79) --
	(219.88, 39.48) --
	(221.23, 39.48) --
	(222.58, 38.45) --
	(223.93, 40.70) --
	(225.28, 40.77) --
	(226.63, 39.48) --
	(227.98, 43.17) --
	(229.33, 43.78) --
	(230.68, 43.24) --
	(232.03, 43.81) --
	(233.38, 43.81) --
	(234.73, 43.81) --
	(236.08, 43.81) --
	(237.43, 43.81) --
	(238.78, 39.57) --
	(240.13, 40.14) --
	(241.48, 40.57) --
	(242.83, 39.42) --
	(244.18, 37.52) --
	(245.53, 37.52) --
	(246.88, 36.98) --
	(248.23, 36.49) --
	(249.58, 36.49) --
	(250.93, 36.49) --
	(252.28, 36.49) --
	(253.63, 36.49) --
	(254.98, 36.07) --
	(256.33, 36.38) --
	(257.68, 36.49) --
	(259.03, 36.49) --
	(260.38, 37.34) --
	(261.73, 37.85) --
	(263.08, 37.83) --
	(264.43, 37.00) --
	(265.78, 37.55) --
	(267.13, 37.79) --
	(268.48, 37.52) --
	(269.83, 37.68) --
	(271.18, 36.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (161.30, 28.21) rectangle (291.96,162.23);
\end{scope}
\begin{scope}
\path[clip] (297.46, 28.21) rectangle (428.12,162.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (297.46, 28.21) rectangle (428.12,162.23);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (297.46, 56.10) --
	(428.12, 56.10);

\path[draw=drawColor,line width= 0.3pt,line join=round] (297.46,101.47) --
	(428.12,101.47);

\path[draw=drawColor,line width= 0.3pt,line join=round] (297.46,146.83) --
	(428.12,146.83);

\path[draw=drawColor,line width= 0.3pt,line join=round] (320.95, 28.21) --
	(320.95,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (347.94, 28.21) --
	(347.94,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (374.94, 28.21) --
	(374.94,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (401.93, 28.21) --
	(401.93,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (297.46, 33.42) --
	(428.12, 33.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (297.46, 78.78) --
	(428.12, 78.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (297.46,124.15) --
	(428.12,124.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] (307.45, 28.21) --
	(307.45,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (334.45, 28.21) --
	(334.45,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.44, 28.21) --
	(361.44,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (388.44, 28.21) --
	(388.44,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (415.43, 28.21) --
	(415.43,162.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (308.80, 44.87) --
	(310.15, 44.87) --
	(311.50, 44.87) --
	(312.85, 42.13) --
	(314.20, 42.13) --
	(315.55, 44.87) --
	(316.90, 44.87) --
	(318.25, 44.27) --
	(319.60, 42.32) --
	(320.95, 48.60) --
	(322.30, 50.46) --
	(323.65, 69.90) --
	(325.00, 69.90) --
	(326.35, 69.90) --
	(327.70, 42.93) --
	(329.05, 42.93) --
	(330.40, 50.03) --
	(331.75,116.91) --
	(333.10,120.23) --
	(334.45, 42.93) --
	(335.80, 42.93) --
	(337.15, 64.53) --
	(338.50, 76.72) --
	(339.85, 64.53) --
	(341.19, 53.70) --
	(342.54, 64.53) --
	(343.89, 53.70) --
	(345.24, 43.76) --
	(346.59, 43.30) --
	(347.94, 40.78) --
	(349.29, 50.31) --
	(350.64, 51.13) --
	(351.99, 38.87) --
	(353.34, 38.87) --
	(354.69, 39.55) --
	(356.04, 38.90) --
	(357.39, 40.47) --
	(358.74, 38.87) --
	(360.09, 38.87) --
	(361.44, 38.93) --
	(362.79, 38.87) --
	(364.14, 38.87) --
	(365.49, 38.88) --
	(366.84, 39.40) --
	(368.19, 40.34) --
	(369.54, 39.02) --
	(370.89, 40.49) --
	(372.24, 40.05) --
	(373.59, 41.08) --
	(374.94, 35.04) --
	(376.29, 35.19) --
	(377.64, 35.19) --
	(378.99, 35.03) --
	(380.34, 34.30) --
	(381.69, 34.31) --
	(383.04, 34.53) --
	(384.39, 34.50) --
	(385.74, 34.43) --
	(387.09, 34.76) --
	(388.44, 34.80) --
	(389.79, 34.95) --
	(391.14, 34.71) --
	(392.49, 34.93) --
	(393.84, 35.19) --
	(395.19, 35.19) --
	(396.54, 36.02) --
	(397.89, 36.03) --
	(399.23, 36.02) --
	(400.58, 35.81) --
	(401.93, 36.04) --
	(403.28, 35.88) --
	(404.63, 35.23) --
	(405.98, 36.84) --
	(407.33, 35.58);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.46, 28.21) rectangle (428.12,162.23);
\end{scope}
\begin{scope}
\path[clip] ( 25.15,162.23) rectangle (155.80,175.17);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 25.15,162.23) rectangle (155.80,175.17);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 90.48,165.60) {Soluble};
\end{scope}
\begin{scope}
\path[clip] (161.30,162.23) rectangle (291.96,175.17);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (161.30,162.23) rectangle (291.96,175.17);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (226.63,165.60) {Straight};
\end{scope}
\begin{scope}
\path[clip] (297.46,162.23) rectangle (428.12,175.17);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (297.46,162.23) rectangle (428.12,175.17);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (362.79,165.60) {Synthetic};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.14, 25.46) --
	( 35.14, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.13, 25.46) --
	( 62.13, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.13, 25.46) --
	( 89.13, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.12, 25.46) --
	(116.12, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.12, 25.46) --
	(143.12, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 35.14, 17.75) {1920};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 62.13, 17.75) {1940};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 89.13, 17.75) {1960};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (116.12, 17.75) {1980};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (143.12, 17.75) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (171.29, 25.46) --
	(171.29, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.29, 25.46) --
	(198.29, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.28, 25.46) --
	(225.28, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (252.28, 25.46) --
	(252.28, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.27, 25.46) --
	(279.27, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (171.29, 17.75) {1920};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (198.29, 17.75) {1940};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (225.28, 17.75) {1960};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (252.28, 17.75) {1980};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (279.27, 17.75) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (307.45, 25.46) --
	(307.45, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (334.45, 25.46) --
	(334.45, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.44, 25.46) --
	(361.44, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (388.44, 25.46) --
	(388.44, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (415.43, 25.46) --
	(415.43, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (307.45, 17.75) {1920};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (334.45, 17.75) {1940};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (361.44, 17.75) {1960};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (388.44, 17.75) {1980};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (415.43, 17.75) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20, 30.66) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20, 76.03) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20,121.39) {2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40, 33.42) --
	( 25.15, 33.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40, 78.78) --
	( 25.15, 78.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,124.15) --
	( 25.15,124.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (226.63,  7.25) {Calendar year};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70, 95.22) {Median exposure (mg/m$^3$) among exposed};
\end{scope}
\end{tikzpicture}

\end{document}