### 2022-06-03 resources/exposure_long.tex

% Created by tikzDevice version 0.12.3.1 on 2022-05-11 14:49:57
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (216.81,397.48);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (216.81,397.48);
\end{scope}
\begin{scope}
\path[clip] ( 25.15,274.39) rectangle (211.31,379.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 25.15,274.39) rectangle (211.31,379.04);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,296.17) --
	(211.31,296.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,331.59) --
	(211.31,331.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,367.02) --
	(211.31,367.02);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 58.61,274.39) --
	( 58.61,379.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 97.07,274.39) --
	( 97.07,379.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (135.54,274.39) --
	(135.54,379.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (174.00,274.39) --
	(174.00,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15,278.46) --
	(211.31,278.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15,313.88) --
	(211.31,313.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15,349.30) --
	(211.31,349.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.38,274.39) --
	( 39.38,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 77.84,274.39) --
	( 77.84,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.31,274.39) --
	(116.31,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.77,274.39) --
	(154.77,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.23,274.39) --
	(193.23,379.04);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.38,343.10) --
	( 41.30,372.87) --
	( 43.22,365.57) --
	( 45.15,344.83) --
	( 47.07,347.42) --
	( 48.99,347.42) --
	( 50.92,347.68) --
	( 52.84,356.93) --
	( 54.76,372.01) --
	( 56.69,372.87) --
	( 58.61,372.87) --
	( 60.53,374.28) --
	( 62.46,372.87) --
	( 64.38,356.93) --
	( 66.30,372.87) --
	( 68.23,374.28) --
	( 70.15,374.28) --
	( 72.07,374.28) --
	( 74.00,356.93) --
	( 75.92,372.87) --
	( 77.84,366.04) --
	( 79.77,363.15) --
	( 81.69,346.08) --
	( 83.61,356.93) --
	( 85.53,357.36) --
	( 87.46,347.63) --
	( 89.38,326.13) --
	( 91.30,333.31) --
	( 93.23,326.55) --
	( 95.15,326.13) --
	( 97.07,324.41) --
	( 99.00,324.28) --
	(100.92,317.43) --
	(102.84,317.43) --
	(104.77,319.07) --
	(106.69,319.07) --
	(108.61,317.43) --
	(110.54,319.07) --
	(112.46,315.04) --
	(114.38,317.78) --
	(116.31,321.06) --
	(118.23,318.54) --
	(120.15,322.65) --
	(122.07,323.14) --
	(124.00,320.99) --
	(125.92,319.88) --
	(127.84,319.01) --
	(129.77,317.43) --
	(131.69,317.43) --
	(133.61,318.88) --
	(135.54,292.99) --
	(137.46,293.20) --
	(139.38,291.87) --
	(141.31,292.14) --
	(143.23,289.56) --
	(145.15,289.99) --
	(147.08,290.95) --
	(149.00,291.24) --
	(150.92,290.98) --
	(152.85,291.24) --
	(154.77,287.38) --
	(156.69,288.92) --
	(158.62,286.69) --
	(160.54,286.98) --
	(162.46,290.20) --
	(164.38,288.87) --
	(166.31,287.41) --
	(168.23,287.24) --
	(170.15,286.34) --
	(172.08,286.23) --
	(174.00,286.21) --
	(175.92,285.91) --
	(177.85,285.70) --
	(179.77,285.12) --
	(181.69,282.88);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 25.15,274.39) rectangle (211.31,379.04);
\end{scope}
\begin{scope}
\path[clip] ( 25.15,151.30) rectangle (211.31,255.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 25.15,151.30) rectangle (211.31,255.95);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,173.08) --
	(211.31,173.08);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,208.50) --
	(211.31,208.50);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,243.93) --
	(211.31,243.93);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 58.61,151.30) --
	( 58.61,255.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 97.07,151.30) --
	( 97.07,255.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (135.54,151.30) --
	(135.54,255.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (174.00,151.30) --
	(174.00,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15,155.37) --
	(211.31,155.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15,190.79) --
	(211.31,190.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15,226.21) --
	(211.31,226.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.38,151.30) --
	( 39.38,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 77.84,151.30) --
	( 77.84,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.31,151.30) --
	(116.31,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.77,151.30) --
	(154.77,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.23,151.30) --
	(193.23,255.95);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.38,159.28) --
	( 41.30,161.29) --
	( 43.22,164.06) --
	( 45.15,161.63) --
	( 47.07,161.61) --
	( 48.99,163.31) --
	( 50.92,160.18) --
	( 52.84,164.79) --
	( 54.76,165.43) --
	( 56.69,167.10) --
	( 58.61,164.17) --
	( 60.53,167.10) --
	( 62.46,167.10) --
	( 64.38,161.09) --
	( 66.30,163.03) --
	( 68.23,164.08) --
	( 70.15,164.31) --
	( 72.07,167.10) --
	( 74.00,160.10) --
	( 75.92,167.10) --
	( 77.84,162.10) --
	( 79.77,169.71) --
	( 81.69,167.10) --
	( 83.61,169.16) --
	( 85.53,169.71) --
	( 87.46,166.30) --
	( 89.38,172.13) --
	( 91.30,174.56) --
	( 93.23,170.57) --
	( 95.15,170.73) --
	( 97.07,169.26) --
	( 99.00,163.43) --
	(100.92,159.99) --
	(102.84,159.53) --
	(104.77,160.13) --
	(106.69,160.34) --
	(108.61,160.10) --
	(110.54,160.10) --
	(112.46,159.30) --
	(114.38,161.05) --
	(116.31,161.11) --
	(118.23,160.10) --
	(120.15,162.98) --
	(122.07,163.46) --
	(124.00,163.03) --
	(125.92,163.48) --
	(127.84,163.48) --
	(129.77,163.48) --
	(131.69,163.48) --
	(133.61,163.48) --
	(135.54,160.17) --
	(137.46,160.62) --
	(139.38,160.95) --
	(141.31,160.05) --
	(143.23,158.57) --
	(145.15,158.57) --
	(147.08,158.15) --
	(149.00,157.76) --
	(150.92,157.76) --
	(152.85,157.76) --
	(154.77,157.76) --
	(156.69,157.76) --
	(158.62,157.44) --
	(160.54,157.68) --
	(162.46,157.76) --
	(164.38,157.76) --
	(166.31,158.43) --
	(168.23,158.83) --
	(170.15,158.81) --
	(172.08,158.16) --
	(174.00,158.59) --
	(175.92,158.78) --
	(177.85,158.57) --
	(179.77,158.70) --
	(181.69,158.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 25.15,151.30) rectangle (211.31,255.95);
\end{scope}
\begin{scope}
\path[clip] ( 25.15, 28.21) rectangle (211.31,132.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 25.15, 28.21) rectangle (211.31,132.86);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15, 49.99) --
	(211.31, 49.99);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15, 85.41) --
	(211.31, 85.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 25.15,120.84) --
	(211.31,120.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 58.61, 28.21) --
	( 58.61,132.86);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 97.07, 28.21) --
	( 97.07,132.86);

\path[draw=drawColor,line width= 0.3pt,line join=round] (135.54, 28.21) --
	(135.54,132.86);

\path[draw=drawColor,line width= 0.3pt,line join=round] (174.00, 28.21) --
	(174.00,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15, 32.28) --
	(211.31, 32.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15, 67.70) --
	(211.31, 67.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15,103.12) --
	(211.31,103.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.38, 28.21) --
	( 39.38,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 77.84, 28.21) --
	( 77.84,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.31, 28.21) --
	(116.31,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.77, 28.21) --
	(154.77,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.23, 28.21) --
	(193.23,132.86);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.30, 41.22) --
	( 43.22, 41.22) --
	( 45.15, 41.22) --
	( 47.07, 39.08) --
	( 48.99, 39.08) --
	( 50.92, 41.22) --
	( 52.84, 41.22) --
	( 54.76, 40.75) --
	( 56.69, 39.23) --
	( 58.61, 44.13) --
	( 60.53, 45.59) --
	( 62.46, 60.76) --
	( 64.38, 60.76) --
	( 66.30, 60.76) --
	( 68.23, 39.70) --
	( 70.15, 39.70) --
	( 72.07, 45.25) --
	( 74.00, 97.47) --
	( 75.92,100.06) --
	( 77.84, 39.70) --
	( 79.77, 39.70) --
	( 81.69, 56.57) --
	( 83.61, 66.09) --
	( 85.53, 56.57) --
	( 87.46, 48.11) --
	( 89.38, 56.57) --
	( 91.30, 48.11) --
	( 93.23, 40.35) --
	( 95.15, 39.99) --
	( 97.07, 38.02) --
	( 99.00, 45.46) --
	(100.92, 46.11) --
	(102.84, 36.53) --
	(104.77, 36.53) --
	(106.69, 37.07) --
	(108.61, 36.56) --
	(110.54, 37.79) --
	(112.46, 36.53) --
	(114.38, 36.53) --
	(116.31, 36.58) --
	(118.23, 36.53) --
	(120.15, 36.53) --
	(122.07, 36.54) --
	(124.00, 36.95) --
	(125.92, 37.68) --
	(127.84, 36.65) --
	(129.77, 37.80) --
	(131.69, 37.46) --
	(133.61, 38.26) --
	(135.54, 33.55) --
	(137.46, 33.66) --
	(139.38, 33.66) --
	(141.31, 33.53) --
	(143.23, 32.97) --
	(145.15, 32.97) --
	(147.08, 33.14) --
	(149.00, 33.12) --
	(150.92, 33.07) --
	(152.85, 33.33) --
	(154.77, 33.36) --
	(156.69, 33.47) --
	(158.62, 33.29) --
	(160.54, 33.46) --
	(162.46, 33.66) --
	(164.38, 33.66) --
	(166.31, 34.31) --
	(168.23, 34.31) --
	(170.15, 34.31) --
	(172.08, 34.14) --
	(174.00, 34.32) --
	(175.92, 34.20) --
	(177.85, 33.69) --
	(179.77, 34.95) --
	(181.69, 33.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 25.15, 28.21) rectangle (211.31,132.86);
\end{scope}
\begin{scope}
\path[clip] ( 25.15,132.86) rectangle (211.31,145.80);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 25.15,132.86) rectangle (211.31,145.80);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (118.23,136.23) {Synthetic};
\end{scope}
\begin{scope}
\path[clip] ( 25.15,255.95) rectangle (211.31,268.89);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 25.15,255.95) rectangle (211.31,268.89);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (118.23,259.32) {Straight};
\end{scope}
\begin{scope}
\path[clip] ( 25.15,379.04) rectangle (211.31,391.99);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 25.15,379.04) rectangle (211.31,391.98);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (118.23,382.41) {Soluble};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.38, 25.46) --
	( 39.38, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 77.84, 25.46) --
	( 77.84, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.31, 25.46) --
	(116.31, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.77, 25.46) --
	(154.77, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.23, 25.46) --
	(193.23, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 39.38, 17.75) {1920};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 77.84, 17.75) {1940};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (116.31, 17.75) {1960};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (154.77, 17.75) {1980};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (193.23, 17.75) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20,275.70) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20,311.13) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20,346.55) {2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,278.46) --
	( 25.15,278.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,313.88) --
	( 25.15,313.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,349.30) --
	( 25.15,349.30);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20,152.61) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20,188.04) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20,223.46) {2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,155.37) --
	( 25.15,155.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,190.79) --
	( 25.15,190.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,226.21) --
	( 25.15,226.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20, 29.52) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20, 64.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20,100.37) {2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40, 32.28) --
	( 25.15, 32.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40, 67.70) --
	( 25.15, 67.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,103.12) --
	( 25.15,103.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (118.23,  7.25) {Calendar year};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,203.62) {Median exposure (mg/m$^3$) among exposed};
\end{scope}
\end{tikzpicture}

\end{document}