### 2022-06-03 resources/shift-soluble-dynamic.tex

% Created by tikzDevice version 0.12.3.1 on 2022-06-01 15:12:18
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,198.74);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,198.74);
\end{scope}
\begin{scope}
\path[clip] ( 31.37, 51.21) rectangle (283.58,193.24);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37, 51.21) rectangle (283.58,193.24);

\path[] ( 42.83, 57.70) --
	( 43.28, 57.69) --
	( 43.73, 57.69) --
	( 44.18, 57.69) --
	( 44.63, 57.68) --
	( 45.08, 57.68) --
	( 45.52, 57.67) --
	( 45.97, 57.67) --
	( 46.42, 57.67) --
	( 46.87, 57.67) --
	( 47.32, 57.67) --
	( 47.77, 57.67) --
	( 48.22, 57.67) --
	( 48.67, 57.67) --
	( 49.11, 57.67) --
	( 49.56, 57.67) --
	( 50.01, 57.67) --
	( 50.46, 57.67) --
	( 50.91, 57.67) --
	( 51.36, 57.67) --
	( 51.81, 57.67) --
	( 52.25, 57.67) --
	( 52.70, 57.67) --
	( 53.15, 57.67) --
	( 53.60, 57.67) --
	( 54.05, 57.67) --
	( 54.50, 57.67) --
	( 54.95, 57.67) --
	( 55.40, 57.67) --
	( 55.84, 57.67) --
	( 56.29, 57.67) --
	( 56.74, 57.67) --
	( 57.19, 57.67) --
	( 57.64, 57.67) --
	( 58.09, 57.67) --
	( 58.54, 57.67) --
	( 58.99, 57.67) --
	( 59.43, 57.67) --
	( 59.88, 57.67) --
	( 60.33, 57.67) --
	( 60.78, 57.67) --
	( 61.23, 57.67) --
	( 61.68, 57.67) --
	( 62.13, 57.67) --
	( 62.57, 57.67) --
	( 63.02, 57.67) --
	( 63.47, 57.67) --
	( 63.92, 57.67) --
	( 64.37, 57.67) --
	( 64.82, 57.67) --
	( 65.27, 57.67) --
	( 65.72, 57.67) --
	( 66.16, 57.67) --
	( 66.61, 57.67) --
	( 67.06, 57.67) --
	( 67.51, 57.67) --
	( 67.96, 57.67) --
	( 68.41, 57.67) --
	( 68.86, 57.67) --
	( 69.31, 57.67) --
	( 69.75, 57.67) --
	( 70.20, 57.67) --
	( 70.65, 57.67) --
	( 71.10, 57.67) --
	( 71.55, 57.67) --
	( 72.00, 57.67) --
	( 72.45, 57.67) --
	( 72.89, 57.67) --
	( 73.34, 57.67) --
	( 73.79, 57.67) --
	( 74.24, 57.67) --
	( 74.69, 57.67) --
	( 75.14, 57.67) --
	( 75.59, 57.67) --
	( 76.04, 57.67) --
	( 76.48, 57.67) --
	( 76.93, 57.67) --
	( 77.38, 57.67) --
	( 77.83, 57.67) --
	( 78.28, 57.67) --
	( 78.73, 57.67) --
	( 79.18, 57.67) --
	( 79.63, 57.67) --
	( 80.07, 57.67) --
	( 80.52, 57.67) --
	( 80.97, 57.67) --
	( 81.42, 57.67) --
	( 81.87, 57.67) --
	( 82.32, 57.67) --
	( 82.77, 57.67) --
	( 83.21, 57.67) --
	( 83.66, 57.67) --
	( 84.11, 57.67) --
	( 84.56, 57.67) --
	( 85.01, 57.67) --
	( 85.46, 57.67) --
	( 85.91, 57.67) --
	( 86.36, 57.67) --
	( 86.80, 57.67) --
	( 87.25, 57.67) --
	( 87.70, 57.67) --
	( 88.15, 57.67) --
	( 88.60, 57.67) --
	( 89.05, 57.67) --
	( 89.50, 57.67) --
	( 89.95, 57.67) --
	( 90.39, 57.67) --
	( 90.84, 57.67) --
	( 91.29, 57.67) --
	( 91.74, 57.67) --
	( 92.19, 57.67) --
	( 92.64, 57.67) --
	( 93.09, 57.67) --
	( 93.53, 57.67) --
	( 93.98, 57.67) --
	( 94.43, 57.67) --
	( 94.88, 57.67) --
	( 95.33, 57.67) --
	( 95.78, 57.67) --
	( 96.23, 57.67) --
	( 96.68, 57.67) --
	( 97.12, 57.67) --
	( 97.57, 57.67) --
	( 98.02, 57.67) --
	( 98.47, 57.67) --
	( 98.92, 57.68) --
	( 99.37, 57.68) --
	( 99.82, 57.69) --
	(100.27, 57.69) --
	(100.71, 57.69) --
	(101.16, 57.69) --
	(101.61, 57.69) --
	(102.06, 57.69) --
	(102.51, 57.69) --
	(102.96, 57.68) --
	(103.41, 57.68) --
	(103.85, 57.68) --
	(104.30, 57.67) --
	(104.75, 57.67) --
	(105.20, 57.67) --
	(105.65, 57.68) --
	(106.10, 57.68) --
	(106.55, 57.69) --
	(107.00, 57.69) --
	(107.44, 57.70) --
	(107.89, 57.70) --
	(108.34, 57.71) --
	(108.79, 57.71) --
	(109.24, 57.71) --
	(109.69, 57.71) --
	(110.14, 57.71) --
	(110.59, 57.71) --
	(111.03, 57.72) --
	(111.48, 57.72) --
	(111.93, 57.73) --
	(112.38, 57.74) --
	(112.83, 57.75) --
	(113.28, 57.76) --
	(113.73, 57.76) --
	(114.17, 57.77) --
	(114.62, 57.78) --
	(115.07, 57.78) --
	(115.52, 57.78) --
	(115.97, 57.79) --
	(116.42, 57.79) --
	(116.87, 57.79) --
	(117.32, 57.80) --
	(117.76, 57.81) --
	(118.21, 57.82) --
	(118.66, 57.83) --
	(119.11, 57.85) --
	(119.56, 57.86) --
	(120.01, 57.87) --
	(120.46, 57.88) --
	(120.91, 57.88) --
	(121.35, 57.88) --
	(121.80, 57.87) --
	(122.25, 57.87) --
	(122.70, 57.87) --
	(123.15, 57.87) --
	(123.60, 57.88) --
	(124.05, 57.89) --
	(124.49, 57.91) --
	(124.94, 57.92) --
	(125.39, 57.92) --
	(125.84, 57.93) --
	(126.29, 57.93) --
	(126.74, 57.94) --
	(127.19, 57.95) --
	(127.64, 57.96) --
	(128.08, 57.97) --
	(128.53, 57.98) --
	(128.98, 58.00) --
	(129.43, 58.01) --
	(129.88, 58.03) --
	(130.33, 58.05) --
	(130.78, 58.08) --
	(131.23, 58.11) --
	(131.67, 58.15) --
	(132.12, 58.19) --
	(132.57, 58.23) --
	(133.02, 58.27) --
	(133.47, 58.30) --
	(133.92, 58.33) --
	(134.37, 58.35) --
	(134.81, 58.36) --
	(135.26, 58.37) --
	(135.71, 58.39) --
	(136.16, 58.40) --
	(136.61, 58.42) --
	(137.06, 58.45) --
	(137.51, 58.50) --
	(137.96, 58.56) --
	(138.40, 58.62) --
	(138.85, 58.70) --
	(139.30, 58.78) --
	(139.75, 58.87) --
	(140.20, 58.94) --
	(140.65, 59.00) --
	(141.10, 59.03) --
	(141.55, 59.03) --
	(141.99, 59.01) --
	(142.44, 58.98) --
	(142.89, 58.94) --
	(143.34, 58.92) --
	(143.79, 58.93) --
	(144.24, 58.97) --
	(144.69, 59.04) --
	(145.13, 59.13) --
	(145.58, 59.24) --
	(146.03, 59.36) --
	(146.48, 59.46) --
	(146.93, 59.56) --
	(147.38, 59.65) --
	(147.83, 59.73) --
	(148.28, 59.80) --
	(148.72, 59.88) --
	(149.17, 59.97) --
	(149.62, 60.07) --
	(150.07, 60.17) --
	(150.52, 60.28) --
	(150.97, 60.39) --
	(151.42, 60.51) --
	(151.87, 60.62) --
	(152.31, 60.74) --
	(152.76, 60.84) --
	(153.21, 60.93) --
	(153.66, 61.00) --
	(154.11, 61.03) --
	(154.56, 61.04) --
	(155.01, 61.01) --
	(155.45, 60.97) --
	(155.90, 60.93) --
	(156.35, 60.91) --
	(156.80, 60.94) --
	(157.25, 61.02) --
	(157.70, 61.15) --
	(158.15, 61.31) --
	(158.60, 61.49) --
	(159.04, 61.66) --
	(159.49, 61.80) --
	(159.94, 61.91) --
	(160.39, 61.99) --
	(160.84, 62.04) --
	(161.29, 62.09) --
	(161.74, 62.14) --
	(162.19, 62.21) --
	(162.63, 62.33) --
	(163.08, 62.48) --
	(163.53, 62.66) --
	(163.98, 62.86) --
	(164.43, 63.07) --
	(164.88, 63.27) --
	(165.33, 63.47) --
	(165.77, 63.65) --
	(166.22, 63.83) --
	(166.67, 64.01) --
	(167.12, 64.20) --
	(167.57, 64.40) --
	(168.02, 64.62) --
	(168.47, 64.84) --
	(168.92, 65.06) --
	(169.36, 65.29) --
	(169.81, 65.51) --
	(170.26, 65.72) --
	(170.71, 65.91) --
	(171.16, 66.07) --
	(171.61, 66.20) --
	(172.06, 66.29) --
	(172.51, 66.34) --
	(172.95, 66.37) --
	(173.40, 66.40) --
	(173.85, 66.44) --
	(174.30, 66.53) --
	(174.75, 66.68) --
	(175.20, 66.91) --
	(175.65, 67.22) --
	(176.09, 67.58) --
	(176.54, 67.99) --
	(176.99, 68.40) --
	(177.44, 68.83) --
	(177.89, 69.26) --
	(178.34, 69.71) --
	(178.79, 70.20) --
	(179.24, 70.76) --
	(179.68, 71.47) --
	(180.13, 72.40) --
	(180.58, 73.62) --
	(181.03, 75.20) --
	(181.48, 77.17) --
	(181.93, 79.44) --
	(182.38, 81.84) --
	(182.83, 84.13) --
	(183.27, 86.00) --
	(183.72, 87.21) --
	(184.17, 87.63) --
	(184.62, 87.40) --
	(185.07, 86.74) --
	(185.52, 85.93) --
	(185.97, 85.22) --
	(186.41, 84.76) --
	(186.86, 84.61) --
	(187.31, 84.77) --
	(187.76, 85.24) --
	(188.21, 86.01) --
	(188.66, 87.13) --
	(189.11, 88.65) --
	(189.56, 90.65) --
	(190.00, 93.13) --
	(190.45, 96.00) --
	(190.90, 99.05) --
	(191.35,101.96) --
	(191.80,104.39) --
	(192.25,106.10) --
	(192.70,106.97) --
	(193.15,107.20) --
	(193.59,107.13) --
	(194.04,107.15) --
	(194.49,107.59) --
	(194.94,108.62) --
	(195.39,110.24) --
	(195.84,112.29) --
	(196.29,114.53) --
	(196.73,116.73) --
	(197.18,118.74) --
	(197.63,120.57) --
	(198.08,122.36) --
	(198.53,124.29) --
	(198.98,126.50) --
	(199.43,128.98) --
	(199.88,131.59) --
	(200.32,134.10) --
	(200.77,136.29) --
	(201.22,138.06) --
	(201.67,139.44) --
	(202.12,140.59) --
	(202.57,141.61) --
	(203.02,142.55) --
	(203.47,143.40) --
	(203.91,144.16) --
	(204.36,144.83) --
	(204.81,145.43) --
	(205.26,146.01) --
	(205.71,146.57) --
	(206.16,147.10) --
	(206.61,147.59) --
	(207.05,148.05) --
	(207.50,148.47) --
	(207.95,148.88) --
	(208.40,149.30) --
	(208.85,149.72) --
	(209.30,150.15) --
	(209.75,150.55) --
	(210.20,150.92) --
	(210.64,151.26) --
	(211.09,151.62) --
	(211.54,152.06) --
	(211.99,152.66) --
	(212.44,153.43) --
	(212.89,154.35) --
	(213.34,155.30) --
	(213.79,156.12) --
	(214.23,156.59) --
	(214.68,156.52) --
	(215.13,155.77) --
	(215.58,154.29) --
	(216.03,152.16) --
	(216.48,149.49) --
	(216.93,146.47) --
	(217.37,143.23) --
	(217.82,139.88) --
	(218.27,136.49) --
	(218.72,133.09) --
	(219.17,129.74) --
	(219.62,126.52) --
	(220.07,123.54) --
	(220.52,120.89) --
	(220.96,118.60) --
	(221.41,116.64) --
	(221.86,114.91) --
	(222.31,113.25) --
	(222.76,111.46) --
	(223.21,109.40) --
	(223.66,106.93) --
	(224.11,103.99) --
	(224.55,100.54) --
	(225.00, 96.66) --
	(225.45, 92.45) --
	(225.90, 88.06) --
	(226.35, 83.70) --
	(226.80, 79.52) --
	(227.25, 75.68) --
	(227.69, 72.29) --
	(228.14, 69.34) --
	(228.59, 66.83) --
	(229.04, 64.72) --
	(229.49, 62.97) --
	(229.94, 61.55) --
	(230.39, 60.41) --
	(230.84, 59.53) --
	(231.28, 58.88) --
	(231.73, 58.42) --
	(232.18, 58.11) --
	(232.63, 57.92) --
	(233.08, 57.80) --
	(233.53, 57.73) --
	(233.98, 57.70) --
	(234.43, 57.68) --
	(234.87, 57.67) --
	(235.32, 57.67) --
	(235.77, 57.67) --
	(236.22, 57.67) --
	(236.67, 57.67) --
	(237.12, 57.67) --
	(237.57, 57.67) --
	(238.01, 57.67) --
	(238.46, 57.67) --
	(238.91, 57.67) --
	(239.36, 57.67) --
	(239.81, 57.67) --
	(240.26, 57.67) --
	(240.71, 57.67) --
	(241.16, 57.67) --
	(241.60, 57.67) --
	(242.05, 57.67) --
	(242.50, 57.67) --
	(242.95, 57.67) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67) --
	(272.12, 57.67) --
	(271.67, 57.67) --
	(271.22, 57.67) --
	(270.77, 57.67) --
	(270.32, 57.67) --
	(269.87, 57.67) --
	(269.42, 57.67) --
	(268.97, 57.67) --
	(268.53, 57.67) --
	(268.08, 57.67) --
	(267.63, 57.67) --
	(267.18, 57.67) --
	(266.73, 57.67) --
	(266.28, 57.67) --
	(265.83, 57.67) --
	(265.39, 57.67) --
	(264.94, 57.67) --
	(264.49, 57.67) --
	(264.04, 57.67) --
	(263.59, 57.67) --
	(263.14, 57.67) --
	(262.69, 57.67) --
	(262.24, 57.67) --
	(261.80, 57.67) --
	(261.35, 57.67) --
	(260.90, 57.67) --
	(260.45, 57.67) --
	(260.00, 57.67) --
	(259.55, 57.67) --
	(259.10, 57.67) --
	(258.65, 57.67) --
	(258.21, 57.67) --
	(257.76, 57.67) --
	(257.31, 57.67) --
	(256.86, 57.67) --
	(256.41, 57.67) --
	(255.96, 57.67) --
	(255.51, 57.67) --
	(255.07, 57.67) --
	(254.62, 57.67) --
	(254.17, 57.67) --
	(253.72, 57.67) --
	(253.27, 57.67) --
	(252.82, 57.67) --
	(252.37, 57.67) --
	(251.92, 57.67) --
	(251.48, 57.67) --
	(251.03, 57.67) --
	(250.58, 57.67) --
	(250.13, 57.67) --
	(249.68, 57.67) --
	(249.23, 57.67) --
	(248.78, 57.67) --
	(248.33, 57.67) --
	(247.89, 57.67) --
	(247.44, 57.67) --
	(246.99, 57.67) --
	(246.54, 57.67) --
	(246.09, 57.67) --
	(245.64, 57.67) --
	(245.19, 57.67) --
	(244.75, 57.67) --
	(244.30, 57.67) --
	(243.85, 57.67) --
	(243.40, 57.67) --
	(242.95, 57.67) --
	(242.50, 57.67) --
	(242.05, 57.67) --
	(241.60, 57.67) --
	(241.16, 57.67) --
	(240.71, 57.67) --
	(240.26, 57.67) --
	(239.81, 57.67) --
	(239.36, 57.67) --
	(238.91, 57.67) --
	(238.46, 57.67) --
	(238.01, 57.67) --
	(237.57, 57.67) --
	(237.12, 57.67) --
	(236.67, 57.67) --
	(236.22, 57.67) --
	(235.77, 57.67) --
	(235.32, 57.67) --
	(234.87, 57.67) --
	(234.43, 57.67) --
	(233.98, 57.67) --
	(233.53, 57.67) --
	(233.08, 57.67) --
	(232.63, 57.67) --
	(232.18, 57.67) --
	(231.73, 57.67) --
	(231.28, 57.67) --
	(230.84, 57.67) --
	(230.39, 57.67) --
	(229.94, 57.67) --
	(229.49, 57.67) --
	(229.04, 57.67) --
	(228.59, 57.67) --
	(228.14, 57.67) --
	(227.69, 57.67) --
	(227.25, 57.67) --
	(226.80, 57.67) --
	(226.35, 57.67) --
	(225.90, 57.67) --
	(225.45, 57.67) --
	(225.00, 57.67) --
	(224.55, 57.67) --
	(224.11, 57.67) --
	(223.66, 57.67) --
	(223.21, 57.67) --
	(222.76, 57.67) --
	(222.31, 57.67) --
	(221.86, 57.67) --
	(221.41, 57.67) --
	(220.96, 57.67) --
	(220.52, 57.67) --
	(220.07, 57.67) --
	(219.62, 57.67) --
	(219.17, 57.67) --
	(218.72, 57.67) --
	(218.27, 57.67) --
	(217.82, 57.67) --
	(217.37, 57.67) --
	(216.93, 57.67) --
	(216.48, 57.67) --
	(216.03, 57.67) --
	(215.58, 57.67) --
	(215.13, 57.67) --
	(214.68, 57.67) --
	(214.23, 57.67) --
	(213.79, 57.67) --
	(213.34, 57.67) --
	(212.89, 57.67) --
	(212.44, 57.67) --
	(211.99, 57.67) --
	(211.54, 57.67) --
	(211.09, 57.67) --
	(210.64, 57.67) --
	(210.20, 57.67) --
	(209.75, 57.67) --
	(209.30, 57.67) --
	(208.85, 57.67) --
	(208.40, 57.67) --
	(207.95, 57.67) --
	(207.50, 57.67) --
	(207.05, 57.67) --
	(206.61, 57.67) --
	(206.16, 57.67) --
	(205.71, 57.67) --
	(205.26, 57.67) --
	(204.81, 57.67) --
	(204.36, 57.67) --
	(203.91, 57.67) --
	(203.47, 57.67) --
	(203.02, 57.67) --
	(202.57, 57.67) --
	(202.12, 57.67) --
	(201.67, 57.67) --
	(201.22, 57.67) --
	(200.77, 57.67) --
	(200.32, 57.67) --
	(199.88, 57.67) --
	(199.43, 57.67) --
	(198.98, 57.67) --
	(198.53, 57.67) --
	(198.08, 57.67) --
	(197.63, 57.67) --
	(197.18, 57.67) --
	(196.73, 57.67) --
	(196.29, 57.67) --
	(195.84, 57.67) --
	(195.39, 57.67) --
	(194.94, 57.67) --
	(194.49, 57.67) --
	(194.04, 57.67) --
	(193.59, 57.67) --
	(193.15, 57.67) --
	(192.70, 57.67) --
	(192.25, 57.67) --
	(191.80, 57.67) --
	(191.35, 57.67) --
	(190.90, 57.67) --
	(190.45, 57.67) --
	(190.00, 57.67) --
	(189.56, 57.67) --
	(189.11, 57.67) --
	(188.66, 57.67) --
	(188.21, 57.67) --
	(187.76, 57.67) --
	(187.31, 57.67) --
	(186.86, 57.67) --
	(186.41, 57.67) --
	(185.97, 57.67) --
	(185.52, 57.67) --
	(185.07, 57.67) --
	(184.62, 57.67) --
	(184.17, 57.67) --
	(183.72, 57.67) --
	(183.27, 57.67) --
	(182.83, 57.67) --
	(182.38, 57.67) --
	(181.93, 57.67) --
	(181.48, 57.67) --
	(181.03, 57.67) --
	(180.58, 57.67) --
	(180.13, 57.67) --
	(179.68, 57.67) --
	(179.24, 57.67) --
	(178.79, 57.67) --
	(178.34, 57.67) --
	(177.89, 57.67) --
	(177.44, 57.67) --
	(176.99, 57.67) --
	(176.54, 57.67) --
	(176.09, 57.67) --
	(175.65, 57.67) --
	(175.20, 57.67) --
	(174.75, 57.67) --
	(174.30, 57.67) --
	(173.85, 57.67) --
	(173.40, 57.67) --
	(172.95, 57.67) --
	(172.51, 57.67) --
	(172.06, 57.67) --
	(171.61, 57.67) --
	(171.16, 57.67) --
	(170.71, 57.67) --
	(170.26, 57.67) --
	(169.81, 57.67) --
	(169.36, 57.67) --
	(168.92, 57.67) --
	(168.47, 57.67) --
	(168.02, 57.67) --
	(167.57, 57.67) --
	(167.12, 57.67) --
	(166.67, 57.67) --
	(166.22, 57.67) --
	(165.77, 57.67) --
	(165.33, 57.67) --
	(164.88, 57.67) --
	(164.43, 57.67) --
	(163.98, 57.67) --
	(163.53, 57.67) --
	(163.08, 57.67) --
	(162.63, 57.67) --
	(162.19, 57.67) --
	(161.74, 57.67) --
	(161.29, 57.67) --
	(160.84, 57.67) --
	(160.39, 57.67) --
	(159.94, 57.67) --
	(159.49, 57.67) --
	(159.04, 57.67) --
	(158.60, 57.67) --
	(158.15, 57.67) --
	(157.70, 57.67) --
	(157.25, 57.67) --
	(156.80, 57.67) --
	(156.35, 57.67) --
	(155.90, 57.67) --
	(155.45, 57.67) --
	(155.01, 57.67) --
	(154.56, 57.67) --
	(154.11, 57.67) --
	(153.66, 57.67) --
	(153.21, 57.67) --
	(152.76, 57.67) --
	(152.31, 57.67) --
	(151.87, 57.67) --
	(151.42, 57.67) --
	(150.97, 57.67) --
	(150.52, 57.67) --
	(150.07, 57.67) --
	(149.62, 57.67) --
	(149.17, 57.67) --
	(148.72, 57.67) --
	(148.28, 57.67) --
	(147.83, 57.67) --
	(147.38, 57.67) --
	(146.93, 57.67) --
	(146.48, 57.67) --
	(146.03, 57.67) --
	(145.58, 57.67) --
	(145.13, 57.67) --
	(144.69, 57.67) --
	(144.24, 57.67) --
	(143.79, 57.67) --
	(143.34, 57.67) --
	(142.89, 57.67) --
	(142.44, 57.67) --
	(141.99, 57.67) --
	(141.55, 57.67) --
	(141.10, 57.67) --
	(140.65, 57.67) --
	(140.20, 57.67) --
	(139.75, 57.67) --
	(139.30, 57.67) --
	(138.85, 57.67) --
	(138.40, 57.67) --
	(137.96, 57.67) --
	(137.51, 57.67) --
	(137.06, 57.67) --
	(136.61, 57.67) --
	(136.16, 57.67) --
	(135.71, 57.67) --
	(135.26, 57.67) --
	(134.81, 57.67) --
	(134.37, 57.67) --
	(133.92, 57.67) --
	(133.47, 57.67) --
	(133.02, 57.67) --
	(132.57, 57.67) --
	(132.12, 57.67) --
	(131.67, 57.67) --
	(131.23, 57.67) --
	(130.78, 57.67) --
	(130.33, 57.67) --
	(129.88, 57.67) --
	(129.43, 57.67) --
	(128.98, 57.67) --
	(128.53, 57.67) --
	(128.08, 57.67) --
	(127.64, 57.67) --
	(127.19, 57.67) --
	(126.74, 57.67) --
	(126.29, 57.67) --
	(125.84, 57.67) --
	(125.39, 57.67) --
	(124.94, 57.67) --
	(124.49, 57.67) --
	(124.05, 57.67) --
	(123.60, 57.67) --
	(123.15, 57.67) --
	(122.70, 57.67) --
	(122.25, 57.67) --
	(121.80, 57.67) --
	(121.35, 57.67) --
	(120.91, 57.67) --
	(120.46, 57.67) --
	(120.01, 57.67) --
	(119.56, 57.67) --
	(119.11, 57.67) --
	(118.66, 57.67) --
	(118.21, 57.67) --
	(117.76, 57.67) --
	(117.32, 57.67) --
	(116.87, 57.67) --
	(116.42, 57.67) --
	(115.97, 57.67) --
	(115.52, 57.67) --
	(115.07, 57.67) --
	(114.62, 57.67) --
	(114.17, 57.67) --
	(113.73, 57.67) --
	(113.28, 57.67) --
	(112.83, 57.67) --
	(112.38, 57.67) --
	(111.93, 57.67) --
	(111.48, 57.67) --
	(111.03, 57.67) --
	(110.59, 57.67) --
	(110.14, 57.67) --
	(109.69, 57.67) --
	(109.24, 57.67) --
	(108.79, 57.67) --
	(108.34, 57.67) --
	(107.89, 57.67) --
	(107.44, 57.67) --
	(107.00, 57.67) --
	(106.55, 57.67) --
	(106.10, 57.67) --
	(105.65, 57.67) --
	(105.20, 57.67) --
	(104.75, 57.67) --
	(104.30, 57.67) --
	(103.85, 57.67) --
	(103.41, 57.67) --
	(102.96, 57.67) --
	(102.51, 57.67) --
	(102.06, 57.67) --
	(101.61, 57.67) --
	(101.16, 57.67) --
	(100.71, 57.67) --
	(100.27, 57.67) --
	( 99.82, 57.67) --
	( 99.37, 57.67) --
	( 98.92, 57.67) --
	( 98.47, 57.67) --
	( 98.02, 57.67) --
	( 97.57, 57.67) --
	( 97.12, 57.67) --
	( 96.68, 57.67) --
	( 96.23, 57.67) --
	( 95.78, 57.67) --
	( 95.33, 57.67) --
	( 94.88, 57.67) --
	( 94.43, 57.67) --
	( 93.98, 57.67) --
	( 93.53, 57.67) --
	( 93.09, 57.67) --
	( 92.64, 57.67) --
	( 92.19, 57.67) --
	( 91.74, 57.67) --
	( 91.29, 57.67) --
	( 90.84, 57.67) --
	( 90.39, 57.67) --
	( 89.95, 57.67) --
	( 89.50, 57.67) --
	( 89.05, 57.67) --
	( 88.60, 57.67) --
	( 88.15, 57.67) --
	( 87.70, 57.67) --
	( 87.25, 57.67) --
	( 86.80, 57.67) --
	( 86.36, 57.67) --
	( 85.91, 57.67) --
	( 85.46, 57.67) --
	( 85.01, 57.67) --
	( 84.56, 57.67) --
	( 84.11, 57.67) --
	( 83.66, 57.67) --
	( 83.21, 57.67) --
	( 82.77, 57.67) --
	( 82.32, 57.67) --
	( 81.87, 57.67) --
	( 81.42, 57.67) --
	( 80.97, 57.67) --
	( 80.52, 57.67) --
	( 80.07, 57.67) --
	( 79.63, 57.67) --
	( 79.18, 57.67) --
	( 78.73, 57.67) --
	( 78.28, 57.67) --
	( 77.83, 57.67) --
	( 77.38, 57.67) --
	( 76.93, 57.67) --
	( 76.48, 57.67) --
	( 76.04, 57.67) --
	( 75.59, 57.67) --
	( 75.14, 57.67) --
	( 74.69, 57.67) --
	( 74.24, 57.67) --
	( 73.79, 57.67) --
	( 73.34, 57.67) --
	( 72.89, 57.67) --
	( 72.45, 57.67) --
	( 72.00, 57.67) --
	( 71.55, 57.67) --
	( 71.10, 57.67) --
	( 70.65, 57.67) --
	( 70.20, 57.67) --
	( 69.75, 57.67) --
	( 69.31, 57.67) --
	( 68.86, 57.67) --
	( 68.41, 57.67) --
	( 67.96, 57.67) --
	( 67.51, 57.67) --
	( 67.06, 57.67) --
	( 66.61, 57.67) --
	( 66.16, 57.67) --
	( 65.72, 57.67) --
	( 65.27, 57.67) --
	( 64.82, 57.67) --
	( 64.37, 57.67) --
	( 63.92, 57.67) --
	( 63.47, 57.67) --
	( 63.02, 57.67) --
	( 62.57, 57.67) --
	( 62.13, 57.67) --
	( 61.68, 57.67) --
	( 61.23, 57.67) --
	( 60.78, 57.67) --
	( 60.33, 57.67) --
	( 59.88, 57.67) --
	( 59.43, 57.67) --
	( 58.99, 57.67) --
	( 58.54, 57.67) --
	( 58.09, 57.67) --
	( 57.64, 57.67) --
	( 57.19, 57.67) --
	( 56.74, 57.67) --
	( 56.29, 57.67) --
	( 55.84, 57.67) --
	( 55.40, 57.67) --
	( 54.95, 57.67) --
	( 54.50, 57.67) --
	( 54.05, 57.67) --
	( 53.60, 57.67) --
	( 53.15, 57.67) --
	( 52.70, 57.67) --
	( 52.25, 57.67) --
	( 51.81, 57.67) --
	( 51.36, 57.67) --
	( 50.91, 57.67) --
	( 50.46, 57.67) --
	( 50.01, 57.67) --
	( 49.56, 57.67) --
	( 49.11, 57.67) --
	( 48.67, 57.67) --
	( 48.22, 57.67) --
	( 47.77, 57.67) --
	( 47.32, 57.67) --
	( 46.87, 57.67) --
	( 46.42, 57.67) --
	( 45.97, 57.67) --
	( 45.52, 57.67) --
	( 45.08, 57.67) --
	( 44.63, 57.67) --
	( 44.18, 57.67) --
	( 43.73, 57.67) --
	( 43.28, 57.67) --
	( 42.83, 57.67) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 42.83, 57.70) --
	( 43.28, 57.69) --
	( 43.73, 57.69) --
	( 44.18, 57.69) --
	( 44.63, 57.68) --
	( 45.08, 57.68) --
	( 45.52, 57.67) --
	( 45.97, 57.67) --
	( 46.42, 57.67) --
	( 46.87, 57.67) --
	( 47.32, 57.67) --
	( 47.77, 57.67) --
	( 48.22, 57.67) --
	( 48.67, 57.67) --
	( 49.11, 57.67) --
	( 49.56, 57.67) --
	( 50.01, 57.67) --
	( 50.46, 57.67) --
	( 50.91, 57.67) --
	( 51.36, 57.67) --
	( 51.81, 57.67) --
	( 52.25, 57.67) --
	( 52.70, 57.67) --
	( 53.15, 57.67) --
	( 53.60, 57.67) --
	( 54.05, 57.67) --
	( 54.50, 57.67) --
	( 54.95, 57.67) --
	( 55.40, 57.67) --
	( 55.84, 57.67) --
	( 56.29, 57.67) --
	( 56.74, 57.67) --
	( 57.19, 57.67) --
	( 57.64, 57.67) --
	( 58.09, 57.67) --
	( 58.54, 57.67) --
	( 58.99, 57.67) --
	( 59.43, 57.67) --
	( 59.88, 57.67) --
	( 60.33, 57.67) --
	( 60.78, 57.67) --
	( 61.23, 57.67) --
	( 61.68, 57.67) --
	( 62.13, 57.67) --
	( 62.57, 57.67) --
	( 63.02, 57.67) --
	( 63.47, 57.67) --
	( 63.92, 57.67) --
	( 64.37, 57.67) --
	( 64.82, 57.67) --
	( 65.27, 57.67) --
	( 65.72, 57.67) --
	( 66.16, 57.67) --
	( 66.61, 57.67) --
	( 67.06, 57.67) --
	( 67.51, 57.67) --
	( 67.96, 57.67) --
	( 68.41, 57.67) --
	( 68.86, 57.67) --
	( 69.31, 57.67) --
	( 69.75, 57.67) --
	( 70.20, 57.67) --
	( 70.65, 57.67) --
	( 71.10, 57.67) --
	( 71.55, 57.67) --
	( 72.00, 57.67) --
	( 72.45, 57.67) --
	( 72.89, 57.67) --
	( 73.34, 57.67) --
	( 73.79, 57.67) --
	( 74.24, 57.67) --
	( 74.69, 57.67) --
	( 75.14, 57.67) --
	( 75.59, 57.67) --
	( 76.04, 57.67) --
	( 76.48, 57.67) --
	( 76.93, 57.67) --
	( 77.38, 57.67) --
	( 77.83, 57.67) --
	( 78.28, 57.67) --
	( 78.73, 57.67) --
	( 79.18, 57.67) --
	( 79.63, 57.67) --
	( 80.07, 57.67) --
	( 80.52, 57.67) --
	( 80.97, 57.67) --
	( 81.42, 57.67) --
	( 81.87, 57.67) --
	( 82.32, 57.67) --
	( 82.77, 57.67) --
	( 83.21, 57.67) --
	( 83.66, 57.67) --
	( 84.11, 57.67) --
	( 84.56, 57.67) --
	( 85.01, 57.67) --
	( 85.46, 57.67) --
	( 85.91, 57.67) --
	( 86.36, 57.67) --
	( 86.80, 57.67) --
	( 87.25, 57.67) --
	( 87.70, 57.67) --
	( 88.15, 57.67) --
	( 88.60, 57.67) --
	( 89.05, 57.67) --
	( 89.50, 57.67) --
	( 89.95, 57.67) --
	( 90.39, 57.67) --
	( 90.84, 57.67) --
	( 91.29, 57.67) --
	( 91.74, 57.67) --
	( 92.19, 57.67) --
	( 92.64, 57.67) --
	( 93.09, 57.67) --
	( 93.53, 57.67) --
	( 93.98, 57.67) --
	( 94.43, 57.67) --
	( 94.88, 57.67) --
	( 95.33, 57.67) --
	( 95.78, 57.67) --
	( 96.23, 57.67) --
	( 96.68, 57.67) --
	( 97.12, 57.67) --
	( 97.57, 57.67) --
	( 98.02, 57.67) --
	( 98.47, 57.67) --
	( 98.92, 57.68) --
	( 99.37, 57.68) --
	( 99.82, 57.69) --
	(100.27, 57.69) --
	(100.71, 57.69) --
	(101.16, 57.69) --
	(101.61, 57.69) --
	(102.06, 57.69) --
	(102.51, 57.69) --
	(102.96, 57.68) --
	(103.41, 57.68) --
	(103.85, 57.68) --
	(104.30, 57.67) --
	(104.75, 57.67) --
	(105.20, 57.67) --
	(105.65, 57.68) --
	(106.10, 57.68) --
	(106.55, 57.69) --
	(107.00, 57.69) --
	(107.44, 57.70) --
	(107.89, 57.70) --
	(108.34, 57.71) --
	(108.79, 57.71) --
	(109.24, 57.71) --
	(109.69, 57.71) --
	(110.14, 57.71) --
	(110.59, 57.71) --
	(111.03, 57.72) --
	(111.48, 57.72) --
	(111.93, 57.73) --
	(112.38, 57.74) --
	(112.83, 57.75) --
	(113.28, 57.76) --
	(113.73, 57.76) --
	(114.17, 57.77) --
	(114.62, 57.78) --
	(115.07, 57.78) --
	(115.52, 57.78) --
	(115.97, 57.79) --
	(116.42, 57.79) --
	(116.87, 57.79) --
	(117.32, 57.80) --
	(117.76, 57.81) --
	(118.21, 57.82) --
	(118.66, 57.83) --
	(119.11, 57.85) --
	(119.56, 57.86) --
	(120.01, 57.87) --
	(120.46, 57.88) --
	(120.91, 57.88) --
	(121.35, 57.88) --
	(121.80, 57.87) --
	(122.25, 57.87) --
	(122.70, 57.87) --
	(123.15, 57.87) --
	(123.60, 57.88) --
	(124.05, 57.89) --
	(124.49, 57.91) --
	(124.94, 57.92) --
	(125.39, 57.92) --
	(125.84, 57.93) --
	(126.29, 57.93) --
	(126.74, 57.94) --
	(127.19, 57.95) --
	(127.64, 57.96) --
	(128.08, 57.97) --
	(128.53, 57.98) --
	(128.98, 58.00) --
	(129.43, 58.01) --
	(129.88, 58.03) --
	(130.33, 58.05) --
	(130.78, 58.08) --
	(131.23, 58.11) --
	(131.67, 58.15) --
	(132.12, 58.19) --
	(132.57, 58.23) --
	(133.02, 58.27) --
	(133.47, 58.30) --
	(133.92, 58.33) --
	(134.37, 58.35) --
	(134.81, 58.36) --
	(135.26, 58.37) --
	(135.71, 58.39) --
	(136.16, 58.40) --
	(136.61, 58.42) --
	(137.06, 58.45) --
	(137.51, 58.50) --
	(137.96, 58.56) --
	(138.40, 58.62) --
	(138.85, 58.70) --
	(139.30, 58.78) --
	(139.75, 58.87) --
	(140.20, 58.94) --
	(140.65, 59.00) --
	(141.10, 59.03) --
	(141.55, 59.03) --
	(141.99, 59.01) --
	(142.44, 58.98) --
	(142.89, 58.94) --
	(143.34, 58.92) --
	(143.79, 58.93) --
	(144.24, 58.97) --
	(144.69, 59.04) --
	(145.13, 59.13) --
	(145.58, 59.24) --
	(146.03, 59.36) --
	(146.48, 59.46) --
	(146.93, 59.56) --
	(147.38, 59.65) --
	(147.83, 59.73) --
	(148.28, 59.80) --
	(148.72, 59.88) --
	(149.17, 59.97) --
	(149.62, 60.07) --
	(150.07, 60.17) --
	(150.52, 60.28) --
	(150.97, 60.39) --
	(151.42, 60.51) --
	(151.87, 60.62) --
	(152.31, 60.74) --
	(152.76, 60.84) --
	(153.21, 60.93) --
	(153.66, 61.00) --
	(154.11, 61.03) --
	(154.56, 61.04) --
	(155.01, 61.01) --
	(155.45, 60.97) --
	(155.90, 60.93) --
	(156.35, 60.91) --
	(156.80, 60.94) --
	(157.25, 61.02) --
	(157.70, 61.15) --
	(158.15, 61.31) --
	(158.60, 61.49) --
	(159.04, 61.66) --
	(159.49, 61.80) --
	(159.94, 61.91) --
	(160.39, 61.99) --
	(160.84, 62.04) --
	(161.29, 62.09) --
	(161.74, 62.14) --
	(162.19, 62.21) --
	(162.63, 62.33) --
	(163.08, 62.48) --
	(163.53, 62.66) --
	(163.98, 62.86) --
	(164.43, 63.07) --
	(164.88, 63.27) --
	(165.33, 63.47) --
	(165.77, 63.65) --
	(166.22, 63.83) --
	(166.67, 64.01) --
	(167.12, 64.20) --
	(167.57, 64.40) --
	(168.02, 64.62) --
	(168.47, 64.84) --
	(168.92, 65.06) --
	(169.36, 65.29) --
	(169.81, 65.51) --
	(170.26, 65.72) --
	(170.71, 65.91) --
	(171.16, 66.07) --
	(171.61, 66.20) --
	(172.06, 66.29) --
	(172.51, 66.34) --
	(172.95, 66.37) --
	(173.40, 66.40) --
	(173.85, 66.44) --
	(174.30, 66.53) --
	(174.75, 66.68) --
	(175.20, 66.91) --
	(175.65, 67.22) --
	(176.09, 67.58) --
	(176.54, 67.99) --
	(176.99, 68.40) --
	(177.44, 68.83) --
	(177.89, 69.26) --
	(178.34, 69.71) --
	(178.79, 70.20) --
	(179.24, 70.76) --
	(179.68, 71.47) --
	(180.13, 72.40) --
	(180.58, 73.62) --
	(181.03, 75.20) --
	(181.48, 77.17) --
	(181.93, 79.44) --
	(182.38, 81.84) --
	(182.83, 84.13) --
	(183.27, 86.00) --
	(183.72, 87.21) --
	(184.17, 87.63) --
	(184.62, 87.40) --
	(185.07, 86.74) --
	(185.52, 85.93) --
	(185.97, 85.22) --
	(186.41, 84.76) --
	(186.86, 84.61) --
	(187.31, 84.77) --
	(187.76, 85.24) --
	(188.21, 86.01) --
	(188.66, 87.13) --
	(189.11, 88.65) --
	(189.56, 90.65) --
	(190.00, 93.13) --
	(190.45, 96.00) --
	(190.90, 99.05) --
	(191.35,101.96) --
	(191.80,104.39) --
	(192.25,106.10) --
	(192.70,106.97) --
	(193.15,107.20) --
	(193.59,107.13) --
	(194.04,107.15) --
	(194.49,107.59) --
	(194.94,108.62) --
	(195.39,110.24) --
	(195.84,112.29) --
	(196.29,114.53) --
	(196.73,116.73) --
	(197.18,118.74) --
	(197.63,120.57) --
	(198.08,122.36) --
	(198.53,124.29) --
	(198.98,126.50) --
	(199.43,128.98) --
	(199.88,131.59) --
	(200.32,134.10) --
	(200.77,136.29) --
	(201.22,138.06) --
	(201.67,139.44) --
	(202.12,140.59) --
	(202.57,141.61) --
	(203.02,142.55) --
	(203.47,143.40) --
	(203.91,144.16) --
	(204.36,144.83) --
	(204.81,145.43) --
	(205.26,146.01) --
	(205.71,146.57) --
	(206.16,147.10) --
	(206.61,147.59) --
	(207.05,148.05) --
	(207.50,148.47) --
	(207.95,148.88) --
	(208.40,149.30) --
	(208.85,149.72) --
	(209.30,150.15) --
	(209.75,150.55) --
	(210.20,150.92) --
	(210.64,151.26) --
	(211.09,151.62) --
	(211.54,152.06) --
	(211.99,152.66) --
	(212.44,153.43) --
	(212.89,154.35) --
	(213.34,155.30) --
	(213.79,156.12) --
	(214.23,156.59) --
	(214.68,156.52) --
	(215.13,155.77) --
	(215.58,154.29) --
	(216.03,152.16) --
	(216.48,149.49) --
	(216.93,146.47) --
	(217.37,143.23) --
	(217.82,139.88) --
	(218.27,136.49) --
	(218.72,133.09) --
	(219.17,129.74) --
	(219.62,126.52) --
	(220.07,123.54) --
	(220.52,120.89) --
	(220.96,118.60) --
	(221.41,116.64) --
	(221.86,114.91) --
	(222.31,113.25) --
	(222.76,111.46) --
	(223.21,109.40) --
	(223.66,106.93) --
	(224.11,103.99) --
	(224.55,100.54) --
	(225.00, 96.66) --
	(225.45, 92.45) --
	(225.90, 88.06) --
	(226.35, 83.70) --
	(226.80, 79.52) --
	(227.25, 75.68) --
	(227.69, 72.29) --
	(228.14, 69.34) --
	(228.59, 66.83) --
	(229.04, 64.72) --
	(229.49, 62.97) --
	(229.94, 61.55) --
	(230.39, 60.41) --
	(230.84, 59.53) --
	(231.28, 58.88) --
	(231.73, 58.42) --
	(232.18, 58.11) --
	(232.63, 57.92) --
	(233.08, 57.80) --
	(233.53, 57.73) --
	(233.98, 57.70) --
	(234.43, 57.68) --
	(234.87, 57.67) --
	(235.32, 57.67) --
	(235.77, 57.67) --
	(236.22, 57.67) --
	(236.67, 57.67) --
	(237.12, 57.67) --
	(237.57, 57.67) --
	(238.01, 57.67) --
	(238.46, 57.67) --
	(238.91, 57.67) --
	(239.36, 57.67) --
	(239.81, 57.67) --
	(240.26, 57.67) --
	(240.71, 57.67) --
	(241.16, 57.67) --
	(241.60, 57.67) --
	(242.05, 57.67) --
	(242.50, 57.67) --
	(242.95, 57.67) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67);

\path[] ( 42.83, 57.67) --
	( 43.28, 57.67) --
	( 43.73, 57.67) --
	( 44.18, 57.67) --
	( 44.63, 57.67) --
	( 45.08, 57.67) --
	( 45.52, 57.67) --
	( 45.97, 57.67) --
	( 46.42, 57.67) --
	( 46.87, 57.67) --
	( 47.32, 57.67) --
	( 47.77, 57.67) --
	( 48.22, 57.67) --
	( 48.67, 57.67) --
	( 49.11, 57.67) --
	( 49.56, 57.67) --
	( 50.01, 57.67) --
	( 50.46, 57.67) --
	( 50.91, 57.67) --
	( 51.36, 57.67) --
	( 51.81, 57.67) --
	( 52.25, 57.67) --
	( 52.70, 57.67) --
	( 53.15, 57.67) --
	( 53.60, 57.67) --
	( 54.05, 57.67) --
	( 54.50, 57.67) --
	( 54.95, 57.67) --
	( 55.40, 57.67) --
	( 55.84, 57.67) --
	( 56.29, 57.67) --
	( 56.74, 57.67) --
	( 57.19, 57.67) --
	( 57.64, 57.67) --
	( 58.09, 57.67) --
	( 58.54, 57.67) --
	( 58.99, 57.67) --
	( 59.43, 57.67) --
	( 59.88, 57.67) --
	( 60.33, 57.67) --
	( 60.78, 57.67) --
	( 61.23, 57.67) --
	( 61.68, 57.67) --
	( 62.13, 57.67) --
	( 62.57, 57.67) --
	( 63.02, 57.67) --
	( 63.47, 57.67) --
	( 63.92, 57.67) --
	( 64.37, 57.67) --
	( 64.82, 57.67) --
	( 65.27, 57.67) --
	( 65.72, 57.67) --
	( 66.16, 57.67) --
	( 66.61, 57.67) --
	( 67.06, 57.67) --
	( 67.51, 57.67) --
	( 67.96, 57.67) --
	( 68.41, 57.67) --
	( 68.86, 57.67) --
	( 69.31, 57.67) --
	( 69.75, 57.67) --
	( 70.20, 57.67) --
	( 70.65, 57.67) --
	( 71.10, 57.67) --
	( 71.55, 57.67) --
	( 72.00, 57.67) --
	( 72.45, 57.67) --
	( 72.89, 57.67) --
	( 73.34, 57.67) --
	( 73.79, 57.67) --
	( 74.24, 57.67) --
	( 74.69, 57.67) --
	( 75.14, 57.67) --
	( 75.59, 57.67) --
	( 76.04, 57.67) --
	( 76.48, 57.67) --
	( 76.93, 57.67) --
	( 77.38, 57.67) --
	( 77.83, 57.67) --
	( 78.28, 57.67) --
	( 78.73, 57.67) --
	( 79.18, 57.67) --
	( 79.63, 57.67) --
	( 80.07, 57.67) --
	( 80.52, 57.67) --
	( 80.97, 57.67) --
	( 81.42, 57.67) --
	( 81.87, 57.67) --
	( 82.32, 57.67) --
	( 82.77, 57.67) --
	( 83.21, 57.67) --
	( 83.66, 57.67) --
	( 84.11, 57.67) --
	( 84.56, 57.67) --
	( 85.01, 57.67) --
	( 85.46, 57.67) --
	( 85.91, 57.67) --
	( 86.36, 57.67) --
	( 86.80, 57.67) --
	( 87.25, 57.67) --
	( 87.70, 57.67) --
	( 88.15, 57.67) --
	( 88.60, 57.67) --
	( 89.05, 57.67) --
	( 89.50, 57.67) --
	( 89.95, 57.67) --
	( 90.39, 57.67) --
	( 90.84, 57.67) --
	( 91.29, 57.67) --
	( 91.74, 57.67) --
	( 92.19, 57.67) --
	( 92.64, 57.67) --
	( 93.09, 57.67) --
	( 93.53, 57.67) --
	( 93.98, 57.68) --
	( 94.43, 57.68) --
	( 94.88, 57.68) --
	( 95.33, 57.69) --
	( 95.78, 57.69) --
	( 96.23, 57.69) --
	( 96.68, 57.69) --
	( 97.12, 57.69) --
	( 97.57, 57.69) --
	( 98.02, 57.69) --
	( 98.47, 57.68) --
	( 98.92, 57.68) --
	( 99.37, 57.67) --
	( 99.82, 57.67) --
	(100.27, 57.67) --
	(100.71, 57.67) --
	(101.16, 57.67) --
	(101.61, 57.67) --
	(102.06, 57.67) --
	(102.51, 57.67) --
	(102.96, 57.67) --
	(103.41, 57.67) --
	(103.85, 57.67) --
	(104.30, 57.67) --
	(104.75, 57.67) --
	(105.20, 57.67) --
	(105.65, 57.68) --
	(106.10, 57.68) --
	(106.55, 57.69) --
	(107.00, 57.69) --
	(107.44, 57.70) --
	(107.89, 57.70) --
	(108.34, 57.70) --
	(108.79, 57.71) --
	(109.24, 57.71) --
	(109.69, 57.71) --
	(110.14, 57.71) --
	(110.59, 57.71) --
	(111.03, 57.72) --
	(111.48, 57.72) --
	(111.93, 57.73) --
	(112.38, 57.74) --
	(112.83, 57.74) --
	(113.28, 57.75) --
	(113.73, 57.76) --
	(114.17, 57.77) --
	(114.62, 57.77) --
	(115.07, 57.77) --
	(115.52, 57.77) --
	(115.97, 57.77) --
	(116.42, 57.77) --
	(116.87, 57.77) --
	(117.32, 57.77) --
	(117.76, 57.78) --
	(118.21, 57.78) --
	(118.66, 57.79) --
	(119.11, 57.81) --
	(119.56, 57.82) --
	(120.01, 57.83) --
	(120.46, 57.84) --
	(120.91, 57.85) --
	(121.35, 57.85) --
	(121.80, 57.85) --
	(122.25, 57.85) --
	(122.70, 57.85) --
	(123.15, 57.85) --
	(123.60, 57.86) --
	(124.05, 57.86) --
	(124.49, 57.87) --
	(124.94, 57.88) --
	(125.39, 57.89) --
	(125.84, 57.89) --
	(126.29, 57.90) --
	(126.74, 57.92) --
	(127.19, 57.94) --
	(127.64, 57.97) --
	(128.08, 58.01) --
	(128.53, 58.05) --
	(128.98, 58.09) --
	(129.43, 58.13) --
	(129.88, 58.15) --
	(130.33, 58.17) --
	(130.78, 58.18) --
	(131.23, 58.19) --
	(131.67, 58.20) --
	(132.12, 58.21) --
	(132.57, 58.23) --
	(133.02, 58.25) --
	(133.47, 58.28) --
	(133.92, 58.30) --
	(134.37, 58.33) --
	(134.81, 58.36) --
	(135.26, 58.38) --
	(135.71, 58.41) --
	(136.16, 58.43) --
	(136.61, 58.44) --
	(137.06, 58.45) --
	(137.51, 58.47) --
	(137.96, 58.49) --
	(138.40, 58.52) --
	(138.85, 58.57) --
	(139.30, 58.64) --
	(139.75, 58.73) --
	(140.20, 58.82) --
	(140.65, 58.90) --
	(141.10, 58.95) --
	(141.55, 58.97) --
	(141.99, 58.97) --
	(142.44, 58.94) --
	(142.89, 58.91) --
	(143.34, 58.89) --
	(143.79, 58.88) --
	(144.24, 58.90) --
	(144.69, 58.95) --
	(145.13, 59.01) --
	(145.58, 59.08) --
	(146.03, 59.16) --
	(146.48, 59.24) --
	(146.93, 59.33) --
	(147.38, 59.42) --
	(147.83, 59.53) --
	(148.28, 59.63) --
	(148.72, 59.73) --
	(149.17, 59.82) --
	(149.62, 59.90) --
	(150.07, 59.97) --
	(150.52, 60.03) --
	(150.97, 60.10) --
	(151.42, 60.18) --
	(151.87, 60.27) --
	(152.31, 60.36) --
	(152.76, 60.46) --
	(153.21, 60.54) --
	(153.66, 60.59) --
	(154.11, 60.60) --
	(154.56, 60.58) --
	(155.01, 60.53) --
	(155.45, 60.46) --
	(155.90, 60.40) --
	(156.35, 60.36) --
	(156.80, 60.36) --
	(157.25, 60.40) --
	(157.70, 60.48) --
	(158.15, 60.59) --
	(158.60, 60.72) --
	(159.04, 60.85) --
	(159.49, 60.98) --
	(159.94, 61.09) --
	(160.39, 61.18) --
	(160.84, 61.26) --
	(161.29, 61.34) --
	(161.74, 61.42) --
	(162.19, 61.51) --
	(162.63, 61.62) --
	(163.08, 61.76) --
	(163.53, 61.92) --
	(163.98, 62.09) --
	(164.43, 62.25) --
	(164.88, 62.38) --
	(165.33, 62.49) --
	(165.77, 62.55) --
	(166.22, 62.59) --
	(166.67, 62.60) --
	(167.12, 62.62) --
	(167.57, 62.66) --
	(168.02, 62.71) --
	(168.47, 62.80) --
	(168.92, 62.91) --
	(169.36, 63.03) --
	(169.81, 63.17) --
	(170.26, 63.32) --
	(170.71, 63.48) --
	(171.16, 63.64) --
	(171.61, 63.80) --
	(172.06, 63.95) --
	(172.51, 64.10) --
	(172.95, 64.23) --
	(173.40, 64.33) --
	(173.85, 64.42) --
	(174.30, 64.52) --
	(174.75, 64.63) --
	(175.20, 64.77) --
	(175.65, 64.98) --
	(176.09, 65.23) --
	(176.54, 65.54) --
	(176.99, 65.87) --
	(177.44, 66.22) --
	(177.89, 66.55) --
	(178.34, 66.86) --
	(178.79, 67.12) --
	(179.24, 67.36) --
	(179.68, 67.57) --
	(180.13, 67.79) --
	(180.58, 68.05) --
	(181.03, 68.36) --
	(181.48, 68.73) --
	(181.93, 69.16) --
	(182.38, 69.63) --
	(182.83, 70.13) --
	(183.27, 70.62) --
	(183.72, 71.07) --
	(184.17, 71.46) --
	(184.62, 71.77) --
	(185.07, 72.00) --
	(185.52, 72.16) --
	(185.97, 72.27) --
	(186.41, 72.38) --
	(186.86, 72.51) --
	(187.31, 72.72) --
	(187.76, 73.06) --
	(188.21, 73.59) --
	(188.66, 74.34) --
	(189.11, 75.38) --
	(189.56, 76.73) --
	(190.00, 78.37) --
	(190.45, 80.24) --
	(190.90, 82.21) --
	(191.35, 84.10) --
	(191.80, 85.72) --
	(192.25, 86.91) --
	(192.70, 87.61) --
	(193.15, 87.88) --
	(193.59, 87.88) --
	(194.04, 87.79) --
	(194.49, 87.74) --
	(194.94, 87.84) --
	(195.39, 88.13) --
	(195.84, 88.59) --
	(196.29, 89.21) --
	(196.73, 89.97) --
	(197.18, 90.90) --
	(197.63, 92.03) --
	(198.08, 93.42) --
	(198.53, 95.10) --
	(198.98, 97.07) --
	(199.43, 99.25) --
	(199.88,101.51) --
	(200.32,103.68) --
	(200.77,105.60) --
	(201.22,107.22) --
	(201.67,108.57) --
	(202.12,109.78) --
	(202.57,111.05) --
	(203.02,112.51) --
	(203.47,114.27) --
	(203.91,116.32) --
	(204.36,118.62) --
	(204.81,121.05) --
	(205.26,123.48) --
	(205.71,125.84) --
	(206.16,128.09) --
	(206.61,130.27) --
	(207.05,132.48) --
	(207.50,134.74) --
	(207.95,137.07) --
	(208.40,139.36) --
	(208.85,141.51) --
	(209.30,143.40) --
	(209.75,145.01) --
	(210.20,146.39) --
	(210.64,147.64) --
	(211.09,148.86) --
	(211.54,150.09) --
	(211.99,151.28) --
	(212.44,152.40) --
	(212.89,153.38) --
	(213.34,154.16) --
	(213.79,154.72) --
	(214.23,155.06) --
	(214.68,155.21) --
	(215.13,155.24) --
	(215.58,155.20) --
	(216.03,155.11) --
	(216.48,154.98) --
	(216.93,154.77) --
	(217.37,154.41) --
	(217.82,153.86) --
	(218.27,153.09) --
	(218.72,152.08) --
	(219.17,150.85) --
	(219.62,149.46) --
	(220.07,147.95) --
	(220.52,146.36) --
	(220.96,144.71) --
	(221.41,143.03) --
	(221.86,141.30) --
	(222.31,139.53) --
	(222.76,137.70) --
	(223.21,135.84) --
	(223.66,133.95) --
	(224.11,132.04) --
	(224.55,130.11) --
	(225.00,128.15) --
	(225.45,126.16) --
	(225.90,124.13) --
	(226.35,122.06) --
	(226.80,119.97) --
	(227.25,117.88) --
	(227.69,115.85) --
	(228.14,113.92) --
	(228.59,112.14) --
	(229.04,110.54) --
	(229.49,109.08) --
	(229.94,107.70) --
	(230.39,106.31) --
	(230.84,104.79) --
	(231.28,103.03) --
	(231.73,100.95) --
	(232.18, 98.50) --
	(232.63, 95.66) --
	(233.08, 92.47) --
	(233.53, 89.01) --
	(233.98, 85.40) --
	(234.43, 81.77) --
	(234.87, 78.25) --
	(235.32, 74.97) --
	(235.77, 72.00) --
	(236.22, 69.38) --
	(236.67, 67.11) --
	(237.12, 65.15) --
	(237.57, 63.48) --
	(238.01, 62.06) --
	(238.46, 60.89) --
	(238.91, 59.94) --
	(239.36, 59.21) --
	(239.81, 58.66) --
	(240.26, 58.28) --
	(240.71, 58.02) --
	(241.16, 57.86) --
	(241.60, 57.77) --
	(242.05, 57.72) --
	(242.50, 57.69) --
	(242.95, 57.68) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67) --
	(272.12, 57.67) --
	(271.67, 57.67) --
	(271.22, 57.67) --
	(270.77, 57.67) --
	(270.32, 57.67) --
	(269.87, 57.67) --
	(269.42, 57.67) --
	(268.97, 57.67) --
	(268.53, 57.67) --
	(268.08, 57.67) --
	(267.63, 57.67) --
	(267.18, 57.67) --
	(266.73, 57.67) --
	(266.28, 57.67) --
	(265.83, 57.67) --
	(265.39, 57.67) --
	(264.94, 57.67) --
	(264.49, 57.67) --
	(264.04, 57.67) --
	(263.59, 57.67) --
	(263.14, 57.67) --
	(262.69, 57.67) --
	(262.24, 57.67) --
	(261.80, 57.67) --
	(261.35, 57.67) --
	(260.90, 57.67) --
	(260.45, 57.67) --
	(260.00, 57.67) --
	(259.55, 57.67) --
	(259.10, 57.67) --
	(258.65, 57.67) --
	(258.21, 57.67) --
	(257.76, 57.67) --
	(257.31, 57.67) --
	(256.86, 57.67) --
	(256.41, 57.67) --
	(255.96, 57.67) --
	(255.51, 57.67) --
	(255.07, 57.67) --
	(254.62, 57.67) --
	(254.17, 57.67) --
	(253.72, 57.67) --
	(253.27, 57.67) --
	(252.82, 57.67) --
	(252.37, 57.67) --
	(251.92, 57.67) --
	(251.48, 57.67) --
	(251.03, 57.67) --
	(250.58, 57.67) --
	(250.13, 57.67) --
	(249.68, 57.67) --
	(249.23, 57.67) --
	(248.78, 57.67) --
	(248.33, 57.67) --
	(247.89, 57.67) --
	(247.44, 57.67) --
	(246.99, 57.67) --
	(246.54, 57.67) --
	(246.09, 57.67) --
	(245.64, 57.67) --
	(245.19, 57.67) --
	(244.75, 57.67) --
	(244.30, 57.67) --
	(243.85, 57.67) --
	(243.40, 57.67) --
	(242.95, 57.67) --
	(242.50, 57.67) --
	(242.05, 57.67) --
	(241.60, 57.67) --
	(241.16, 57.67) --
	(240.71, 57.67) --
	(240.26, 57.67) --
	(239.81, 57.67) --
	(239.36, 57.67) --
	(238.91, 57.67) --
	(238.46, 57.67) --
	(238.01, 57.67) --
	(237.57, 57.67) --
	(237.12, 57.67) --
	(236.67, 57.67) --
	(236.22, 57.67) --
	(235.77, 57.67) --
	(235.32, 57.67) --
	(234.87, 57.67) --
	(234.43, 57.67) --
	(233.98, 57.67) --
	(233.53, 57.67) --
	(233.08, 57.67) --
	(232.63, 57.67) --
	(232.18, 57.67) --
	(231.73, 57.67) --
	(231.28, 57.67) --
	(230.84, 57.67) --
	(230.39, 57.67) --
	(229.94, 57.67) --
	(229.49, 57.67) --
	(229.04, 57.67) --
	(228.59, 57.67) --
	(228.14, 57.67) --
	(227.69, 57.67) --
	(227.25, 57.67) --
	(226.80, 57.67) --
	(226.35, 57.67) --
	(225.90, 57.67) --
	(225.45, 57.67) --
	(225.00, 57.67) --
	(224.55, 57.67) --
	(224.11, 57.67) --
	(223.66, 57.67) --
	(223.21, 57.67) --
	(222.76, 57.67) --
	(222.31, 57.67) --
	(221.86, 57.67) --
	(221.41, 57.67) --
	(220.96, 57.67) --
	(220.52, 57.67) --
	(220.07, 57.67) --
	(219.62, 57.67) --
	(219.17, 57.67) --
	(218.72, 57.67) --
	(218.27, 57.67) --
	(217.82, 57.67) --
	(217.37, 57.67) --
	(216.93, 57.67) --
	(216.48, 57.67) --
	(216.03, 57.67) --
	(215.58, 57.67) --
	(215.13, 57.67) --
	(214.68, 57.67) --
	(214.23, 57.67) --
	(213.79, 57.67) --
	(213.34, 57.67) --
	(212.89, 57.67) --
	(212.44, 57.67) --
	(211.99, 57.67) --
	(211.54, 57.67) --
	(211.09, 57.67) --
	(210.64, 57.67) --
	(210.20, 57.67) --
	(209.75, 57.67) --
	(209.30, 57.67) --
	(208.85, 57.67) --
	(208.40, 57.67) --
	(207.95, 57.67) --
	(207.50, 57.67) --
	(207.05, 57.67) --
	(206.61, 57.67) --
	(206.16, 57.67) --
	(205.71, 57.67) --
	(205.26, 57.67) --
	(204.81, 57.67) --
	(204.36, 57.67) --
	(203.91, 57.67) --
	(203.47, 57.67) --
	(203.02, 57.67) --
	(202.57, 57.67) --
	(202.12, 57.67) --
	(201.67, 57.67) --
	(201.22, 57.67) --
	(200.77, 57.67) --
	(200.32, 57.67) --
	(199.88, 57.67) --
	(199.43, 57.67) --
	(198.98, 57.67) --
	(198.53, 57.67) --
	(198.08, 57.67) --
	(197.63, 57.67) --
	(197.18, 57.67) --
	(196.73, 57.67) --
	(196.29, 57.67) --
	(195.84, 57.67) --
	(195.39, 57.67) --
	(194.94, 57.67) --
	(194.49, 57.67) --
	(194.04, 57.67) --
	(193.59, 57.67) --
	(193.15, 57.67) --
	(192.70, 57.67) --
	(192.25, 57.67) --
	(191.80, 57.67) --
	(191.35, 57.67) --
	(190.90, 57.67) --
	(190.45, 57.67) --
	(190.00, 57.67) --
	(189.56, 57.67) --
	(189.11, 57.67) --
	(188.66, 57.67) --
	(188.21, 57.67) --
	(187.76, 57.67) --
	(187.31, 57.67) --
	(186.86, 57.67) --
	(186.41, 57.67) --
	(185.97, 57.67) --
	(185.52, 57.67) --
	(185.07, 57.67) --
	(184.62, 57.67) --
	(184.17, 57.67) --
	(183.72, 57.67) --
	(183.27, 57.67) --
	(182.83, 57.67) --
	(182.38, 57.67) --
	(181.93, 57.67) --
	(181.48, 57.67) --
	(181.03, 57.67) --
	(180.58, 57.67) --
	(180.13, 57.67) --
	(179.68, 57.67) --
	(179.24, 57.67) --
	(178.79, 57.67) --
	(178.34, 57.67) --
	(177.89, 57.67) --
	(177.44, 57.67) --
	(176.99, 57.67) --
	(176.54, 57.67) --
	(176.09, 57.67) --
	(175.65, 57.67) --
	(175.20, 57.67) --
	(174.75, 57.67) --
	(174.30, 57.67) --
	(173.85, 57.67) --
	(173.40, 57.67) --
	(172.95, 57.67) --
	(172.51, 57.67) --
	(172.06, 57.67) --
	(171.61, 57.67) --
	(171.16, 57.67) --
	(170.71, 57.67) --
	(170.26, 57.67) --
	(169.81, 57.67) --
	(169.36, 57.67) --
	(168.92, 57.67) --
	(168.47, 57.67) --
	(168.02, 57.67) --
	(167.57, 57.67) --
	(167.12, 57.67) --
	(166.67, 57.67) --
	(166.22, 57.67) --
	(165.77, 57.67) --
	(165.33, 57.67) --
	(164.88, 57.67) --
	(164.43, 57.67) --
	(163.98, 57.67) --
	(163.53, 57.67) --
	(163.08, 57.67) --
	(162.63, 57.67) --
	(162.19, 57.67) --
	(161.74, 57.67) --
	(161.29, 57.67) --
	(160.84, 57.67) --
	(160.39, 57.67) --
	(159.94, 57.67) --
	(159.49, 57.67) --
	(159.04, 57.67) --
	(158.60, 57.67) --
	(158.15, 57.67) --
	(157.70, 57.67) --
	(157.25, 57.67) --
	(156.80, 57.67) --
	(156.35, 57.67) --
	(155.90, 57.67) --
	(155.45, 57.67) --
	(155.01, 57.67) --
	(154.56, 57.67) --
	(154.11, 57.67) --
	(153.66, 57.67) --
	(153.21, 57.67) --
	(152.76, 57.67) --
	(152.31, 57.67) --
	(151.87, 57.67) --
	(151.42, 57.67) --
	(150.97, 57.67) --
	(150.52, 57.67) --
	(150.07, 57.67) --
	(149.62, 57.67) --
	(149.17, 57.67) --
	(148.72, 57.67) --
	(148.28, 57.67) --
	(147.83, 57.67) --
	(147.38, 57.67) --
	(146.93, 57.67) --
	(146.48, 57.67) --
	(146.03, 57.67) --
	(145.58, 57.67) --
	(145.13, 57.67) --
	(144.69, 57.67) --
	(144.24, 57.67) --
	(143.79, 57.67) --
	(143.34, 57.67) --
	(142.89, 57.67) --
	(142.44, 57.67) --
	(141.99, 57.67) --
	(141.55, 57.67) --
	(141.10, 57.67) --
	(140.65, 57.67) --
	(140.20, 57.67) --
	(139.75, 57.67) --
	(139.30, 57.67) --
	(138.85, 57.67) --
	(138.40, 57.67) --
	(137.96, 57.67) --
	(137.51, 57.67) --
	(137.06, 57.67) --
	(136.61, 57.67) --
	(136.16, 57.67) --
	(135.71, 57.67) --
	(135.26, 57.67) --
	(134.81, 57.67) --
	(134.37, 57.67) --
	(133.92, 57.67) --
	(133.47, 57.67) --
	(133.02, 57.67) --
	(132.57, 57.67) --
	(132.12, 57.67) --
	(131.67, 57.67) --
	(131.23, 57.67) --
	(130.78, 57.67) --
	(130.33, 57.67) --
	(129.88, 57.67) --
	(129.43, 57.67) --
	(128.98, 57.67) --
	(128.53, 57.67) --
	(128.08, 57.67) --
	(127.64, 57.67) --
	(127.19, 57.67) --
	(126.74, 57.67) --
	(126.29, 57.67) --
	(125.84, 57.67) --
	(125.39, 57.67) --
	(124.94, 57.67) --
	(124.49, 57.67) --
	(124.05, 57.67) --
	(123.60, 57.67) --
	(123.15, 57.67) --
	(122.70, 57.67) --
	(122.25, 57.67) --
	(121.80, 57.67) --
	(121.35, 57.67) --
	(120.91, 57.67) --
	(120.46, 57.67) --
	(120.01, 57.67) --
	(119.56, 57.67) --
	(119.11, 57.67) --
	(118.66, 57.67) --
	(118.21, 57.67) --
	(117.76, 57.67) --
	(117.32, 57.67) --
	(116.87, 57.67) --
	(116.42, 57.67) --
	(115.97, 57.67) --
	(115.52, 57.67) --
	(115.07, 57.67) --
	(114.62, 57.67) --
	(114.17, 57.67) --
	(113.73, 57.67) --
	(113.28, 57.67) --
	(112.83, 57.67) --
	(112.38, 57.67) --
	(111.93, 57.67) --
	(111.48, 57.67) --
	(111.03, 57.67) --
	(110.59, 57.67) --
	(110.14, 57.67) --
	(109.69, 57.67) --
	(109.24, 57.67) --
	(108.79, 57.67) --
	(108.34, 57.67) --
	(107.89, 57.67) --
	(107.44, 57.67) --
	(107.00, 57.67) --
	(106.55, 57.67) --
	(106.10, 57.67) --
	(105.65, 57.67) --
	(105.20, 57.67) --
	(104.75, 57.67) --
	(104.30, 57.67) --
	(103.85, 57.67) --
	(103.41, 57.67) --
	(102.96, 57.67) --
	(102.51, 57.67) --
	(102.06, 57.67) --
	(101.61, 57.67) --
	(101.16, 57.67) --
	(100.71, 57.67) --
	(100.27, 57.67) --
	( 99.82, 57.67) --
	( 99.37, 57.67) --
	( 98.92, 57.67) --
	( 98.47, 57.67) --
	( 98.02, 57.67) --
	( 97.57, 57.67) --
	( 97.12, 57.67) --
	( 96.68, 57.67) --
	( 96.23, 57.67) --
	( 95.78, 57.67) --
	( 95.33, 57.67) --
	( 94.88, 57.67) --
	( 94.43, 57.67) --
	( 93.98, 57.67) --
	( 93.53, 57.67) --
	( 93.09, 57.67) --
	( 92.64, 57.67) --
	( 92.19, 57.67) --
	( 91.74, 57.67) --
	( 91.29, 57.67) --
	( 90.84, 57.67) --
	( 90.39, 57.67) --
	( 89.95, 57.67) --
	( 89.50, 57.67) --
	( 89.05, 57.67) --
	( 88.60, 57.67) --
	( 88.15, 57.67) --
	( 87.70, 57.67) --
	( 87.25, 57.67) --
	( 86.80, 57.67) --
	( 86.36, 57.67) --
	( 85.91, 57.67) --
	( 85.46, 57.67) --
	( 85.01, 57.67) --
	( 84.56, 57.67) --
	( 84.11, 57.67) --
	( 83.66, 57.67) --
	( 83.21, 57.67) --
	( 82.77, 57.67) --
	( 82.32, 57.67) --
	( 81.87, 57.67) --
	( 81.42, 57.67) --
	( 80.97, 57.67) --
	( 80.52, 57.67) --
	( 80.07, 57.67) --
	( 79.63, 57.67) --
	( 79.18, 57.67) --
	( 78.73, 57.67) --
	( 78.28, 57.67) --
	( 77.83, 57.67) --
	( 77.38, 57.67) --
	( 76.93, 57.67) --
	( 76.48, 57.67) --
	( 76.04, 57.67) --
	( 75.59, 57.67) --
	( 75.14, 57.67) --
	( 74.69, 57.67) --
	( 74.24, 57.67) --
	( 73.79, 57.67) --
	( 73.34, 57.67) --
	( 72.89, 57.67) --
	( 72.45, 57.67) --
	( 72.00, 57.67) --
	( 71.55, 57.67) --
	( 71.10, 57.67) --
	( 70.65, 57.67) --
	( 70.20, 57.67) --
	( 69.75, 57.67) --
	( 69.31, 57.67) --
	( 68.86, 57.67) --
	( 68.41, 57.67) --
	( 67.96, 57.67) --
	( 67.51, 57.67) --
	( 67.06, 57.67) --
	( 66.61, 57.67) --
	( 66.16, 57.67) --
	( 65.72, 57.67) --
	( 65.27, 57.67) --
	( 64.82, 57.67) --
	( 64.37, 57.67) --
	( 63.92, 57.67) --
	( 63.47, 57.67) --
	( 63.02, 57.67) --
	( 62.57, 57.67) --
	( 62.13, 57.67) --
	( 61.68, 57.67) --
	( 61.23, 57.67) --
	( 60.78, 57.67) --
	( 60.33, 57.67) --
	( 59.88, 57.67) --
	( 59.43, 57.67) --
	( 58.99, 57.67) --
	( 58.54, 57.67) --
	( 58.09, 57.67) --
	( 57.64, 57.67) --
	( 57.19, 57.67) --
	( 56.74, 57.67) --
	( 56.29, 57.67) --
	( 55.84, 57.67) --
	( 55.40, 57.67) --
	( 54.95, 57.67) --
	( 54.50, 57.67) --
	( 54.05, 57.67) --
	( 53.60, 57.67) --
	( 53.15, 57.67) --
	( 52.70, 57.67) --
	( 52.25, 57.67) --
	( 51.81, 57.67) --
	( 51.36, 57.67) --
	( 50.91, 57.67) --
	( 50.46, 57.67) --
	( 50.01, 57.67) --
	( 49.56, 57.67) --
	( 49.11, 57.67) --
	( 48.67, 57.67) --
	( 48.22, 57.67) --
	( 47.77, 57.67) --
	( 47.32, 57.67) --
	( 46.87, 57.67) --
	( 46.42, 57.67) --
	( 45.97, 57.67) --
	( 45.52, 57.67) --
	( 45.08, 57.67) --
	( 44.63, 57.67) --
	( 44.18, 57.67) --
	( 43.73, 57.67) --
	( 43.28, 57.67) --
	( 42.83, 57.67) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 7pt off 3pt ,line join=round,line cap=round] ( 42.83, 57.67) --
	( 43.28, 57.67) --
	( 43.73, 57.67) --
	( 44.18, 57.67) --
	( 44.63, 57.67) --
	( 45.08, 57.67) --
	( 45.52, 57.67) --
	( 45.97, 57.67) --
	( 46.42, 57.67) --
	( 46.87, 57.67) --
	( 47.32, 57.67) --
	( 47.77, 57.67) --
	( 48.22, 57.67) --
	( 48.67, 57.67) --
	( 49.11, 57.67) --
	( 49.56, 57.67) --
	( 50.01, 57.67) --
	( 50.46, 57.67) --
	( 50.91, 57.67) --
	( 51.36, 57.67) --
	( 51.81, 57.67) --
	( 52.25, 57.67) --
	( 52.70, 57.67) --
	( 53.15, 57.67) --
	( 53.60, 57.67) --
	( 54.05, 57.67) --
	( 54.50, 57.67) --
	( 54.95, 57.67) --
	( 55.40, 57.67) --
	( 55.84, 57.67) --
	( 56.29, 57.67) --
	( 56.74, 57.67) --
	( 57.19, 57.67) --
	( 57.64, 57.67) --
	( 58.09, 57.67) --
	( 58.54, 57.67) --
	( 58.99, 57.67) --
	( 59.43, 57.67) --
	( 59.88, 57.67) --
	( 60.33, 57.67) --
	( 60.78, 57.67) --
	( 61.23, 57.67) --
	( 61.68, 57.67) --
	( 62.13, 57.67) --
	( 62.57, 57.67) --
	( 63.02, 57.67) --
	( 63.47, 57.67) --
	( 63.92, 57.67) --
	( 64.37, 57.67) --
	( 64.82, 57.67) --
	( 65.27, 57.67) --
	( 65.72, 57.67) --
	( 66.16, 57.67) --
	( 66.61, 57.67) --
	( 67.06, 57.67) --
	( 67.51, 57.67) --
	( 67.96, 57.67) --
	( 68.41, 57.67) --
	( 68.86, 57.67) --
	( 69.31, 57.67) --
	( 69.75, 57.67) --
	( 70.20, 57.67) --
	( 70.65, 57.67) --
	( 71.10, 57.67) --
	( 71.55, 57.67) --
	( 72.00, 57.67) --
	( 72.45, 57.67) --
	( 72.89, 57.67) --
	( 73.34, 57.67) --
	( 73.79, 57.67) --
	( 74.24, 57.67) --
	( 74.69, 57.67) --
	( 75.14, 57.67) --
	( 75.59, 57.67) --
	( 76.04, 57.67) --
	( 76.48, 57.67) --
	( 76.93, 57.67) --
	( 77.38, 57.67) --
	( 77.83, 57.67) --
	( 78.28, 57.67) --
	( 78.73, 57.67) --
	( 79.18, 57.67) --
	( 79.63, 57.67) --
	( 80.07, 57.67) --
	( 80.52, 57.67) --
	( 80.97, 57.67) --
	( 81.42, 57.67) --
	( 81.87, 57.67) --
	( 82.32, 57.67) --
	( 82.77, 57.67) --
	( 83.21, 57.67) --
	( 83.66, 57.67) --
	( 84.11, 57.67) --
	( 84.56, 57.67) --
	( 85.01, 57.67) --
	( 85.46, 57.67) --
	( 85.91, 57.67) --
	( 86.36, 57.67) --
	( 86.80, 57.67) --
	( 87.25, 57.67) --
	( 87.70, 57.67) --
	( 88.15, 57.67) --
	( 88.60, 57.67) --
	( 89.05, 57.67) --
	( 89.50, 57.67) --
	( 89.95, 57.67) --
	( 90.39, 57.67) --
	( 90.84, 57.67) --
	( 91.29, 57.67) --
	( 91.74, 57.67) --
	( 92.19, 57.67) --
	( 92.64, 57.67) --
	( 93.09, 57.67) --
	( 93.53, 57.67) --
	( 93.98, 57.68) --
	( 94.43, 57.68) --
	( 94.88, 57.68) --
	( 95.33, 57.69) --
	( 95.78, 57.69) --
	( 96.23, 57.69) --
	( 96.68, 57.69) --
	( 97.12, 57.69) --
	( 97.57, 57.69) --
	( 98.02, 57.69) --
	( 98.47, 57.68) --
	( 98.92, 57.68) --
	( 99.37, 57.67) --
	( 99.82, 57.67) --
	(100.27, 57.67) --
	(100.71, 57.67) --
	(101.16, 57.67) --
	(101.61, 57.67) --
	(102.06, 57.67) --
	(102.51, 57.67) --
	(102.96, 57.67) --
	(103.41, 57.67) --
	(103.85, 57.67) --
	(104.30, 57.67) --
	(104.75, 57.67) --
	(105.20, 57.67) --
	(105.65, 57.68) --
	(106.10, 57.68) --
	(106.55, 57.69) --
	(107.00, 57.69) --
	(107.44, 57.70) --
	(107.89, 57.70) --
	(108.34, 57.70) --
	(108.79, 57.71) --
	(109.24, 57.71) --
	(109.69, 57.71) --
	(110.14, 57.71) --
	(110.59, 57.71) --
	(111.03, 57.72) --
	(111.48, 57.72) --
	(111.93, 57.73) --
	(112.38, 57.74) --
	(112.83, 57.74) --
	(113.28, 57.75) --
	(113.73, 57.76) --
	(114.17, 57.77) --
	(114.62, 57.77) --
	(115.07, 57.77) --
	(115.52, 57.77) --
	(115.97, 57.77) --
	(116.42, 57.77) --
	(116.87, 57.77) --
	(117.32, 57.77) --
	(117.76, 57.78) --
	(118.21, 57.78) --
	(118.66, 57.79) --
	(119.11, 57.81) --
	(119.56, 57.82) --
	(120.01, 57.83) --
	(120.46, 57.84) --
	(120.91, 57.85) --
	(121.35, 57.85) --
	(121.80, 57.85) --
	(122.25, 57.85) --
	(122.70, 57.85) --
	(123.15, 57.85) --
	(123.60, 57.86) --
	(124.05, 57.86) --
	(124.49, 57.87) --
	(124.94, 57.88) --
	(125.39, 57.89) --
	(125.84, 57.89) --
	(126.29, 57.90) --
	(126.74, 57.92) --
	(127.19, 57.94) --
	(127.64, 57.97) --
	(128.08, 58.01) --
	(128.53, 58.05) --
	(128.98, 58.09) --
	(129.43, 58.13) --
	(129.88, 58.15) --
	(130.33, 58.17) --
	(130.78, 58.18) --
	(131.23, 58.19) --
	(131.67, 58.20) --
	(132.12, 58.21) --
	(132.57, 58.23) --
	(133.02, 58.25) --
	(133.47, 58.28) --
	(133.92, 58.30) --
	(134.37, 58.33) --
	(134.81, 58.36) --
	(135.26, 58.38) --
	(135.71, 58.41) --
	(136.16, 58.43) --
	(136.61, 58.44) --
	(137.06, 58.45) --
	(137.51, 58.47) --
	(137.96, 58.49) --
	(138.40, 58.52) --
	(138.85, 58.57) --
	(139.30, 58.64) --
	(139.75, 58.73) --
	(140.20, 58.82) --
	(140.65, 58.90) --
	(141.10, 58.95) --
	(141.55, 58.97) --
	(141.99, 58.97) --
	(142.44, 58.94) --
	(142.89, 58.91) --
	(143.34, 58.89) --
	(143.79, 58.88) --
	(144.24, 58.90) --
	(144.69, 58.95) --
	(145.13, 59.01) --
	(145.58, 59.08) --
	(146.03, 59.16) --
	(146.48, 59.24) --
	(146.93, 59.33) --
	(147.38, 59.42) --
	(147.83, 59.53) --
	(148.28, 59.63) --
	(148.72, 59.73) --
	(149.17, 59.82) --
	(149.62, 59.90) --
	(150.07, 59.97) --
	(150.52, 60.03) --
	(150.97, 60.10) --
	(151.42, 60.18) --
	(151.87, 60.27) --
	(152.31, 60.36) --
	(152.76, 60.46) --
	(153.21, 60.54) --
	(153.66, 60.59) --
	(154.11, 60.60) --
	(154.56, 60.58) --
	(155.01, 60.53) --
	(155.45, 60.46) --
	(155.90, 60.40) --
	(156.35, 60.36) --
	(156.80, 60.36) --
	(157.25, 60.40) --
	(157.70, 60.48) --
	(158.15, 60.59) --
	(158.60, 60.72) --
	(159.04, 60.85) --
	(159.49, 60.98) --
	(159.94, 61.09) --
	(160.39, 61.18) --
	(160.84, 61.26) --
	(161.29, 61.34) --
	(161.74, 61.42) --
	(162.19, 61.51) --
	(162.63, 61.62) --
	(163.08, 61.76) --
	(163.53, 61.92) --
	(163.98, 62.09) --
	(164.43, 62.25) --
	(164.88, 62.38) --
	(165.33, 62.49) --
	(165.77, 62.55) --
	(166.22, 62.59) --
	(166.67, 62.60) --
	(167.12, 62.62) --
	(167.57, 62.66) --
	(168.02, 62.71) --
	(168.47, 62.80) --
	(168.92, 62.91) --
	(169.36, 63.03) --
	(169.81, 63.17) --
	(170.26, 63.32) --
	(170.71, 63.48) --
	(171.16, 63.64) --
	(171.61, 63.80) --
	(172.06, 63.95) --
	(172.51, 64.10) --
	(172.95, 64.23) --
	(173.40, 64.33) --
	(173.85, 64.42) --
	(174.30, 64.52) --
	(174.75, 64.63) --
	(175.20, 64.77) --
	(175.65, 64.98) --
	(176.09, 65.23) --
	(176.54, 65.54) --
	(176.99, 65.87) --
	(177.44, 66.22) --
	(177.89, 66.55) --
	(178.34, 66.86) --
	(178.79, 67.12) --
	(179.24, 67.36) --
	(179.68, 67.57) --
	(180.13, 67.79) --
	(180.58, 68.05) --
	(181.03, 68.36) --
	(181.48, 68.73) --
	(181.93, 69.16) --
	(182.38, 69.63) --
	(182.83, 70.13) --
	(183.27, 70.62) --
	(183.72, 71.07) --
	(184.17, 71.46) --
	(184.62, 71.77) --
	(185.07, 72.00) --
	(185.52, 72.16) --
	(185.97, 72.27) --
	(186.41, 72.38) --
	(186.86, 72.51) --
	(187.31, 72.72) --
	(187.76, 73.06) --
	(188.21, 73.59) --
	(188.66, 74.34) --
	(189.11, 75.38) --
	(189.56, 76.73) --
	(190.00, 78.37) --
	(190.45, 80.24) --
	(190.90, 82.21) --
	(191.35, 84.10) --
	(191.80, 85.72) --
	(192.25, 86.91) --
	(192.70, 87.61) --
	(193.15, 87.88) --
	(193.59, 87.88) --
	(194.04, 87.79) --
	(194.49, 87.74) --
	(194.94, 87.84) --
	(195.39, 88.13) --
	(195.84, 88.59) --
	(196.29, 89.21) --
	(196.73, 89.97) --
	(197.18, 90.90) --
	(197.63, 92.03) --
	(198.08, 93.42) --
	(198.53, 95.10) --
	(198.98, 97.07) --
	(199.43, 99.25) --
	(199.88,101.51) --
	(200.32,103.68) --
	(200.77,105.60) --
	(201.22,107.22) --
	(201.67,108.57) --
	(202.12,109.78) --
	(202.57,111.05) --
	(203.02,112.51) --
	(203.47,114.27) --
	(203.91,116.32) --
	(204.36,118.62) --
	(204.81,121.05) --
	(205.26,123.48) --
	(205.71,125.84) --
	(206.16,128.09) --
	(206.61,130.27) --
	(207.05,132.48) --
	(207.50,134.74) --
	(207.95,137.07) --
	(208.40,139.36) --
	(208.85,141.51) --
	(209.30,143.40) --
	(209.75,145.01) --
	(210.20,146.39) --
	(210.64,147.64) --
	(211.09,148.86) --
	(211.54,150.09) --
	(211.99,151.28) --
	(212.44,152.40) --
	(212.89,153.38) --
	(213.34,154.16) --
	(213.79,154.72) --
	(214.23,155.06) --
	(214.68,155.21) --
	(215.13,155.24) --
	(215.58,155.20) --
	(216.03,155.11) --
	(216.48,154.98) --
	(216.93,154.77) --
	(217.37,154.41) --
	(217.82,153.86) --
	(218.27,153.09) --
	(218.72,152.08) --
	(219.17,150.85) --
	(219.62,149.46) --
	(220.07,147.95) --
	(220.52,146.36) --
	(220.96,144.71) --
	(221.41,143.03) --
	(221.86,141.30) --
	(222.31,139.53) --
	(222.76,137.70) --
	(223.21,135.84) --
	(223.66,133.95) --
	(224.11,132.04) --
	(224.55,130.11) --
	(225.00,128.15) --
	(225.45,126.16) --
	(225.90,124.13) --
	(226.35,122.06) --
	(226.80,119.97) --
	(227.25,117.88) --
	(227.69,115.85) --
	(228.14,113.92) --
	(228.59,112.14) --
	(229.04,110.54) --
	(229.49,109.08) --
	(229.94,107.70) --
	(230.39,106.31) --
	(230.84,104.79) --
	(231.28,103.03) --
	(231.73,100.95) --
	(232.18, 98.50) --
	(232.63, 95.66) --
	(233.08, 92.47) --
	(233.53, 89.01) --
	(233.98, 85.40) --
	(234.43, 81.77) --
	(234.87, 78.25) --
	(235.32, 74.97) --
	(235.77, 72.00) --
	(236.22, 69.38) --
	(236.67, 67.11) --
	(237.12, 65.15) --
	(237.57, 63.48) --
	(238.01, 62.06) --
	(238.46, 60.89) --
	(238.91, 59.94) --
	(239.36, 59.21) --
	(239.81, 58.66) --
	(240.26, 58.28) --
	(240.71, 58.02) --
	(241.16, 57.86) --
	(241.60, 57.77) --
	(242.05, 57.72) --
	(242.50, 57.69) --
	(242.95, 57.68) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67);

\path[] ( 42.83, 57.67) --
	( 43.28, 57.67) --
	( 43.73, 57.67) --
	( 44.18, 57.67) --
	( 44.63, 57.67) --
	( 45.08, 57.67) --
	( 45.52, 57.67) --
	( 45.97, 57.67) --
	( 46.42, 57.67) --
	( 46.87, 57.67) --
	( 47.32, 57.67) --
	( 47.77, 57.67) --
	( 48.22, 57.67) --
	( 48.67, 57.67) --
	( 49.11, 57.67) --
	( 49.56, 57.67) --
	( 50.01, 57.67) --
	( 50.46, 57.67) --
	( 50.91, 57.67) --
	( 51.36, 57.67) --
	( 51.81, 57.67) --
	( 52.25, 57.67) --
	( 52.70, 57.67) --
	( 53.15, 57.67) --
	( 53.60, 57.67) --
	( 54.05, 57.67) --
	( 54.50, 57.67) --
	( 54.95, 57.67) --
	( 55.40, 57.67) --
	( 55.84, 57.67) --
	( 56.29, 57.67) --
	( 56.74, 57.67) --
	( 57.19, 57.67) --
	( 57.64, 57.67) --
	( 58.09, 57.67) --
	( 58.54, 57.67) --
	( 58.99, 57.67) --
	( 59.43, 57.67) --
	( 59.88, 57.67) --
	( 60.33, 57.67) --
	( 60.78, 57.67) --
	( 61.23, 57.67) --
	( 61.68, 57.67) --
	( 62.13, 57.67) --
	( 62.57, 57.67) --
	( 63.02, 57.67) --
	( 63.47, 57.67) --
	( 63.92, 57.67) --
	( 64.37, 57.67) --
	( 64.82, 57.67) --
	( 65.27, 57.67) --
	( 65.72, 57.67) --
	( 66.16, 57.67) --
	( 66.61, 57.67) --
	( 67.06, 57.67) --
	( 67.51, 57.67) --
	( 67.96, 57.67) --
	( 68.41, 57.67) --
	( 68.86, 57.67) --
	( 69.31, 57.67) --
	( 69.75, 57.67) --
	( 70.20, 57.67) --
	( 70.65, 57.67) --
	( 71.10, 57.67) --
	( 71.55, 57.67) --
	( 72.00, 57.67) --
	( 72.45, 57.67) --
	( 72.89, 57.67) --
	( 73.34, 57.67) --
	( 73.79, 57.67) --
	( 74.24, 57.67) --
	( 74.69, 57.67) --
	( 75.14, 57.67) --
	( 75.59, 57.67) --
	( 76.04, 57.67) --
	( 76.48, 57.67) --
	( 76.93, 57.67) --
	( 77.38, 57.67) --
	( 77.83, 57.67) --
	( 78.28, 57.67) --
	( 78.73, 57.67) --
	( 79.18, 57.67) --
	( 79.63, 57.67) --
	( 80.07, 57.67) --
	( 80.52, 57.67) --
	( 80.97, 57.67) --
	( 81.42, 57.67) --
	( 81.87, 57.67) --
	( 82.32, 57.67) --
	( 82.77, 57.67) --
	( 83.21, 57.67) --
	( 83.66, 57.67) --
	( 84.11, 57.67) --
	( 84.56, 57.67) --
	( 85.01, 57.67) --
	( 85.46, 57.67) --
	( 85.91, 57.67) --
	( 86.36, 57.67) --
	( 86.80, 57.67) --
	( 87.25, 57.67) --
	( 87.70, 57.67) --
	( 88.15, 57.67) --
	( 88.60, 57.67) --
	( 89.05, 57.67) --
	( 89.50, 57.67) --
	( 89.95, 57.67) --
	( 90.39, 57.67) --
	( 90.84, 57.67) --
	( 91.29, 57.67) --
	( 91.74, 57.67) --
	( 92.19, 57.67) --
	( 92.64, 57.67) --
	( 93.09, 57.67) --
	( 93.53, 57.67) --
	( 93.98, 57.67) --
	( 94.43, 57.67) --
	( 94.88, 57.67) --
	( 95.33, 57.67) --
	( 95.78, 57.67) --
	( 96.23, 57.67) --
	( 96.68, 57.67) --
	( 97.12, 57.67) --
	( 97.57, 57.67) --
	( 98.02, 57.67) --
	( 98.47, 57.67) --
	( 98.92, 57.67) --
	( 99.37, 57.67) --
	( 99.82, 57.67) --
	(100.27, 57.67) --
	(100.71, 57.67) --
	(101.16, 57.67) --
	(101.61, 57.67) --
	(102.06, 57.67) --
	(102.51, 57.67) --
	(102.96, 57.67) --
	(103.41, 57.67) --
	(103.85, 57.67) --
	(104.30, 57.67) --
	(104.75, 57.67) --
	(105.20, 57.68) --
	(105.65, 57.68) --
	(106.10, 57.68) --
	(106.55, 57.69) --
	(107.00, 57.69) --
	(107.44, 57.69) --
	(107.89, 57.70) --
	(108.34, 57.70) --
	(108.79, 57.71) --
	(109.24, 57.71) --
	(109.69, 57.71) --
	(110.14, 57.72) --
	(110.59, 57.72) --
	(111.03, 57.72) --
	(111.48, 57.73) --
	(111.93, 57.74) --
	(112.38, 57.75) --
	(112.83, 57.75) --
	(113.28, 57.76) --
	(113.73, 57.76) --
	(114.17, 57.77) --
	(114.62, 57.77) --
	(115.07, 57.77) --
	(115.52, 57.77) --
	(115.97, 57.76) --
	(116.42, 57.76) --
	(116.87, 57.76) --
	(117.32, 57.77) --
	(117.76, 57.77) --
	(118.21, 57.78) --
	(118.66, 57.78) --
	(119.11, 57.79) --
	(119.56, 57.80) --
	(120.01, 57.81) --
	(120.46, 57.82) --
	(120.91, 57.83) --
	(121.35, 57.83) --
	(121.80, 57.84) --
	(122.25, 57.85) --
	(122.70, 57.86) --
	(123.15, 57.87) --
	(123.60, 57.87) --
	(124.05, 57.88) --
	(124.49, 57.89) --
	(124.94, 57.89) --
	(125.39, 57.90) --
	(125.84, 57.90) --
	(126.29, 57.90) --
	(126.74, 57.90) --
	(127.19, 57.91) --
	(127.64, 57.91) --
	(128.08, 57.92) --
	(128.53, 57.93) --
	(128.98, 57.95) --
	(129.43, 57.97) --
	(129.88, 57.99) --
	(130.33, 58.02) --
	(130.78, 58.05) --
	(131.23, 58.08) --
	(131.67, 58.11) --
	(132.12, 58.14) --
	(132.57, 58.17) --
	(133.02, 58.20) --
	(133.47, 58.22) --
	(133.92, 58.23) --
	(134.37, 58.24) --
	(134.81, 58.24) --
	(135.26, 58.24) --
	(135.71, 58.25) --
	(136.16, 58.25) --
	(136.61, 58.25) --
	(137.06, 58.26) --
	(137.51, 58.28) --
	(137.96, 58.31) --
	(138.40, 58.35) --
	(138.85, 58.40) --
	(139.30, 58.45) --
	(139.75, 58.51) --
	(140.20, 58.56) --
	(140.65, 58.61) --
	(141.10, 58.65) --
	(141.55, 58.68) --
	(141.99, 58.70) --
	(142.44, 58.72) --
	(142.89, 58.73) --
	(143.34, 58.75) --
	(143.79, 58.77) --
	(144.24, 58.80) --
	(144.69, 58.84) --
	(145.13, 58.89) --
	(145.58, 58.94) --
	(146.03, 59.00) --
	(146.48, 59.06) --
	(146.93, 59.12) --
	(147.38, 59.18) --
	(147.83, 59.24) --
	(148.28, 59.30) --
	(148.72, 59.36) --
	(149.17, 59.42) --
	(149.62, 59.48) --
	(150.07, 59.55) --
	(150.52, 59.62) --
	(150.97, 59.70) --
	(151.42, 59.78) --
	(151.87, 59.87) --
	(152.31, 59.95) --
	(152.76, 60.03) --
	(153.21, 60.09) --
	(153.66, 60.14) --
	(154.11, 60.18) --
	(154.56, 60.19) --
	(155.01, 60.19) --
	(155.45, 60.18) --
	(155.90, 60.18) --
	(156.35, 60.18) --
	(156.80, 60.20) --
	(157.25, 60.24) --
	(157.70, 60.30) --
	(158.15, 60.38) --
	(158.60, 60.47) --
	(159.04, 60.57) --
	(159.49, 60.66) --
	(159.94, 60.74) --
	(160.39, 60.81) --
	(160.84, 60.87) --
	(161.29, 60.92) --
	(161.74, 60.96) --
	(162.19, 61.01) --
	(162.63, 61.07) --
	(163.08, 61.14) --
	(163.53, 61.23) --
	(163.98, 61.33) --
	(164.43, 61.44) --
	(164.88, 61.55) --
	(165.33, 61.67) --
	(165.77, 61.78) --
	(166.22, 61.88) --
	(166.67, 61.97) --
	(167.12, 62.06) --
	(167.57, 62.15) --
	(168.02, 62.24) --
	(168.47, 62.33) --
	(168.92, 62.42) --
	(169.36, 62.52) --
	(169.81, 62.61) --
	(170.26, 62.71) --
	(170.71, 62.82) --
	(171.16, 62.92) --
	(171.61, 63.03) --
	(172.06, 63.13) --
	(172.51, 63.23) --
	(172.95, 63.32) --
	(173.40, 63.41) --
	(173.85, 63.49) --
	(174.30, 63.57) --
	(174.75, 63.67) --
	(175.20, 63.77) --
	(175.65, 63.90) --
	(176.09, 64.04) --
	(176.54, 64.21) --
	(176.99, 64.39) --
	(177.44, 64.58) --
	(177.89, 64.78) --
	(178.34, 64.98) --
	(178.79, 65.19) --
	(179.24, 65.40) --
	(179.68, 65.61) --
	(180.13, 65.83) --
	(180.58, 66.06) --
	(181.03, 66.30) --
	(181.48, 66.56) --
	(181.93, 66.82) --
	(182.38, 67.10) --
	(182.83, 67.38) --
	(183.27, 67.66) --
	(183.72, 67.95) --
	(184.17, 68.23) --
	(184.62, 68.50) --
	(185.07, 68.75) --
	(185.52, 69.00) --
	(185.97, 69.24) --
	(186.41, 69.48) --
	(186.86, 69.74) --
	(187.31, 70.03) --
	(187.76, 70.38) --
	(188.21, 70.80) --
	(188.66, 71.30) --
	(189.11, 71.87) --
	(189.56, 72.52) --
	(190.00, 73.22) --
	(190.45, 73.96) --
	(190.90, 74.71) --
	(191.35, 75.45) --
	(191.80, 76.14) --
	(192.25, 76.78) --
	(192.70, 77.35) --
	(193.15, 77.85) --
	(193.59, 78.28) --
	(194.04, 78.66) --
	(194.49, 79.02) --
	(194.94, 79.38) --
	(195.39, 79.75) --
	(195.84, 80.16) --
	(196.29, 80.61) --
	(196.73, 81.10) --
	(197.18, 81.65) --
	(197.63, 82.25) --
	(198.08, 82.90) --
	(198.53, 83.59) --
	(198.98, 84.33) --
	(199.43, 85.11) --
	(199.88, 85.93) --
	(200.32, 86.80) --
	(200.77, 87.70) --
	(201.22, 88.64) --
	(201.67, 89.62) --
	(202.12, 90.64) --
	(202.57, 91.70) --
	(203.02, 92.81) --
	(203.47, 93.99) --
	(203.91, 95.23) --
	(204.36, 96.53) --
	(204.81, 97.88) --
	(205.26, 99.26) --
	(205.71,100.65) --
	(206.16,102.02) --
	(206.61,103.34) --
	(207.05,104.59) --
	(207.50,105.79) --
	(207.95,106.95) --
	(208.40,108.12) --
	(208.85,109.33) --
	(209.30,110.64) --
	(209.75,112.07) --
	(210.20,113.65) --
	(210.64,115.38) --
	(211.09,117.25) --
	(211.54,119.23) --
	(211.99,121.28) --
	(212.44,123.37) --
	(212.89,125.45) --
	(213.34,127.51) --
	(213.79,129.49) --
	(214.23,131.38) --
	(214.68,133.14) --
	(215.13,134.75) --
	(215.58,136.17) --
	(216.03,137.38) --
	(216.48,138.36) --
	(216.93,139.09) --
	(217.37,139.59) --
	(217.82,139.86) --
	(218.27,139.93) --
	(218.72,139.81) --
	(219.17,139.53) --
	(219.62,139.11) --
	(220.07,138.54) --
	(220.52,137.83) --
	(220.96,136.99) --
	(221.41,136.02) --
	(221.86,134.93) --
	(222.31,133.73) --
	(222.76,132.45) --
	(223.21,131.09) --
	(223.66,129.68) --
	(224.11,128.22) --
	(224.55,126.72) --
	(225.00,125.21) --
	(225.45,123.69) --
	(225.90,122.20) --
	(226.35,120.74) --
	(226.80,119.37) --
	(227.25,118.09) --
	(227.69,116.95) --
	(228.14,115.97) --
	(228.59,115.14) --
	(229.04,114.44) --
	(229.49,113.84) --
	(229.94,113.31) --
	(230.39,112.81) --
	(230.84,112.29) --
	(231.28,111.73) --
	(231.73,111.12) --
	(232.18,110.45) --
	(232.63,109.73) --
	(233.08,108.98) --
	(233.53,108.23) --
	(233.98,107.49) --
	(234.43,106.78) --
	(234.87,106.11) --
	(235.32,105.45) --
	(235.77,104.81) --
	(236.22,104.15) --
	(236.67,103.46) --
	(237.12,102.73) --
	(237.57,101.95) --
	(238.01,101.11) --
	(238.46,100.21) --
	(238.91, 99.24) --
	(239.36, 98.22) --
	(239.81, 97.15) --
	(240.26, 96.04) --
	(240.71, 94.90) --
	(241.16, 93.74) --
	(241.60, 92.57) --
	(242.05, 91.39) --
	(242.50, 90.22) --
	(242.95, 89.06) --
	(243.40, 87.91) --
	(243.85, 86.78) --
	(244.30, 85.67) --
	(244.75, 84.57) --
	(245.19, 83.50) --
	(245.64, 82.44) --
	(246.09, 81.39) --
	(246.54, 80.35) --
	(246.99, 79.31) --
	(247.44, 78.29) --
	(247.89, 77.29) --
	(248.33, 76.32) --
	(248.78, 75.38) --
	(249.23, 74.49) --
	(249.68, 73.65) --
	(250.13, 72.86) --
	(250.58, 72.10) --
	(251.03, 71.40) --
	(251.48, 70.73) --
	(251.92, 70.10) --
	(252.37, 69.50) --
	(252.82, 68.94) --
	(253.27, 68.40) --
	(253.72, 67.90) --
	(254.17, 67.43) --
	(254.62, 66.98) --
	(255.07, 66.56) --
	(255.51, 66.16) --
	(255.96, 65.77) --
	(256.41, 65.41) --
	(256.86, 65.04) --
	(257.31, 64.68) --
	(257.76, 64.29) --
	(258.21, 63.89) --
	(258.65, 63.46) --
	(259.10, 63.01) --
	(259.55, 62.54) --
	(260.00, 62.07) --
	(260.45, 61.60) --
	(260.90, 61.15) --
	(261.35, 60.73) --
	(261.80, 60.34) --
	(262.24, 59.99) --
	(262.69, 59.67) --
	(263.14, 59.40) --
	(263.59, 59.16) --
	(264.04, 58.96) --
	(264.49, 58.79) --
	(264.94, 58.65) --
	(265.39, 58.53) --
	(265.83, 58.44) --
	(266.28, 58.38) --
	(266.73, 58.33) --
	(267.18, 58.30) --
	(267.63, 58.28) --
	(268.08, 58.27) --
	(268.53, 58.27) --
	(268.97, 58.26) --
	(269.42, 58.24) --
	(269.87, 58.22) --
	(270.32, 58.19) --
	(270.77, 58.14) --
	(271.22, 58.09) --
	(271.67, 58.03) --
	(272.12, 57.97) --
	(272.12, 57.67) --
	(271.67, 57.67) --
	(271.22, 57.67) --
	(270.77, 57.67) --
	(270.32, 57.67) --
	(269.87, 57.67) --
	(269.42, 57.67) --
	(268.97, 57.67) --
	(268.53, 57.67) --
	(268.08, 57.67) --
	(267.63, 57.67) --
	(267.18, 57.67) --
	(266.73, 57.67) --
	(266.28, 57.67) --
	(265.83, 57.67) --
	(265.39, 57.67) --
	(264.94, 57.67) --
	(264.49, 57.67) --
	(264.04, 57.67) --
	(263.59, 57.67) --
	(263.14, 57.67) --
	(262.69, 57.67) --
	(262.24, 57.67) --
	(261.80, 57.67) --
	(261.35, 57.67) --
	(260.90, 57.67) --
	(260.45, 57.67) --
	(260.00, 57.67) --
	(259.55, 57.67) --
	(259.10, 57.67) --
	(258.65, 57.67) --
	(258.21, 57.67) --
	(257.76, 57.67) --
	(257.31, 57.67) --
	(256.86, 57.67) --
	(256.41, 57.67) --
	(255.96, 57.67) --
	(255.51, 57.67) --
	(255.07, 57.67) --
	(254.62, 57.67) --
	(254.17, 57.67) --
	(253.72, 57.67) --
	(253.27, 57.67) --
	(252.82, 57.67) --
	(252.37, 57.67) --
	(251.92, 57.67) --
	(251.48, 57.67) --
	(251.03, 57.67) --
	(250.58, 57.67) --
	(250.13, 57.67) --
	(249.68, 57.67) --
	(249.23, 57.67) --
	(248.78, 57.67) --
	(248.33, 57.67) --
	(247.89, 57.67) --
	(247.44, 57.67) --
	(246.99, 57.67) --
	(246.54, 57.67) --
	(246.09, 57.67) --
	(245.64, 57.67) --
	(245.19, 57.67) --
	(244.75, 57.67) --
	(244.30, 57.67) --
	(243.85, 57.67) --
	(243.40, 57.67) --
	(242.95, 57.67) --
	(242.50, 57.67) --
	(242.05, 57.67) --
	(241.60, 57.67) --
	(241.16, 57.67) --
	(240.71, 57.67) --
	(240.26, 57.67) --
	(239.81, 57.67) --
	(239.36, 57.67) --
	(238.91, 57.67) --
	(238.46, 57.67) --
	(238.01, 57.67) --
	(237.57, 57.67) --
	(237.12, 57.67) --
	(236.67, 57.67) --
	(236.22, 57.67) --
	(235.77, 57.67) --
	(235.32, 57.67) --
	(234.87, 57.67) --
	(234.43, 57.67) --
	(233.98, 57.67) --
	(233.53, 57.67) --
	(233.08, 57.67) --
	(232.63, 57.67) --
	(232.18, 57.67) --
	(231.73, 57.67) --
	(231.28, 57.67) --
	(230.84, 57.67) --
	(230.39, 57.67) --
	(229.94, 57.67) --
	(229.49, 57.67) --
	(229.04, 57.67) --
	(228.59, 57.67) --
	(228.14, 57.67) --
	(227.69, 57.67) --
	(227.25, 57.67) --
	(226.80, 57.67) --
	(226.35, 57.67) --
	(225.90, 57.67) --
	(225.45, 57.67) --
	(225.00, 57.67) --
	(224.55, 57.67) --
	(224.11, 57.67) --
	(223.66, 57.67) --
	(223.21, 57.67) --
	(222.76, 57.67) --
	(222.31, 57.67) --
	(221.86, 57.67) --
	(221.41, 57.67) --
	(220.96, 57.67) --
	(220.52, 57.67) --
	(220.07, 57.67) --
	(219.62, 57.67) --
	(219.17, 57.67) --
	(218.72, 57.67) --
	(218.27, 57.67) --
	(217.82, 57.67) --
	(217.37, 57.67) --
	(216.93, 57.67) --
	(216.48, 57.67) --
	(216.03, 57.67) --
	(215.58, 57.67) --
	(215.13, 57.67) --
	(214.68, 57.67) --
	(214.23, 57.67) --
	(213.79, 57.67) --
	(213.34, 57.67) --
	(212.89, 57.67) --
	(212.44, 57.67) --
	(211.99, 57.67) --
	(211.54, 57.67) --
	(211.09, 57.67) --
	(210.64, 57.67) --
	(210.20, 57.67) --
	(209.75, 57.67) --
	(209.30, 57.67) --
	(208.85, 57.67) --
	(208.40, 57.67) --
	(207.95, 57.67) --
	(207.50, 57.67) --
	(207.05, 57.67) --
	(206.61, 57.67) --
	(206.16, 57.67) --
	(205.71, 57.67) --
	(205.26, 57.67) --
	(204.81, 57.67) --
	(204.36, 57.67) --
	(203.91, 57.67) --
	(203.47, 57.67) --
	(203.02, 57.67) --
	(202.57, 57.67) --
	(202.12, 57.67) --
	(201.67, 57.67) --
	(201.22, 57.67) --
	(200.77, 57.67) --
	(200.32, 57.67) --
	(199.88, 57.67) --
	(199.43, 57.67) --
	(198.98, 57.67) --
	(198.53, 57.67) --
	(198.08, 57.67) --
	(197.63, 57.67) --
	(197.18, 57.67) --
	(196.73, 57.67) --
	(196.29, 57.67) --
	(195.84, 57.67) --
	(195.39, 57.67) --
	(194.94, 57.67) --
	(194.49, 57.67) --
	(194.04, 57.67) --
	(193.59, 57.67) --
	(193.15, 57.67) --
	(192.70, 57.67) --
	(192.25, 57.67) --
	(191.80, 57.67) --
	(191.35, 57.67) --
	(190.90, 57.67) --
	(190.45, 57.67) --
	(190.00, 57.67) --
	(189.56, 57.67) --
	(189.11, 57.67) --
	(188.66, 57.67) --
	(188.21, 57.67) --
	(187.76, 57.67) --
	(187.31, 57.67) --
	(186.86, 57.67) --
	(186.41, 57.67) --
	(185.97, 57.67) --
	(185.52, 57.67) --
	(185.07, 57.67) --
	(184.62, 57.67) --
	(184.17, 57.67) --
	(183.72, 57.67) --
	(183.27, 57.67) --
	(182.83, 57.67) --
	(182.38, 57.67) --
	(181.93, 57.67) --
	(181.48, 57.67) --
	(181.03, 57.67) --
	(180.58, 57.67) --
	(180.13, 57.67) --
	(179.68, 57.67) --
	(179.24, 57.67) --
	(178.79, 57.67) --
	(178.34, 57.67) --
	(177.89, 57.67) --
	(177.44, 57.67) --
	(176.99, 57.67) --
	(176.54, 57.67) --
	(176.09, 57.67) --
	(175.65, 57.67) --
	(175.20, 57.67) --
	(174.75, 57.67) --
	(174.30, 57.67) --
	(173.85, 57.67) --
	(173.40, 57.67) --
	(172.95, 57.67) --
	(172.51, 57.67) --
	(172.06, 57.67) --
	(171.61, 57.67) --
	(171.16, 57.67) --
	(170.71, 57.67) --
	(170.26, 57.67) --
	(169.81, 57.67) --
	(169.36, 57.67) --
	(168.92, 57.67) --
	(168.47, 57.67) --
	(168.02, 57.67) --
	(167.57, 57.67) --
	(167.12, 57.67) --
	(166.67, 57.67) --
	(166.22, 57.67) --
	(165.77, 57.67) --
	(165.33, 57.67) --
	(164.88, 57.67) --
	(164.43, 57.67) --
	(163.98, 57.67) --
	(163.53, 57.67) --
	(163.08, 57.67) --
	(162.63, 57.67) --
	(162.19, 57.67) --
	(161.74, 57.67) --
	(161.29, 57.67) --
	(160.84, 57.67) --
	(160.39, 57.67) --
	(159.94, 57.67) --
	(159.49, 57.67) --
	(159.04, 57.67) --
	(158.60, 57.67) --
	(158.15, 57.67) --
	(157.70, 57.67) --
	(157.25, 57.67) --
	(156.80, 57.67) --
	(156.35, 57.67) --
	(155.90, 57.67) --
	(155.45, 57.67) --
	(155.01, 57.67) --
	(154.56, 57.67) --
	(154.11, 57.67) --
	(153.66, 57.67) --
	(153.21, 57.67) --
	(152.76, 57.67) --
	(152.31, 57.67) --
	(151.87, 57.67) --
	(151.42, 57.67) --
	(150.97, 57.67) --
	(150.52, 57.67) --
	(150.07, 57.67) --
	(149.62, 57.67) --
	(149.17, 57.67) --
	(148.72, 57.67) --
	(148.28, 57.67) --
	(147.83, 57.67) --
	(147.38, 57.67) --
	(146.93, 57.67) --
	(146.48, 57.67) --
	(146.03, 57.67) --
	(145.58, 57.67) --
	(145.13, 57.67) --
	(144.69, 57.67) --
	(144.24, 57.67) --
	(143.79, 57.67) --
	(143.34, 57.67) --
	(142.89, 57.67) --
	(142.44, 57.67) --
	(141.99, 57.67) --
	(141.55, 57.67) --
	(141.10, 57.67) --
	(140.65, 57.67) --
	(140.20, 57.67) --
	(139.75, 57.67) --
	(139.30, 57.67) --
	(138.85, 57.67) --
	(138.40, 57.67) --
	(137.96, 57.67) --
	(137.51, 57.67) --
	(137.06, 57.67) --
	(136.61, 57.67) --
	(136.16, 57.67) --
	(135.71, 57.67) --
	(135.26, 57.67) --
	(134.81, 57.67) --
	(134.37, 57.67) --
	(133.92, 57.67) --
	(133.47, 57.67) --
	(133.02, 57.67) --
	(132.57, 57.67) --
	(132.12, 57.67) --
	(131.67, 57.67) --
	(131.23, 57.67) --
	(130.78, 57.67) --
	(130.33, 57.67) --
	(129.88, 57.67) --
	(129.43, 57.67) --
	(128.98, 57.67) --
	(128.53, 57.67) --
	(128.08, 57.67) --
	(127.64, 57.67) --
	(127.19, 57.67) --
	(126.74, 57.67) --
	(126.29, 57.67) --
	(125.84, 57.67) --
	(125.39, 57.67) --
	(124.94, 57.67) --
	(124.49, 57.67) --
	(124.05, 57.67) --
	(123.60, 57.67) --
	(123.15, 57.67) --
	(122.70, 57.67) --
	(122.25, 57.67) --
	(121.80, 57.67) --
	(121.35, 57.67) --
	(120.91, 57.67) --
	(120.46, 57.67) --
	(120.01, 57.67) --
	(119.56, 57.67) --
	(119.11, 57.67) --
	(118.66, 57.67) --
	(118.21, 57.67) --
	(117.76, 57.67) --
	(117.32, 57.67) --
	(116.87, 57.67) --
	(116.42, 57.67) --
	(115.97, 57.67) --
	(115.52, 57.67) --
	(115.07, 57.67) --
	(114.62, 57.67) --
	(114.17, 57.67) --
	(113.73, 57.67) --
	(113.28, 57.67) --
	(112.83, 57.67) --
	(112.38, 57.67) --
	(111.93, 57.67) --
	(111.48, 57.67) --
	(111.03, 57.67) --
	(110.59, 57.67) --
	(110.14, 57.67) --
	(109.69, 57.67) --
	(109.24, 57.67) --
	(108.79, 57.67) --
	(108.34, 57.67) --
	(107.89, 57.67) --
	(107.44, 57.67) --
	(107.00, 57.67) --
	(106.55, 57.67) --
	(106.10, 57.67) --
	(105.65, 57.67) --
	(105.20, 57.67) --
	(104.75, 57.67) --
	(104.30, 57.67) --
	(103.85, 57.67) --
	(103.41, 57.67) --
	(102.96, 57.67) --
	(102.51, 57.67) --
	(102.06, 57.67) --
	(101.61, 57.67) --
	(101.16, 57.67) --
	(100.71, 57.67) --
	(100.27, 57.67) --
	( 99.82, 57.67) --
	( 99.37, 57.67) --
	( 98.92, 57.67) --
	( 98.47, 57.67) --
	( 98.02, 57.67) --
	( 97.57, 57.67) --
	( 97.12, 57.67) --
	( 96.68, 57.67) --
	( 96.23, 57.67) --
	( 95.78, 57.67) --
	( 95.33, 57.67) --
	( 94.88, 57.67) --
	( 94.43, 57.67) --
	( 93.98, 57.67) --
	( 93.53, 57.67) --
	( 93.09, 57.67) --
	( 92.64, 57.67) --
	( 92.19, 57.67) --
	( 91.74, 57.67) --
	( 91.29, 57.67) --
	( 90.84, 57.67) --
	( 90.39, 57.67) --
	( 89.95, 57.67) --
	( 89.50, 57.67) --
	( 89.05, 57.67) --
	( 88.60, 57.67) --
	( 88.15, 57.67) --
	( 87.70, 57.67) --
	( 87.25, 57.67) --
	( 86.80, 57.67) --
	( 86.36, 57.67) --
	( 85.91, 57.67) --
	( 85.46, 57.67) --
	( 85.01, 57.67) --
	( 84.56, 57.67) --
	( 84.11, 57.67) --
	( 83.66, 57.67) --
	( 83.21, 57.67) --
	( 82.77, 57.67) --
	( 82.32, 57.67) --
	( 81.87, 57.67) --
	( 81.42, 57.67) --
	( 80.97, 57.67) --
	( 80.52, 57.67) --
	( 80.07, 57.67) --
	( 79.63, 57.67) --
	( 79.18, 57.67) --
	( 78.73, 57.67) --
	( 78.28, 57.67) --
	( 77.83, 57.67) --
	( 77.38, 57.67) --
	( 76.93, 57.67) --
	( 76.48, 57.67) --
	( 76.04, 57.67) --
	( 75.59, 57.67) --
	( 75.14, 57.67) --
	( 74.69, 57.67) --
	( 74.24, 57.67) --
	( 73.79, 57.67) --
	( 73.34, 57.67) --
	( 72.89, 57.67) --
	( 72.45, 57.67) --
	( 72.00, 57.67) --
	( 71.55, 57.67) --
	( 71.10, 57.67) --
	( 70.65, 57.67) --
	( 70.20, 57.67) --
	( 69.75, 57.67) --
	( 69.31, 57.67) --
	( 68.86, 57.67) --
	( 68.41, 57.67) --
	( 67.96, 57.67) --
	( 67.51, 57.67) --
	( 67.06, 57.67) --
	( 66.61, 57.67) --
	( 66.16, 57.67) --
	( 65.72, 57.67) --
	( 65.27, 57.67) --
	( 64.82, 57.67) --
	( 64.37, 57.67) --
	( 63.92, 57.67) --
	( 63.47, 57.67) --
	( 63.02, 57.67) --
	( 62.57, 57.67) --
	( 62.13, 57.67) --
	( 61.68, 57.67) --
	( 61.23, 57.67) --
	( 60.78, 57.67) --
	( 60.33, 57.67) --
	( 59.88, 57.67) --
	( 59.43, 57.67) --
	( 58.99, 57.67) --
	( 58.54, 57.67) --
	( 58.09, 57.67) --
	( 57.64, 57.67) --
	( 57.19, 57.67) --
	( 56.74, 57.67) --
	( 56.29, 57.67) --
	( 55.84, 57.67) --
	( 55.40, 57.67) --
	( 54.95, 57.67) --
	( 54.50, 57.67) --
	( 54.05, 57.67) --
	( 53.60, 57.67) --
	( 53.15, 57.67) --
	( 52.70, 57.67) --
	( 52.25, 57.67) --
	( 51.81, 57.67) --
	( 51.36, 57.67) --
	( 50.91, 57.67) --
	( 50.46, 57.67) --
	( 50.01, 57.67) --
	( 49.56, 57.67) --
	( 49.11, 57.67) --
	( 48.67, 57.67) --
	( 48.22, 57.67) --
	( 47.77, 57.67) --
	( 47.32, 57.67) --
	( 46.87, 57.67) --
	( 46.42, 57.67) --
	( 45.97, 57.67) --
	( 45.52, 57.67) --
	( 45.08, 57.67) --
	( 44.63, 57.67) --
	( 44.18, 57.67) --
	( 43.73, 57.67) --
	( 43.28, 57.67) --
	( 42.83, 57.67) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 42.83, 57.67) --
	( 43.28, 57.67) --
	( 43.73, 57.67) --
	( 44.18, 57.67) --
	( 44.63, 57.67) --
	( 45.08, 57.67) --
	( 45.52, 57.67) --
	( 45.97, 57.67) --
	( 46.42, 57.67) --
	( 46.87, 57.67) --
	( 47.32, 57.67) --
	( 47.77, 57.67) --
	( 48.22, 57.67) --
	( 48.67, 57.67) --
	( 49.11, 57.67) --
	( 49.56, 57.67) --
	( 50.01, 57.67) --
	( 50.46, 57.67) --
	( 50.91, 57.67) --
	( 51.36, 57.67) --
	( 51.81, 57.67) --
	( 52.25, 57.67) --
	( 52.70, 57.67) --
	( 53.15, 57.67) --
	( 53.60, 57.67) --
	( 54.05, 57.67) --
	( 54.50, 57.67) --
	( 54.95, 57.67) --
	( 55.40, 57.67) --
	( 55.84, 57.67) --
	( 56.29, 57.67) --
	( 56.74, 57.67) --
	( 57.19, 57.67) --
	( 57.64, 57.67) --
	( 58.09, 57.67) --
	( 58.54, 57.67) --
	( 58.99, 57.67) --
	( 59.43, 57.67) --
	( 59.88, 57.67) --
	( 60.33, 57.67) --
	( 60.78, 57.67) --
	( 61.23, 57.67) --
	( 61.68, 57.67) --
	( 62.13, 57.67) --
	( 62.57, 57.67) --
	( 63.02, 57.67) --
	( 63.47, 57.67) --
	( 63.92, 57.67) --
	( 64.37, 57.67) --
	( 64.82, 57.67) --
	( 65.27, 57.67) --
	( 65.72, 57.67) --
	( 66.16, 57.67) --
	( 66.61, 57.67) --
	( 67.06, 57.67) --
	( 67.51, 57.67) --
	( 67.96, 57.67) --
	( 68.41, 57.67) --
	( 68.86, 57.67) --
	( 69.31, 57.67) --
	( 69.75, 57.67) --
	( 70.20, 57.67) --
	( 70.65, 57.67) --
	( 71.10, 57.67) --
	( 71.55, 57.67) --
	( 72.00, 57.67) --
	( 72.45, 57.67) --
	( 72.89, 57.67) --
	( 73.34, 57.67) --
	( 73.79, 57.67) --
	( 74.24, 57.67) --
	( 74.69, 57.67) --
	( 75.14, 57.67) --
	( 75.59, 57.67) --
	( 76.04, 57.67) --
	( 76.48, 57.67) --
	( 76.93, 57.67) --
	( 77.38, 57.67) --
	( 77.83, 57.67) --
	( 78.28, 57.67) --
	( 78.73, 57.67) --
	( 79.18, 57.67) --
	( 79.63, 57.67) --
	( 80.07, 57.67) --
	( 80.52, 57.67) --
	( 80.97, 57.67) --
	( 81.42, 57.67) --
	( 81.87, 57.67) --
	( 82.32, 57.67) --
	( 82.77, 57.67) --
	( 83.21, 57.67) --
	( 83.66, 57.67) --
	( 84.11, 57.67) --
	( 84.56, 57.67) --
	( 85.01, 57.67) --
	( 85.46, 57.67) --
	( 85.91, 57.67) --
	( 86.36, 57.67) --
	( 86.80, 57.67) --
	( 87.25, 57.67) --
	( 87.70, 57.67) --
	( 88.15, 57.67) --
	( 88.60, 57.67) --
	( 89.05, 57.67) --
	( 89.50, 57.67) --
	( 89.95, 57.67) --
	( 90.39, 57.67) --
	( 90.84, 57.67) --
	( 91.29, 57.67) --
	( 91.74, 57.67) --
	( 92.19, 57.67) --
	( 92.64, 57.67) --
	( 93.09, 57.67) --
	( 93.53, 57.67) --
	( 93.98, 57.67) --
	( 94.43, 57.67) --
	( 94.88, 57.67) --
	( 95.33, 57.67) --
	( 95.78, 57.67) --
	( 96.23, 57.67) --
	( 96.68, 57.67) --
	( 97.12, 57.67) --
	( 97.57, 57.67) --
	( 98.02, 57.67) --
	( 98.47, 57.67) --
	( 98.92, 57.67) --
	( 99.37, 57.67) --
	( 99.82, 57.67) --
	(100.27, 57.67) --
	(100.71, 57.67) --
	(101.16, 57.67) --
	(101.61, 57.67) --
	(102.06, 57.67) --
	(102.51, 57.67) --
	(102.96, 57.67) --
	(103.41, 57.67) --
	(103.85, 57.67) --
	(104.30, 57.67) --
	(104.75, 57.67) --
	(105.20, 57.68) --
	(105.65, 57.68) --
	(106.10, 57.68) --
	(106.55, 57.69) --
	(107.00, 57.69) --
	(107.44, 57.69) --
	(107.89, 57.70) --
	(108.34, 57.70) --
	(108.79, 57.71) --
	(109.24, 57.71) --
	(109.69, 57.71) --
	(110.14, 57.72) --
	(110.59, 57.72) --
	(111.03, 57.72) --
	(111.48, 57.73) --
	(111.93, 57.74) --
	(112.38, 57.75) --
	(112.83, 57.75) --
	(113.28, 57.76) --
	(113.73, 57.76) --
	(114.17, 57.77) --
	(114.62, 57.77) --
	(115.07, 57.77) --
	(115.52, 57.77) --
	(115.97, 57.76) --
	(116.42, 57.76) --
	(116.87, 57.76) --
	(117.32, 57.77) --
	(117.76, 57.77) --
	(118.21, 57.78) --
	(118.66, 57.78) --
	(119.11, 57.79) --
	(119.56, 57.80) --
	(120.01, 57.81) --
	(120.46, 57.82) --
	(120.91, 57.83) --
	(121.35, 57.83) --
	(121.80, 57.84) --
	(122.25, 57.85) --
	(122.70, 57.86) --
	(123.15, 57.87) --
	(123.60, 57.87) --
	(124.05, 57.88) --
	(124.49, 57.89) --
	(124.94, 57.89) --
	(125.39, 57.90) --
	(125.84, 57.90) --
	(126.29, 57.90) --
	(126.74, 57.90) --
	(127.19, 57.91) --
	(127.64, 57.91) --
	(128.08, 57.92) --
	(128.53, 57.93) --
	(128.98, 57.95) --
	(129.43, 57.97) --
	(129.88, 57.99) --
	(130.33, 58.02) --
	(130.78, 58.05) --
	(131.23, 58.08) --
	(131.67, 58.11) --
	(132.12, 58.14) --
	(132.57, 58.17) --
	(133.02, 58.20) --
	(133.47, 58.22) --
	(133.92, 58.23) --
	(134.37, 58.24) --
	(134.81, 58.24) --
	(135.26, 58.24) --
	(135.71, 58.25) --
	(136.16, 58.25) --
	(136.61, 58.25) --
	(137.06, 58.26) --
	(137.51, 58.28) --
	(137.96, 58.31) --
	(138.40, 58.35) --
	(138.85, 58.40) --
	(139.30, 58.45) --
	(139.75, 58.51) --
	(140.20, 58.56) --
	(140.65, 58.61) --
	(141.10, 58.65) --
	(141.55, 58.68) --
	(141.99, 58.70) --
	(142.44, 58.72) --
	(142.89, 58.73) --
	(143.34, 58.75) --
	(143.79, 58.77) --
	(144.24, 58.80) --
	(144.69, 58.84) --
	(145.13, 58.89) --
	(145.58, 58.94) --
	(146.03, 59.00) --
	(146.48, 59.06) --
	(146.93, 59.12) --
	(147.38, 59.18) --
	(147.83, 59.24) --
	(148.28, 59.30) --
	(148.72, 59.36) --
	(149.17, 59.42) --
	(149.62, 59.48) --
	(150.07, 59.55) --
	(150.52, 59.62) --
	(150.97, 59.70) --
	(151.42, 59.78) --
	(151.87, 59.87) --
	(152.31, 59.95) --
	(152.76, 60.03) --
	(153.21, 60.09) --
	(153.66, 60.14) --
	(154.11, 60.18) --
	(154.56, 60.19) --
	(155.01, 60.19) --
	(155.45, 60.18) --
	(155.90, 60.18) --
	(156.35, 60.18) --
	(156.80, 60.20) --
	(157.25, 60.24) --
	(157.70, 60.30) --
	(158.15, 60.38) --
	(158.60, 60.47) --
	(159.04, 60.57) --
	(159.49, 60.66) --
	(159.94, 60.74) --
	(160.39, 60.81) --
	(160.84, 60.87) --
	(161.29, 60.92) --
	(161.74, 60.96) --
	(162.19, 61.01) --
	(162.63, 61.07) --
	(163.08, 61.14) --
	(163.53, 61.23) --
	(163.98, 61.33) --
	(164.43, 61.44) --
	(164.88, 61.55) --
	(165.33, 61.67) --
	(165.77, 61.78) --
	(166.22, 61.88) --
	(166.67, 61.97) --
	(167.12, 62.06) --
	(167.57, 62.15) --
	(168.02, 62.24) --
	(168.47, 62.33) --
	(168.92, 62.42) --
	(169.36, 62.52) --
	(169.81, 62.61) --
	(170.26, 62.71) --
	(170.71, 62.82) --
	(171.16, 62.92) --
	(171.61, 63.03) --
	(172.06, 63.13) --
	(172.51, 63.23) --
	(172.95, 63.32) --
	(173.40, 63.41) --
	(173.85, 63.49) --
	(174.30, 63.57) --
	(174.75, 63.67) --
	(175.20, 63.77) --
	(175.65, 63.90) --
	(176.09, 64.04) --
	(176.54, 64.21) --
	(176.99, 64.39) --
	(177.44, 64.58) --
	(177.89, 64.78) --
	(178.34, 64.98) --
	(178.79, 65.19) --
	(179.24, 65.40) --
	(179.68, 65.61) --
	(180.13, 65.83) --
	(180.58, 66.06) --
	(181.03, 66.30) --
	(181.48, 66.56) --
	(181.93, 66.82) --
	(182.38, 67.10) --
	(182.83, 67.38) --
	(183.27, 67.66) --
	(183.72, 67.95) --
	(184.17, 68.23) --
	(184.62, 68.50) --
	(185.07, 68.75) --
	(185.52, 69.00) --
	(185.97, 69.24) --
	(186.41, 69.48) --
	(186.86, 69.74) --
	(187.31, 70.03) --
	(187.76, 70.38) --
	(188.21, 70.80) --
	(188.66, 71.30) --
	(189.11, 71.87) --
	(189.56, 72.52) --
	(190.00, 73.22) --
	(190.45, 73.96) --
	(190.90, 74.71) --
	(191.35, 75.45) --
	(191.80, 76.14) --
	(192.25, 76.78) --
	(192.70, 77.35) --
	(193.15, 77.85) --
	(193.59, 78.28) --
	(194.04, 78.66) --
	(194.49, 79.02) --
	(194.94, 79.38) --
	(195.39, 79.75) --
	(195.84, 80.16) --
	(196.29, 80.61) --
	(196.73, 81.10) --
	(197.18, 81.65) --
	(197.63, 82.25) --
	(198.08, 82.90) --
	(198.53, 83.59) --
	(198.98, 84.33) --
	(199.43, 85.11) --
	(199.88, 85.93) --
	(200.32, 86.80) --
	(200.77, 87.70) --
	(201.22, 88.64) --
	(201.67, 89.62) --
	(202.12, 90.64) --
	(202.57, 91.70) --
	(203.02, 92.81) --
	(203.47, 93.99) --
	(203.91, 95.23) --
	(204.36, 96.53) --
	(204.81, 97.88) --
	(205.26, 99.26) --
	(205.71,100.65) --
	(206.16,102.02) --
	(206.61,103.34) --
	(207.05,104.59) --
	(207.50,105.79) --
	(207.95,106.95) --
	(208.40,108.12) --
	(208.85,109.33) --
	(209.30,110.64) --
	(209.75,112.07) --
	(210.20,113.65) --
	(210.64,115.38) --
	(211.09,117.25) --
	(211.54,119.23) --
	(211.99,121.28) --
	(212.44,123.37) --
	(212.89,125.45) --
	(213.34,127.51) --
	(213.79,129.49) --
	(214.23,131.38) --
	(214.68,133.14) --
	(215.13,134.75) --
	(215.58,136.17) --
	(216.03,137.38) --
	(216.48,138.36) --
	(216.93,139.09) --
	(217.37,139.59) --
	(217.82,139.86) --
	(218.27,139.93) --
	(218.72,139.81) --
	(219.17,139.53) --
	(219.62,139.11) --
	(220.07,138.54) --
	(220.52,137.83) --
	(220.96,136.99) --
	(221.41,136.02) --
	(221.86,134.93) --
	(222.31,133.73) --
	(222.76,132.45) --
	(223.21,131.09) --
	(223.66,129.68) --
	(224.11,128.22) --
	(224.55,126.72) --
	(225.00,125.21) --
	(225.45,123.69) --
	(225.90,122.20) --
	(226.35,120.74) --
	(226.80,119.37) --
	(227.25,118.09) --
	(227.69,116.95) --
	(228.14,115.97) --
	(228.59,115.14) --
	(229.04,114.44) --
	(229.49,113.84) --
	(229.94,113.31) --
	(230.39,112.81) --
	(230.84,112.29) --
	(231.28,111.73) --
	(231.73,111.12) --
	(232.18,110.45) --
	(232.63,109.73) --
	(233.08,108.98) --
	(233.53,108.23) --
	(233.98,107.49) --
	(234.43,106.78) --
	(234.87,106.11) --
	(235.32,105.45) --
	(235.77,104.81) --
	(236.22,104.15) --
	(236.67,103.46) --
	(237.12,102.73) --
	(237.57,101.95) --
	(238.01,101.11) --
	(238.46,100.21) --
	(238.91, 99.24) --
	(239.36, 98.22) --
	(239.81, 97.15) --
	(240.26, 96.04) --
	(240.71, 94.90) --
	(241.16, 93.74) --
	(241.60, 92.57) --
	(242.05, 91.39) --
	(242.50, 90.22) --
	(242.95, 89.06) --
	(243.40, 87.91) --
	(243.85, 86.78) --
	(244.30, 85.67) --
	(244.75, 84.57) --
	(245.19, 83.50) --
	(245.64, 82.44) --
	(246.09, 81.39) --
	(246.54, 80.35) --
	(246.99, 79.31) --
	(247.44, 78.29) --
	(247.89, 77.29) --
	(248.33, 76.32) --
	(248.78, 75.38) --
	(249.23, 74.49) --
	(249.68, 73.65) --
	(250.13, 72.86) --
	(250.58, 72.10) --
	(251.03, 71.40) --
	(251.48, 70.73) --
	(251.92, 70.10) --
	(252.37, 69.50) --
	(252.82, 68.94) --
	(253.27, 68.40) --
	(253.72, 67.90) --
	(254.17, 67.43) --
	(254.62, 66.98) --
	(255.07, 66.56) --
	(255.51, 66.16) --
	(255.96, 65.77) --
	(256.41, 65.41) --
	(256.86, 65.04) --
	(257.31, 64.68) --
	(257.76, 64.29) --
	(258.21, 63.89) --
	(258.65, 63.46) --
	(259.10, 63.01) --
	(259.55, 62.54) --
	(260.00, 62.07) --
	(260.45, 61.60) --
	(260.90, 61.15) --
	(261.35, 60.73) --
	(261.80, 60.34) --
	(262.24, 59.99) --
	(262.69, 59.67) --
	(263.14, 59.40) --
	(263.59, 59.16) --
	(264.04, 58.96) --
	(264.49, 58.79) --
	(264.94, 58.65) --
	(265.39, 58.53) --
	(265.83, 58.44) --
	(266.28, 58.38) --
	(266.73, 58.33) --
	(267.18, 58.30) --
	(267.63, 58.28) --
	(268.08, 58.27) --
	(268.53, 58.27) --
	(268.97, 58.26) --
	(269.42, 58.24) --
	(269.87, 58.22) --
	(270.32, 58.19) --
	(270.77, 58.14) --
	(271.22, 58.09) --
	(271.67, 58.03) --
	(272.12, 57.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.62, 51.21) -- (234.62, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.46, 51.21) -- (211.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (238.29, 51.21) -- (238.29, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.27, 51.21) -- (218.27, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.56, 51.21) -- (216.56, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.20, 51.21) -- (231.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.58, 51.21) -- (225.58, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (250.36, 51.21) -- (250.36, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (239.90, 51.21) -- (239.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.10, 51.21) -- (234.10, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (255.24, 51.21) -- (255.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (239.30, 51.21) -- (239.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.51, 51.21) -- (202.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (255.32, 51.21) -- (255.32, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (249.26, 51.21) -- (249.26, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.87, 51.21) -- (221.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.16, 51.21) -- (231.16, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.87, 51.21) -- (236.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (240.60, 51.21) -- (240.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.64, 51.21) -- (212.64, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.50, 51.21) -- (206.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.52, 51.21) -- (219.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.30, 51.21) -- (232.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.96, 51.21) -- (222.96, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.65, 51.21) -- (220.65, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.75, 51.21) -- (229.75, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.17, 51.21) -- (226.17, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.33, 51.21) -- (221.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.12, 51.21) -- (216.12, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.09, 51.21) -- (218.09, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.64, 51.21) -- (199.64, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.60, 51.21) -- (203.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.46, 51.21) -- (201.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (245.61, 51.21) -- (245.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (242.81, 51.21) -- (242.81, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.51, 51.21) -- (232.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (235.83, 51.21) -- (235.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (246.43, 51.21) -- (246.43, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.14, 51.21) -- (205.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (244.87, 51.21) -- (244.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.05, 51.21) -- (210.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (242.98, 51.21) -- (242.98, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.52, 51.21) -- (189.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.89, 51.21) -- (184.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.35, 51.21) -- (224.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.48, 51.21) -- (215.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.03, 51.21) -- (224.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (255.79, 51.21) -- (255.79, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (251.38, 51.21) -- (251.38, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.02, 51.21) -- (208.02, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.72, 51.21) -- (211.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (260.05, 51.21) -- (260.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.50, 51.21) -- (210.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.46, 51.21) -- (169.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (239.18, 51.21) -- (239.18, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.99, 51.21) -- (219.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.91, 51.21) -- (223.91, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.77, 51.21) -- (218.77, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.63, 51.21) -- (217.63, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.85, 51.21) -- (203.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.86, 51.21) -- (225.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.98, 51.21) -- (172.98, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.95, 51.21) -- (226.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.97, 51.21) -- (234.97, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.29, 51.21) -- (194.29, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (255.49, 51.21) -- (255.49, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (239.09, 51.21) -- (239.09, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.28, 51.21) -- (191.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (246.42, 51.21) -- (246.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.34, 51.21) -- (221.34, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.40, 51.21) -- (216.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.72, 51.21) -- (226.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (237.10, 51.21) -- (237.10, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.89, 51.21) -- (194.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (248.46, 51.21) -- (248.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.29, 51.21) -- (236.29, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.76, 51.21) -- (216.76, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.68, 51.21) -- (232.68, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (243.97, 51.21) -- (243.97, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.44, 51.21) -- (224.44, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (253.54, 51.21) -- (253.54, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (238.66, 51.21) -- (238.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.99, 51.21) -- (231.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.80, 51.21) -- (181.80, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.20, 51.21) -- (222.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.15, 51.21) -- (225.15, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.28, 51.21) -- (229.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.94, 51.21) -- (213.94, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (261.40, 51.21) -- (261.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (240.29, 51.21) -- (240.29, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (256.86, 51.21) -- (256.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.22, 51.21) -- (217.22, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (157.46, 51.21) -- (157.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (250.66, 51.21) -- (250.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.20, 51.21) -- (189.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (245.15, 51.21) -- (245.15, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.28, 51.21) -- (221.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (235.97, 51.21) -- (235.97, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.30, 51.21) -- (222.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.85, 51.21) -- (201.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (255.81, 51.21) -- (255.81, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.70, 51.21) -- (230.70, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.82, 51.21) -- (233.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.31, 51.21) -- (195.31, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (245.41, 51.21) -- (245.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.96, 51.21) -- (227.96, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.85, 51.21) -- (199.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.86, 51.21) -- (227.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.40, 51.21) -- (212.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.66, 51.21) -- (216.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.51, 51.21) -- (231.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.48, 51.21) -- (218.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (239.24, 51.21) -- (239.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (243.78, 51.21) -- (243.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.30, 51.21) -- (224.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.23, 51.21) -- (217.23, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.45, 51.21) -- (205.45, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.47, 51.21) -- (232.47, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (242.88, 51.21) -- (242.88, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.79, 51.21) -- (236.79, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.10, 51.21) -- (229.10, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.50, 51.21) -- (202.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.43, 51.21) -- (223.43, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.08, 51.21) -- (185.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.12, 51.21) -- (236.12, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.62, 51.21) -- (202.62, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.41, 51.21) -- (217.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (238.50, 51.21) -- (238.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (237.28, 51.21) -- (237.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.66, 51.21) -- (214.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.23, 51.21) -- (211.23, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.48, 51.21) -- (219.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.98, 51.21) -- (196.98, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.93, 51.21) -- (220.93, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.52, 51.21) -- (223.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (241.68, 51.21) -- (241.68, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.41, 51.21) -- (222.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.14, 51.21) -- (210.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.60, 51.21) -- (219.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (240.87, 51.21) -- (240.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.46, 51.21) -- (225.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.53, 51.21) -- (189.53, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.78, 51.21) -- (209.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.58, 51.21) -- (217.58, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.84, 51.21) -- (202.84, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.28, 51.21) -- (231.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.55, 51.21) -- (224.55, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.69, 51.21) -- (204.69, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.73, 51.21) -- (221.73, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.92, 51.21) -- (224.92, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (238.10, 51.21) -- (238.10, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.19, 51.21) -- (218.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.03, 51.21) -- (213.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (165.24, 51.21) -- (165.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.35, 51.21) -- (218.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.40, 51.21) -- (207.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (249.49, 51.21) -- (249.49, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.55, 51.21) -- (229.55, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.02, 51.21) -- (211.02, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.21, 51.21) -- (218.21, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.69, 51.21) -- (219.69, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (238.48, 51.21) -- (238.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.66, 51.21) -- (208.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (247.82, 51.21) -- (247.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.61, 51.21) -- (225.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.90, 51.21) -- (203.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.66, 51.21) -- (228.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.74, 51.21) -- (231.74, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.47, 51.21) -- (191.47, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.26, 51.21) -- (236.26, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.41, 51.21) -- (230.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (247.22, 51.21) -- (247.22, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.12, 51.21) -- (218.12, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (242.19, 51.21) -- (242.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.95, 51.21) -- (223.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.72, 51.21) -- (207.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (237.66, 51.21) -- (237.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.34, 51.21) -- (227.34, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (247.59, 51.21) -- (247.59, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.82, 51.21) -- (220.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.25, 51.21) -- (220.25, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.48, 51.21) -- (195.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.10, 51.21) -- (214.10, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.63, 51.21) -- (212.63, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.30, 51.21) -- (226.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.45, 51.21) -- (231.45, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (237.91, 51.21) -- (237.91, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (244.73, 51.21) -- (244.73, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (254.53, 51.21) -- (254.53, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.32, 51.21) -- (187.32, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.53, 51.21) -- (221.53, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.20, 51.21) -- (216.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.38, 51.21) -- (213.38, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (246.95, 51.21) -- (246.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.66, 51.21) -- (210.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (237.86, 51.21) -- (237.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (240.92, 51.21) -- (240.92, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (243.62, 51.21) -- (243.62, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.08, 51.21) -- (182.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.31, 51.21) -- (226.31, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.68, 51.21) -- (236.68, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.87, 51.21) -- (224.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.18, 51.21) -- (220.18, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (238.49, 51.21) -- (238.49, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.02, 51.21) -- (227.02, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.43, 51.21) -- (230.43, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (244.37, 51.21) -- (244.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.79, 51.21) -- (232.79, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.57, 51.21) -- (202.57, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (235.34, 51.21) -- (235.34, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.65, 51.21) -- (217.65, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.14, 51.21) -- (203.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.00, 51.21) -- (217.00, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.54, 51.21) -- (208.54, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.20, 51.21) -- (208.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.64, 51.21) -- (225.64, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.88, 51.21) -- (219.88, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.17, 51.21) -- (228.17, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.05, 51.21) -- (229.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.81, 51.21) -- (226.81, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.15, 51.21) -- (224.15, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.51, 51.21) -- (229.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.83, 51.21) -- (201.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.15, 51.21) -- (234.15, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.75, 51.21) -- (219.75, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.22, 51.21) -- (210.22, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.59, 51.21) -- (227.59, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.69, 51.21) -- (229.69, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.29, 51.21) -- (231.29, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.86, 51.21) -- (211.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.91, 51.21) -- (200.91, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.15, 51.21) -- (212.15, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.58, 51.21) -- (229.58, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.73, 51.21) -- (211.73, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.12, 51.21) -- (218.12, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.79, 51.21) -- (225.79, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.52, 51.21) -- (223.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.17, 51.21) -- (213.17, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.64, 51.21) -- (210.64, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.03, 51.21) -- (216.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.63, 51.21) -- (197.63, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.60, 51.21) -- (203.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.32, 51.21) -- (197.32, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.67, 51.21) -- (227.67, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.03, 51.21) -- (223.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.86, 51.21) -- (225.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.54, 51.21) -- (226.54, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.65, 51.21) -- (231.65, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.14, 51.21) -- (205.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.01, 51.21) -- (229.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.78, 51.21) -- (208.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (235.04, 51.21) -- (235.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.52, 51.21) -- (189.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.89, 51.21) -- (184.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.70, 51.21) -- (215.70, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.13, 51.21) -- (209.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.69, 51.21) -- (213.69, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.71, 51.21) -- (228.71, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.67, 51.21) -- (229.67, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.54, 51.21) -- (205.54, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.72, 51.21) -- (211.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.64, 51.21) -- (236.64, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.50, 51.21) -- (210.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.46, 51.21) -- (169.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.80, 51.21) -- (233.80, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.59, 51.21) -- (214.59, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.54, 51.21) -- (221.54, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.41, 51.21) -- (213.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.46, 51.21) -- (217.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.85, 51.21) -- (203.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.35, 51.21) -- (208.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.98, 51.21) -- (172.98, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.97, 51.21) -- (221.97, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.72, 51.21) -- (229.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.15, 51.21) -- (234.15, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.59, 51.21) -- (217.59, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.28, 51.21) -- (191.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.19, 51.21) -- (234.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.85, 51.21) -- (216.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.40, 51.21) -- (216.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.78, 51.21) -- (221.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.58, 51.21) -- (230.58, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.42, 51.21) -- (232.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.43, 51.21) -- (228.43, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.13, 51.21) -- (209.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.92, 51.21) -- (222.92, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.49, 51.21) -- (231.49, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.87, 51.21) -- (218.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.33, 51.21) -- (226.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.04, 51.21) -- (219.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.56, 51.21) -- (231.56, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.80, 51.21) -- (181.80, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.95, 51.21) -- (217.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.66, 51.21) -- (210.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.11, 51.21) -- (211.11, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.89, 51.21) -- (207.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (238.15, 51.21) -- (238.15, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.82, 51.21) -- (230.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.06, 51.21) -- (233.06, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.66, 51.21) -- (215.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (157.46, 51.21) -- (157.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.61, 51.21) -- (233.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.20, 51.21) -- (189.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.60, 51.21) -- (234.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.56, 51.21) -- (220.56, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.54, 51.21) -- (215.54, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.95, 51.21) -- (213.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.85, 51.21) -- (201.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.75, 51.21) -- (233.75, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.95, 51.21) -- (228.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.48, 51.21) -- (230.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.31, 51.21) -- (195.31, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.24, 51.21) -- (233.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.66, 51.21) -- (220.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.85, 51.21) -- (199.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.45, 51.21) -- (225.45, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.40, 51.21) -- (212.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.66, 51.21) -- (216.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.04, 51.21) -- (219.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.50, 51.21) -- (211.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.88, 51.21) -- (225.88, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.33, 51.21) -- (232.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.36, 51.21) -- (221.36, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.91, 51.21) -- (214.91, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.45, 51.21) -- (205.45, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.26, 51.21) -- (228.26, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.81, 51.21) -- (231.81, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.20, 51.21) -- (229.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.70, 51.21) -- (208.70, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.04, 51.21) -- (202.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.77, 51.21) -- (221.77, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.08, 51.21) -- (185.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.38, 51.21) -- (225.38, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.58, 51.21) -- (199.58, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.99, 51.21) -- (216.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.51, 51.21) -- (232.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.43, 51.21) -- (228.43, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.66, 51.21) -- (214.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.62, 51.21) -- (208.62, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.48, 51.21) -- (219.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.88, 51.21) -- (194.88, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.19, 51.21) -- (208.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.31, 51.21) -- (219.31, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.91, 51.21) -- (232.91, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.41, 51.21) -- (222.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.14, 51.21) -- (210.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.18, 51.21) -- (207.18, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.58, 51.21) -- (230.58, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.03, 51.21) -- (220.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.53, 51.21) -- (189.53, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.78, 51.21) -- (209.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.14, 51.21) -- (210.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.84, 51.21) -- (202.84, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.68, 51.21) -- (222.68, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.90, 51.21) -- (215.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.69, 51.21) -- (204.69, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.78, 51.21) -- (219.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.46, 51.21) -- (219.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.12, 51.21) -- (222.12, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.68, 51.21) -- (202.68, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.72, 51.21) -- (212.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (165.24, 51.21) -- (165.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.35, 51.21) -- (218.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.84, 51.21) -- (233.84, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.13, 51.21) -- (209.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.77, 51.21) -- (207.77, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.21, 51.21) -- (218.21, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.29, 51.21) -- (213.29, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.60, 51.21) -- (217.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.07, 51.21) -- (201.07, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.44, 51.21) -- (228.44, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.18, 51.21) -- (223.18, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.48, 51.21) -- (194.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.00, 51.21) -- (217.00, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.50, 51.21) -- (223.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.47, 51.21) -- (191.47, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.80, 51.21) -- (228.80, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.59, 51.21) -- (223.59, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.43, 51.21) -- (233.43, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.13, 51.21) -- (209.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.35, 51.21) -- (227.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.13, 51.21) -- (209.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.72, 51.21) -- (207.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.55, 51.21) -- (222.55, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.60, 51.21) -- (225.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.41, 51.21) -- (229.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.86, 51.21) -- (211.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.65, 51.21) -- (209.65, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.48, 51.21) -- (195.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.13, 51.21) -- (208.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.27, 51.21) -- (210.27, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.42, 51.21) -- (217.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.42, 51.21) -- (220.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.87, 51.21) -- (230.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.98, 51.21) -- (232.98, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.91, 51.21) -- (230.91, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.32, 51.21) -- (187.32, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.24, 51.21) -- (221.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.50, 51.21) -- (211.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.38, 51.21) -- (213.38, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.51, 51.21) -- (227.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.66, 51.21) -- (210.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.27, 51.21) -- (229.27, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.49, 51.21) -- (231.49, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.48, 51.21) -- (225.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.08, 51.21) -- (182.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.86, 51.21) -- (220.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.95, 51.21) -- (231.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.51, 51.21) -- (222.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.99, 51.21) -- (218.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.06, 51.21) -- (232.06, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.96, 51.21) -- (223.96, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.22, 51.21) -- (224.22, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.19, 51.21) -- (225.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.51, 51.21) -- (221.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.57, 51.21) -- (202.57, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.06, 51.21) -- (227.06, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.24, 51.21) -- (209.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.73, 51.21) -- (196.73, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.13, 51.21) -- (209.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.46, 51.21) -- (201.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.34, 51.21) -- (218.34, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.86, 51.21) -- (211.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.70, 51.21) -- (219.70, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.69, 51.21) -- (224.69, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.63, 51.21) -- (219.63, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.97, 51.21) -- (215.97, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.16, 51.21) -- (221.16, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.60, 51.21) -- (195.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.68, 51.21) -- (225.68, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.86, 51.21) -- (211.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.76, 51.21) -- (202.76, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.87, 51.21) -- (220.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.14, 51.21) -- (222.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.05, 51.21) -- (223.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.81, 51.21) -- (206.81, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.74, 51.21) -- (195.74, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.97, 51.21) -- (203.97, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.58, 51.21) -- (222.58, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.33, 51.21) -- (204.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.18, 51.21) -- (213.18, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.30, 51.21) -- (220.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.08, 51.21) -- (219.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.61, 51.21) -- (205.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.39, 51.21) -- (203.39, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.02, 51.21) -- (209.02, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.27, 51.21) -- (203.27, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.45, 51.21) -- (192.45, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.36, 51.21) -- (219.36, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.11, 51.21) -- (215.11, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.58, 51.21) -- (222.58, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.10, 51.21) -- (221.10, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.80, 51.21) -- (223.80, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.30, 51.21) -- (200.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.16, 51.21) -- (221.16, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.71, 51.21) -- (205.71, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.51, 51.21) -- (227.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.52, 51.21) -- (189.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.89, 51.21) -- (184.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.50, 51.21) -- (207.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.61, 51.21) -- (205.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.33, 51.21) -- (220.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.47, 51.21) -- (221.47, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.14, 51.21) -- (197.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.52, 51.21) -- (211.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.17, 51.21) -- (228.17, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.50, 51.21) -- (210.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.46, 51.21) -- (169.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.38, 51.21) -- (225.38, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.42, 51.21) -- (207.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.42, 51.21) -- (216.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.70, 51.21) -- (207.70, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.36, 51.21) -- (212.36, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.08, 51.21) -- (202.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.98, 51.21) -- (172.98, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.52, 51.21) -- (215.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.78, 51.21) -- (221.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.68, 51.21) -- (225.68, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.13, 51.21) -- (209.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.07, 51.21) -- (226.07, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.63, 51.21) -- (210.63, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.86, 51.21) -- (210.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.61, 51.21) -- (214.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.01, 51.21) -- (224.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.99, 51.21) -- (223.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.21, 51.21) -- (222.21, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.94, 51.21) -- (214.94, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.85, 51.21) -- (223.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.26, 51.21) -- (212.26, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.14, 51.21) -- (220.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.57, 51.21) -- (210.57, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.16, 51.21) -- (227.16, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.80, 51.21) -- (181.80, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.19, 51.21) -- (210.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.55, 51.21) -- (203.55, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.39, 51.21) -- (203.39, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.68, 51.21) -- (229.68, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.41, 51.21) -- (223.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.73, 51.21) -- (224.73, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.30, 51.21) -- (209.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (157.46, 51.21) -- (157.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.28, 51.21) -- (225.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.20, 51.21) -- (189.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.30, 51.21) -- (227.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.28, 51.21) -- (219.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.50, 51.21) -- (207.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.61, 51.21) -- (205.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.85, 51.21) -- (201.85, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.28, 51.21) -- (225.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.48, 51.21) -- (223.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.87, 51.21) -- (223.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.77, 51.21) -- (193.77, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.87, 51.21) -- (224.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.41, 51.21) -- (219.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.10, 51.21) -- (198.10, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.04, 51.21) -- (219.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.21, 51.21) -- (209.21, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.70, 51.21) -- (215.70, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.57, 51.21) -- (210.57, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.37, 51.21) -- (206.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.60, 51.21) -- (217.60, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.35, 51.21) -- (226.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.13, 51.21) -- (215.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.65, 51.21) -- (211.65, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.55, 51.21) -- (198.55, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.87, 51.21) -- (219.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.62, 51.21) -- (223.62, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.92, 51.21) -- (220.92, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.66, 51.21) -- (201.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.77, 51.21) -- (197.77, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.26, 51.21) -- (214.26, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.08, 51.21) -- (185.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.71, 51.21) -- (220.71, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.03, 51.21) -- (193.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.82, 51.21) -- (210.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.79, 51.21) -- (224.79, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.39, 51.21) -- (221.39, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.66, 51.21) -- (214.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.05, 51.21) -- (219.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.62, 51.21) -- (188.62, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.52, 51.21) -- (212.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.44, 51.21) -- (224.44, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.65, 51.21) -- (218.65, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.56, 51.21) -- (209.56, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.10, 51.21) -- (200.10, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.55, 51.21) -- (222.55, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.95, 51.21) -- (213.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.53, 51.21) -- (189.53, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.78, 51.21) -- (209.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.87, 51.21) -- (202.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.84, 51.21) -- (202.84, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.18, 51.21) -- (216.18, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.92, 51.21) -- (208.92, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.29, 51.21) -- (202.29, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.63, 51.21) -- (212.63, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.86, 51.21) -- (211.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.08, 51.21) -- (214.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.94, 51.21) -- (195.94, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.61, 51.21) -- (205.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (165.24, 51.21) -- (165.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.32, 51.21) -- (216.32, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.45, 51.21) -- (225.45, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.16, 51.21) -- (203.16, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.38, 51.21) -- (215.38, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.62, 51.21) -- (206.62, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.13, 51.21) -- (209.13, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.42, 51.21) -- (197.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.24, 51.21) -- (220.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.96, 51.21) -- (220.96, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.94, 51.21) -- (187.94, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.95, 51.21) -- (211.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.55, 51.21) -- (219.55, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.22, 51.21) -- (184.22, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.41, 51.21) -- (221.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.45, 51.21) -- (215.45, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.97, 51.21) -- (225.97, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.04, 51.21) -- (219.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.72, 51.21) -- (207.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.08, 51.21) -- (214.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.45, 51.21) -- (218.45, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.56, 51.21) -- (221.56, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.40, 51.21) -- (203.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.52, 51.21) -- (205.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.37, 51.21) -- (203.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.52, 51.21) -- (205.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.18, 51.21) -- (210.18, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.04, 51.21) -- (212.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.90, 51.21) -- (223.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.73, 51.21) -- (224.73, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.01, 51.21) -- (223.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.17, 51.21) -- (217.17, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.39, 51.21) -- (203.39, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.21, 51.21) -- (210.21, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.04, 51.21) -- (219.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.41, 51.21) -- (210.41, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.33, 51.21) -- (221.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.56, 51.21) -- (224.56, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.83, 51.21) -- (217.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.08, 51.21) -- (182.08, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.65, 51.21) -- (214.65, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.91, 51.21) -- (223.91, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.75, 51.21) -- (214.75, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.14, 51.21) -- (217.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.28, 51.21) -- (225.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.76, 51.21) -- (218.76, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.40, 51.21) -- (216.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.89, 51.21) -- (216.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.76, 51.21) -- (213.76, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.92, 51.21) -- (197.92, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.89, 51.21) -- (219.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.00, 51.21) -- (190.00, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.48, 51.21) -- (177.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.46, 51.21) -- (189.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.67, 51.21) -- (198.67, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.03, 51.21) -- (200.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.40, 51.21) -- (206.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.03, 51.21) -- (200.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.30, 51.21) -- (196.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.82, 51.21) -- (201.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.48, 51.21) -- (177.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.01, 51.21) -- (206.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.95, 51.21) -- (184.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.87, 51.21) -- (203.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.89, 51.21) -- (202.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.39, 51.21) -- (203.39, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.83, 51.21) -- (187.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.48, 51.21) -- (177.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.28, 51.21) -- (204.28, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.37, 51.21) -- (202.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.46, 51.21) -- (189.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.52, 51.21) -- (172.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.67, 51.21) -- (190.67, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.30, 51.21) -- (176.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.03, 51.21) -- (200.03, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.30, 51.21) -- (196.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.77, 51.21) -- (204.77, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.89, 51.21) -- (202.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.33, 51.21) -- (204.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.24, 51.21) -- (184.24, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.82, 51.21) -- (201.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.30, 51.21) -- (198.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.84, 51.21) -- (207.84, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.52, 51.21) -- (172.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.48, 51.21) -- (177.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.83, 51.21) -- (187.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.37, 51.21) -- (202.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.48, 51.21) -- (177.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.51, 51.21) -- (196.51, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.50, 51.21) -- (208.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.93, 51.21) -- (195.93, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.05, 51.21) -- (164.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.01, 51.21) -- (206.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.83, 51.21) -- (187.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.67, 51.21) -- (198.67, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.42, 51.21) -- (189.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.82, 51.21) -- (196.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.05, 51.21) -- (164.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.30, 51.21) -- (196.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.37, 51.21) -- (202.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.05, 51.21) -- (164.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.01, 51.21) -- (206.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.46, 51.21) -- (189.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.05, 51.21) -- (164.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.40, 51.21) -- (206.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.35, 51.21) -- (193.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.30, 51.21) -- (196.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.86, 51.21) -- (204.86, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.05, 51.21) -- (164.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.33, 51.21) -- (204.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.82, 51.21) -- (203.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.30, 51.21) -- (196.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.33, 51.21) -- (204.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.35, 51.21) -- (193.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.05, 51.21) -- (207.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.90, 51.21) -- (190.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.50, 51.21) -- (207.50, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.52, 51.21) -- (172.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.90, 51.21) -- (190.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.88, 51.21) -- (185.88, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.01, 51.21) -- (210.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.61, 51.21) -- (204.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.20, 51.21) -- (205.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.31, 51.21) -- (190.31, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (157.46, 51.21) -- (157.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.61, 51.21) -- (205.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.32, 51.21) -- (180.32, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.84, 51.21) -- (207.84, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.37, 51.21) -- (202.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.83, 51.21) -- (187.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.19, 51.21) -- (192.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.61, 51.21) -- (205.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.01, 51.21) -- (206.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.54, 51.21) -- (204.54, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.93, 51.21) -- (178.93, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.20, 51.21) -- (205.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.89, 51.21) -- (202.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.23, 51.21) -- (182.23, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.37, 51.21) -- (199.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.90, 51.21) -- (190.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.19, 51.21) -- (199.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.90, 51.21) -- (190.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.57, 51.21) -- (190.57, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.93, 51.21) -- (197.93, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.14, 51.21) -- (207.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.80, 51.21) -- (196.80, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.19, 51.21) -- (183.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.33, 51.21) -- (204.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.26, 51.21) -- (201.26, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.31, 51.21) -- (180.31, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.19, 51.21) -- (196.19, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.52, 51.21) -- (172.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.47, 51.21) -- (203.47, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.25, 51.21) -- (176.25, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.35, 51.21) -- (193.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.20, 51.21) -- (205.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.37, 51.21) -- (202.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.67, 51.21) -- (198.67, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.37, 51.21) -- (202.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.25, 51.21) -- (177.25, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.82, 51.21) -- (194.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.77, 51.21) -- (204.77, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.37, 51.21) -- (199.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.89, 51.21) -- (202.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.18, 51.21) -- (195.18, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.78, 51.21) -- (178.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.67, 51.21) -- (198.67, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.46, 51.21) -- (189.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.04, 51.21) -- (194.04, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.35, 51.21) -- (193.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.48, 51.21) -- (177.48, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.05, 51.21) -- (164.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.57, 51.21) -- (199.57, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.52, 51.21) -- (172.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.01, 51.21) -- (206.01, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.70, 51.21) -- (201.70, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.83, 51.21) -- (187.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.46, 51.21) -- (189.46, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.66, 51.21) -- (200.66, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.87, 51.21) -- (203.87, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.52, 51.21) -- (172.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.14, 51.21) -- (197.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.82, 51.21) -- (201.82, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.36, 51.21) -- (166.36, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.25, 51.21) -- (202.25, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.30, 51.21) -- (196.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.40, 51.21) -- (206.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.37, 51.21) -- (199.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.99, 51.21) -- (180.99, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.35, 51.21) -- (193.35, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.65, 51.21) -- (199.65, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.89, 51.21) -- (202.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.78, 51.21) -- (183.78, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.95, 51.21) -- (185.95, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.52, 51.21) -- (172.52, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.40, 51.21) -- (186.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.83, 51.21) -- (187.83, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.90, 51.21) -- (190.90, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.07, 51.21) -- (193.07, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.71, 51.21) -- (204.71, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.20, 51.21) -- (205.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.39, 51.21) -- (203.39, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.05, 51.21) -- (164.05, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.21, 51.21) -- (200.21, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.72, 51.21) -- (183.72, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.92, 51.21) -- (192.92, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.37, 51.21) -- (199.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.37, 51.21) -- (202.37, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.20, 51.21) -- (205.20, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.67, 51.21) -- (198.67, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.07, 51.21) -- (177.07, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.30, 51.21) -- (196.30, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.33, 51.21) -- (204.33, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.40, 51.21) -- (195.40, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.89, 51.21) -- (200.89, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.61, 51.21) -- (205.61, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.92, 51.21) -- (201.92, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.14, 51.21) -- (197.14, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.53, 51.21) -- (197.53, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.42, 51.21) -- (194.42, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.93, 51.21) -- (180.93, 57.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.82, 51.21) -- (201.82, 57.71);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37, 51.21) rectangle (283.58,193.24);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 54.91) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 80.74) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,106.56) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,132.38) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,158.21) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,184.03) {0.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 57.67) --
	( 31.37, 57.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 83.49) --
	( 31.37, 83.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,109.32) --
	( 31.37,109.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,135.14) --
	( 31.37,135.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,160.96) --
	( 31.37,160.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,186.79) --
	( 31.37,186.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.46, 48.46) --
	( 78.46, 51.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (127.34, 48.46) --
	(127.34, 51.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (176.22, 48.46) --
	(176.22, 51.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.10, 48.46) --
	(225.10, 51.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.98, 48.46) --
	(273.98, 51.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 78.46, 40.75) {0.000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (127.34, 40.75) {0.002};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (176.22, 40.75) {0.135};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (225.10, 40.75) {7.389};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (273.98, 40.75) {403.429};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (157.47, 30.25) {Cumulative exposure (mg/m$^3\cdot\,$years)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,122.23) {Density};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 90.36,  5.50) rectangle (224.59, 26.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (101.36, 11.00) rectangle (111.36, 21.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line cap=rect] (102.07, 11.71) rectangle (110.65, 20.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (138.35, 11.00) rectangle (148.35, 21.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 7pt off 3pt ,line cap=rect] (139.06, 11.71) rectangle (147.64, 20.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (170.85, 11.00) rectangle (180.85, 21.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line cap=rect] (171.56, 11.71) rectangle (180.14, 20.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (116.86, 12.90) {0.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (153.85, 12.90) {0.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (186.35, 12.90) {No limit};
\end{scope}
\end{tikzpicture}

\end{document}