### 2022-06-03 resources/shift-soluble.tex

% Created by tikzDevice version 0.12.3.1 on 2022-06-02 13:14:22
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,198.74);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,198.74);
\end{scope}
\begin{scope}
\path[clip] ( 31.37, 51.21) rectangle (283.58,193.24);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37, 51.21) rectangle (283.58,193.24);

\path[] ( 42.83, 57.70) --
	( 43.28, 57.71) --
	( 43.73, 57.71) --
	( 44.18, 57.71) --
	( 44.63, 57.71) --
	( 45.08, 57.71) --
	( 45.52, 57.71) --
	( 45.97, 57.71) --
	( 46.42, 57.71) --
	( 46.87, 57.72) --
	( 47.32, 57.72) --
	( 47.77, 57.72) --
	( 48.22, 57.73) --
	( 48.67, 57.74) --
	( 49.11, 57.75) --
	( 49.56, 57.76) --
	( 50.01, 57.77) --
	( 50.46, 57.78) --
	( 50.91, 57.78) --
	( 51.36, 57.79) --
	( 51.81, 57.78) --
	( 52.25, 57.78) --
	( 52.70, 57.77) --
	( 53.15, 57.77) --
	( 53.60, 57.76) --
	( 54.05, 57.75) --
	( 54.50, 57.75) --
	( 54.95, 57.75) --
	( 55.40, 57.75) --
	( 55.84, 57.75) --
	( 56.29, 57.76) --
	( 56.74, 57.77) --
	( 57.19, 57.77) --
	( 57.64, 57.78) --
	( 58.09, 57.79) --
	( 58.54, 57.80) --
	( 58.99, 57.81) --
	( 59.43, 57.81) --
	( 59.88, 57.82) --
	( 60.33, 57.83) --
	( 60.78, 57.83) --
	( 61.23, 57.84) --
	( 61.68, 57.84) --
	( 62.13, 57.84) --
	( 62.57, 57.84) --
	( 63.02, 57.85) --
	( 63.47, 57.85) --
	( 63.92, 57.86) --
	( 64.37, 57.87) --
	( 64.82, 57.88) --
	( 65.27, 57.89) --
	( 65.72, 57.90) --
	( 66.16, 57.91) --
	( 66.61, 57.91) --
	( 67.06, 57.91) --
	( 67.51, 57.91) --
	( 67.96, 57.91) --
	( 68.41, 57.90) --
	( 68.86, 57.90) --
	( 69.31, 57.90) --
	( 69.75, 57.90) --
	( 70.20, 57.90) --
	( 70.65, 57.90) --
	( 71.10, 57.90) --
	( 71.55, 57.91) --
	( 72.00, 57.92) --
	( 72.45, 57.94) --
	( 72.89, 57.95) --
	( 73.34, 57.97) --
	( 73.79, 58.00) --
	( 74.24, 58.02) --
	( 74.69, 58.05) --
	( 75.14, 58.07) --
	( 75.59, 58.10) --
	( 76.04, 58.13) --
	( 76.48, 58.16) --
	( 76.93, 58.19) --
	( 77.38, 58.22) --
	( 77.83, 58.24) --
	( 78.28, 58.25) --
	( 78.73, 58.26) --
	( 79.18, 58.27) --
	( 79.63, 58.27) --
	( 80.07, 58.26) --
	( 80.52, 58.26) --
	( 80.97, 58.25) --
	( 81.42, 58.24) --
	( 81.87, 58.23) --
	( 82.32, 58.22) --
	( 82.77, 58.22) --
	( 83.21, 58.22) --
	( 83.66, 58.22) --
	( 84.11, 58.24) --
	( 84.56, 58.25) --
	( 85.01, 58.28) --
	( 85.46, 58.31) --
	( 85.91, 58.35) --
	( 86.36, 58.40) --
	( 86.80, 58.46) --
	( 87.25, 58.52) --
	( 87.70, 58.59) --
	( 88.15, 58.65) --
	( 88.60, 58.71) --
	( 89.05, 58.75) --
	( 89.50, 58.77) --
	( 89.95, 58.78) --
	( 90.39, 58.76) --
	( 90.84, 58.74) --
	( 91.29, 58.71) --
	( 91.74, 58.69) --
	( 92.19, 58.68) --
	( 92.64, 58.68) --
	( 93.09, 58.70) --
	( 93.53, 58.73) --
	( 93.98, 58.78) --
	( 94.43, 58.84) --
	( 94.88, 58.89) --
	( 95.33, 58.95) --
	( 95.78, 59.00) --
	( 96.23, 59.04) --
	( 96.68, 59.09) --
	( 97.12, 59.12) --
	( 97.57, 59.16) --
	( 98.02, 59.20) --
	( 98.47, 59.24) --
	( 98.92, 59.29) --
	( 99.37, 59.33) --
	( 99.82, 59.38) --
	(100.27, 59.42) --
	(100.71, 59.46) --
	(101.16, 59.49) --
	(101.61, 59.53) --
	(102.06, 59.57) --
	(102.51, 59.62) --
	(102.96, 59.68) --
	(103.41, 59.74) --
	(103.85, 59.82) --
	(104.30, 59.90) --
	(104.75, 59.98) --
	(105.20, 60.07) --
	(105.65, 60.16) --
	(106.10, 60.23) --
	(106.55, 60.29) --
	(107.00, 60.33) --
	(107.44, 60.33) --
	(107.89, 60.32) --
	(108.34, 60.27) --
	(108.79, 60.21) --
	(109.24, 60.15) --
	(109.69, 60.09) --
	(110.14, 60.05) --
	(110.59, 60.03) --
	(111.03, 60.04) --
	(111.48, 60.07) --
	(111.93, 60.13) --
	(112.38, 60.21) --
	(112.83, 60.31) --
	(113.28, 60.42) --
	(113.73, 60.54) --
	(114.17, 60.68) --
	(114.62, 60.83) --
	(115.07, 61.03) --
	(115.52, 61.32) --
	(115.97, 61.75) --
	(116.42, 62.41) --
	(116.87, 63.41) --
	(117.32, 64.87) --
	(117.76, 66.87) --
	(118.21, 69.43) --
	(118.66, 72.50) --
	(119.11, 75.89) --
	(119.56, 79.33) --
	(120.01, 82.43) --
	(120.46, 84.82) --
	(120.91, 86.17) --
	(121.35, 86.30) --
	(121.80, 85.21) --
	(122.25, 83.04) --
	(122.70, 80.20) --
	(123.15, 77.10) --
	(123.60, 74.08) --
	(124.05, 71.39) --
	(124.49, 69.17) --
	(124.94, 67.48) --
	(125.39, 66.29) --
	(125.84, 65.51) --
	(126.29, 65.07) --
	(126.74, 64.91) --
	(127.19, 64.99) --
	(127.64, 65.33) --
	(128.08, 65.97) --
	(128.53, 67.02) --
	(128.98, 68.59) --
	(129.43, 70.86) --
	(129.88, 73.86) --
	(130.33, 77.54) --
	(130.78, 81.72) --
	(131.23, 86.09) --
	(131.67, 90.19) --
	(132.12, 93.54) --
	(132.57, 95.71) --
	(133.02, 96.43) --
	(133.47, 95.65) --
	(133.92, 93.57) --
	(134.37, 90.61) --
	(134.81, 87.32) --
	(135.26, 84.28) --
	(135.71, 82.02) --
	(136.16, 80.96) --
	(136.61, 81.39) --
	(137.06, 83.20) --
	(137.51, 86.19) --
	(137.96, 89.97) --
	(138.40, 94.06) --
	(138.85, 97.92) --
	(139.30,101.10) --
	(139.75,103.30) --
	(140.20,104.46) --
	(140.65,104.81) --
	(141.10,104.78) --
	(141.55,104.91) --
	(141.99,105.73) --
	(142.44,107.60) --
	(142.89,110.57) --
	(143.34,114.42) --
	(143.79,118.64) --
	(144.24,122.65) --
	(144.69,126.06) --
	(145.13,128.71) --
	(145.58,130.72) --
	(146.03,132.38) --
	(146.48,134.01) --
	(146.93,135.86) --
	(147.38,137.97) --
	(147.83,140.22) --
	(148.28,142.36) --
	(148.72,144.16) --
	(149.17,145.46) --
	(149.62,146.24) --
	(150.07,146.61) --
	(150.52,146.73) --
	(150.97,146.72) --
	(151.42,146.65) --
	(151.87,146.50) --
	(152.31,146.24) --
	(152.76,145.86) --
	(153.21,145.36) --
	(153.66,144.76) --
	(154.11,144.13) --
	(154.56,143.52) --
	(155.01,143.00) --
	(155.45,142.61) --
	(155.90,142.39) --
	(156.35,142.38) --
	(156.80,142.59) --
	(157.25,142.99) --
	(157.70,143.56) --
	(158.15,144.27) --
	(158.60,145.06) --
	(159.04,145.91) --
	(159.49,146.80) --
	(159.94,147.73) --
	(160.39,148.67) --
	(160.84,149.62) --
	(161.29,150.54) --
	(161.74,151.39) --
	(162.19,152.14) --
	(162.63,152.78) --
	(163.08,153.32) --
	(163.53,153.81) --
	(163.98,154.31) --
	(164.43,154.92) --
	(164.88,155.75) --
	(165.33,156.89) --
	(165.77,158.39) --
	(166.22,160.28) --
	(166.67,162.54) --
	(167.12,165.11) --
	(167.57,167.88) --
	(168.02,170.70) --
	(168.47,173.39) --
	(168.92,175.76) --
	(169.36,177.63) --
	(169.81,178.86) --
	(170.26,179.34) --
	(170.71,179.06) --
	(171.16,178.08) --
	(171.61,176.54) --
	(172.06,174.61) --
	(172.51,172.54) --
	(172.95,170.52) --
	(173.40,168.69) --
	(173.85,167.12) --
	(174.30,165.82) --
	(174.75,164.74) --
	(175.20,163.78) --
	(175.65,162.80) --
	(176.09,161.68) --
	(176.54,160.26) --
	(176.99,158.41) --
	(177.44,156.03) --
	(177.89,153.00) --
	(178.34,149.26) --
	(178.79,144.76) --
	(179.24,139.49) --
	(179.68,133.57) --
	(180.13,127.15) --
	(180.58,120.44) --
	(181.03,113.65) --
	(181.48,107.01) --
	(181.93,100.68) --
	(182.38, 94.80) --
	(182.83, 89.43) --
	(183.27, 84.61) --
	(183.72, 80.31) --
	(184.17, 76.50) --
	(184.62, 73.14) --
	(185.07, 70.18) --
	(185.52, 67.62) --
	(185.97, 65.41) --
	(186.41, 63.58) --
	(186.86, 62.07) --
	(187.31, 60.87) --
	(187.76, 59.93) --
	(188.21, 59.22) --
	(188.66, 58.71) --
	(189.11, 58.34) --
	(189.56, 58.09) --
	(190.00, 57.92) --
	(190.45, 57.82) --
	(190.90, 57.75) --
	(191.35, 57.72) --
	(191.80, 57.69) --
	(192.25, 57.68) --
	(192.70, 57.67) --
	(193.15, 57.67) --
	(193.59, 57.67) --
	(194.04, 57.67) --
	(194.49, 57.67) --
	(194.94, 57.67) --
	(195.39, 57.67) --
	(195.84, 57.67) --
	(196.29, 57.67) --
	(196.73, 57.67) --
	(197.18, 57.67) --
	(197.63, 57.67) --
	(198.08, 57.67) --
	(198.53, 57.67) --
	(198.98, 57.67) --
	(199.43, 57.67) --
	(199.88, 57.67) --
	(200.32, 57.67) --
	(200.77, 57.67) --
	(201.22, 57.67) --
	(201.67, 57.67) --
	(202.12, 57.67) --
	(202.57, 57.67) --
	(203.02, 57.67) --
	(203.47, 57.67) --
	(203.91, 57.67) --
	(204.36, 57.67) --
	(204.81, 57.67) --
	(205.26, 57.67) --
	(205.71, 57.67) --
	(206.16, 57.67) --
	(206.61, 57.67) --
	(207.05, 57.67) --
	(207.50, 57.67) --
	(207.95, 57.67) --
	(208.40, 57.67) --
	(208.85, 57.67) --
	(209.30, 57.67) --
	(209.75, 57.67) --
	(210.20, 57.67) --
	(210.64, 57.67) --
	(211.09, 57.67) --
	(211.54, 57.67) --
	(211.99, 57.67) --
	(212.44, 57.67) --
	(212.89, 57.67) --
	(213.34, 57.67) --
	(213.79, 57.67) --
	(214.23, 57.67) --
	(214.68, 57.67) --
	(215.13, 57.67) --
	(215.58, 57.67) --
	(216.03, 57.67) --
	(216.48, 57.67) --
	(216.93, 57.67) --
	(217.37, 57.67) --
	(217.82, 57.67) --
	(218.27, 57.67) --
	(218.72, 57.67) --
	(219.17, 57.67) --
	(219.62, 57.67) --
	(220.07, 57.67) --
	(220.52, 57.67) --
	(220.96, 57.67) --
	(221.41, 57.67) --
	(221.86, 57.67) --
	(222.31, 57.67) --
	(222.76, 57.67) --
	(223.21, 57.67) --
	(223.66, 57.67) --
	(224.11, 57.67) --
	(224.55, 57.67) --
	(225.00, 57.67) --
	(225.45, 57.67) --
	(225.90, 57.67) --
	(226.35, 57.67) --
	(226.80, 57.67) --
	(227.25, 57.67) --
	(227.69, 57.67) --
	(228.14, 57.67) --
	(228.59, 57.67) --
	(229.04, 57.67) --
	(229.49, 57.67) --
	(229.94, 57.67) --
	(230.39, 57.67) --
	(230.84, 57.67) --
	(231.28, 57.67) --
	(231.73, 57.67) --
	(232.18, 57.67) --
	(232.63, 57.67) --
	(233.08, 57.67) --
	(233.53, 57.67) --
	(233.98, 57.67) --
	(234.43, 57.67) --
	(234.87, 57.67) --
	(235.32, 57.67) --
	(235.77, 57.67) --
	(236.22, 57.67) --
	(236.67, 57.67) --
	(237.12, 57.67) --
	(237.57, 57.67) --
	(238.01, 57.67) --
	(238.46, 57.67) --
	(238.91, 57.67) --
	(239.36, 57.67) --
	(239.81, 57.67) --
	(240.26, 57.67) --
	(240.71, 57.67) --
	(241.16, 57.67) --
	(241.60, 57.67) --
	(242.05, 57.67) --
	(242.50, 57.67) --
	(242.95, 57.67) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67) --
	(272.12, 57.67) --
	(271.67, 57.67) --
	(271.22, 57.67) --
	(270.77, 57.67) --
	(270.32, 57.67) --
	(269.87, 57.67) --
	(269.42, 57.67) --
	(268.97, 57.67) --
	(268.53, 57.67) --
	(268.08, 57.67) --
	(267.63, 57.67) --
	(267.18, 57.67) --
	(266.73, 57.67) --
	(266.28, 57.67) --
	(265.83, 57.67) --
	(265.39, 57.67) --
	(264.94, 57.67) --
	(264.49, 57.67) --
	(264.04, 57.67) --
	(263.59, 57.67) --
	(263.14, 57.67) --
	(262.69, 57.67) --
	(262.24, 57.67) --
	(261.80, 57.67) --
	(261.35, 57.67) --
	(260.90, 57.67) --
	(260.45, 57.67) --
	(260.00, 57.67) --
	(259.55, 57.67) --
	(259.10, 57.67) --
	(258.65, 57.67) --
	(258.21, 57.67) --
	(257.76, 57.67) --
	(257.31, 57.67) --
	(256.86, 57.67) --
	(256.41, 57.67) --
	(255.96, 57.67) --
	(255.51, 57.67) --
	(255.07, 57.67) --
	(254.62, 57.67) --
	(254.17, 57.67) --
	(253.72, 57.67) --
	(253.27, 57.67) --
	(252.82, 57.67) --
	(252.37, 57.67) --
	(251.92, 57.67) --
	(251.48, 57.67) --
	(251.03, 57.67) --
	(250.58, 57.67) --
	(250.13, 57.67) --
	(249.68, 57.67) --
	(249.23, 57.67) --
	(248.78, 57.67) --
	(248.33, 57.67) --
	(247.89, 57.67) --
	(247.44, 57.67) --
	(246.99, 57.67) --
	(246.54, 57.67) --
	(246.09, 57.67) --
	(245.64, 57.67) --
	(245.19, 57.67) --
	(244.75, 57.67) --
	(244.30, 57.67) --
	(243.85, 57.67) --
	(243.40, 57.67) --
	(242.95, 57.67) --
	(242.50, 57.67) --
	(242.05, 57.67) --
	(241.60, 57.67) --
	(241.16, 57.67) --
	(240.71, 57.67) --
	(240.26, 57.67) --
	(239.81, 57.67) --
	(239.36, 57.67) --
	(238.91, 57.67) --
	(238.46, 57.67) --
	(238.01, 57.67) --
	(237.57, 57.67) --
	(237.12, 57.67) --
	(236.67, 57.67) --
	(236.22, 57.67) --
	(235.77, 57.67) --
	(235.32, 57.67) --
	(234.87, 57.67) --
	(234.43, 57.67) --
	(233.98, 57.67) --
	(233.53, 57.67) --
	(233.08, 57.67) --
	(232.63, 57.67) --
	(232.18, 57.67) --
	(231.73, 57.67) --
	(231.28, 57.67) --
	(230.84, 57.67) --
	(230.39, 57.67) --
	(229.94, 57.67) --
	(229.49, 57.67) --
	(229.04, 57.67) --
	(228.59, 57.67) --
	(228.14, 57.67) --
	(227.69, 57.67) --
	(227.25, 57.67) --
	(226.80, 57.67) --
	(226.35, 57.67) --
	(225.90, 57.67) --
	(225.45, 57.67) --
	(225.00, 57.67) --
	(224.55, 57.67) --
	(224.11, 57.67) --
	(223.66, 57.67) --
	(223.21, 57.67) --
	(222.76, 57.67) --
	(222.31, 57.67) --
	(221.86, 57.67) --
	(221.41, 57.67) --
	(220.96, 57.67) --
	(220.52, 57.67) --
	(220.07, 57.67) --
	(219.62, 57.67) --
	(219.17, 57.67) --
	(218.72, 57.67) --
	(218.27, 57.67) --
	(217.82, 57.67) --
	(217.37, 57.67) --
	(216.93, 57.67) --
	(216.48, 57.67) --
	(216.03, 57.67) --
	(215.58, 57.67) --
	(215.13, 57.67) --
	(214.68, 57.67) --
	(214.23, 57.67) --
	(213.79, 57.67) --
	(213.34, 57.67) --
	(212.89, 57.67) --
	(212.44, 57.67) --
	(211.99, 57.67) --
	(211.54, 57.67) --
	(211.09, 57.67) --
	(210.64, 57.67) --
	(210.20, 57.67) --
	(209.75, 57.67) --
	(209.30, 57.67) --
	(208.85, 57.67) --
	(208.40, 57.67) --
	(207.95, 57.67) --
	(207.50, 57.67) --
	(207.05, 57.67) --
	(206.61, 57.67) --
	(206.16, 57.67) --
	(205.71, 57.67) --
	(205.26, 57.67) --
	(204.81, 57.67) --
	(204.36, 57.67) --
	(203.91, 57.67) --
	(203.47, 57.67) --
	(203.02, 57.67) --
	(202.57, 57.67) --
	(202.12, 57.67) --
	(201.67, 57.67) --
	(201.22, 57.67) --
	(200.77, 57.67) --
	(200.32, 57.67) --
	(199.88, 57.67) --
	(199.43, 57.67) --
	(198.98, 57.67) --
	(198.53, 57.67) --
	(198.08, 57.67) --
	(197.63, 57.67) --
	(197.18, 57.67) --
	(196.73, 57.67) --
	(196.29, 57.67) --
	(195.84, 57.67) --
	(195.39, 57.67) --
	(194.94, 57.67) --
	(194.49, 57.67) --
	(194.04, 57.67) --
	(193.59, 57.67) --
	(193.15, 57.67) --
	(192.70, 57.67) --
	(192.25, 57.67) --
	(191.80, 57.67) --
	(191.35, 57.67) --
	(190.90, 57.67) --
	(190.45, 57.67) --
	(190.00, 57.67) --
	(189.56, 57.67) --
	(189.11, 57.67) --
	(188.66, 57.67) --
	(188.21, 57.67) --
	(187.76, 57.67) --
	(187.31, 57.67) --
	(186.86, 57.67) --
	(186.41, 57.67) --
	(185.97, 57.67) --
	(185.52, 57.67) --
	(185.07, 57.67) --
	(184.62, 57.67) --
	(184.17, 57.67) --
	(183.72, 57.67) --
	(183.27, 57.67) --
	(182.83, 57.67) --
	(182.38, 57.67) --
	(181.93, 57.67) --
	(181.48, 57.67) --
	(181.03, 57.67) --
	(180.58, 57.67) --
	(180.13, 57.67) --
	(179.68, 57.67) --
	(179.24, 57.67) --
	(178.79, 57.67) --
	(178.34, 57.67) --
	(177.89, 57.67) --
	(177.44, 57.67) --
	(176.99, 57.67) --
	(176.54, 57.67) --
	(176.09, 57.67) --
	(175.65, 57.67) --
	(175.20, 57.67) --
	(174.75, 57.67) --
	(174.30, 57.67) --
	(173.85, 57.67) --
	(173.40, 57.67) --
	(172.95, 57.67) --
	(172.51, 57.67) --
	(172.06, 57.67) --
	(171.61, 57.67) --
	(171.16, 57.67) --
	(170.71, 57.67) --
	(170.26, 57.67) --
	(169.81, 57.67) --
	(169.36, 57.67) --
	(168.92, 57.67) --
	(168.47, 57.67) --
	(168.02, 57.67) --
	(167.57, 57.67) --
	(167.12, 57.67) --
	(166.67, 57.67) --
	(166.22, 57.67) --
	(165.77, 57.67) --
	(165.33, 57.67) --
	(164.88, 57.67) --
	(164.43, 57.67) --
	(163.98, 57.67) --
	(163.53, 57.67) --
	(163.08, 57.67) --
	(162.63, 57.67) --
	(162.19, 57.67) --
	(161.74, 57.67) --
	(161.29, 57.67) --
	(160.84, 57.67) --
	(160.39, 57.67) --
	(159.94, 57.67) --
	(159.49, 57.67) --
	(159.04, 57.67) --
	(158.60, 57.67) --
	(158.15, 57.67) --
	(157.70, 57.67) --
	(157.25, 57.67) --
	(156.80, 57.67) --
	(156.35, 57.67) --
	(155.90, 57.67) --
	(155.45, 57.67) --
	(155.01, 57.67) --
	(154.56, 57.67) --
	(154.11, 57.67) --
	(153.66, 57.67) --
	(153.21, 57.67) --
	(152.76, 57.67) --
	(152.31, 57.67) --
	(151.87, 57.67) --
	(151.42, 57.67) --
	(150.97, 57.67) --
	(150.52, 57.67) --
	(150.07, 57.67) --
	(149.62, 57.67) --
	(149.17, 57.67) --
	(148.72, 57.67) --
	(148.28, 57.67) --
	(147.83, 57.67) --
	(147.38, 57.67) --
	(146.93, 57.67) --
	(146.48, 57.67) --
	(146.03, 57.67) --
	(145.58, 57.67) --
	(145.13, 57.67) --
	(144.69, 57.67) --
	(144.24, 57.67) --
	(143.79, 57.67) --
	(143.34, 57.67) --
	(142.89, 57.67) --
	(142.44, 57.67) --
	(141.99, 57.67) --
	(141.55, 57.67) --
	(141.10, 57.67) --
	(140.65, 57.67) --
	(140.20, 57.67) --
	(139.75, 57.67) --
	(139.30, 57.67) --
	(138.85, 57.67) --
	(138.40, 57.67) --
	(137.96, 57.67) --
	(137.51, 57.67) --
	(137.06, 57.67) --
	(136.61, 57.67) --
	(136.16, 57.67) --
	(135.71, 57.67) --
	(135.26, 57.67) --
	(134.81, 57.67) --
	(134.37, 57.67) --
	(133.92, 57.67) --
	(133.47, 57.67) --
	(133.02, 57.67) --
	(132.57, 57.67) --
	(132.12, 57.67) --
	(131.67, 57.67) --
	(131.23, 57.67) --
	(130.78, 57.67) --
	(130.33, 57.67) --
	(129.88, 57.67) --
	(129.43, 57.67) --
	(128.98, 57.67) --
	(128.53, 57.67) --
	(128.08, 57.67) --
	(127.64, 57.67) --
	(127.19, 57.67) --
	(126.74, 57.67) --
	(126.29, 57.67) --
	(125.84, 57.67) --
	(125.39, 57.67) --
	(124.94, 57.67) --
	(124.49, 57.67) --
	(124.05, 57.67) --
	(123.60, 57.67) --
	(123.15, 57.67) --
	(122.70, 57.67) --
	(122.25, 57.67) --
	(121.80, 57.67) --
	(121.35, 57.67) --
	(120.91, 57.67) --
	(120.46, 57.67) --
	(120.01, 57.67) --
	(119.56, 57.67) --
	(119.11, 57.67) --
	(118.66, 57.67) --
	(118.21, 57.67) --
	(117.76, 57.67) --
	(117.32, 57.67) --
	(116.87, 57.67) --
	(116.42, 57.67) --
	(115.97, 57.67) --
	(115.52, 57.67) --
	(115.07, 57.67) --
	(114.62, 57.67) --
	(114.17, 57.67) --
	(113.73, 57.67) --
	(113.28, 57.67) --
	(112.83, 57.67) --
	(112.38, 57.67) --
	(111.93, 57.67) --
	(111.48, 57.67) --
	(111.03, 57.67) --
	(110.59, 57.67) --
	(110.14, 57.67) --
	(109.69, 57.67) --
	(109.24, 57.67) --
	(108.79, 57.67) --
	(108.34, 57.67) --
	(107.89, 57.67) --
	(107.44, 57.67) --
	(107.00, 57.67) --
	(106.55, 57.67) --
	(106.10, 57.67) --
	(105.65, 57.67) --
	(105.20, 57.67) --
	(104.75, 57.67) --
	(104.30, 57.67) --
	(103.85, 57.67) --
	(103.41, 57.67) --
	(102.96, 57.67) --
	(102.51, 57.67) --
	(102.06, 57.67) --
	(101.61, 57.67) --
	(101.16, 57.67) --
	(100.71, 57.67) --
	(100.27, 57.67) --
	( 99.82, 57.67) --
	( 99.37, 57.67) --
	( 98.92, 57.67) --
	( 98.47, 57.67) --
	( 98.02, 57.67) --
	( 97.57, 57.67) --
	( 97.12, 57.67) --
	( 96.68, 57.67) --
	( 96.23, 57.67) --
	( 95.78, 57.67) --
	( 95.33, 57.67) --
	( 94.88, 57.67) --
	( 94.43, 57.67) --
	( 93.98, 57.67) --
	( 93.53, 57.67) --
	( 93.09, 57.67) --
	( 92.64, 57.67) --
	( 92.19, 57.67) --
	( 91.74, 57.67) --
	( 91.29, 57.67) --
	( 90.84, 57.67) --
	( 90.39, 57.67) --
	( 89.95, 57.67) --
	( 89.50, 57.67) --
	( 89.05, 57.67) --
	( 88.60, 57.67) --
	( 88.15, 57.67) --
	( 87.70, 57.67) --
	( 87.25, 57.67) --
	( 86.80, 57.67) --
	( 86.36, 57.67) --
	( 85.91, 57.67) --
	( 85.46, 57.67) --
	( 85.01, 57.67) --
	( 84.56, 57.67) --
	( 84.11, 57.67) --
	( 83.66, 57.67) --
	( 83.21, 57.67) --
	( 82.77, 57.67) --
	( 82.32, 57.67) --
	( 81.87, 57.67) --
	( 81.42, 57.67) --
	( 80.97, 57.67) --
	( 80.52, 57.67) --
	( 80.07, 57.67) --
	( 79.63, 57.67) --
	( 79.18, 57.67) --
	( 78.73, 57.67) --
	( 78.28, 57.67) --
	( 77.83, 57.67) --
	( 77.38, 57.67) --
	( 76.93, 57.67) --
	( 76.48, 57.67) --
	( 76.04, 57.67) --
	( 75.59, 57.67) --
	( 75.14, 57.67) --
	( 74.69, 57.67) --
	( 74.24, 57.67) --
	( 73.79, 57.67) --
	( 73.34, 57.67) --
	( 72.89, 57.67) --
	( 72.45, 57.67) --
	( 72.00, 57.67) --
	( 71.55, 57.67) --
	( 71.10, 57.67) --
	( 70.65, 57.67) --
	( 70.20, 57.67) --
	( 69.75, 57.67) --
	( 69.31, 57.67) --
	( 68.86, 57.67) --
	( 68.41, 57.67) --
	( 67.96, 57.67) --
	( 67.51, 57.67) --
	( 67.06, 57.67) --
	( 66.61, 57.67) --
	( 66.16, 57.67) --
	( 65.72, 57.67) --
	( 65.27, 57.67) --
	( 64.82, 57.67) --
	( 64.37, 57.67) --
	( 63.92, 57.67) --
	( 63.47, 57.67) --
	( 63.02, 57.67) --
	( 62.57, 57.67) --
	( 62.13, 57.67) --
	( 61.68, 57.67) --
	( 61.23, 57.67) --
	( 60.78, 57.67) --
	( 60.33, 57.67) --
	( 59.88, 57.67) --
	( 59.43, 57.67) --
	( 58.99, 57.67) --
	( 58.54, 57.67) --
	( 58.09, 57.67) --
	( 57.64, 57.67) --
	( 57.19, 57.67) --
	( 56.74, 57.67) --
	( 56.29, 57.67) --
	( 55.84, 57.67) --
	( 55.40, 57.67) --
	( 54.95, 57.67) --
	( 54.50, 57.67) --
	( 54.05, 57.67) --
	( 53.60, 57.67) --
	( 53.15, 57.67) --
	( 52.70, 57.67) --
	( 52.25, 57.67) --
	( 51.81, 57.67) --
	( 51.36, 57.67) --
	( 50.91, 57.67) --
	( 50.46, 57.67) --
	( 50.01, 57.67) --
	( 49.56, 57.67) --
	( 49.11, 57.67) --
	( 48.67, 57.67) --
	( 48.22, 57.67) --
	( 47.77, 57.67) --
	( 47.32, 57.67) --
	( 46.87, 57.67) --
	( 46.42, 57.67) --
	( 45.97, 57.67) --
	( 45.52, 57.67) --
	( 45.08, 57.67) --
	( 44.63, 57.67) --
	( 44.18, 57.67) --
	( 43.73, 57.67) --
	( 43.28, 57.67) --
	( 42.83, 57.67) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 42.83, 57.70) --
	( 43.28, 57.71) --
	( 43.73, 57.71) --
	( 44.18, 57.71) --
	( 44.63, 57.71) --
	( 45.08, 57.71) --
	( 45.52, 57.71) --
	( 45.97, 57.71) --
	( 46.42, 57.71) --
	( 46.87, 57.72) --
	( 47.32, 57.72) --
	( 47.77, 57.72) --
	( 48.22, 57.73) --
	( 48.67, 57.74) --
	( 49.11, 57.75) --
	( 49.56, 57.76) --
	( 50.01, 57.77) --
	( 50.46, 57.78) --
	( 50.91, 57.78) --
	( 51.36, 57.79) --
	( 51.81, 57.78) --
	( 52.25, 57.78) --
	( 52.70, 57.77) --
	( 53.15, 57.77) --
	( 53.60, 57.76) --
	( 54.05, 57.75) --
	( 54.50, 57.75) --
	( 54.95, 57.75) --
	( 55.40, 57.75) --
	( 55.84, 57.75) --
	( 56.29, 57.76) --
	( 56.74, 57.77) --
	( 57.19, 57.77) --
	( 57.64, 57.78) --
	( 58.09, 57.79) --
	( 58.54, 57.80) --
	( 58.99, 57.81) --
	( 59.43, 57.81) --
	( 59.88, 57.82) --
	( 60.33, 57.83) --
	( 60.78, 57.83) --
	( 61.23, 57.84) --
	( 61.68, 57.84) --
	( 62.13, 57.84) --
	( 62.57, 57.84) --
	( 63.02, 57.85) --
	( 63.47, 57.85) --
	( 63.92, 57.86) --
	( 64.37, 57.87) --
	( 64.82, 57.88) --
	( 65.27, 57.89) --
	( 65.72, 57.90) --
	( 66.16, 57.91) --
	( 66.61, 57.91) --
	( 67.06, 57.91) --
	( 67.51, 57.91) --
	( 67.96, 57.91) --
	( 68.41, 57.90) --
	( 68.86, 57.90) --
	( 69.31, 57.90) --
	( 69.75, 57.90) --
	( 70.20, 57.90) --
	( 70.65, 57.90) --
	( 71.10, 57.90) --
	( 71.55, 57.91) --
	( 72.00, 57.92) --
	( 72.45, 57.94) --
	( 72.89, 57.95) --
	( 73.34, 57.97) --
	( 73.79, 58.00) --
	( 74.24, 58.02) --
	( 74.69, 58.05) --
	( 75.14, 58.07) --
	( 75.59, 58.10) --
	( 76.04, 58.13) --
	( 76.48, 58.16) --
	( 76.93, 58.19) --
	( 77.38, 58.22) --
	( 77.83, 58.24) --
	( 78.28, 58.25) --
	( 78.73, 58.26) --
	( 79.18, 58.27) --
	( 79.63, 58.27) --
	( 80.07, 58.26) --
	( 80.52, 58.26) --
	( 80.97, 58.25) --
	( 81.42, 58.24) --
	( 81.87, 58.23) --
	( 82.32, 58.22) --
	( 82.77, 58.22) --
	( 83.21, 58.22) --
	( 83.66, 58.22) --
	( 84.11, 58.24) --
	( 84.56, 58.25) --
	( 85.01, 58.28) --
	( 85.46, 58.31) --
	( 85.91, 58.35) --
	( 86.36, 58.40) --
	( 86.80, 58.46) --
	( 87.25, 58.52) --
	( 87.70, 58.59) --
	( 88.15, 58.65) --
	( 88.60, 58.71) --
	( 89.05, 58.75) --
	( 89.50, 58.77) --
	( 89.95, 58.78) --
	( 90.39, 58.76) --
	( 90.84, 58.74) --
	( 91.29, 58.71) --
	( 91.74, 58.69) --
	( 92.19, 58.68) --
	( 92.64, 58.68) --
	( 93.09, 58.70) --
	( 93.53, 58.73) --
	( 93.98, 58.78) --
	( 94.43, 58.84) --
	( 94.88, 58.89) --
	( 95.33, 58.95) --
	( 95.78, 59.00) --
	( 96.23, 59.04) --
	( 96.68, 59.09) --
	( 97.12, 59.12) --
	( 97.57, 59.16) --
	( 98.02, 59.20) --
	( 98.47, 59.24) --
	( 98.92, 59.29) --
	( 99.37, 59.33) --
	( 99.82, 59.38) --
	(100.27, 59.42) --
	(100.71, 59.46) --
	(101.16, 59.49) --
	(101.61, 59.53) --
	(102.06, 59.57) --
	(102.51, 59.62) --
	(102.96, 59.68) --
	(103.41, 59.74) --
	(103.85, 59.82) --
	(104.30, 59.90) --
	(104.75, 59.98) --
	(105.20, 60.07) --
	(105.65, 60.16) --
	(106.10, 60.23) --
	(106.55, 60.29) --
	(107.00, 60.33) --
	(107.44, 60.33) --
	(107.89, 60.32) --
	(108.34, 60.27) --
	(108.79, 60.21) --
	(109.24, 60.15) --
	(109.69, 60.09) --
	(110.14, 60.05) --
	(110.59, 60.03) --
	(111.03, 60.04) --
	(111.48, 60.07) --
	(111.93, 60.13) --
	(112.38, 60.21) --
	(112.83, 60.31) --
	(113.28, 60.42) --
	(113.73, 60.54) --
	(114.17, 60.68) --
	(114.62, 60.83) --
	(115.07, 61.03) --
	(115.52, 61.32) --
	(115.97, 61.75) --
	(116.42, 62.41) --
	(116.87, 63.41) --
	(117.32, 64.87) --
	(117.76, 66.87) --
	(118.21, 69.43) --
	(118.66, 72.50) --
	(119.11, 75.89) --
	(119.56, 79.33) --
	(120.01, 82.43) --
	(120.46, 84.82) --
	(120.91, 86.17) --
	(121.35, 86.30) --
	(121.80, 85.21) --
	(122.25, 83.04) --
	(122.70, 80.20) --
	(123.15, 77.10) --
	(123.60, 74.08) --
	(124.05, 71.39) --
	(124.49, 69.17) --
	(124.94, 67.48) --
	(125.39, 66.29) --
	(125.84, 65.51) --
	(126.29, 65.07) --
	(126.74, 64.91) --
	(127.19, 64.99) --
	(127.64, 65.33) --
	(128.08, 65.97) --
	(128.53, 67.02) --
	(128.98, 68.59) --
	(129.43, 70.86) --
	(129.88, 73.86) --
	(130.33, 77.54) --
	(130.78, 81.72) --
	(131.23, 86.09) --
	(131.67, 90.19) --
	(132.12, 93.54) --
	(132.57, 95.71) --
	(133.02, 96.43) --
	(133.47, 95.65) --
	(133.92, 93.57) --
	(134.37, 90.61) --
	(134.81, 87.32) --
	(135.26, 84.28) --
	(135.71, 82.02) --
	(136.16, 80.96) --
	(136.61, 81.39) --
	(137.06, 83.20) --
	(137.51, 86.19) --
	(137.96, 89.97) --
	(138.40, 94.06) --
	(138.85, 97.92) --
	(139.30,101.10) --
	(139.75,103.30) --
	(140.20,104.46) --
	(140.65,104.81) --
	(141.10,104.78) --
	(141.55,104.91) --
	(141.99,105.73) --
	(142.44,107.60) --
	(142.89,110.57) --
	(143.34,114.42) --
	(143.79,118.64) --
	(144.24,122.65) --
	(144.69,126.06) --
	(145.13,128.71) --
	(145.58,130.72) --
	(146.03,132.38) --
	(146.48,134.01) --
	(146.93,135.86) --
	(147.38,137.97) --
	(147.83,140.22) --
	(148.28,142.36) --
	(148.72,144.16) --
	(149.17,145.46) --
	(149.62,146.24) --
	(150.07,146.61) --
	(150.52,146.73) --
	(150.97,146.72) --
	(151.42,146.65) --
	(151.87,146.50) --
	(152.31,146.24) --
	(152.76,145.86) --
	(153.21,145.36) --
	(153.66,144.76) --
	(154.11,144.13) --
	(154.56,143.52) --
	(155.01,143.00) --
	(155.45,142.61) --
	(155.90,142.39) --
	(156.35,142.38) --
	(156.80,142.59) --
	(157.25,142.99) --
	(157.70,143.56) --
	(158.15,144.27) --
	(158.60,145.06) --
	(159.04,145.91) --
	(159.49,146.80) --
	(159.94,147.73) --
	(160.39,148.67) --
	(160.84,149.62) --
	(161.29,150.54) --
	(161.74,151.39) --
	(162.19,152.14) --
	(162.63,152.78) --
	(163.08,153.32) --
	(163.53,153.81) --
	(163.98,154.31) --
	(164.43,154.92) --
	(164.88,155.75) --
	(165.33,156.89) --
	(165.77,158.39) --
	(166.22,160.28) --
	(166.67,162.54) --
	(167.12,165.11) --
	(167.57,167.88) --
	(168.02,170.70) --
	(168.47,173.39) --
	(168.92,175.76) --
	(169.36,177.63) --
	(169.81,178.86) --
	(170.26,179.34) --
	(170.71,179.06) --
	(171.16,178.08) --
	(171.61,176.54) --
	(172.06,174.61) --
	(172.51,172.54) --
	(172.95,170.52) --
	(173.40,168.69) --
	(173.85,167.12) --
	(174.30,165.82) --
	(174.75,164.74) --
	(175.20,163.78) --
	(175.65,162.80) --
	(176.09,161.68) --
	(176.54,160.26) --
	(176.99,158.41) --
	(177.44,156.03) --
	(177.89,153.00) --
	(178.34,149.26) --
	(178.79,144.76) --
	(179.24,139.49) --
	(179.68,133.57) --
	(180.13,127.15) --
	(180.58,120.44) --
	(181.03,113.65) --
	(181.48,107.01) --
	(181.93,100.68) --
	(182.38, 94.80) --
	(182.83, 89.43) --
	(183.27, 84.61) --
	(183.72, 80.31) --
	(184.17, 76.50) --
	(184.62, 73.14) --
	(185.07, 70.18) --
	(185.52, 67.62) --
	(185.97, 65.41) --
	(186.41, 63.58) --
	(186.86, 62.07) --
	(187.31, 60.87) --
	(187.76, 59.93) --
	(188.21, 59.22) --
	(188.66, 58.71) --
	(189.11, 58.34) --
	(189.56, 58.09) --
	(190.00, 57.92) --
	(190.45, 57.82) --
	(190.90, 57.75) --
	(191.35, 57.72) --
	(191.80, 57.69) --
	(192.25, 57.68) --
	(192.70, 57.67) --
	(193.15, 57.67) --
	(193.59, 57.67) --
	(194.04, 57.67) --
	(194.49, 57.67) --
	(194.94, 57.67) --
	(195.39, 57.67) --
	(195.84, 57.67) --
	(196.29, 57.67) --
	(196.73, 57.67) --
	(197.18, 57.67) --
	(197.63, 57.67) --
	(198.08, 57.67) --
	(198.53, 57.67) --
	(198.98, 57.67) --
	(199.43, 57.67) --
	(199.88, 57.67) --
	(200.32, 57.67) --
	(200.77, 57.67) --
	(201.22, 57.67) --
	(201.67, 57.67) --
	(202.12, 57.67) --
	(202.57, 57.67) --
	(203.02, 57.67) --
	(203.47, 57.67) --
	(203.91, 57.67) --
	(204.36, 57.67) --
	(204.81, 57.67) --
	(205.26, 57.67) --
	(205.71, 57.67) --
	(206.16, 57.67) --
	(206.61, 57.67) --
	(207.05, 57.67) --
	(207.50, 57.67) --
	(207.95, 57.67) --
	(208.40, 57.67) --
	(208.85, 57.67) --
	(209.30, 57.67) --
	(209.75, 57.67) --
	(210.20, 57.67) --
	(210.64, 57.67) --
	(211.09, 57.67) --
	(211.54, 57.67) --
	(211.99, 57.67) --
	(212.44, 57.67) --
	(212.89, 57.67) --
	(213.34, 57.67) --
	(213.79, 57.67) --
	(214.23, 57.67) --
	(214.68, 57.67) --
	(215.13, 57.67) --
	(215.58, 57.67) --
	(216.03, 57.67) --
	(216.48, 57.67) --
	(216.93, 57.67) --
	(217.37, 57.67) --
	(217.82, 57.67) --
	(218.27, 57.67) --
	(218.72, 57.67) --
	(219.17, 57.67) --
	(219.62, 57.67) --
	(220.07, 57.67) --
	(220.52, 57.67) --
	(220.96, 57.67) --
	(221.41, 57.67) --
	(221.86, 57.67) --
	(222.31, 57.67) --
	(222.76, 57.67) --
	(223.21, 57.67) --
	(223.66, 57.67) --
	(224.11, 57.67) --
	(224.55, 57.67) --
	(225.00, 57.67) --
	(225.45, 57.67) --
	(225.90, 57.67) --
	(226.35, 57.67) --
	(226.80, 57.67) --
	(227.25, 57.67) --
	(227.69, 57.67) --
	(228.14, 57.67) --
	(228.59, 57.67) --
	(229.04, 57.67) --
	(229.49, 57.67) --
	(229.94, 57.67) --
	(230.39, 57.67) --
	(230.84, 57.67) --
	(231.28, 57.67) --
	(231.73, 57.67) --
	(232.18, 57.67) --
	(232.63, 57.67) --
	(233.08, 57.67) --
	(233.53, 57.67) --
	(233.98, 57.67) --
	(234.43, 57.67) --
	(234.87, 57.67) --
	(235.32, 57.67) --
	(235.77, 57.67) --
	(236.22, 57.67) --
	(236.67, 57.67) --
	(237.12, 57.67) --
	(237.57, 57.67) --
	(238.01, 57.67) --
	(238.46, 57.67) --
	(238.91, 57.67) --
	(239.36, 57.67) --
	(239.81, 57.67) --
	(240.26, 57.67) --
	(240.71, 57.67) --
	(241.16, 57.67) --
	(241.60, 57.67) --
	(242.05, 57.67) --
	(242.50, 57.67) --
	(242.95, 57.67) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67);

\path[] ( 42.83, 57.70) --
	( 43.28, 57.71) --
	( 43.73, 57.71) --
	( 44.18, 57.71) --
	( 44.63, 57.71) --
	( 45.08, 57.71) --
	( 45.52, 57.71) --
	( 45.97, 57.71) --
	( 46.42, 57.71) --
	( 46.87, 57.72) --
	( 47.32, 57.72) --
	( 47.77, 57.73) --
	( 48.22, 57.73) --
	( 48.67, 57.74) --
	( 49.11, 57.75) --
	( 49.56, 57.76) --
	( 50.01, 57.77) --
	( 50.46, 57.77) --
	( 50.91, 57.78) --
	( 51.36, 57.78) --
	( 51.81, 57.78) --
	( 52.25, 57.78) --
	( 52.70, 57.77) --
	( 53.15, 57.77) --
	( 53.60, 57.76) --
	( 54.05, 57.76) --
	( 54.50, 57.75) --
	( 54.95, 57.75) --
	( 55.40, 57.75) --
	( 55.84, 57.76) --
	( 56.29, 57.76) --
	( 56.74, 57.77) --
	( 57.19, 57.77) --
	( 57.64, 57.78) --
	( 58.09, 57.79) --
	( 58.54, 57.80) --
	( 58.99, 57.80) --
	( 59.43, 57.81) --
	( 59.88, 57.82) --
	( 60.33, 57.82) --
	( 60.78, 57.83) --
	( 61.23, 57.83) --
	( 61.68, 57.84) --
	( 62.13, 57.84) --
	( 62.57, 57.84) --
	( 63.02, 57.85) --
	( 63.47, 57.85) --
	( 63.92, 57.86) --
	( 64.37, 57.87) --
	( 64.82, 57.88) --
	( 65.27, 57.89) --
	( 65.72, 57.90) --
	( 66.16, 57.90) --
	( 66.61, 57.91) --
	( 67.06, 57.91) --
	( 67.51, 57.91) --
	( 67.96, 57.91) --
	( 68.41, 57.90) --
	( 68.86, 57.90) --
	( 69.31, 57.90) --
	( 69.75, 57.90) --
	( 70.20, 57.90) --
	( 70.65, 57.90) --
	( 71.10, 57.91) --
	( 71.55, 57.92) --
	( 72.00, 57.93) --
	( 72.45, 57.94) --
	( 72.89, 57.96) --
	( 73.34, 57.98) --
	( 73.79, 58.00) --
	( 74.24, 58.02) --
	( 74.69, 58.05) --
	( 75.14, 58.07) --
	( 75.59, 58.10) --
	( 76.04, 58.13) --
	( 76.48, 58.16) --
	( 76.93, 58.19) --
	( 77.38, 58.21) --
	( 77.83, 58.23) --
	( 78.28, 58.25) --
	( 78.73, 58.26) --
	( 79.18, 58.26) --
	( 79.63, 58.26) --
	( 80.07, 58.26) --
	( 80.52, 58.26) --
	( 80.97, 58.25) --
	( 81.42, 58.24) --
	( 81.87, 58.23) --
	( 82.32, 58.23) --
	( 82.77, 58.22) --
	( 83.21, 58.22) --
	( 83.66, 58.23) --
	( 84.11, 58.24) --
	( 84.56, 58.26) --
	( 85.01, 58.29) --
	( 85.46, 58.32) --
	( 85.91, 58.36) --
	( 86.36, 58.41) --
	( 86.80, 58.47) --
	( 87.25, 58.52) --
	( 87.70, 58.58) --
	( 88.15, 58.64) --
	( 88.60, 58.69) --
	( 89.05, 58.73) --
	( 89.50, 58.75) --
	( 89.95, 58.76) --
	( 90.39, 58.75) --
	( 90.84, 58.74) --
	( 91.29, 58.72) --
	( 91.74, 58.70) --
	( 92.19, 58.69) --
	( 92.64, 58.70) --
	( 93.09, 58.72) --
	( 93.53, 58.75) --
	( 93.98, 58.79) --
	( 94.43, 58.84) --
	( 94.88, 58.89) --
	( 95.33, 58.94) --
	( 95.78, 58.99) --
	( 96.23, 59.04) --
	( 96.68, 59.08) --
	( 97.12, 59.12) --
	( 97.57, 59.16) --
	( 98.02, 59.20) --
	( 98.47, 59.25) --
	( 98.92, 59.29) --
	( 99.37, 59.33) --
	( 99.82, 59.37) --
	(100.27, 59.41) --
	(100.71, 59.45) --
	(101.16, 59.50) --
	(101.61, 59.54) --
	(102.06, 59.58) --
	(102.51, 59.63) --
	(102.96, 59.69) --
	(103.41, 59.75) --
	(103.85, 59.82) --
	(104.30, 59.90) --
	(104.75, 59.99) --
	(105.20, 60.07) --
	(105.65, 60.14) --
	(106.10, 60.21) --
	(106.55, 60.26) --
	(107.00, 60.29) --
	(107.44, 60.30) --
	(107.89, 60.29) --
	(108.34, 60.25) --
	(108.79, 60.21) --
	(109.24, 60.16) --
	(109.69, 60.11) --
	(110.14, 60.07) --
	(110.59, 60.06) --
	(111.03, 60.07) --
	(111.48, 60.10) --
	(111.93, 60.15) --
	(112.38, 60.23) --
	(112.83, 60.32) --
	(113.28, 60.41) --
	(113.73, 60.52) --
	(114.17, 60.62) --
	(114.62, 60.71) --
	(115.07, 60.80) --
	(115.52, 60.86) --
	(115.97, 60.91) --
	(116.42, 60.94) --
	(116.87, 60.96) --
	(117.32, 60.96) --
	(117.76, 60.95) --
	(118.21, 60.94) --
	(118.66, 60.95) --
	(119.11, 60.97) --
	(119.56, 61.01) --
	(120.01, 61.07) --
	(120.46, 61.16) --
	(120.91, 61.26) --
	(121.35, 61.37) --
	(121.80, 61.49) --
	(122.25, 61.60) --
	(122.70, 61.70) --
	(123.15, 61.78) --
	(123.60, 61.86) --
	(124.05, 61.92) --
	(124.49, 61.97) --
	(124.94, 62.02) --
	(125.39, 62.07) --
	(125.84, 62.12) --
	(126.29, 62.18) --
	(126.74, 62.24) --
	(127.19, 62.31) --
	(127.64, 62.38) --
	(128.08, 62.45) --
	(128.53, 62.52) --
	(128.98, 62.58) --
	(129.43, 62.65) --
	(129.88, 62.72) --
	(130.33, 62.79) --
	(130.78, 62.87) --
	(131.23, 62.95) --
	(131.67, 63.04) --
	(132.12, 63.13) --
	(132.57, 63.22) --
	(133.02, 63.31) --
	(133.47, 63.38) --
	(133.92, 63.44) --
	(134.37, 63.48) --
	(134.81, 63.51) --
	(135.26, 63.53) --
	(135.71, 63.55) --
	(136.16, 63.59) --
	(136.61, 63.64) --
	(137.06, 63.72) --
	(137.51, 63.82) --
	(137.96, 63.95) --
	(138.40, 64.09) --
	(138.85, 64.24) --
	(139.30, 64.40) --
	(139.75, 64.56) --
	(140.20, 64.72) --
	(140.65, 64.88) --
	(141.10, 65.03) --
	(141.55, 65.19) --
	(141.99, 65.36) --
	(142.44, 65.56) --
	(142.89, 65.80) --
	(143.34, 66.12) --
	(143.79, 66.54) --
	(144.24, 67.13) --
	(144.69, 67.91) --
	(145.13, 68.93) --
	(145.58, 70.17) --
	(146.03, 71.63) --
	(146.48, 73.22) --
	(146.93, 74.86) --
	(147.38, 76.42) --
	(147.83, 77.77) --
	(148.28, 78.79) --
	(148.72, 79.40) --
	(149.17, 79.57) --
	(149.62, 79.33) --
	(150.07, 78.77) --
	(150.52, 77.99) --
	(150.97, 77.11) --
	(151.42, 76.28) --
	(151.87, 75.57) --
	(152.31, 75.03) --
	(152.76, 74.67) --
	(153.21, 74.51) --
	(153.66, 74.52) --
	(154.11, 74.68) --
	(154.56, 75.00) --
	(155.01, 75.48) --
	(155.45, 76.14) --
	(155.90, 77.03) --
	(156.35, 78.18) --
	(156.80, 79.65) --
	(157.25, 81.43) --
	(157.70, 83.53) --
	(158.15, 85.86) --
	(158.60, 88.28) --
	(159.04, 90.62) --
	(159.49, 92.70) --
	(159.94, 94.36) --
	(160.39, 95.47) --
	(160.84, 96.01) --
	(161.29, 96.02) --
	(161.74, 95.62) --
	(162.19, 94.99) --
	(162.63, 94.32) --
	(163.08, 93.84) --
	(163.53, 93.70) --
	(163.98, 94.03) --
	(164.43, 94.90) --
	(164.88, 96.25) --
	(165.33, 97.99) --
	(165.77, 99.96) --
	(166.22,102.01) --
	(166.67,104.00) --
	(167.12,105.84) --
	(167.57,107.52) --
	(168.02,109.08) --
	(168.47,110.65) --
	(168.92,112.37) --
	(169.36,114.38) --
	(169.81,116.76) --
	(170.26,119.52) --
	(170.71,122.57) --
	(171.16,125.76) --
	(171.61,128.87) --
	(172.06,131.72) --
	(172.51,134.22) --
	(172.95,136.36) --
	(173.40,138.16) --
	(173.85,139.72) --
	(174.30,141.10) --
	(174.75,142.33) --
	(175.20,143.42) --
	(175.65,144.33) --
	(176.09,145.02) --
	(176.54,145.44) --
	(176.99,145.63) --
	(177.44,145.61) --
	(177.89,145.48) --
	(178.34,145.30) --
	(178.79,145.15) --
	(179.24,145.07) --
	(179.68,145.04) --
	(180.13,145.08) --
	(180.58,145.16) --
	(181.03,145.27) --
	(181.48,145.40) --
	(181.93,145.54) --
	(182.38,145.69) --
	(182.83,145.85) --
	(183.27,146.02) --
	(183.72,146.20) --
	(184.17,146.41) --
	(184.62,146.64) --
	(185.07,146.91) --
	(185.52,147.23) --
	(185.97,147.60) --
	(186.41,148.03) --
	(186.86,148.51) --
	(187.31,149.05) --
	(187.76,149.62) --
	(188.21,150.21) --
	(188.66,150.81) --
	(189.11,151.41) --
	(189.56,151.99) --
	(190.00,152.57) --
	(190.45,153.13) --
	(190.90,153.68) --
	(191.35,154.23) --
	(191.80,154.80) --
	(192.25,155.41) --
	(192.70,156.06) --
	(193.15,156.76) --
	(193.59,157.50) --
	(194.04,158.24) --
	(194.49,158.94) --
	(194.94,159.54) --
	(195.39,159.95) --
	(195.84,160.10) --
	(196.29,159.92) --
	(196.73,159.37) --
	(197.18,158.45) --
	(197.63,157.17) --
	(198.08,155.62) --
	(198.53,153.90) --
	(198.98,152.15) --
	(199.43,150.47) --
	(199.88,148.96) --
	(200.32,147.67) --
	(200.77,146.61) --
	(201.22,145.75) --
	(201.67,145.01) --
	(202.12,144.31) --
	(202.57,143.54) --
	(203.02,142.61) --
	(203.47,141.43) --
	(203.91,139.92) --
	(204.36,138.00) --
	(204.81,135.62) --
	(205.26,132.72) --
	(205.71,129.32) --
	(206.16,125.43) --
	(206.61,121.08) --
	(207.05,116.37) --
	(207.50,111.40) --
	(207.95,106.30) --
	(208.40,101.20) --
	(208.85, 96.24) --
	(209.30, 91.51) --
	(209.75, 87.10) --
	(210.20, 83.04) --
	(210.64, 79.35) --
	(211.09, 76.03) --
	(211.54, 73.07) --
	(211.99, 70.44) --
	(212.44, 68.11) --
	(212.89, 66.07) --
	(213.34, 64.30) --
	(213.79, 62.80) --
	(214.23, 61.55) --
	(214.68, 60.54) --
	(215.13, 59.73) --
	(215.58, 59.12) --
	(216.03, 58.65) --
	(216.48, 58.32) --
	(216.93, 58.09) --
	(217.37, 57.93) --
	(217.82, 57.82) --
	(218.27, 57.76) --
	(218.72, 57.72) --
	(219.17, 57.70) --
	(219.62, 57.68) --
	(220.07, 57.68) --
	(220.52, 57.67) --
	(220.96, 57.67) --
	(221.41, 57.67) --
	(221.86, 57.67) --
	(222.31, 57.67) --
	(222.76, 57.67) --
	(223.21, 57.67) --
	(223.66, 57.67) --
	(224.11, 57.67) --
	(224.55, 57.67) --
	(225.00, 57.67) --
	(225.45, 57.67) --
	(225.90, 57.67) --
	(226.35, 57.67) --
	(226.80, 57.67) --
	(227.25, 57.67) --
	(227.69, 57.67) --
	(228.14, 57.67) --
	(228.59, 57.67) --
	(229.04, 57.67) --
	(229.49, 57.67) --
	(229.94, 57.67) --
	(230.39, 57.67) --
	(230.84, 57.67) --
	(231.28, 57.67) --
	(231.73, 57.67) --
	(232.18, 57.67) --
	(232.63, 57.67) --
	(233.08, 57.67) --
	(233.53, 57.67) --
	(233.98, 57.67) --
	(234.43, 57.67) --
	(234.87, 57.67) --
	(235.32, 57.67) --
	(235.77, 57.67) --
	(236.22, 57.67) --
	(236.67, 57.67) --
	(237.12, 57.67) --
	(237.57, 57.67) --
	(238.01, 57.67) --
	(238.46, 57.67) --
	(238.91, 57.67) --
	(239.36, 57.67) --
	(239.81, 57.67) --
	(240.26, 57.67) --
	(240.71, 57.67) --
	(241.16, 57.67) --
	(241.60, 57.67) --
	(242.05, 57.67) --
	(242.50, 57.67) --
	(242.95, 57.67) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67) --
	(272.12, 57.67) --
	(271.67, 57.67) --
	(271.22, 57.67) --
	(270.77, 57.67) --
	(270.32, 57.67) --
	(269.87, 57.67) --
	(269.42, 57.67) --
	(268.97, 57.67) --
	(268.53, 57.67) --
	(268.08, 57.67) --
	(267.63, 57.67) --
	(267.18, 57.67) --
	(266.73, 57.67) --
	(266.28, 57.67) --
	(265.83, 57.67) --
	(265.39, 57.67) --
	(264.94, 57.67) --
	(264.49, 57.67) --
	(264.04, 57.67) --
	(263.59, 57.67) --
	(263.14, 57.67) --
	(262.69, 57.67) --
	(262.24, 57.67) --
	(261.80, 57.67) --
	(261.35, 57.67) --
	(260.90, 57.67) --
	(260.45, 57.67) --
	(260.00, 57.67) --
	(259.55, 57.67) --
	(259.10, 57.67) --
	(258.65, 57.67) --
	(258.21, 57.67) --
	(257.76, 57.67) --
	(257.31, 57.67) --
	(256.86, 57.67) --
	(256.41, 57.67) --
	(255.96, 57.67) --
	(255.51, 57.67) --
	(255.07, 57.67) --
	(254.62, 57.67) --
	(254.17, 57.67) --
	(253.72, 57.67) --
	(253.27, 57.67) --
	(252.82, 57.67) --
	(252.37, 57.67) --
	(251.92, 57.67) --
	(251.48, 57.67) --
	(251.03, 57.67) --
	(250.58, 57.67) --
	(250.13, 57.67) --
	(249.68, 57.67) --
	(249.23, 57.67) --
	(248.78, 57.67) --
	(248.33, 57.67) --
	(247.89, 57.67) --
	(247.44, 57.67) --
	(246.99, 57.67) --
	(246.54, 57.67) --
	(246.09, 57.67) --
	(245.64, 57.67) --
	(245.19, 57.67) --
	(244.75, 57.67) --
	(244.30, 57.67) --
	(243.85, 57.67) --
	(243.40, 57.67) --
	(242.95, 57.67) --
	(242.50, 57.67) --
	(242.05, 57.67) --
	(241.60, 57.67) --
	(241.16, 57.67) --
	(240.71, 57.67) --
	(240.26, 57.67) --
	(239.81, 57.67) --
	(239.36, 57.67) --
	(238.91, 57.67) --
	(238.46, 57.67) --
	(238.01, 57.67) --
	(237.57, 57.67) --
	(237.12, 57.67) --
	(236.67, 57.67) --
	(236.22, 57.67) --
	(235.77, 57.67) --
	(235.32, 57.67) --
	(234.87, 57.67) --
	(234.43, 57.67) --
	(233.98, 57.67) --
	(233.53, 57.67) --
	(233.08, 57.67) --
	(232.63, 57.67) --
	(232.18, 57.67) --
	(231.73, 57.67) --
	(231.28, 57.67) --
	(230.84, 57.67) --
	(230.39, 57.67) --
	(229.94, 57.67) --
	(229.49, 57.67) --
	(229.04, 57.67) --
	(228.59, 57.67) --
	(228.14, 57.67) --
	(227.69, 57.67) --
	(227.25, 57.67) --
	(226.80, 57.67) --
	(226.35, 57.67) --
	(225.90, 57.67) --
	(225.45, 57.67) --
	(225.00, 57.67) --
	(224.55, 57.67) --
	(224.11, 57.67) --
	(223.66, 57.67) --
	(223.21, 57.67) --
	(222.76, 57.67) --
	(222.31, 57.67) --
	(221.86, 57.67) --
	(221.41, 57.67) --
	(220.96, 57.67) --
	(220.52, 57.67) --
	(220.07, 57.67) --
	(219.62, 57.67) --
	(219.17, 57.67) --
	(218.72, 57.67) --
	(218.27, 57.67) --
	(217.82, 57.67) --
	(217.37, 57.67) --
	(216.93, 57.67) --
	(216.48, 57.67) --
	(216.03, 57.67) --
	(215.58, 57.67) --
	(215.13, 57.67) --
	(214.68, 57.67) --
	(214.23, 57.67) --
	(213.79, 57.67) --
	(213.34, 57.67) --
	(212.89, 57.67) --
	(212.44, 57.67) --
	(211.99, 57.67) --
	(211.54, 57.67) --
	(211.09, 57.67) --
	(210.64, 57.67) --
	(210.20, 57.67) --
	(209.75, 57.67) --
	(209.30, 57.67) --
	(208.85, 57.67) --
	(208.40, 57.67) --
	(207.95, 57.67) --
	(207.50, 57.67) --
	(207.05, 57.67) --
	(206.61, 57.67) --
	(206.16, 57.67) --
	(205.71, 57.67) --
	(205.26, 57.67) --
	(204.81, 57.67) --
	(204.36, 57.67) --
	(203.91, 57.67) --
	(203.47, 57.67) --
	(203.02, 57.67) --
	(202.57, 57.67) --
	(202.12, 57.67) --
	(201.67, 57.67) --
	(201.22, 57.67) --
	(200.77, 57.67) --
	(200.32, 57.67) --
	(199.88, 57.67) --
	(199.43, 57.67) --
	(198.98, 57.67) --
	(198.53, 57.67) --
	(198.08, 57.67) --
	(197.63, 57.67) --
	(197.18, 57.67) --
	(196.73, 57.67) --
	(196.29, 57.67) --
	(195.84, 57.67) --
	(195.39, 57.67) --
	(194.94, 57.67) --
	(194.49, 57.67) --
	(194.04, 57.67) --
	(193.59, 57.67) --
	(193.15, 57.67) --
	(192.70, 57.67) --
	(192.25, 57.67) --
	(191.80, 57.67) --
	(191.35, 57.67) --
	(190.90, 57.67) --
	(190.45, 57.67) --
	(190.00, 57.67) --
	(189.56, 57.67) --
	(189.11, 57.67) --
	(188.66, 57.67) --
	(188.21, 57.67) --
	(187.76, 57.67) --
	(187.31, 57.67) --
	(186.86, 57.67) --
	(186.41, 57.67) --
	(185.97, 57.67) --
	(185.52, 57.67) --
	(185.07, 57.67) --
	(184.62, 57.67) --
	(184.17, 57.67) --
	(183.72, 57.67) --
	(183.27, 57.67) --
	(182.83, 57.67) --
	(182.38, 57.67) --
	(181.93, 57.67) --
	(181.48, 57.67) --
	(181.03, 57.67) --
	(180.58, 57.67) --
	(180.13, 57.67) --
	(179.68, 57.67) --
	(179.24, 57.67) --
	(178.79, 57.67) --
	(178.34, 57.67) --
	(177.89, 57.67) --
	(177.44, 57.67) --
	(176.99, 57.67) --
	(176.54, 57.67) --
	(176.09, 57.67) --
	(175.65, 57.67) --
	(175.20, 57.67) --
	(174.75, 57.67) --
	(174.30, 57.67) --
	(173.85, 57.67) --
	(173.40, 57.67) --
	(172.95, 57.67) --
	(172.51, 57.67) --
	(172.06, 57.67) --
	(171.61, 57.67) --
	(171.16, 57.67) --
	(170.71, 57.67) --
	(170.26, 57.67) --
	(169.81, 57.67) --
	(169.36, 57.67) --
	(168.92, 57.67) --
	(168.47, 57.67) --
	(168.02, 57.67) --
	(167.57, 57.67) --
	(167.12, 57.67) --
	(166.67, 57.67) --
	(166.22, 57.67) --
	(165.77, 57.67) --
	(165.33, 57.67) --
	(164.88, 57.67) --
	(164.43, 57.67) --
	(163.98, 57.67) --
	(163.53, 57.67) --
	(163.08, 57.67) --
	(162.63, 57.67) --
	(162.19, 57.67) --
	(161.74, 57.67) --
	(161.29, 57.67) --
	(160.84, 57.67) --
	(160.39, 57.67) --
	(159.94, 57.67) --
	(159.49, 57.67) --
	(159.04, 57.67) --
	(158.60, 57.67) --
	(158.15, 57.67) --
	(157.70, 57.67) --
	(157.25, 57.67) --
	(156.80, 57.67) --
	(156.35, 57.67) --
	(155.90, 57.67) --
	(155.45, 57.67) --
	(155.01, 57.67) --
	(154.56, 57.67) --
	(154.11, 57.67) --
	(153.66, 57.67) --
	(153.21, 57.67) --
	(152.76, 57.67) --
	(152.31, 57.67) --
	(151.87, 57.67) --
	(151.42, 57.67) --
	(150.97, 57.67) --
	(150.52, 57.67) --
	(150.07, 57.67) --
	(149.62, 57.67) --
	(149.17, 57.67) --
	(148.72, 57.67) --
	(148.28, 57.67) --
	(147.83, 57.67) --
	(147.38, 57.67) --
	(146.93, 57.67) --
	(146.48, 57.67) --
	(146.03, 57.67) --
	(145.58, 57.67) --
	(145.13, 57.67) --
	(144.69, 57.67) --
	(144.24, 57.67) --
	(143.79, 57.67) --
	(143.34, 57.67) --
	(142.89, 57.67) --
	(142.44, 57.67) --
	(141.99, 57.67) --
	(141.55, 57.67) --
	(141.10, 57.67) --
	(140.65, 57.67) --
	(140.20, 57.67) --
	(139.75, 57.67) --
	(139.30, 57.67) --
	(138.85, 57.67) --
	(138.40, 57.67) --
	(137.96, 57.67) --
	(137.51, 57.67) --
	(137.06, 57.67) --
	(136.61, 57.67) --
	(136.16, 57.67) --
	(135.71, 57.67) --
	(135.26, 57.67) --
	(134.81, 57.67) --
	(134.37, 57.67) --
	(133.92, 57.67) --
	(133.47, 57.67) --
	(133.02, 57.67) --
	(132.57, 57.67) --
	(132.12, 57.67) --
	(131.67, 57.67) --
	(131.23, 57.67) --
	(130.78, 57.67) --
	(130.33, 57.67) --
	(129.88, 57.67) --
	(129.43, 57.67) --
	(128.98, 57.67) --
	(128.53, 57.67) --
	(128.08, 57.67) --
	(127.64, 57.67) --
	(127.19, 57.67) --
	(126.74, 57.67) --
	(126.29, 57.67) --
	(125.84, 57.67) --
	(125.39, 57.67) --
	(124.94, 57.67) --
	(124.49, 57.67) --
	(124.05, 57.67) --
	(123.60, 57.67) --
	(123.15, 57.67) --
	(122.70, 57.67) --
	(122.25, 57.67) --
	(121.80, 57.67) --
	(121.35, 57.67) --
	(120.91, 57.67) --
	(120.46, 57.67) --
	(120.01, 57.67) --
	(119.56, 57.67) --
	(119.11, 57.67) --
	(118.66, 57.67) --
	(118.21, 57.67) --
	(117.76, 57.67) --
	(117.32, 57.67) --
	(116.87, 57.67) --
	(116.42, 57.67) --
	(115.97, 57.67) --
	(115.52, 57.67) --
	(115.07, 57.67) --
	(114.62, 57.67) --
	(114.17, 57.67) --
	(113.73, 57.67) --
	(113.28, 57.67) --
	(112.83, 57.67) --
	(112.38, 57.67) --
	(111.93, 57.67) --
	(111.48, 57.67) --
	(111.03, 57.67) --
	(110.59, 57.67) --
	(110.14, 57.67) --
	(109.69, 57.67) --
	(109.24, 57.67) --
	(108.79, 57.67) --
	(108.34, 57.67) --
	(107.89, 57.67) --
	(107.44, 57.67) --
	(107.00, 57.67) --
	(106.55, 57.67) --
	(106.10, 57.67) --
	(105.65, 57.67) --
	(105.20, 57.67) --
	(104.75, 57.67) --
	(104.30, 57.67) --
	(103.85, 57.67) --
	(103.41, 57.67) --
	(102.96, 57.67) --
	(102.51, 57.67) --
	(102.06, 57.67) --
	(101.61, 57.67) --
	(101.16, 57.67) --
	(100.71, 57.67) --
	(100.27, 57.67) --
	( 99.82, 57.67) --
	( 99.37, 57.67) --
	( 98.92, 57.67) --
	( 98.47, 57.67) --
	( 98.02, 57.67) --
	( 97.57, 57.67) --
	( 97.12, 57.67) --
	( 96.68, 57.67) --
	( 96.23, 57.67) --
	( 95.78, 57.67) --
	( 95.33, 57.67) --
	( 94.88, 57.67) --
	( 94.43, 57.67) --
	( 93.98, 57.67) --
	( 93.53, 57.67) --
	( 93.09, 57.67) --
	( 92.64, 57.67) --
	( 92.19, 57.67) --
	( 91.74, 57.67) --
	( 91.29, 57.67) --
	( 90.84, 57.67) --
	( 90.39, 57.67) --
	( 89.95, 57.67) --
	( 89.50, 57.67) --
	( 89.05, 57.67) --
	( 88.60, 57.67) --
	( 88.15, 57.67) --
	( 87.70, 57.67) --
	( 87.25, 57.67) --
	( 86.80, 57.67) --
	( 86.36, 57.67) --
	( 85.91, 57.67) --
	( 85.46, 57.67) --
	( 85.01, 57.67) --
	( 84.56, 57.67) --
	( 84.11, 57.67) --
	( 83.66, 57.67) --
	( 83.21, 57.67) --
	( 82.77, 57.67) --
	( 82.32, 57.67) --
	( 81.87, 57.67) --
	( 81.42, 57.67) --
	( 80.97, 57.67) --
	( 80.52, 57.67) --
	( 80.07, 57.67) --
	( 79.63, 57.67) --
	( 79.18, 57.67) --
	( 78.73, 57.67) --
	( 78.28, 57.67) --
	( 77.83, 57.67) --
	( 77.38, 57.67) --
	( 76.93, 57.67) --
	( 76.48, 57.67) --
	( 76.04, 57.67) --
	( 75.59, 57.67) --
	( 75.14, 57.67) --
	( 74.69, 57.67) --
	( 74.24, 57.67) --
	( 73.79, 57.67) --
	( 73.34, 57.67) --
	( 72.89, 57.67) --
	( 72.45, 57.67) --
	( 72.00, 57.67) --
	( 71.55, 57.67) --
	( 71.10, 57.67) --
	( 70.65, 57.67) --
	( 70.20, 57.67) --
	( 69.75, 57.67) --
	( 69.31, 57.67) --
	( 68.86, 57.67) --
	( 68.41, 57.67) --
	( 67.96, 57.67) --
	( 67.51, 57.67) --
	( 67.06, 57.67) --
	( 66.61, 57.67) --
	( 66.16, 57.67) --
	( 65.72, 57.67) --
	( 65.27, 57.67) --
	( 64.82, 57.67) --
	( 64.37, 57.67) --
	( 63.92, 57.67) --
	( 63.47, 57.67) --
	( 63.02, 57.67) --
	( 62.57, 57.67) --
	( 62.13, 57.67) --
	( 61.68, 57.67) --
	( 61.23, 57.67) --
	( 60.78, 57.67) --
	( 60.33, 57.67) --
	( 59.88, 57.67) --
	( 59.43, 57.67) --
	( 58.99, 57.67) --
	( 58.54, 57.67) --
	( 58.09, 57.67) --
	( 57.64, 57.67) --
	( 57.19, 57.67) --
	( 56.74, 57.67) --
	( 56.29, 57.67) --
	( 55.84, 57.67) --
	( 55.40, 57.67) --
	( 54.95, 57.67) --
	( 54.50, 57.67) --
	( 54.05, 57.67) --
	( 53.60, 57.67) --
	( 53.15, 57.67) --
	( 52.70, 57.67) --
	( 52.25, 57.67) --
	( 51.81, 57.67) --
	( 51.36, 57.67) --
	( 50.91, 57.67) --
	( 50.46, 57.67) --
	( 50.01, 57.67) --
	( 49.56, 57.67) --
	( 49.11, 57.67) --
	( 48.67, 57.67) --
	( 48.22, 57.67) --
	( 47.77, 57.67) --
	( 47.32, 57.67) --
	( 46.87, 57.67) --
	( 46.42, 57.67) --
	( 45.97, 57.67) --
	( 45.52, 57.67) --
	( 45.08, 57.67) --
	( 44.63, 57.67) --
	( 44.18, 57.67) --
	( 43.73, 57.67) --
	( 43.28, 57.67) --
	( 42.83, 57.67) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 7pt off 3pt ,line join=round,line cap=round] ( 42.83, 57.70) --
	( 43.28, 57.71) --
	( 43.73, 57.71) --
	( 44.18, 57.71) --
	( 44.63, 57.71) --
	( 45.08, 57.71) --
	( 45.52, 57.71) --
	( 45.97, 57.71) --
	( 46.42, 57.71) --
	( 46.87, 57.72) --
	( 47.32, 57.72) --
	( 47.77, 57.73) --
	( 48.22, 57.73) --
	( 48.67, 57.74) --
	( 49.11, 57.75) --
	( 49.56, 57.76) --
	( 50.01, 57.77) --
	( 50.46, 57.77) --
	( 50.91, 57.78) --
	( 51.36, 57.78) --
	( 51.81, 57.78) --
	( 52.25, 57.78) --
	( 52.70, 57.77) --
	( 53.15, 57.77) --
	( 53.60, 57.76) --
	( 54.05, 57.76) --
	( 54.50, 57.75) --
	( 54.95, 57.75) --
	( 55.40, 57.75) --
	( 55.84, 57.76) --
	( 56.29, 57.76) --
	( 56.74, 57.77) --
	( 57.19, 57.77) --
	( 57.64, 57.78) --
	( 58.09, 57.79) --
	( 58.54, 57.80) --
	( 58.99, 57.80) --
	( 59.43, 57.81) --
	( 59.88, 57.82) --
	( 60.33, 57.82) --
	( 60.78, 57.83) --
	( 61.23, 57.83) --
	( 61.68, 57.84) --
	( 62.13, 57.84) --
	( 62.57, 57.84) --
	( 63.02, 57.85) --
	( 63.47, 57.85) --
	( 63.92, 57.86) --
	( 64.37, 57.87) --
	( 64.82, 57.88) --
	( 65.27, 57.89) --
	( 65.72, 57.90) --
	( 66.16, 57.90) --
	( 66.61, 57.91) --
	( 67.06, 57.91) --
	( 67.51, 57.91) --
	( 67.96, 57.91) --
	( 68.41, 57.90) --
	( 68.86, 57.90) --
	( 69.31, 57.90) --
	( 69.75, 57.90) --
	( 70.20, 57.90) --
	( 70.65, 57.90) --
	( 71.10, 57.91) --
	( 71.55, 57.92) --
	( 72.00, 57.93) --
	( 72.45, 57.94) --
	( 72.89, 57.96) --
	( 73.34, 57.98) --
	( 73.79, 58.00) --
	( 74.24, 58.02) --
	( 74.69, 58.05) --
	( 75.14, 58.07) --
	( 75.59, 58.10) --
	( 76.04, 58.13) --
	( 76.48, 58.16) --
	( 76.93, 58.19) --
	( 77.38, 58.21) --
	( 77.83, 58.23) --
	( 78.28, 58.25) --
	( 78.73, 58.26) --
	( 79.18, 58.26) --
	( 79.63, 58.26) --
	( 80.07, 58.26) --
	( 80.52, 58.26) --
	( 80.97, 58.25) --
	( 81.42, 58.24) --
	( 81.87, 58.23) --
	( 82.32, 58.23) --
	( 82.77, 58.22) --
	( 83.21, 58.22) --
	( 83.66, 58.23) --
	( 84.11, 58.24) --
	( 84.56, 58.26) --
	( 85.01, 58.29) --
	( 85.46, 58.32) --
	( 85.91, 58.36) --
	( 86.36, 58.41) --
	( 86.80, 58.47) --
	( 87.25, 58.52) --
	( 87.70, 58.58) --
	( 88.15, 58.64) --
	( 88.60, 58.69) --
	( 89.05, 58.73) --
	( 89.50, 58.75) --
	( 89.95, 58.76) --
	( 90.39, 58.75) --
	( 90.84, 58.74) --
	( 91.29, 58.72) --
	( 91.74, 58.70) --
	( 92.19, 58.69) --
	( 92.64, 58.70) --
	( 93.09, 58.72) --
	( 93.53, 58.75) --
	( 93.98, 58.79) --
	( 94.43, 58.84) --
	( 94.88, 58.89) --
	( 95.33, 58.94) --
	( 95.78, 58.99) --
	( 96.23, 59.04) --
	( 96.68, 59.08) --
	( 97.12, 59.12) --
	( 97.57, 59.16) --
	( 98.02, 59.20) --
	( 98.47, 59.25) --
	( 98.92, 59.29) --
	( 99.37, 59.33) --
	( 99.82, 59.37) --
	(100.27, 59.41) --
	(100.71, 59.45) --
	(101.16, 59.50) --
	(101.61, 59.54) --
	(102.06, 59.58) --
	(102.51, 59.63) --
	(102.96, 59.69) --
	(103.41, 59.75) --
	(103.85, 59.82) --
	(104.30, 59.90) --
	(104.75, 59.99) --
	(105.20, 60.07) --
	(105.65, 60.14) --
	(106.10, 60.21) --
	(106.55, 60.26) --
	(107.00, 60.29) --
	(107.44, 60.30) --
	(107.89, 60.29) --
	(108.34, 60.25) --
	(108.79, 60.21) --
	(109.24, 60.16) --
	(109.69, 60.11) --
	(110.14, 60.07) --
	(110.59, 60.06) --
	(111.03, 60.07) --
	(111.48, 60.10) --
	(111.93, 60.15) --
	(112.38, 60.23) --
	(112.83, 60.32) --
	(113.28, 60.41) --
	(113.73, 60.52) --
	(114.17, 60.62) --
	(114.62, 60.71) --
	(115.07, 60.80) --
	(115.52, 60.86) --
	(115.97, 60.91) --
	(116.42, 60.94) --
	(116.87, 60.96) --
	(117.32, 60.96) --
	(117.76, 60.95) --
	(118.21, 60.94) --
	(118.66, 60.95) --
	(119.11, 60.97) --
	(119.56, 61.01) --
	(120.01, 61.07) --
	(120.46, 61.16) --
	(120.91, 61.26) --
	(121.35, 61.37) --
	(121.80, 61.49) --
	(122.25, 61.60) --
	(122.70, 61.70) --
	(123.15, 61.78) --
	(123.60, 61.86) --
	(124.05, 61.92) --
	(124.49, 61.97) --
	(124.94, 62.02) --
	(125.39, 62.07) --
	(125.84, 62.12) --
	(126.29, 62.18) --
	(126.74, 62.24) --
	(127.19, 62.31) --
	(127.64, 62.38) --
	(128.08, 62.45) --
	(128.53, 62.52) --
	(128.98, 62.58) --
	(129.43, 62.65) --
	(129.88, 62.72) --
	(130.33, 62.79) --
	(130.78, 62.87) --
	(131.23, 62.95) --
	(131.67, 63.04) --
	(132.12, 63.13) --
	(132.57, 63.22) --
	(133.02, 63.31) --
	(133.47, 63.38) --
	(133.92, 63.44) --
	(134.37, 63.48) --
	(134.81, 63.51) --
	(135.26, 63.53) --
	(135.71, 63.55) --
	(136.16, 63.59) --
	(136.61, 63.64) --
	(137.06, 63.72) --
	(137.51, 63.82) --
	(137.96, 63.95) --
	(138.40, 64.09) --
	(138.85, 64.24) --
	(139.30, 64.40) --
	(139.75, 64.56) --
	(140.20, 64.72) --
	(140.65, 64.88) --
	(141.10, 65.03) --
	(141.55, 65.19) --
	(141.99, 65.36) --
	(142.44, 65.56) --
	(142.89, 65.80) --
	(143.34, 66.12) --
	(143.79, 66.54) --
	(144.24, 67.13) --
	(144.69, 67.91) --
	(145.13, 68.93) --
	(145.58, 70.17) --
	(146.03, 71.63) --
	(146.48, 73.22) --
	(146.93, 74.86) --
	(147.38, 76.42) --
	(147.83, 77.77) --
	(148.28, 78.79) --
	(148.72, 79.40) --
	(149.17, 79.57) --
	(149.62, 79.33) --
	(150.07, 78.77) --
	(150.52, 77.99) --
	(150.97, 77.11) --
	(151.42, 76.28) --
	(151.87, 75.57) --
	(152.31, 75.03) --
	(152.76, 74.67) --
	(153.21, 74.51) --
	(153.66, 74.52) --
	(154.11, 74.68) --
	(154.56, 75.00) --
	(155.01, 75.48) --
	(155.45, 76.14) --
	(155.90, 77.03) --
	(156.35, 78.18) --
	(156.80, 79.65) --
	(157.25, 81.43) --
	(157.70, 83.53) --
	(158.15, 85.86) --
	(158.60, 88.28) --
	(159.04, 90.62) --
	(159.49, 92.70) --
	(159.94, 94.36) --
	(160.39, 95.47) --
	(160.84, 96.01) --
	(161.29, 96.02) --
	(161.74, 95.62) --
	(162.19, 94.99) --
	(162.63, 94.32) --
	(163.08, 93.84) --
	(163.53, 93.70) --
	(163.98, 94.03) --
	(164.43, 94.90) --
	(164.88, 96.25) --
	(165.33, 97.99) --
	(165.77, 99.96) --
	(166.22,102.01) --
	(166.67,104.00) --
	(167.12,105.84) --
	(167.57,107.52) --
	(168.02,109.08) --
	(168.47,110.65) --
	(168.92,112.37) --
	(169.36,114.38) --
	(169.81,116.76) --
	(170.26,119.52) --
	(170.71,122.57) --
	(171.16,125.76) --
	(171.61,128.87) --
	(172.06,131.72) --
	(172.51,134.22) --
	(172.95,136.36) --
	(173.40,138.16) --
	(173.85,139.72) --
	(174.30,141.10) --
	(174.75,142.33) --
	(175.20,143.42) --
	(175.65,144.33) --
	(176.09,145.02) --
	(176.54,145.44) --
	(176.99,145.63) --
	(177.44,145.61) --
	(177.89,145.48) --
	(178.34,145.30) --
	(178.79,145.15) --
	(179.24,145.07) --
	(179.68,145.04) --
	(180.13,145.08) --
	(180.58,145.16) --
	(181.03,145.27) --
	(181.48,145.40) --
	(181.93,145.54) --
	(182.38,145.69) --
	(182.83,145.85) --
	(183.27,146.02) --
	(183.72,146.20) --
	(184.17,146.41) --
	(184.62,146.64) --
	(185.07,146.91) --
	(185.52,147.23) --
	(185.97,147.60) --
	(186.41,148.03) --
	(186.86,148.51) --
	(187.31,149.05) --
	(187.76,149.62) --
	(188.21,150.21) --
	(188.66,150.81) --
	(189.11,151.41) --
	(189.56,151.99) --
	(190.00,152.57) --
	(190.45,153.13) --
	(190.90,153.68) --
	(191.35,154.23) --
	(191.80,154.80) --
	(192.25,155.41) --
	(192.70,156.06) --
	(193.15,156.76) --
	(193.59,157.50) --
	(194.04,158.24) --
	(194.49,158.94) --
	(194.94,159.54) --
	(195.39,159.95) --
	(195.84,160.10) --
	(196.29,159.92) --
	(196.73,159.37) --
	(197.18,158.45) --
	(197.63,157.17) --
	(198.08,155.62) --
	(198.53,153.90) --
	(198.98,152.15) --
	(199.43,150.47) --
	(199.88,148.96) --
	(200.32,147.67) --
	(200.77,146.61) --
	(201.22,145.75) --
	(201.67,145.01) --
	(202.12,144.31) --
	(202.57,143.54) --
	(203.02,142.61) --
	(203.47,141.43) --
	(203.91,139.92) --
	(204.36,138.00) --
	(204.81,135.62) --
	(205.26,132.72) --
	(205.71,129.32) --
	(206.16,125.43) --
	(206.61,121.08) --
	(207.05,116.37) --
	(207.50,111.40) --
	(207.95,106.30) --
	(208.40,101.20) --
	(208.85, 96.24) --
	(209.30, 91.51) --
	(209.75, 87.10) --
	(210.20, 83.04) --
	(210.64, 79.35) --
	(211.09, 76.03) --
	(211.54, 73.07) --
	(211.99, 70.44) --
	(212.44, 68.11) --
	(212.89, 66.07) --
	(213.34, 64.30) --
	(213.79, 62.80) --
	(214.23, 61.55) --
	(214.68, 60.54) --
	(215.13, 59.73) --
	(215.58, 59.12) --
	(216.03, 58.65) --
	(216.48, 58.32) --
	(216.93, 58.09) --
	(217.37, 57.93) --
	(217.82, 57.82) --
	(218.27, 57.76) --
	(218.72, 57.72) --
	(219.17, 57.70) --
	(219.62, 57.68) --
	(220.07, 57.68) --
	(220.52, 57.67) --
	(220.96, 57.67) --
	(221.41, 57.67) --
	(221.86, 57.67) --
	(222.31, 57.67) --
	(222.76, 57.67) --
	(223.21, 57.67) --
	(223.66, 57.67) --
	(224.11, 57.67) --
	(224.55, 57.67) --
	(225.00, 57.67) --
	(225.45, 57.67) --
	(225.90, 57.67) --
	(226.35, 57.67) --
	(226.80, 57.67) --
	(227.25, 57.67) --
	(227.69, 57.67) --
	(228.14, 57.67) --
	(228.59, 57.67) --
	(229.04, 57.67) --
	(229.49, 57.67) --
	(229.94, 57.67) --
	(230.39, 57.67) --
	(230.84, 57.67) --
	(231.28, 57.67) --
	(231.73, 57.67) --
	(232.18, 57.67) --
	(232.63, 57.67) --
	(233.08, 57.67) --
	(233.53, 57.67) --
	(233.98, 57.67) --
	(234.43, 57.67) --
	(234.87, 57.67) --
	(235.32, 57.67) --
	(235.77, 57.67) --
	(236.22, 57.67) --
	(236.67, 57.67) --
	(237.12, 57.67) --
	(237.57, 57.67) --
	(238.01, 57.67) --
	(238.46, 57.67) --
	(238.91, 57.67) --
	(239.36, 57.67) --
	(239.81, 57.67) --
	(240.26, 57.67) --
	(240.71, 57.67) --
	(241.16, 57.67) --
	(241.60, 57.67) --
	(242.05, 57.67) --
	(242.50, 57.67) --
	(242.95, 57.67) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67);

\path[] ( 42.83, 57.70) --
	( 43.28, 57.70) --
	( 43.73, 57.71) --
	( 44.18, 57.71) --
	( 44.63, 57.71) --
	( 45.08, 57.71) --
	( 45.52, 57.71) --
	( 45.97, 57.71) --
	( 46.42, 57.72) --
	( 46.87, 57.72) --
	( 47.32, 57.72) --
	( 47.77, 57.73) --
	( 48.22, 57.73) --
	( 48.67, 57.74) --
	( 49.11, 57.75) --
	( 49.56, 57.76) --
	( 50.01, 57.77) --
	( 50.46, 57.77) --
	( 50.91, 57.78) --
	( 51.36, 57.78) --
	( 51.81, 57.78) --
	( 52.25, 57.78) --
	( 52.70, 57.77) --
	( 53.15, 57.77) --
	( 53.60, 57.76) --
	( 54.05, 57.76) --
	( 54.50, 57.76) --
	( 54.95, 57.75) --
	( 55.40, 57.76) --
	( 55.84, 57.76) --
	( 56.29, 57.76) --
	( 56.74, 57.77) --
	( 57.19, 57.77) --
	( 57.64, 57.78) --
	( 58.09, 57.79) --
	( 58.54, 57.80) --
	( 58.99, 57.80) --
	( 59.43, 57.81) --
	( 59.88, 57.82) --
	( 60.33, 57.82) --
	( 60.78, 57.83) --
	( 61.23, 57.83) --
	( 61.68, 57.84) --
	( 62.13, 57.84) --
	( 62.57, 57.85) --
	( 63.02, 57.85) --
	( 63.47, 57.86) --
	( 63.92, 57.86) --
	( 64.37, 57.87) --
	( 64.82, 57.88) --
	( 65.27, 57.89) --
	( 65.72, 57.90) --
	( 66.16, 57.90) --
	( 66.61, 57.91) --
	( 67.06, 57.91) --
	( 67.51, 57.91) --
	( 67.96, 57.90) --
	( 68.41, 57.90) --
	( 68.86, 57.90) --
	( 69.31, 57.90) --
	( 69.75, 57.90) --
	( 70.20, 57.90) --
	( 70.65, 57.90) --
	( 71.10, 57.91) --
	( 71.55, 57.92) --
	( 72.00, 57.93) --
	( 72.45, 57.94) --
	( 72.89, 57.96) --
	( 73.34, 57.98) --
	( 73.79, 58.00) --
	( 74.24, 58.02) --
	( 74.69, 58.05) --
	( 75.14, 58.07) --
	( 75.59, 58.10) --
	( 76.04, 58.13) --
	( 76.48, 58.16) --
	( 76.93, 58.18) --
	( 77.38, 58.21) --
	( 77.83, 58.23) --
	( 78.28, 58.24) --
	( 78.73, 58.25) --
	( 79.18, 58.26) --
	( 79.63, 58.26) --
	( 80.07, 58.26) --
	( 80.52, 58.26) --
	( 80.97, 58.25) --
	( 81.42, 58.24) --
	( 81.87, 58.24) --
	( 82.32, 58.23) --
	( 82.77, 58.23) --
	( 83.21, 58.23) --
	( 83.66, 58.24) --
	( 84.11, 58.25) --
	( 84.56, 58.27) --
	( 85.01, 58.29) --
	( 85.46, 58.33) --
	( 85.91, 58.37) --
	( 86.36, 58.41) --
	( 86.80, 58.47) --
	( 87.25, 58.52) --
	( 87.70, 58.58) --
	( 88.15, 58.64) --
	( 88.60, 58.68) --
	( 89.05, 58.72) --
	( 89.50, 58.74) --
	( 89.95, 58.75) --
	( 90.39, 58.75) --
	( 90.84, 58.73) --
	( 91.29, 58.72) --
	( 91.74, 58.71) --
	( 92.19, 58.70) --
	( 92.64, 58.71) --
	( 93.09, 58.73) --
	( 93.53, 58.76) --
	( 93.98, 58.80) --
	( 94.43, 58.84) --
	( 94.88, 58.89) --
	( 95.33, 58.94) --
	( 95.78, 58.99) --
	( 96.23, 59.04) --
	( 96.68, 59.08) --
	( 97.12, 59.12) --
	( 97.57, 59.16) --
	( 98.02, 59.21) --
	( 98.47, 59.25) --
	( 98.92, 59.29) --
	( 99.37, 59.33) --
	( 99.82, 59.37) --
	(100.27, 59.41) --
	(100.71, 59.46) --
	(101.16, 59.50) --
	(101.61, 59.54) --
	(102.06, 59.58) --
	(102.51, 59.64) --
	(102.96, 59.69) --
	(103.41, 59.76) --
	(103.85, 59.83) --
	(104.30, 59.90) --
	(104.75, 59.98) --
	(105.20, 60.06) --
	(105.65, 60.14) --
	(106.10, 60.20) --
	(106.55, 60.25) --
	(107.00, 60.28) --
	(107.44, 60.29) --
	(107.89, 60.27) --
	(108.34, 60.24) --
	(108.79, 60.20) --
	(109.24, 60.16) --
	(109.69, 60.12) --
	(110.14, 60.09) --
	(110.59, 60.08) --
	(111.03, 60.08) --
	(111.48, 60.11) --
	(111.93, 60.16) --
	(112.38, 60.24) --
	(112.83, 60.32) --
	(113.28, 60.42) --
	(113.73, 60.52) --
	(114.17, 60.62) --
	(114.62, 60.71) --
	(115.07, 60.79) --
	(115.52, 60.85) --
	(115.97, 60.90) --
	(116.42, 60.93) --
	(116.87, 60.95) --
	(117.32, 60.95) --
	(117.76, 60.95) --
	(118.21, 60.95) --
	(118.66, 60.96) --
	(119.11, 60.98) --
	(119.56, 61.02) --
	(120.01, 61.08) --
	(120.46, 61.17) --
	(120.91, 61.26) --
	(121.35, 61.37) --
	(121.80, 61.48) --
	(122.25, 61.59) --
	(122.70, 61.69) --
	(123.15, 61.77) --
	(123.60, 61.85) --
	(124.05, 61.91) --
	(124.49, 61.97) --
	(124.94, 62.02) --
	(125.39, 62.07) --
	(125.84, 62.12) --
	(126.29, 62.18) --
	(126.74, 62.25) --
	(127.19, 62.31) --
	(127.64, 62.38) --
	(128.08, 62.45) --
	(128.53, 62.51) --
	(128.98, 62.58) --
	(129.43, 62.65) --
	(129.88, 62.72) --
	(130.33, 62.79) --
	(130.78, 62.87) --
	(131.23, 62.95) --
	(131.67, 63.04) --
	(132.12, 63.13) --
	(132.57, 63.22) --
	(133.02, 63.30) --
	(133.47, 63.37) --
	(133.92, 63.43) --
	(134.37, 63.47) --
	(134.81, 63.50) --
	(135.26, 63.53) --
	(135.71, 63.56) --
	(136.16, 63.60) --
	(136.61, 63.66) --
	(137.06, 63.73) --
	(137.51, 63.84) --
	(137.96, 63.96) --
	(138.40, 64.10) --
	(138.85, 64.25) --
	(139.30, 64.40) --
	(139.75, 64.56) --
	(140.20, 64.71) --
	(140.65, 64.87) --
	(141.10, 65.01) --
	(141.55, 65.16) --
	(141.99, 65.30) --
	(142.44, 65.44) --
	(142.89, 65.58) --
	(143.34, 65.72) --
	(143.79, 65.87) --
	(144.24, 66.03) --
	(144.69, 66.20) --
	(145.13, 66.38) --
	(145.58, 66.57) --
	(146.03, 66.77) --
	(146.48, 66.97) --
	(146.93, 67.18) --
	(147.38, 67.39) --
	(147.83, 67.61) --
	(148.28, 67.83) --
	(148.72, 68.05) --
	(149.17, 68.27) --
	(149.62, 68.49) --
	(150.07, 68.70) --
	(150.52, 68.89) --
	(150.97, 69.05) --
	(151.42, 69.19) --
	(151.87, 69.32) --
	(152.31, 69.43) --
	(152.76, 69.55) --
	(153.21, 69.69) --
	(153.66, 69.86) --
	(154.11, 70.09) --
	(154.56, 70.38) --
	(155.01, 70.77) --
	(155.45, 71.26) --
	(155.90, 71.88) --
	(156.35, 72.64) --
	(156.80, 73.55) --
	(157.25, 74.61) --
	(157.70, 75.81) --
	(158.15, 77.11) --
	(158.60, 78.46) --
	(159.04, 79.78) --
	(159.49, 80.99) --
	(159.94, 82.04) --
	(160.39, 82.88) --
	(160.84, 83.47) --
	(161.29, 83.81) --
	(161.74, 83.94) --
	(162.19, 83.91) --
	(162.63, 83.78) --
	(163.08, 83.61) --
	(163.53, 83.45) --
	(163.98, 83.35) --
	(164.43, 83.34) --
	(164.88, 83.42) --
	(165.33, 83.61) --
	(165.77, 83.89) --
	(166.22, 84.25) --
	(166.67, 84.72) --
	(167.12, 85.28) --
	(167.57, 85.95) --
	(168.02, 86.76) --
	(168.47, 87.72) --
	(168.92, 88.84) --
	(169.36, 90.12) --
	(169.81, 91.55) --
	(170.26, 93.09) --
	(170.71, 94.68) --
	(171.16, 96.24) --
	(171.61, 97.68) --
	(172.06, 98.94) --
	(172.51,100.00) --
	(172.95,100.85) --
	(173.40,101.53) --
	(173.85,102.09) --
	(174.30,102.63) --
	(174.75,103.22) --
	(175.20,103.94) --
	(175.65,104.86) --
	(176.09,106.03) --
	(176.54,107.45) --
	(176.99,109.11) --
	(177.44,110.98) --
	(177.89,112.99) --
	(178.34,115.06) --
	(178.79,117.14) --
	(179.24,119.19) --
	(179.68,121.19) --
	(180.13,123.16) --
	(180.58,125.12) --
	(181.03,127.11) --
	(181.48,129.17) --
	(181.93,131.29) --
	(182.38,133.46) --
	(182.83,135.63) --
	(183.27,137.75) --
	(183.72,139.75) --
	(184.17,141.58) --
	(184.62,143.26) --
	(185.07,144.80) --
	(185.52,146.26) --
	(185.97,147.65) --
	(186.41,149.01) --
	(186.86,150.33) --
	(187.31,151.57) --
	(187.76,152.70) --
	(188.21,153.69) --
	(188.66,154.49) --
	(189.11,155.09) --
	(189.56,155.49) --
	(190.00,155.73) --
	(190.45,155.80) --
	(190.90,155.77) --
	(191.35,155.66) --
	(191.80,155.50) --
	(192.25,155.31) --
	(192.70,155.11) --
	(193.15,154.89) --
	(193.59,154.66) --
	(194.04,154.40) --
	(194.49,154.09) --
	(194.94,153.72) --
	(195.39,153.27) --
	(195.84,152.73) --
	(196.29,152.11) --
	(196.73,151.42) --
	(197.18,150.67) --
	(197.63,149.88) --
	(198.08,149.07) --
	(198.53,148.24) --
	(198.98,147.39) --
	(199.43,146.52) --
	(199.88,145.60) --
	(200.32,144.64) --
	(200.77,143.64) --
	(201.22,142.59) --
	(201.67,141.51) --
	(202.12,140.40) --
	(202.57,139.28) --
	(203.02,138.17) --
	(203.47,137.08) --
	(203.91,136.02) --
	(204.36,135.01) --
	(204.81,134.04) --
	(205.26,133.12) --
	(205.71,132.25) --
	(206.16,131.42) --
	(206.61,130.65) --
	(207.05,129.93) --
	(207.50,129.27) --
	(207.95,128.68) --
	(208.40,128.16) --
	(208.85,127.70) --
	(209.30,127.32) --
	(209.75,127.01) --
	(210.20,126.77) --
	(210.64,126.58) --
	(211.09,126.43) --
	(211.54,126.28) --
	(211.99,126.12) --
	(212.44,125.90) --
	(212.89,125.59) --
	(213.34,125.14) --
	(213.79,124.52) --
	(214.23,123.67) --
	(214.68,122.57) --
	(215.13,121.18) --
	(215.58,119.49) --
	(216.03,117.48) --
	(216.48,115.13) --
	(216.93,112.46) --
	(217.37,109.50) --
	(217.82,106.29) --
	(218.27,102.88) --
	(218.72, 99.33) --
	(219.17, 95.72) --
	(219.62, 92.11) --
	(220.07, 88.59) --
	(220.52, 85.21) --
	(220.96, 82.01) --
	(221.41, 79.03) --
	(221.86, 76.28) --
	(222.31, 73.77) --
	(222.76, 71.47) --
	(223.21, 69.41) --
	(223.66, 67.54) --
	(224.11, 65.86) --
	(224.55, 64.36) --
	(225.00, 63.03) --
	(225.45, 61.89) --
	(225.90, 60.92) --
	(226.35, 60.11) --
	(226.80, 59.46) --
	(227.25, 58.94) --
	(227.69, 58.55) --
	(228.14, 58.26) --
	(228.59, 58.06) --
	(229.04, 57.91) --
	(229.49, 57.82) --
	(229.94, 57.76) --
	(230.39, 57.72) --
	(230.84, 57.70) --
	(231.28, 57.69) --
	(231.73, 57.68) --
	(232.18, 57.67) --
	(232.63, 57.67) --
	(233.08, 57.67) --
	(233.53, 57.67) --
	(233.98, 57.67) --
	(234.43, 57.67) --
	(234.87, 57.67) --
	(235.32, 57.67) --
	(235.77, 57.67) --
	(236.22, 57.67) --
	(236.67, 57.67) --
	(237.12, 57.67) --
	(237.57, 57.67) --
	(238.01, 57.67) --
	(238.46, 57.67) --
	(238.91, 57.67) --
	(239.36, 57.67) --
	(239.81, 57.67) --
	(240.26, 57.67) --
	(240.71, 57.67) --
	(241.16, 57.67) --
	(241.60, 57.67) --
	(242.05, 57.67) --
	(242.50, 57.67) --
	(242.95, 57.67) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67) --
	(272.12, 57.67) --
	(271.67, 57.67) --
	(271.22, 57.67) --
	(270.77, 57.67) --
	(270.32, 57.67) --
	(269.87, 57.67) --
	(269.42, 57.67) --
	(268.97, 57.67) --
	(268.53, 57.67) --
	(268.08, 57.67) --
	(267.63, 57.67) --
	(267.18, 57.67) --
	(266.73, 57.67) --
	(266.28, 57.67) --
	(265.83, 57.67) --
	(265.39, 57.67) --
	(264.94, 57.67) --
	(264.49, 57.67) --
	(264.04, 57.67) --
	(263.59, 57.67) --
	(263.14, 57.67) --
	(262.69, 57.67) --
	(262.24, 57.67) --
	(261.80, 57.67) --
	(261.35, 57.67) --
	(260.90, 57.67) --
	(260.45, 57.67) --
	(260.00, 57.67) --
	(259.55, 57.67) --
	(259.10, 57.67) --
	(258.65, 57.67) --
	(258.21, 57.67) --
	(257.76, 57.67) --
	(257.31, 57.67) --
	(256.86, 57.67) --
	(256.41, 57.67) --
	(255.96, 57.67) --
	(255.51, 57.67) --
	(255.07, 57.67) --
	(254.62, 57.67) --
	(254.17, 57.67) --
	(253.72, 57.67) --
	(253.27, 57.67) --
	(252.82, 57.67) --
	(252.37, 57.67) --
	(251.92, 57.67) --
	(251.48, 57.67) --
	(251.03, 57.67) --
	(250.58, 57.67) --
	(250.13, 57.67) --
	(249.68, 57.67) --
	(249.23, 57.67) --
	(248.78, 57.67) --
	(248.33, 57.67) --
	(247.89, 57.67) --
	(247.44, 57.67) --
	(246.99, 57.67) --
	(246.54, 57.67) --
	(246.09, 57.67) --
	(245.64, 57.67) --
	(245.19, 57.67) --
	(244.75, 57.67) --
	(244.30, 57.67) --
	(243.85, 57.67) --
	(243.40, 57.67) --
	(242.95, 57.67) --
	(242.50, 57.67) --
	(242.05, 57.67) --
	(241.60, 57.67) --
	(241.16, 57.67) --
	(240.71, 57.67) --
	(240.26, 57.67) --
	(239.81, 57.67) --
	(239.36, 57.67) --
	(238.91, 57.67) --
	(238.46, 57.67) --
	(238.01, 57.67) --
	(237.57, 57.67) --
	(237.12, 57.67) --
	(236.67, 57.67) --
	(236.22, 57.67) --
	(235.77, 57.67) --
	(235.32, 57.67) --
	(234.87, 57.67) --
	(234.43, 57.67) --
	(233.98, 57.67) --
	(233.53, 57.67) --
	(233.08, 57.67) --
	(232.63, 57.67) --
	(232.18, 57.67) --
	(231.73, 57.67) --
	(231.28, 57.67) --
	(230.84, 57.67) --
	(230.39, 57.67) --
	(229.94, 57.67) --
	(229.49, 57.67) --
	(229.04, 57.67) --
	(228.59, 57.67) --
	(228.14, 57.67) --
	(227.69, 57.67) --
	(227.25, 57.67) --
	(226.80, 57.67) --
	(226.35, 57.67) --
	(225.90, 57.67) --
	(225.45, 57.67) --
	(225.00, 57.67) --
	(224.55, 57.67) --
	(224.11, 57.67) --
	(223.66, 57.67) --
	(223.21, 57.67) --
	(222.76, 57.67) --
	(222.31, 57.67) --
	(221.86, 57.67) --
	(221.41, 57.67) --
	(220.96, 57.67) --
	(220.52, 57.67) --
	(220.07, 57.67) --
	(219.62, 57.67) --
	(219.17, 57.67) --
	(218.72, 57.67) --
	(218.27, 57.67) --
	(217.82, 57.67) --
	(217.37, 57.67) --
	(216.93, 57.67) --
	(216.48, 57.67) --
	(216.03, 57.67) --
	(215.58, 57.67) --
	(215.13, 57.67) --
	(214.68, 57.67) --
	(214.23, 57.67) --
	(213.79, 57.67) --
	(213.34, 57.67) --
	(212.89, 57.67) --
	(212.44, 57.67) --
	(211.99, 57.67) --
	(211.54, 57.67) --
	(211.09, 57.67) --
	(210.64, 57.67) --
	(210.20, 57.67) --
	(209.75, 57.67) --
	(209.30, 57.67) --
	(208.85, 57.67) --
	(208.40, 57.67) --
	(207.95, 57.67) --
	(207.50, 57.67) --
	(207.05, 57.67) --
	(206.61, 57.67) --
	(206.16, 57.67) --
	(205.71, 57.67) --
	(205.26, 57.67) --
	(204.81, 57.67) --
	(204.36, 57.67) --
	(203.91, 57.67) --
	(203.47, 57.67) --
	(203.02, 57.67) --
	(202.57, 57.67) --
	(202.12, 57.67) --
	(201.67, 57.67) --
	(201.22, 57.67) --
	(200.77, 57.67) --
	(200.32, 57.67) --
	(199.88, 57.67) --
	(199.43, 57.67) --
	(198.98, 57.67) --
	(198.53, 57.67) --
	(198.08, 57.67) --
	(197.63, 57.67) --
	(197.18, 57.67) --
	(196.73, 57.67) --
	(196.29, 57.67) --
	(195.84, 57.67) --
	(195.39, 57.67) --
	(194.94, 57.67) --
	(194.49, 57.67) --
	(194.04, 57.67) --
	(193.59, 57.67) --
	(193.15, 57.67) --
	(192.70, 57.67) --
	(192.25, 57.67) --
	(191.80, 57.67) --
	(191.35, 57.67) --
	(190.90, 57.67) --
	(190.45, 57.67) --
	(190.00, 57.67) --
	(189.56, 57.67) --
	(189.11, 57.67) --
	(188.66, 57.67) --
	(188.21, 57.67) --
	(187.76, 57.67) --
	(187.31, 57.67) --
	(186.86, 57.67) --
	(186.41, 57.67) --
	(185.97, 57.67) --
	(185.52, 57.67) --
	(185.07, 57.67) --
	(184.62, 57.67) --
	(184.17, 57.67) --
	(183.72, 57.67) --
	(183.27, 57.67) --
	(182.83, 57.67) --
	(182.38, 57.67) --
	(181.93, 57.67) --
	(181.48, 57.67) --
	(181.03, 57.67) --
	(180.58, 57.67) --
	(180.13, 57.67) --
	(179.68, 57.67) --
	(179.24, 57.67) --
	(178.79, 57.67) --
	(178.34, 57.67) --
	(177.89, 57.67) --
	(177.44, 57.67) --
	(176.99, 57.67) --
	(176.54, 57.67) --
	(176.09, 57.67) --
	(175.65, 57.67) --
	(175.20, 57.67) --
	(174.75, 57.67) --
	(174.30, 57.67) --
	(173.85, 57.67) --
	(173.40, 57.67) --
	(172.95, 57.67) --
	(172.51, 57.67) --
	(172.06, 57.67) --
	(171.61, 57.67) --
	(171.16, 57.67) --
	(170.71, 57.67) --
	(170.26, 57.67) --
	(169.81, 57.67) --
	(169.36, 57.67) --
	(168.92, 57.67) --
	(168.47, 57.67) --
	(168.02, 57.67) --
	(167.57, 57.67) --
	(167.12, 57.67) --
	(166.67, 57.67) --
	(166.22, 57.67) --
	(165.77, 57.67) --
	(165.33, 57.67) --
	(164.88, 57.67) --
	(164.43, 57.67) --
	(163.98, 57.67) --
	(163.53, 57.67) --
	(163.08, 57.67) --
	(162.63, 57.67) --
	(162.19, 57.67) --
	(161.74, 57.67) --
	(161.29, 57.67) --
	(160.84, 57.67) --
	(160.39, 57.67) --
	(159.94, 57.67) --
	(159.49, 57.67) --
	(159.04, 57.67) --
	(158.60, 57.67) --
	(158.15, 57.67) --
	(157.70, 57.67) --
	(157.25, 57.67) --
	(156.80, 57.67) --
	(156.35, 57.67) --
	(155.90, 57.67) --
	(155.45, 57.67) --
	(155.01, 57.67) --
	(154.56, 57.67) --
	(154.11, 57.67) --
	(153.66, 57.67) --
	(153.21, 57.67) --
	(152.76, 57.67) --
	(152.31, 57.67) --
	(151.87, 57.67) --
	(151.42, 57.67) --
	(150.97, 57.67) --
	(150.52, 57.67) --
	(150.07, 57.67) --
	(149.62, 57.67) --
	(149.17, 57.67) --
	(148.72, 57.67) --
	(148.28, 57.67) --
	(147.83, 57.67) --
	(147.38, 57.67) --
	(146.93, 57.67) --
	(146.48, 57.67) --
	(146.03, 57.67) --
	(145.58, 57.67) --
	(145.13, 57.67) --
	(144.69, 57.67) --
	(144.24, 57.67) --
	(143.79, 57.67) --
	(143.34, 57.67) --
	(142.89, 57.67) --
	(142.44, 57.67) --
	(141.99, 57.67) --
	(141.55, 57.67) --
	(141.10, 57.67) --
	(140.65, 57.67) --
	(140.20, 57.67) --
	(139.75, 57.67) --
	(139.30, 57.67) --
	(138.85, 57.67) --
	(138.40, 57.67) --
	(137.96, 57.67) --
	(137.51, 57.67) --
	(137.06, 57.67) --
	(136.61, 57.67) --
	(136.16, 57.67) --
	(135.71, 57.67) --
	(135.26, 57.67) --
	(134.81, 57.67) --
	(134.37, 57.67) --
	(133.92, 57.67) --
	(133.47, 57.67) --
	(133.02, 57.67) --
	(132.57, 57.67) --
	(132.12, 57.67) --
	(131.67, 57.67) --
	(131.23, 57.67) --
	(130.78, 57.67) --
	(130.33, 57.67) --
	(129.88, 57.67) --
	(129.43, 57.67) --
	(128.98, 57.67) --
	(128.53, 57.67) --
	(128.08, 57.67) --
	(127.64, 57.67) --
	(127.19, 57.67) --
	(126.74, 57.67) --
	(126.29, 57.67) --
	(125.84, 57.67) --
	(125.39, 57.67) --
	(124.94, 57.67) --
	(124.49, 57.67) --
	(124.05, 57.67) --
	(123.60, 57.67) --
	(123.15, 57.67) --
	(122.70, 57.67) --
	(122.25, 57.67) --
	(121.80, 57.67) --
	(121.35, 57.67) --
	(120.91, 57.67) --
	(120.46, 57.67) --
	(120.01, 57.67) --
	(119.56, 57.67) --
	(119.11, 57.67) --
	(118.66, 57.67) --
	(118.21, 57.67) --
	(117.76, 57.67) --
	(117.32, 57.67) --
	(116.87, 57.67) --
	(116.42, 57.67) --
	(115.97, 57.67) --
	(115.52, 57.67) --
	(115.07, 57.67) --
	(114.62, 57.67) --
	(114.17, 57.67) --
	(113.73, 57.67) --
	(113.28, 57.67) --
	(112.83, 57.67) --
	(112.38, 57.67) --
	(111.93, 57.67) --
	(111.48, 57.67) --
	(111.03, 57.67) --
	(110.59, 57.67) --
	(110.14, 57.67) --
	(109.69, 57.67) --
	(109.24, 57.67) --
	(108.79, 57.67) --
	(108.34, 57.67) --
	(107.89, 57.67) --
	(107.44, 57.67) --
	(107.00, 57.67) --
	(106.55, 57.67) --
	(106.10, 57.67) --
	(105.65, 57.67) --
	(105.20, 57.67) --
	(104.75, 57.67) --
	(104.30, 57.67) --
	(103.85, 57.67) --
	(103.41, 57.67) --
	(102.96, 57.67) --
	(102.51, 57.67) --
	(102.06, 57.67) --
	(101.61, 57.67) --
	(101.16, 57.67) --
	(100.71, 57.67) --
	(100.27, 57.67) --
	( 99.82, 57.67) --
	( 99.37, 57.67) --
	( 98.92, 57.67) --
	( 98.47, 57.67) --
	( 98.02, 57.67) --
	( 97.57, 57.67) --
	( 97.12, 57.67) --
	( 96.68, 57.67) --
	( 96.23, 57.67) --
	( 95.78, 57.67) --
	( 95.33, 57.67) --
	( 94.88, 57.67) --
	( 94.43, 57.67) --
	( 93.98, 57.67) --
	( 93.53, 57.67) --
	( 93.09, 57.67) --
	( 92.64, 57.67) --
	( 92.19, 57.67) --
	( 91.74, 57.67) --
	( 91.29, 57.67) --
	( 90.84, 57.67) --
	( 90.39, 57.67) --
	( 89.95, 57.67) --
	( 89.50, 57.67) --
	( 89.05, 57.67) --
	( 88.60, 57.67) --
	( 88.15, 57.67) --
	( 87.70, 57.67) --
	( 87.25, 57.67) --
	( 86.80, 57.67) --
	( 86.36, 57.67) --
	( 85.91, 57.67) --
	( 85.46, 57.67) --
	( 85.01, 57.67) --
	( 84.56, 57.67) --
	( 84.11, 57.67) --
	( 83.66, 57.67) --
	( 83.21, 57.67) --
	( 82.77, 57.67) --
	( 82.32, 57.67) --
	( 81.87, 57.67) --
	( 81.42, 57.67) --
	( 80.97, 57.67) --
	( 80.52, 57.67) --
	( 80.07, 57.67) --
	( 79.63, 57.67) --
	( 79.18, 57.67) --
	( 78.73, 57.67) --
	( 78.28, 57.67) --
	( 77.83, 57.67) --
	( 77.38, 57.67) --
	( 76.93, 57.67) --
	( 76.48, 57.67) --
	( 76.04, 57.67) --
	( 75.59, 57.67) --
	( 75.14, 57.67) --
	( 74.69, 57.67) --
	( 74.24, 57.67) --
	( 73.79, 57.67) --
	( 73.34, 57.67) --
	( 72.89, 57.67) --
	( 72.45, 57.67) --
	( 72.00, 57.67) --
	( 71.55, 57.67) --
	( 71.10, 57.67) --
	( 70.65, 57.67) --
	( 70.20, 57.67) --
	( 69.75, 57.67) --
	( 69.31, 57.67) --
	( 68.86, 57.67) --
	( 68.41, 57.67) --
	( 67.96, 57.67) --
	( 67.51, 57.67) --
	( 67.06, 57.67) --
	( 66.61, 57.67) --
	( 66.16, 57.67) --
	( 65.72, 57.67) --
	( 65.27, 57.67) --
	( 64.82, 57.67) --
	( 64.37, 57.67) --
	( 63.92, 57.67) --
	( 63.47, 57.67) --
	( 63.02, 57.67) --
	( 62.57, 57.67) --
	( 62.13, 57.67) --
	( 61.68, 57.67) --
	( 61.23, 57.67) --
	( 60.78, 57.67) --
	( 60.33, 57.67) --
	( 59.88, 57.67) --
	( 59.43, 57.67) --
	( 58.99, 57.67) --
	( 58.54, 57.67) --
	( 58.09, 57.67) --
	( 57.64, 57.67) --
	( 57.19, 57.67) --
	( 56.74, 57.67) --
	( 56.29, 57.67) --
	( 55.84, 57.67) --
	( 55.40, 57.67) --
	( 54.95, 57.67) --
	( 54.50, 57.67) --
	( 54.05, 57.67) --
	( 53.60, 57.67) --
	( 53.15, 57.67) --
	( 52.70, 57.67) --
	( 52.25, 57.67) --
	( 51.81, 57.67) --
	( 51.36, 57.67) --
	( 50.91, 57.67) --
	( 50.46, 57.67) --
	( 50.01, 57.67) --
	( 49.56, 57.67) --
	( 49.11, 57.67) --
	( 48.67, 57.67) --
	( 48.22, 57.67) --
	( 47.77, 57.67) --
	( 47.32, 57.67) --
	( 46.87, 57.67) --
	( 46.42, 57.67) --
	( 45.97, 57.67) --
	( 45.52, 57.67) --
	( 45.08, 57.67) --
	( 44.63, 57.67) --
	( 44.18, 57.67) --
	( 43.73, 57.67) --
	( 43.28, 57.67) --
	( 42.83, 57.67) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 42.83, 57.70) --
	( 43.28, 57.70) --
	( 43.73, 57.71) --
	( 44.18, 57.71) --
	( 44.63, 57.71) --
	( 45.08, 57.71) --
	( 45.52, 57.71) --
	( 45.97, 57.71) --
	( 46.42, 57.72) --
	( 46.87, 57.72) --
	( 47.32, 57.72) --
	( 47.77, 57.73) --
	( 48.22, 57.73) --
	( 48.67, 57.74) --
	( 49.11, 57.75) --
	( 49.56, 57.76) --
	( 50.01, 57.77) --
	( 50.46, 57.77) --
	( 50.91, 57.78) --
	( 51.36, 57.78) --
	( 51.81, 57.78) --
	( 52.25, 57.78) --
	( 52.70, 57.77) --
	( 53.15, 57.77) --
	( 53.60, 57.76) --
	( 54.05, 57.76) --
	( 54.50, 57.76) --
	( 54.95, 57.75) --
	( 55.40, 57.76) --
	( 55.84, 57.76) --
	( 56.29, 57.76) --
	( 56.74, 57.77) --
	( 57.19, 57.77) --
	( 57.64, 57.78) --
	( 58.09, 57.79) --
	( 58.54, 57.80) --
	( 58.99, 57.80) --
	( 59.43, 57.81) --
	( 59.88, 57.82) --
	( 60.33, 57.82) --
	( 60.78, 57.83) --
	( 61.23, 57.83) --
	( 61.68, 57.84) --
	( 62.13, 57.84) --
	( 62.57, 57.85) --
	( 63.02, 57.85) --
	( 63.47, 57.86) --
	( 63.92, 57.86) --
	( 64.37, 57.87) --
	( 64.82, 57.88) --
	( 65.27, 57.89) --
	( 65.72, 57.90) --
	( 66.16, 57.90) --
	( 66.61, 57.91) --
	( 67.06, 57.91) --
	( 67.51, 57.91) --
	( 67.96, 57.90) --
	( 68.41, 57.90) --
	( 68.86, 57.90) --
	( 69.31, 57.90) --
	( 69.75, 57.90) --
	( 70.20, 57.90) --
	( 70.65, 57.90) --
	( 71.10, 57.91) --
	( 71.55, 57.92) --
	( 72.00, 57.93) --
	( 72.45, 57.94) --
	( 72.89, 57.96) --
	( 73.34, 57.98) --
	( 73.79, 58.00) --
	( 74.24, 58.02) --
	( 74.69, 58.05) --
	( 75.14, 58.07) --
	( 75.59, 58.10) --
	( 76.04, 58.13) --
	( 76.48, 58.16) --
	( 76.93, 58.18) --
	( 77.38, 58.21) --
	( 77.83, 58.23) --
	( 78.28, 58.24) --
	( 78.73, 58.25) --
	( 79.18, 58.26) --
	( 79.63, 58.26) --
	( 80.07, 58.26) --
	( 80.52, 58.26) --
	( 80.97, 58.25) --
	( 81.42, 58.24) --
	( 81.87, 58.24) --
	( 82.32, 58.23) --
	( 82.77, 58.23) --
	( 83.21, 58.23) --
	( 83.66, 58.24) --
	( 84.11, 58.25) --
	( 84.56, 58.27) --
	( 85.01, 58.29) --
	( 85.46, 58.33) --
	( 85.91, 58.37) --
	( 86.36, 58.41) --
	( 86.80, 58.47) --
	( 87.25, 58.52) --
	( 87.70, 58.58) --
	( 88.15, 58.64) --
	( 88.60, 58.68) --
	( 89.05, 58.72) --
	( 89.50, 58.74) --
	( 89.95, 58.75) --
	( 90.39, 58.75) --
	( 90.84, 58.73) --
	( 91.29, 58.72) --
	( 91.74, 58.71) --
	( 92.19, 58.70) --
	( 92.64, 58.71) --
	( 93.09, 58.73) --
	( 93.53, 58.76) --
	( 93.98, 58.80) --
	( 94.43, 58.84) --
	( 94.88, 58.89) --
	( 95.33, 58.94) --
	( 95.78, 58.99) --
	( 96.23, 59.04) --
	( 96.68, 59.08) --
	( 97.12, 59.12) --
	( 97.57, 59.16) --
	( 98.02, 59.21) --
	( 98.47, 59.25) --
	( 98.92, 59.29) --
	( 99.37, 59.33) --
	( 99.82, 59.37) --
	(100.27, 59.41) --
	(100.71, 59.46) --
	(101.16, 59.50) --
	(101.61, 59.54) --
	(102.06, 59.58) --
	(102.51, 59.64) --
	(102.96, 59.69) --
	(103.41, 59.76) --
	(103.85, 59.83) --
	(104.30, 59.90) --
	(104.75, 59.98) --
	(105.20, 60.06) --
	(105.65, 60.14) --
	(106.10, 60.20) --
	(106.55, 60.25) --
	(107.00, 60.28) --
	(107.44, 60.29) --
	(107.89, 60.27) --
	(108.34, 60.24) --
	(108.79, 60.20) --
	(109.24, 60.16) --
	(109.69, 60.12) --
	(110.14, 60.09) --
	(110.59, 60.08) --
	(111.03, 60.08) --
	(111.48, 60.11) --
	(111.93, 60.16) --
	(112.38, 60.24) --
	(112.83, 60.32) --
	(113.28, 60.42) --
	(113.73, 60.52) --
	(114.17, 60.62) --
	(114.62, 60.71) --
	(115.07, 60.79) --
	(115.52, 60.85) --
	(115.97, 60.90) --
	(116.42, 60.93) --
	(116.87, 60.95) --
	(117.32, 60.95) --
	(117.76, 60.95) --
	(118.21, 60.95) --
	(118.66, 60.96) --
	(119.11, 60.98) --
	(119.56, 61.02) --
	(120.01, 61.08) --
	(120.46, 61.17) --
	(120.91, 61.26) --
	(121.35, 61.37) --
	(121.80, 61.48) --
	(122.25, 61.59) --
	(122.70, 61.69) --
	(123.15, 61.77) --
	(123.60, 61.85) --
	(124.05, 61.91) --
	(124.49, 61.97) --
	(124.94, 62.02) --
	(125.39, 62.07) --
	(125.84, 62.12) --
	(126.29, 62.18) --
	(126.74, 62.25) --
	(127.19, 62.31) --
	(127.64, 62.38) --
	(128.08, 62.45) --
	(128.53, 62.51) --
	(128.98, 62.58) --
	(129.43, 62.65) --
	(129.88, 62.72) --
	(130.33, 62.79) --
	(130.78, 62.87) --
	(131.23, 62.95) --
	(131.67, 63.04) --
	(132.12, 63.13) --
	(132.57, 63.22) --
	(133.02, 63.30) --
	(133.47, 63.37) --
	(133.92, 63.43) --
	(134.37, 63.47) --
	(134.81, 63.50) --
	(135.26, 63.53) --
	(135.71, 63.56) --
	(136.16, 63.60) --
	(136.61, 63.66) --
	(137.06, 63.73) --
	(137.51, 63.84) --
	(137.96, 63.96) --
	(138.40, 64.10) --
	(138.85, 64.25) --
	(139.30, 64.40) --
	(139.75, 64.56) --
	(140.20, 64.71) --
	(140.65, 64.87) --
	(141.10, 65.01) --
	(141.55, 65.16) --
	(141.99, 65.30) --
	(142.44, 65.44) --
	(142.89, 65.58) --
	(143.34, 65.72) --
	(143.79, 65.87) --
	(144.24, 66.03) --
	(144.69, 66.20) --
	(145.13, 66.38) --
	(145.58, 66.57) --
	(146.03, 66.77) --
	(146.48, 66.97) --
	(146.93, 67.18) --
	(147.38, 67.39) --
	(147.83, 67.61) --
	(148.28, 67.83) --
	(148.72, 68.05) --
	(149.17, 68.27) --
	(149.62, 68.49) --
	(150.07, 68.70) --
	(150.52, 68.89) --
	(150.97, 69.05) --
	(151.42, 69.19) --
	(151.87, 69.32) --
	(152.31, 69.43) --
	(152.76, 69.55) --
	(153.21, 69.69) --
	(153.66, 69.86) --
	(154.11, 70.09) --
	(154.56, 70.38) --
	(155.01, 70.77) --
	(155.45, 71.26) --
	(155.90, 71.88) --
	(156.35, 72.64) --
	(156.80, 73.55) --
	(157.25, 74.61) --
	(157.70, 75.81) --
	(158.15, 77.11) --
	(158.60, 78.46) --
	(159.04, 79.78) --
	(159.49, 80.99) --
	(159.94, 82.04) --
	(160.39, 82.88) --
	(160.84, 83.47) --
	(161.29, 83.81) --
	(161.74, 83.94) --
	(162.19, 83.91) --
	(162.63, 83.78) --
	(163.08, 83.61) --
	(163.53, 83.45) --
	(163.98, 83.35) --
	(164.43, 83.34) --
	(164.88, 83.42) --
	(165.33, 83.61) --
	(165.77, 83.89) --
	(166.22, 84.25) --
	(166.67, 84.72) --
	(167.12, 85.28) --
	(167.57, 85.95) --
	(168.02, 86.76) --
	(168.47, 87.72) --
	(168.92, 88.84) --
	(169.36, 90.12) --
	(169.81, 91.55) --
	(170.26, 93.09) --
	(170.71, 94.68) --
	(171.16, 96.24) --
	(171.61, 97.68) --
	(172.06, 98.94) --
	(172.51,100.00) --
	(172.95,100.85) --
	(173.40,101.53) --
	(173.85,102.09) --
	(174.30,102.63) --
	(174.75,103.22) --
	(175.20,103.94) --
	(175.65,104.86) --
	(176.09,106.03) --
	(176.54,107.45) --
	(176.99,109.11) --
	(177.44,110.98) --
	(177.89,112.99) --
	(178.34,115.06) --
	(178.79,117.14) --
	(179.24,119.19) --
	(179.68,121.19) --
	(180.13,123.16) --
	(180.58,125.12) --
	(181.03,127.11) --
	(181.48,129.17) --
	(181.93,131.29) --
	(182.38,133.46) --
	(182.83,135.63) --
	(183.27,137.75) --
	(183.72,139.75) --
	(184.17,141.58) --
	(184.62,143.26) --
	(185.07,144.80) --
	(185.52,146.26) --
	(185.97,147.65) --
	(186.41,149.01) --
	(186.86,150.33) --
	(187.31,151.57) --
	(187.76,152.70) --
	(188.21,153.69) --
	(188.66,154.49) --
	(189.11,155.09) --
	(189.56,155.49) --
	(190.00,155.73) --
	(190.45,155.80) --
	(190.90,155.77) --
	(191.35,155.66) --
	(191.80,155.50) --
	(192.25,155.31) --
	(192.70,155.11) --
	(193.15,154.89) --
	(193.59,154.66) --
	(194.04,154.40) --
	(194.49,154.09) --
	(194.94,153.72) --
	(195.39,153.27) --
	(195.84,152.73) --
	(196.29,152.11) --
	(196.73,151.42) --
	(197.18,150.67) --
	(197.63,149.88) --
	(198.08,149.07) --
	(198.53,148.24) --
	(198.98,147.39) --
	(199.43,146.52) --
	(199.88,145.60) --
	(200.32,144.64) --
	(200.77,143.64) --
	(201.22,142.59) --
	(201.67,141.51) --
	(202.12,140.40) --
	(202.57,139.28) --
	(203.02,138.17) --
	(203.47,137.08) --
	(203.91,136.02) --
	(204.36,135.01) --
	(204.81,134.04) --
	(205.26,133.12) --
	(205.71,132.25) --
	(206.16,131.42) --
	(206.61,130.65) --
	(207.05,129.93) --
	(207.50,129.27) --
	(207.95,128.68) --
	(208.40,128.16) --
	(208.85,127.70) --
	(209.30,127.32) --
	(209.75,127.01) --
	(210.20,126.77) --
	(210.64,126.58) --
	(211.09,126.43) --
	(211.54,126.28) --
	(211.99,126.12) --
	(212.44,125.90) --
	(212.89,125.59) --
	(213.34,125.14) --
	(213.79,124.52) --
	(214.23,123.67) --
	(214.68,122.57) --
	(215.13,121.18) --
	(215.58,119.49) --
	(216.03,117.48) --
	(216.48,115.13) --
	(216.93,112.46) --
	(217.37,109.50) --
	(217.82,106.29) --
	(218.27,102.88) --
	(218.72, 99.33) --
	(219.17, 95.72) --
	(219.62, 92.11) --
	(220.07, 88.59) --
	(220.52, 85.21) --
	(220.96, 82.01) --
	(221.41, 79.03) --
	(221.86, 76.28) --
	(222.31, 73.77) --
	(222.76, 71.47) --
	(223.21, 69.41) --
	(223.66, 67.54) --
	(224.11, 65.86) --
	(224.55, 64.36) --
	(225.00, 63.03) --
	(225.45, 61.89) --
	(225.90, 60.92) --
	(226.35, 60.11) --
	(226.80, 59.46) --
	(227.25, 58.94) --
	(227.69, 58.55) --
	(228.14, 58.26) --
	(228.59, 58.06) --
	(229.04, 57.91) --
	(229.49, 57.82) --
	(229.94, 57.76) --
	(230.39, 57.72) --
	(230.84, 57.70) --
	(231.28, 57.69) --
	(231.73, 57.68) --
	(232.18, 57.67) --
	(232.63, 57.67) --
	(233.08, 57.67) --
	(233.53, 57.67) --
	(233.98, 57.67) --
	(234.43, 57.67) --
	(234.87, 57.67) --
	(235.32, 57.67) --
	(235.77, 57.67) --
	(236.22, 57.67) --
	(236.67, 57.67) --
	(237.12, 57.67) --
	(237.57, 57.67) --
	(238.01, 57.67) --
	(238.46, 57.67) --
	(238.91, 57.67) --
	(239.36, 57.67) --
	(239.81, 57.67) --
	(240.26, 57.67) --
	(240.71, 57.67) --
	(241.16, 57.67) --
	(241.60, 57.67) --
	(242.05, 57.67) --
	(242.50, 57.67) --
	(242.95, 57.67) --
	(243.40, 57.67) --
	(243.85, 57.67) --
	(244.30, 57.67) --
	(244.75, 57.67) --
	(245.19, 57.67) --
	(245.64, 57.67) --
	(246.09, 57.67) --
	(246.54, 57.67) --
	(246.99, 57.67) --
	(247.44, 57.67) --
	(247.89, 57.67) --
	(248.33, 57.67) --
	(248.78, 57.67) --
	(249.23, 57.67) --
	(249.68, 57.67) --
	(250.13, 57.67) --
	(250.58, 57.67) --
	(251.03, 57.67) --
	(251.48, 57.67) --
	(251.92, 57.67) --
	(252.37, 57.67) --
	(252.82, 57.67) --
	(253.27, 57.67) --
	(253.72, 57.67) --
	(254.17, 57.67) --
	(254.62, 57.67) --
	(255.07, 57.67) --
	(255.51, 57.67) --
	(255.96, 57.67) --
	(256.41, 57.67) --
	(256.86, 57.67) --
	(257.31, 57.67) --
	(257.76, 57.67) --
	(258.21, 57.67) --
	(258.65, 57.67) --
	(259.10, 57.67) --
	(259.55, 57.67) --
	(260.00, 57.67) --
	(260.45, 57.67) --
	(260.90, 57.67) --
	(261.35, 57.67) --
	(261.80, 57.67) --
	(262.24, 57.67) --
	(262.69, 57.67) --
	(263.14, 57.67) --
	(263.59, 57.67) --
	(264.04, 57.67) --
	(264.49, 57.67) --
	(264.94, 57.67) --
	(265.39, 57.67) --
	(265.83, 57.67) --
	(266.28, 57.67) --
	(266.73, 57.67) --
	(267.18, 57.67) --
	(267.63, 57.67) --
	(268.08, 57.67) --
	(268.53, 57.67) --
	(268.97, 57.67) --
	(269.42, 57.67) --
	(269.87, 57.67) --
	(270.32, 57.67) --
	(270.77, 57.67) --
	(271.22, 57.67) --
	(271.67, 57.67) --
	(272.12, 57.67);

\path[] ( 42.83, 57.70) --
	( 43.28, 57.70) --
	( 43.73, 57.71) --
	( 44.18, 57.71) --
	( 44.63, 57.71) --
	( 45.08, 57.71) --
	( 45.52, 57.71) --
	( 45.97, 57.72) --
	( 46.42, 57.72) --
	( 46.87, 57.72) --
	( 47.32, 57.73) --
	( 47.77, 57.73) --
	( 48.22, 57.74) --
	( 48.67, 57.74) --
	( 49.11, 57.75) --
	( 49.56, 57.75) --
	( 50.01, 57.76) --
	( 50.46, 57.76) --
	( 50.91, 57.77) --
	( 51.36, 57.77) --
	( 51.81, 57.77) --
	( 52.25, 57.77) --
	( 52.70, 57.77) --
	( 53.15, 57.77) --
	( 53.60, 57.77) --
	( 54.05, 57.76) --
	( 54.50, 57.76) --
	( 54.95, 57.76) --
	( 55.40, 57.77) --
	( 55.84, 57.77) --
	( 56.29, 57.77) --
	( 56.74, 57.77) --
	( 57.19, 57.78) --
	( 57.64, 57.78) --
	( 58.09, 57.79) --
	( 58.54, 57.80) --
	( 58.99, 57.80) --
	( 59.43, 57.81) --
	( 59.88, 57.82) --
	( 60.33, 57.82) --
	( 60.78, 57.83) --
	( 61.23, 57.83) --
	( 61.68, 57.84) --
	( 62.13, 57.84) --
	( 62.57, 57.85) --
	( 63.02, 57.85) --
	( 63.47, 57.86) --
	( 63.92, 57.87) --
	( 64.37, 57.87) --
	( 64.82, 57.88) --
	( 65.27, 57.88) --
	( 65.72, 57.89) --
	( 66.16, 57.89) --
	( 66.61, 57.90) --
	( 67.06, 57.90) --
	( 67.51, 57.90) --
	( 67.96, 57.90) --
	( 68.41, 57.90) --
	( 68.86, 57.90) --
	( 69.31, 57.91) --
	( 69.75, 57.91) --
	( 70.20, 57.91) --
	( 70.65, 57.92) --
	( 71.10, 57.93) --
	( 71.55, 57.94) --
	( 72.00, 57.95) --
	( 72.45, 57.96) --
	( 72.89, 57.98) --
	( 73.34, 57.99) --
	( 73.79, 58.01) --
	( 74.24, 58.03) --
	( 74.69, 58.06) --
	( 75.14, 58.08) --
	( 75.59, 58.10) --
	( 76.04, 58.12) --
	( 76.48, 58.15) --
	( 76.93, 58.17) --
	( 77.38, 58.19) --
	( 77.83, 58.20) --
	( 78.28, 58.22) --
	( 78.73, 58.23) --
	( 79.18, 58.23) --
	( 79.63, 58.24) --
	( 80.07, 58.24) --
	( 80.52, 58.24) --
	( 80.97, 58.25) --
	( 81.42, 58.25) --
	( 81.87, 58.25) --
	( 82.32, 58.25) --
	( 82.77, 58.25) --
	( 83.21, 58.26) --
	( 83.66, 58.27) --
	( 84.11, 58.29) --
	( 84.56, 58.31) --
	( 85.01, 58.34) --
	( 85.46, 58.37) --
	( 85.91, 58.40) --
	( 86.36, 58.44) --
	( 86.80, 58.48) --
	( 87.25, 58.52) --
	( 87.70, 58.56) --
	( 88.15, 58.59) --
	( 88.60, 58.63) --
	( 89.05, 58.65) --
	( 89.50, 58.68) --
	( 89.95, 58.70) --
	( 90.39, 58.71) --
	( 90.84, 58.72) --
	( 91.29, 58.73) --
	( 91.74, 58.74) --
	( 92.19, 58.75) --
	( 92.64, 58.76) --
	( 93.09, 58.78) --
	( 93.53, 58.81) --
	( 93.98, 58.84) --
	( 94.43, 58.87) --
	( 94.88, 58.91) --
	( 95.33, 58.95) --
	( 95.78, 58.99) --
	( 96.23, 59.03) --
	( 96.68, 59.07) --
	( 97.12, 59.12) --
	( 97.57, 59.16) --
	( 98.02, 59.20) --
	( 98.47, 59.25) --
	( 98.92, 59.29) --
	( 99.37, 59.33) --
	( 99.82, 59.38) --
	(100.27, 59.42) --
	(100.71, 59.46) --
	(101.16, 59.51) --
	(101.61, 59.56) --
	(102.06, 59.61) --
	(102.51, 59.67) --
	(102.96, 59.73) --
	(103.41, 59.78) --
	(103.85, 59.85) --
	(104.30, 59.91) --
	(104.75, 59.97) --
	(105.20, 60.02) --
	(105.65, 60.07) --
	(106.10, 60.11) --
	(106.55, 60.15) --
	(107.00, 60.17) --
	(107.44, 60.19) --
	(107.89, 60.19) --
	(108.34, 60.19) --
	(108.79, 60.19) --
	(109.24, 60.18) --
	(109.69, 60.17) --
	(110.14, 60.18) --
	(110.59, 60.18) --
	(111.03, 60.20) --
	(111.48, 60.23) --
	(111.93, 60.27) --
	(112.38, 60.32) --
	(112.83, 60.38) --
	(113.28, 60.45) --
	(113.73, 60.51) --
	(114.17, 60.58) --
	(114.62, 60.65) --
	(115.07, 60.71) --
	(115.52, 60.77) --
	(115.97, 60.82) --
	(116.42, 60.86) --
	(116.87, 60.90) --
	(117.32, 60.93) --
	(117.76, 60.96) --
	(118.21, 60.99) --
	(118.66, 61.03) --
	(119.11, 61.07) --
	(119.56, 61.12) --
	(120.01, 61.18) --
	(120.46, 61.25) --
	(120.91, 61.32) --
	(121.35, 61.40) --
	(121.80, 61.48) --
	(122.25, 61.56) --
	(122.70, 61.64) --
	(123.15, 61.72) --
	(123.60, 61.80) --
	(124.05, 61.87) --
	(124.49, 61.94) --
	(124.94, 62.01) --
	(125.39, 62.07) --
	(125.84, 62.13) --
	(126.29, 62.20) --
	(126.74, 62.26) --
	(127.19, 62.32) --
	(127.64, 62.39) --
	(128.08, 62.45) --
	(128.53, 62.52) --
	(128.98, 62.59) --
	(129.43, 62.66) --
	(129.88, 62.74) --
	(130.33, 62.81) --
	(130.78, 62.89) --
	(131.23, 62.96) --
	(131.67, 63.04) --
	(132.12, 63.11) --
	(132.57, 63.18) --
	(133.02, 63.25) --
	(133.47, 63.32) --
	(133.92, 63.38) --
	(134.37, 63.44) --
	(134.81, 63.50) --
	(135.26, 63.55) --
	(135.71, 63.62) --
	(136.16, 63.69) --
	(136.61, 63.76) --
	(137.06, 63.85) --
	(137.51, 63.95) --
	(137.96, 64.05) --
	(138.40, 64.17) --
	(138.85, 64.30) --
	(139.30, 64.43) --
	(139.75, 64.57) --
	(140.20, 64.71) --
	(140.65, 64.86) --
	(141.10, 65.00) --
	(141.55, 65.15) --
	(141.99, 65.30) --
	(142.44, 65.45) --
	(142.89, 65.60) --
	(143.34, 65.76) --
	(143.79, 65.92) --
	(144.24, 66.09) --
	(144.69, 66.26) --
	(145.13, 66.44) --
	(145.58, 66.62) --
	(146.03, 66.81) --
	(146.48, 67.01) --
	(146.93, 67.21) --
	(147.38, 67.41) --
	(147.83, 67.62) --
	(148.28, 67.82) --
	(148.72, 68.03) --
	(149.17, 68.23) --
	(149.62, 68.42) --
	(150.07, 68.61) --
	(150.52, 68.79) --
	(150.97, 68.96) --
	(151.42, 69.13) --
	(151.87, 69.30) --
	(152.31, 69.47) --
	(152.76, 69.65) --
	(153.21, 69.85) --
	(153.66, 70.07) --
	(154.11, 70.32) --
	(154.56, 70.61) --
	(155.01, 70.93) --
	(155.45, 71.29) --
	(155.90, 71.69) --
	(156.35, 72.13) --
	(156.80, 72.61) --
	(157.25, 73.11) --
	(157.70, 73.64) --
	(158.15, 74.17) --
	(158.60, 74.71) --
	(159.04, 75.24) --
	(159.49, 75.76) --
	(159.94, 76.24) --
	(160.39, 76.70) --
	(160.84, 77.12) --
	(161.29, 77.51) --
	(161.74, 77.85) --
	(162.19, 78.17) --
	(162.63, 78.46) --
	(163.08, 78.72) --
	(163.53, 78.98) --
	(163.98, 79.23) --
	(164.43, 79.49) --
	(164.88, 79.75) --
	(165.33, 80.04) --
	(165.77, 80.35) --
	(166.22, 80.68) --
	(166.67, 81.04) --
	(167.12, 81.42) --
	(167.57, 81.83) --
	(168.02, 82.26) --
	(168.47, 82.72) --
	(168.92, 83.20) --
	(169.36, 83.71) --
	(169.81, 84.24) --
	(170.26, 84.79) --
	(170.71, 85.37) --
	(171.16, 85.96) --
	(171.61, 86.58) --
	(172.06, 87.21) --
	(172.51, 87.86) --
	(172.95, 88.54) --
	(173.40, 89.23) --
	(173.85, 89.94) --
	(174.30, 90.67) --
	(174.75, 91.43) --
	(175.20, 92.21) --
	(175.65, 93.02) --
	(176.09, 93.87) --
	(176.54, 94.74) --
	(176.99, 95.65) --
	(177.44, 96.59) --
	(177.89, 97.55) --
	(178.34, 98.54) --
	(178.79, 99.54) --
	(179.24,100.53) --
	(179.68,101.52) --
	(180.13,102.48) --
	(180.58,103.41) --
	(181.03,104.31) --
	(181.48,105.18) --
	(181.93,106.03) --
	(182.38,106.86) --
	(182.83,107.68) --
	(183.27,108.52) --
	(183.72,109.40) --
	(184.17,110.33) --
	(184.62,111.31) --
	(185.07,112.38) --
	(185.52,113.52) --
	(185.97,114.73) --
	(186.41,116.03) --
	(186.86,117.38) --
	(187.31,118.80) --
	(187.76,120.26) --
	(188.21,121.74) --
	(188.66,123.24) --
	(189.11,124.74) --
	(189.56,126.22) --
	(190.00,127.67) --
	(190.45,129.10) --
	(190.90,130.47) --
	(191.35,131.80) --
	(191.80,133.05) --
	(192.25,134.23) --
	(192.70,135.33) --
	(193.15,136.32) --
	(193.59,137.20) --
	(194.04,137.96) --
	(194.49,138.59) --
	(194.94,139.09) --
	(195.39,139.47) --
	(195.84,139.74) --
	(196.29,139.90) --
	(196.73,139.95) --
	(197.18,139.91) --
	(197.63,139.78) --
	(198.08,139.56) --
	(198.53,139.28) --
	(198.98,138.91) --
	(199.43,138.48) --
	(199.88,137.97) --
	(200.32,137.40) --
	(200.77,136.76) --
	(201.22,136.05) --
	(201.67,135.29) --
	(202.12,134.47) --
	(202.57,133.60) --
	(203.02,132.69) --
	(203.47,131.73) --
	(203.91,130.74) --
	(204.36,129.72) --
	(204.81,128.68) --
	(205.26,127.62) --
	(205.71,126.55) --
	(206.16,125.46) --
	(206.61,124.38) --
	(207.05,123.29) --
	(207.50,122.22) --
	(207.95,121.17) --
	(208.40,120.15) --
	(208.85,119.18) --
	(209.30,118.27) --
	(209.75,117.43) --
	(210.20,116.66) --
	(210.64,115.97) --
	(211.09,115.36) --
	(211.54,114.82) --
	(211.99,114.34) --
	(212.44,113.91) --
	(212.89,113.53) --
	(213.34,113.16) --
	(213.79,112.81) --
	(214.23,112.44) --
	(214.68,112.06) --
	(215.13,111.65) --
	(215.58,111.21) --
	(216.03,110.74) --
	(216.48,110.24) --
	(216.93,109.72) --
	(217.37,109.19) --
	(217.82,108.65) --
	(218.27,108.11) --
	(218.72,107.59) --
	(219.17,107.07) --
	(219.62,106.58) --
	(220.07,106.10) --
	(220.52,105.63) --
	(220.96,105.17) --
	(221.41,104.70) --
	(221.86,104.23) --
	(222.31,103.75) --
	(222.76,103.25) --
	(223.21,102.72) --
	(223.66,102.17) --
	(224.11,101.59) --
	(224.55,100.98) --
	(225.00,100.33) --
	(225.45, 99.65) --
	(225.90, 98.94) --
	(226.35, 98.20) --
	(226.80, 97.44) --
	(227.25, 96.65) --
	(227.69, 95.85) --
	(228.14, 95.03) --
	(228.59, 94.20) --
	(229.04, 93.36) --
	(229.49, 92.52) --
	(229.94, 91.68) --
	(230.39, 90.84) --
	(230.84, 90.00) --
	(231.28, 89.17) --
	(231.73, 88.35) --
	(232.18, 87.53) --
	(232.63, 86.72) --
	(233.08, 85.93) --
	(233.53, 85.14) --
	(233.98, 84.36) --
	(234.43, 83.60) --
	(234.87, 82.84) --
	(235.32, 82.08) --
	(235.77, 81.33) --
	(236.22, 80.59) --
	(236.67, 79.85) --
	(237.12, 79.11) --
	(237.57, 78.38) --
	(238.01, 77.66) --
	(238.46, 76.95) --
	(238.91, 76.26) --
	(239.36, 75.59) --
	(239.81, 74.94) --
	(240.26, 74.31) --
	(240.71, 73.71) --
	(241.16, 73.14) --
	(241.60, 72.58) --
	(242.05, 72.05) --
	(242.50, 71.54) --
	(242.95, 71.05) --
	(243.40, 70.59) --
	(243.85, 70.14) --
	(244.30, 69.70) --
	(244.75, 69.29) --
	(245.19, 68.89) --
	(245.64, 68.51) --
	(246.09, 68.14) --
	(246.54, 67.79) --
	(246.99, 67.45) --
	(247.44, 67.13) --
	(247.89, 66.82) --
	(248.33, 66.52) --
	(248.78, 66.23) --
	(249.23, 65.96) --
	(249.68, 65.69) --
	(250.13, 65.43) --
	(250.58, 65.17) --
	(251.03, 64.91) --
	(251.48, 64.64) --
	(251.92, 64.37) --
	(252.37, 64.09) --
	(252.82, 63.79) --
	(253.27, 63.48) --
	(253.72, 63.16) --
	(254.17, 62.83) --
	(254.62, 62.50) --
	(255.07, 62.16) --
	(255.51, 61.82) --
	(255.96, 61.49) --
	(256.41, 61.17) --
	(256.86, 60.86) --
	(257.31, 60.57) --
	(257.76, 60.30) --
	(258.21, 60.04) --
	(258.65, 59.81) --
	(259.10, 59.60) --
	(259.55, 59.41) --
	(260.00, 59.23) --
	(260.45, 59.08) --
	(260.90, 58.94) --
	(261.35, 58.81) --
	(261.80, 58.71) --
	(262.24, 58.61) --
	(262.69, 58.53) --
	(263.14, 58.47) --
	(263.59, 58.41) --
	(264.04, 58.37) --
	(264.49, 58.33) --
	(264.94, 58.31) --
	(265.39, 58.29) --
	(265.83, 58.28) --
	(266.28, 58.27) --
	(266.73, 58.27) --
	(267.18, 58.27) --
	(267.63, 58.26) --
	(268.08, 58.25) --
	(268.53, 58.24) --
	(268.97, 58.22) --
	(269.42, 58.20) --
	(269.87, 58.17) --
	(270.32, 58.14) --
	(270.77, 58.10) --
	(271.22, 58.06) --
	(271.67, 58.02) --
	(272.12, 57.97) --
	(272.12, 57.67) --
	(271.67, 57.67) --
	(271.22, 57.67) --
	(270.77, 57.67) --
	(270.32, 57.67) --
	(269.87, 57.67) --
	(269.42, 57.67) --
	(268.97, 57.67) --
	(268.53, 57.67) --
	(268.08, 57.67) --
	(267.63, 57.67) --
	(267.18, 57.67) --
	(266.73, 57.67) --
	(266.28, 57.67) --
	(265.83, 57.67) --
	(265.39, 57.67) --
	(264.94, 57.67) --
	(264.49, 57.67) --
	(264.04, 57.67) --
	(263.59, 57.67) --
	(263.14, 57.67) --
	(262.69, 57.67) --
	(262.24, 57.67) --
	(261.80, 57.67) --
	(261.35, 57.67) --
	(260.90, 57.67) --
	(260.45, 57.67) --
	(260.00, 57.67) --
	(259.55, 57.67) --
	(259.10, 57.67) --
	(258.65, 57.67) --
	(258.21, 57.67) --
	(257.76, 57.67) --
	(257.31, 57.67) --
	(256.86, 57.67) --
	(256.41, 57.67) --
	(255.96, 57.67) --
	(255.51, 57.67) --
	(255.07, 57.67) --
	(254.62, 57.67) --
	(254.17, 57.67) --
	(253.72, 57.67) --
	(253.27, 57.67) --
	(252.82, 57.67) --
	(252.37, 57.67) --
	(251.92, 57.67) --
	(251.48, 57.67) --
	(251.03, 57.67) --
	(250.58, 57.67) --
	(250.13, 57.67) --
	(249.68, 57.67) --
	(249.23, 57.67) --
	(248.78, 57.67) --
	(248.33, 57.67) --
	(247.89, 57.67) --
	(247.44, 57.67) --
	(246.99, 57.67) --
	(246.54, 57.67) --
	(246.09, 57.67) --
	(245.64, 57.67) --
	(245.19, 57.67) --
	(244.75, 57.67) --
	(244.30, 57.67) --
	(243.85, 57.67) --
	(243.40, 57.67) --
	(242.95, 57.67) --
	(242.50, 57.67) --
	(242.05, 57.67) --
	(241.60, 57.67) --
	(241.16, 57.67) --
	(240.71, 57.67) --
	(240.26, 57.67) --
	(239.81, 57.67) --
	(239.36, 57.67) --
	(238.91, 57.67) --
	(238.46, 57.67) --
	(238.01, 57.67) --
	(237.57, 57.67) --
	(237.12, 57.67) --
	(236.67, 57.67) --
	(236.22, 57.67) --
	(235.77, 57.67) --
	(235.32, 57.67) --
	(234.87, 57.67) --
	(234.43, 57.67) --
	(233.98, 57.67) --
	(233.53, 57.67) --
	(233.08, 57.67) --
	(232.63, 57.67) --
	(232.18, 57.67) --
	(231.73, 57.67) --
	(231.28, 57.67) --
	(230.84, 57.67) --
	(230.39, 57.67) --
	(229.94, 57.67) --
	(229.49, 57.67) --
	(229.04, 57.67) --
	(228.59, 57.67) --
	(228.14, 57.67) --
	(227.69, 57.67) --
	(227.25, 57.67) --
	(226.80, 57.67) --
	(226.35, 57.67) --
	(225.90, 57.67) --
	(225.45, 57.67) --
	(225.00, 57.67) --
	(224.55, 57.67) --
	(224.11, 57.67) --
	(223.66, 57.67) --
	(223.21, 57.67) --
	(222.76, 57.67) --
	(222.31, 57.67) --
	(221.86, 57.67) --
	(221.41, 57.67) --
	(220.96, 57.67) --
	(220.52, 57.67) --
	(220.07, 57.67) --
	(219.62, 57.67) --
	(219.17, 57.67) --
	(218.72, 57.67) --
	(218.27, 57.67) --
	(217.82, 57.67) --
	(217.37, 57.67) --
	(216.93, 57.67) --
	(216.48, 57.67) --
	(216.03, 57.67) --
	(215.58, 57.67) --
	(215.13, 57.67) --
	(214.68, 57.67) --
	(214.23, 57.67) --
	(213.79, 57.67) --
	(213.34, 57.67) --
	(212.89, 57.67) --
	(212.44, 57.67) --
	(211.99, 57.67) --
	(211.54, 57.67) --
	(211.09, 57.67) --
	(210.64, 57.67) --
	(210.20, 57.67) --
	(209.75, 57.67) --
	(209.30, 57.67) --
	(208.85, 57.67) --
	(208.40, 57.67) --
	(207.95, 57.67) --
	(207.50, 57.67) --
	(207.05, 57.67) --
	(206.61, 57.67) --
	(206.16, 57.67) --
	(205.71, 57.67) --
	(205.26, 57.67) --
	(204.81, 57.67) --
	(204.36, 57.67) --
	(203.91, 57.67) --
	(203.47, 57.67) --
	(203.02, 57.67) --
	(202.57, 57.67) --
	(202.12, 57.67) --
	(201.67, 57.67) --
	(201.22, 57.67) --
	(200.77, 57.67) --
	(200.32, 57.67) --
	(199.88, 57.67) --
	(199.43, 57.67) --
	(198.98, 57.67) --
	(198.53, 57.67) --
	(198.08, 57.67) --
	(197.63, 57.67) --
	(197.18, 57.67) --
	(196.73, 57.67) --
	(196.29, 57.67) --
	(195.84, 57.67) --
	(195.39, 57.67) --
	(194.94, 57.67) --
	(194.49, 57.67) --
	(194.04, 57.67) --
	(193.59, 57.67) --
	(193.15, 57.67) --
	(192.70, 57.67) --
	(192.25, 57.67) --
	(191.80, 57.67) --
	(191.35, 57.67) --
	(190.90, 57.67) --
	(190.45, 57.67) --
	(190.00, 57.67) --
	(189.56, 57.67) --
	(189.11, 57.67) --
	(188.66, 57.67) --
	(188.21, 57.67) --
	(187.76, 57.67) --
	(187.31, 57.67) --
	(186.86, 57.67) --
	(186.41, 57.67) --
	(185.97, 57.67) --
	(185.52, 57.67) --
	(185.07, 57.67) --
	(184.62, 57.67) --
	(184.17, 57.67) --
	(183.72, 57.67) --
	(183.27, 57.67) --
	(182.83, 57.67) --
	(182.38, 57.67) --
	(181.93, 57.67) --
	(181.48, 57.67) --
	(181.03, 57.67) --
	(180.58, 57.67) --
	(180.13, 57.67) --
	(179.68, 57.67) --
	(179.24, 57.67) --
	(178.79, 57.67) --
	(178.34, 57.67) --
	(177.89, 57.67) --
	(177.44, 57.67) --
	(176.99, 57.67) --
	(176.54, 57.67) --
	(176.09, 57.67) --
	(175.65, 57.67) --
	(175.20, 57.67) --
	(174.75, 57.67) --
	(174.30, 57.67) --
	(173.85, 57.67) --
	(173.40, 57.67) --
	(172.95, 57.67) --
	(172.51, 57.67) --
	(172.06, 57.67) --
	(171.61, 57.67) --
	(171.16, 57.67) --
	(170.71, 57.67) --
	(170.26, 57.67) --
	(169.81, 57.67) --
	(169.36, 57.67) --
	(168.92, 57.67) --
	(168.47, 57.67) --
	(168.02, 57.67) --
	(167.57, 57.67) --
	(167.12, 57.67) --
	(166.67, 57.67) --
	(166.22, 57.67) --
	(165.77, 57.67) --
	(165.33, 57.67) --
	(164.88, 57.67) --
	(164.43, 57.67) --
	(163.98, 57.67) --
	(163.53, 57.67) --
	(163.08, 57.67) --
	(162.63, 57.67) --
	(162.19, 57.67) --
	(161.74, 57.67) --
	(161.29, 57.67) --
	(160.84, 57.67) --
	(160.39, 57.67) --
	(159.94, 57.67) --
	(159.49, 57.67) --
	(159.04, 57.67) --
	(158.60, 57.67) --
	(158.15, 57.67) --
	(157.70, 57.67) --
	(157.25, 57.67) --
	(156.80, 57.67) --
	(156.35, 57.67) --
	(155.90, 57.67) --
	(155.45, 57.67) --
	(155.01, 57.67) --
	(154.56, 57.67) --
	(154.11, 57.67) --
	(153.66, 57.67) --
	(153.21, 57.67) --
	(152.76, 57.67) --
	(152.31, 57.67) --
	(151.87, 57.67) --
	(151.42, 57.67) --
	(150.97, 57.67) --
	(150.52, 57.67) --
	(150.07, 57.67) --
	(149.62, 57.67) --
	(149.17, 57.67) --
	(148.72, 57.67) --
	(148.28, 57.67) --
	(147.83, 57.67) --
	(147.38, 57.67) --
	(146.93, 57.67) --
	(146.48, 57.67) --
	(146.03, 57.67) --
	(145.58, 57.67) --
	(145.13, 57.67) --
	(144.69, 57.67) --
	(144.24, 57.67) --
	(143.79, 57.67) --
	(143.34, 57.67) --
	(142.89, 57.67) --
	(142.44, 57.67) --
	(141.99, 57.67) --
	(141.55, 57.67) --
	(141.10, 57.67) --
	(140.65, 57.67) --
	(140.20, 57.67) --
	(139.75, 57.67) --
	(139.30, 57.67) --
	(138.85, 57.67) --
	(138.40, 57.67) --
	(137.96, 57.67) --
	(137.51, 57.67) --
	(137.06, 57.67) --
	(136.61, 57.67) --
	(136.16, 57.67) --
	(135.71, 57.67) --
	(135.26, 57.67) --
	(134.81, 57.67) --
	(134.37, 57.67) --
	(133.92, 57.67) --
	(133.47, 57.67) --
	(133.02, 57.67) --
	(132.57, 57.67) --
	(132.12, 57.67) --
	(131.67, 57.67) --
	(131.23, 57.67) --
	(130.78, 57.67) --
	(130.33, 57.67) --
	(129.88, 57.67) --
	(129.43, 57.67) --
	(128.98, 57.67) --
	(128.53, 57.67) --
	(128.08, 57.67) --
	(127.64, 57.67) --
	(127.19, 57.67) --
	(126.74, 57.67) --
	(126.29, 57.67) --
	(125.84, 57.67) --
	(125.39, 57.67) --
	(124.94, 57.67) --
	(124.49, 57.67) --
	(124.05, 57.67) --
	(123.60, 57.67) --
	(123.15, 57.67) --
	(122.70, 57.67) --
	(122.25, 57.67) --
	(121.80, 57.67) --
	(121.35, 57.67) --
	(120.91, 57.67) --
	(120.46, 57.67) --
	(120.01, 57.67) --
	(119.56, 57.67) --
	(119.11, 57.67) --
	(118.66, 57.67) --
	(118.21, 57.67) --
	(117.76, 57.67) --
	(117.32, 57.67) --
	(116.87, 57.67) --
	(116.42, 57.67) --
	(115.97, 57.67) --
	(115.52, 57.67) --
	(115.07, 57.67) --
	(114.62, 57.67) --
	(114.17, 57.67) --
	(113.73, 57.67) --
	(113.28, 57.67) --
	(112.83, 57.67) --
	(112.38, 57.67) --
	(111.93, 57.67) --
	(111.48, 57.67) --
	(111.03, 57.67) --
	(110.59, 57.67) --
	(110.14, 57.67) --
	(109.69, 57.67) --
	(109.24, 57.67) --
	(108.79, 57.67) --
	(108.34, 57.67) --
	(107.89, 57.67) --
	(107.44, 57.67) --
	(107.00, 57.67) --
	(106.55, 57.67) --
	(106.10, 57.67) --
	(105.65, 57.67) --
	(105.20, 57.67) --
	(104.75, 57.67) --
	(104.30, 57.67) --
	(103.85, 57.67) --
	(103.41, 57.67) --
	(102.96, 57.67) --
	(102.51, 57.67) --
	(102.06, 57.67) --
	(101.61, 57.67) --
	(101.16, 57.67) --
	(100.71, 57.67) --
	(100.27, 57.67) --
	( 99.82, 57.67) --
	( 99.37, 57.67) --
	( 98.92, 57.67) --
	( 98.47, 57.67) --
	( 98.02, 57.67) --
	( 97.57, 57.67) --
	( 97.12, 57.67) --
	( 96.68, 57.67) --
	( 96.23, 57.67) --
	( 95.78, 57.67) --
	( 95.33, 57.67) --
	( 94.88, 57.67) --
	( 94.43, 57.67) --
	( 93.98, 57.67) --
	( 93.53, 57.67) --
	( 93.09, 57.67) --
	( 92.64, 57.67) --
	( 92.19, 57.67) --
	( 91.74, 57.67) --
	( 91.29, 57.67) --
	( 90.84, 57.67) --
	( 90.39, 57.67) --
	( 89.95, 57.67) --
	( 89.50, 57.67) --
	( 89.05, 57.67) --
	( 88.60, 57.67) --
	( 88.15, 57.67) --
	( 87.70, 57.67) --
	( 87.25, 57.67) --
	( 86.80, 57.67) --
	( 86.36, 57.67) --
	( 85.91, 57.67) --
	( 85.46, 57.67) --
	( 85.01, 57.67) --
	( 84.56, 57.67) --
	( 84.11, 57.67) --
	( 83.66, 57.67) --
	( 83.21, 57.67) --
	( 82.77, 57.67) --
	( 82.32, 57.67) --
	( 81.87, 57.67) --
	( 81.42, 57.67) --
	( 80.97, 57.67) --
	( 80.52, 57.67) --
	( 80.07, 57.67) --
	( 79.63, 57.67) --
	( 79.18, 57.67) --
	( 78.73, 57.67) --
	( 78.28, 57.67) --
	( 77.83, 57.67) --
	( 77.38, 57.67) --
	( 76.93, 57.67) --
	( 76.48, 57.67) --
	( 76.04, 57.67) --
	( 75.59, 57.67) --
	( 75.14, 57.67) --
	( 74.69, 57.67) --
	( 74.24, 57.67) --
	( 73.79, 57.67) --
	( 73.34, 57.67) --
	( 72.89, 57.67) --
	( 72.45, 57.67) --
	( 72.00, 57.67) --
	( 71.55, 57.67) --
	( 71.10, 57.67) --
	( 70.65, 57.67) --
	( 70.20, 57.67) --
	( 69.75, 57.67) --
	( 69.31, 57.67) --
	( 68.86, 57.67) --
	( 68.41, 57.67) --
	( 67.96, 57.67) --
	( 67.51, 57.67) --
	( 67.06, 57.67) --
	( 66.61, 57.67) --
	( 66.16, 57.67) --
	( 65.72, 57.67) --
	( 65.27, 57.67) --
	( 64.82, 57.67) --
	( 64.37, 57.67) --
	( 63.92, 57.67) --
	( 63.47, 57.67) --
	( 63.02, 57.67) --
	( 62.57, 57.67) --
	( 62.13, 57.67) --
	( 61.68, 57.67) --
	( 61.23, 57.67) --
	( 60.78, 57.67) --
	( 60.33, 57.67) --
	( 59.88, 57.67) --
	( 59.43, 57.67) --
	( 58.99, 57.67) --
	( 58.54, 57.67) --
	( 58.09, 57.67) --
	( 57.64, 57.67) --
	( 57.19, 57.67) --
	( 56.74, 57.67) --
	( 56.29, 57.67) --
	( 55.84, 57.67) --
	( 55.40, 57.67) --
	( 54.95, 57.67) --
	( 54.50, 57.67) --
	( 54.05, 57.67) --
	( 53.60, 57.67) --
	( 53.15, 57.67) --
	( 52.70, 57.67) --
	( 52.25, 57.67) --
	( 51.81, 57.67) --
	( 51.36, 57.67) --
	( 50.91, 57.67) --
	( 50.46, 57.67) --
	( 50.01, 57.67) --
	( 49.56, 57.67) --
	( 49.11, 57.67) --
	( 48.67, 57.67) --
	( 48.22, 57.67) --
	( 47.77, 57.67) --
	( 47.32, 57.67) --
	( 46.87, 57.67) --
	( 46.42, 57.67) --
	( 45.97, 57.67) --
	( 45.52, 57.67) --
	( 45.08, 57.67) --
	( 44.63, 57.67) --
	( 44.18, 57.67) --
	( 43.73, 57.67) --
	( 43.28, 57.67) --
	( 42.83, 57.67) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 42.83, 57.70) --
	( 43.28, 57.70) --
	( 43.73, 57.71) --
	( 44.18, 57.71) --
	( 44.63, 57.71) --
	( 45.08, 57.71) --
	( 45.52, 57.71) --
	( 45.97, 57.72) --
	( 46.42, 57.72) --
	( 46.87, 57.72) --
	( 47.32, 57.73) --
	( 47.77, 57.73) --
	( 48.22, 57.74) --
	( 48.67, 57.74) --
	( 49.11, 57.75) --
	( 49.56, 57.75) --
	( 50.01, 57.76) --
	( 50.46, 57.76) --
	( 50.91, 57.77) --
	( 51.36, 57.77) --
	( 51.81, 57.77) --
	( 52.25, 57.77) --
	( 52.70, 57.77) --
	( 53.15, 57.77) --
	( 53.60, 57.77) --
	( 54.05, 57.76) --
	( 54.50, 57.76) --
	( 54.95, 57.76) --
	( 55.40, 57.77) --
	( 55.84, 57.77) --
	( 56.29, 57.77) --
	( 56.74, 57.77) --
	( 57.19, 57.78) --
	( 57.64, 57.78) --
	( 58.09, 57.79) --
	( 58.54, 57.80) --
	( 58.99, 57.80) --
	( 59.43, 57.81) --
	( 59.88, 57.82) --
	( 60.33, 57.82) --
	( 60.78, 57.83) --
	( 61.23, 57.83) --
	( 61.68, 57.84) --
	( 62.13, 57.84) --
	( 62.57, 57.85) --
	( 63.02, 57.85) --
	( 63.47, 57.86) --
	( 63.92, 57.87) --
	( 64.37, 57.87) --
	( 64.82, 57.88) --
	( 65.27, 57.88) --
	( 65.72, 57.89) --
	( 66.16, 57.89) --
	( 66.61, 57.90) --
	( 67.06, 57.90) --
	( 67.51, 57.90) --
	( 67.96, 57.90) --
	( 68.41, 57.90) --
	( 68.86, 57.90) --
	( 69.31, 57.91) --
	( 69.75, 57.91) --
	( 70.20, 57.91) --
	( 70.65, 57.92) --
	( 71.10, 57.93) --
	( 71.55, 57.94) --
	( 72.00, 57.95) --
	( 72.45, 57.96) --
	( 72.89, 57.98) --
	( 73.34, 57.99) --
	( 73.79, 58.01) --
	( 74.24, 58.03) --
	( 74.69, 58.06) --
	( 75.14, 58.08) --
	( 75.59, 58.10) --
	( 76.04, 58.12) --
	( 76.48, 58.15) --
	( 76.93, 58.17) --
	( 77.38, 58.19) --
	( 77.83, 58.20) --
	( 78.28, 58.22) --
	( 78.73, 58.23) --
	( 79.18, 58.23) --
	( 79.63, 58.24) --
	( 80.07, 58.24) --
	( 80.52, 58.24) --
	( 80.97, 58.25) --
	( 81.42, 58.25) --
	( 81.87, 58.25) --
	( 82.32, 58.25) --
	( 82.77, 58.25) --
	( 83.21, 58.26) --
	( 83.66, 58.27) --
	( 84.11, 58.29) --
	( 84.56, 58.31) --
	( 85.01, 58.34) --
	( 85.46, 58.37) --
	( 85.91, 58.40) --
	( 86.36, 58.44) --
	( 86.80, 58.48) --
	( 87.25, 58.52) --
	( 87.70, 58.56) --
	( 88.15, 58.59) --
	( 88.60, 58.63) --
	( 89.05, 58.65) --
	( 89.50, 58.68) --
	( 89.95, 58.70) --
	( 90.39, 58.71) --
	( 90.84, 58.72) --
	( 91.29, 58.73) --
	( 91.74, 58.74) --
	( 92.19, 58.75) --
	( 92.64, 58.76) --
	( 93.09, 58.78) --
	( 93.53, 58.81) --
	( 93.98, 58.84) --
	( 94.43, 58.87) --
	( 94.88, 58.91) --
	( 95.33, 58.95) --
	( 95.78, 58.99) --
	( 96.23, 59.03) --
	( 96.68, 59.07) --
	( 97.12, 59.12) --
	( 97.57, 59.16) --
	( 98.02, 59.20) --
	( 98.47, 59.25) --
	( 98.92, 59.29) --
	( 99.37, 59.33) --
	( 99.82, 59.38) --
	(100.27, 59.42) --
	(100.71, 59.46) --
	(101.16, 59.51) --
	(101.61, 59.56) --
	(102.06, 59.61) --
	(102.51, 59.67) --
	(102.96, 59.73) --
	(103.41, 59.78) --
	(103.85, 59.85) --
	(104.30, 59.91) --
	(104.75, 59.97) --
	(105.20, 60.02) --
	(105.65, 60.07) --
	(106.10, 60.11) --
	(106.55, 60.15) --
	(107.00, 60.17) --
	(107.44, 60.19) --
	(107.89, 60.19) --
	(108.34, 60.19) --
	(108.79, 60.19) --
	(109.24, 60.18) --
	(109.69, 60.17) --
	(110.14, 60.18) --
	(110.59, 60.18) --
	(111.03, 60.20) --
	(111.48, 60.23) --
	(111.93, 60.27) --
	(112.38, 60.32) --
	(112.83, 60.38) --
	(113.28, 60.45) --
	(113.73, 60.51) --
	(114.17, 60.58) --
	(114.62, 60.65) --
	(115.07, 60.71) --
	(115.52, 60.77) --
	(115.97, 60.82) --
	(116.42, 60.86) --
	(116.87, 60.90) --
	(117.32, 60.93) --
	(117.76, 60.96) --
	(118.21, 60.99) --
	(118.66, 61.03) --
	(119.11, 61.07) --
	(119.56, 61.12) --
	(120.01, 61.18) --
	(120.46, 61.25) --
	(120.91, 61.32) --
	(121.35, 61.40) --
	(121.80, 61.48) --
	(122.25, 61.56) --
	(122.70, 61.64) --
	(123.15, 61.72) --
	(123.60, 61.80) --
	(124.05, 61.87) --
	(124.49, 61.94) --
	(124.94, 62.01) --
	(125.39, 62.07) --
	(125.84, 62.13) --
	(126.29, 62.20) --
	(126.74, 62.26) --
	(127.19, 62.32) --
	(127.64, 62.39) --
	(128.08, 62.45) --
	(128.53, 62.52) --
	(128.98, 62.59) --
	(129.43, 62.66) --
	(129.88, 62.74) --
	(130.33, 62.81) --
	(130.78, 62.89) --
	(131.23, 62.96) --
	(131.67, 63.04) --
	(132.12, 63.11) --
	(132.57, 63.18) --
	(133.02, 63.25) --
	(133.47, 63.32) --
	(133.92, 63.38) --
	(134.37, 63.44) --
	(134.81, 63.50) --
	(135.26, 63.55) --
	(135.71, 63.62) --
	(136.16, 63.69) --
	(136.61, 63.76) --
	(137.06, 63.85) --
	(137.51, 63.95) --
	(137.96, 64.05) --
	(138.40, 64.17) --
	(138.85, 64.30) --
	(139.30, 64.43) --
	(139.75, 64.57) --
	(140.20, 64.71) --
	(140.65, 64.86) --
	(141.10, 65.00) --
	(141.55, 65.15) --
	(141.99, 65.30) --
	(142.44, 65.45) --
	(142.89, 65.60) --
	(143.34, 65.76) --
	(143.79, 65.92) --
	(144.24, 66.09) --
	(144.69, 66.26) --
	(145.13, 66.44) --
	(145.58, 66.62) --
	(146.03, 66.81) --
	(146.48, 67.01) --
	(146.93, 67.21) --
	(147.38, 67.41) --
	(147.83, 67.62) --
	(148.28, 67.82) --
	(148.72, 68.03) --
	(149.17, 68.23) --
	(149.62, 68.42) --
	(150.07, 68.61) --
	(150.52, 68.79) --
	(150.97, 68.96) --
	(151.42, 69.13) --
	(151.87, 69.30) --
	(152.31, 69.47) --
	(152.76, 69.65) --
	(153.21, 69.85) --
	(153.66, 70.07) --
	(154.11, 70.32) --
	(154.56, 70.61) --
	(155.01, 70.93) --
	(155.45, 71.29) --
	(155.90, 71.69) --
	(156.35, 72.13) --
	(156.80, 72.61) --
	(157.25, 73.11) --
	(157.70, 73.64) --
	(158.15, 74.17) --
	(158.60, 74.71) --
	(159.04, 75.24) --
	(159.49, 75.76) --
	(159.94, 76.24) --
	(160.39, 76.70) --
	(160.84, 77.12) --
	(161.29, 77.51) --
	(161.74, 77.85) --
	(162.19, 78.17) --
	(162.63, 78.46) --
	(163.08, 78.72) --
	(163.53, 78.98) --
	(163.98, 79.23) --
	(164.43, 79.49) --
	(164.88, 79.75) --
	(165.33, 80.04) --
	(165.77, 80.35) --
	(166.22, 80.68) --
	(166.67, 81.04) --
	(167.12, 81.42) --
	(167.57, 81.83) --
	(168.02, 82.26) --
	(168.47, 82.72) --
	(168.92, 83.20) --
	(169.36, 83.71) --
	(169.81, 84.24) --
	(170.26, 84.79) --
	(170.71, 85.37) --
	(171.16, 85.96) --
	(171.61, 86.58) --
	(172.06, 87.21) --
	(172.51, 87.86) --
	(172.95, 88.54) --
	(173.40, 89.23) --
	(173.85, 89.94) --
	(174.30, 90.67) --
	(174.75, 91.43) --
	(175.20, 92.21) --
	(175.65, 93.02) --
	(176.09, 93.87) --
	(176.54, 94.74) --
	(176.99, 95.65) --
	(177.44, 96.59) --
	(177.89, 97.55) --
	(178.34, 98.54) --
	(178.79, 99.54) --
	(179.24,100.53) --
	(179.68,101.52) --
	(180.13,102.48) --
	(180.58,103.41) --
	(181.03,104.31) --
	(181.48,105.18) --
	(181.93,106.03) --
	(182.38,106.86) --
	(182.83,107.68) --
	(183.27,108.52) --
	(183.72,109.40) --
	(184.17,110.33) --
	(184.62,111.31) --
	(185.07,112.38) --
	(185.52,113.52) --
	(185.97,114.73) --
	(186.41,116.03) --
	(186.86,117.38) --
	(187.31,118.80) --
	(187.76,120.26) --
	(188.21,121.74) --
	(188.66,123.24) --
	(189.11,124.74) --
	(189.56,126.22) --
	(190.00,127.67) --
	(190.45,129.10) --
	(190.90,130.47) --
	(191.35,131.80) --
	(191.80,133.05) --
	(192.25,134.23) --
	(192.70,135.33) --
	(193.15,136.32) --
	(193.59,137.20) --
	(194.04,137.96) --
	(194.49,138.59) --
	(194.94,139.09) --
	(195.39,139.47) --
	(195.84,139.74) --
	(196.29,139.90) --
	(196.73,139.95) --
	(197.18,139.91) --
	(197.63,139.78) --
	(198.08,139.56) --
	(198.53,139.28) --
	(198.98,138.91) --
	(199.43,138.48) --
	(199.88,137.97) --
	(200.32,137.40) --
	(200.77,136.76) --
	(201.22,136.05) --
	(201.67,135.29) --
	(202.12,134.47) --
	(202.57,133.60) --
	(203.02,132.69) --
	(203.47,131.73) --
	(203.91,130.74) --
	(204.36,129.72) --
	(204.81,128.68) --
	(205.26,127.62) --
	(205.71,126.55) --
	(206.16,125.46) --
	(206.61,124.38) --
	(207.05,123.29) --
	(207.50,122.22) --
	(207.95,121.17) --
	(208.40,120.15) --
	(208.85,119.18) --
	(209.30,118.27) --
	(209.75,117.43) --
	(210.20,116.66) --
	(210.64,115.97) --
	(211.09,115.36) --
	(211.54,114.82) --
	(211.99,114.34) --
	(212.44,113.91) --
	(212.89,113.53) --
	(213.34,113.16) --
	(213.79,112.81) --
	(214.23,112.44) --
	(214.68,112.06) --
	(215.13,111.65) --
	(215.58,111.21) --
	(216.03,110.74) --
	(216.48,110.24) --
	(216.93,109.72) --
	(217.37,109.19) --
	(217.82,108.65) --
	(218.27,108.11) --
	(218.72,107.59) --
	(219.17,107.07) --
	(219.62,106.58) --
	(220.07,106.10) --
	(220.52,105.63) --
	(220.96,105.17) --
	(221.41,104.70) --
	(221.86,104.23) --
	(222.31,103.75) --
	(222.76,103.25) --
	(223.21,102.72) --
	(223.66,102.17) --
	(224.11,101.59) --
	(224.55,100.98) --
	(225.00,100.33) --
	(225.45, 99.65) --
	(225.90, 98.94) --
	(226.35, 98.20) --
	(226.80, 97.44) --
	(227.25, 96.65) --
	(227.69, 95.85) --
	(228.14, 95.03) --
	(228.59, 94.20) --
	(229.04, 93.36) --
	(229.49, 92.52) --
	(229.94, 91.68) --
	(230.39, 90.84) --
	(230.84, 90.00) --
	(231.28, 89.17) --
	(231.73, 88.35) --
	(232.18, 87.53) --
	(232.63, 86.72) --
	(233.08, 85.93) --
	(233.53, 85.14) --
	(233.98, 84.36) --
	(234.43, 83.60) --
	(234.87, 82.84) --
	(235.32, 82.08) --
	(235.77, 81.33) --
	(236.22, 80.59) --
	(236.67, 79.85) --
	(237.12, 79.11) --
	(237.57, 78.38) --
	(238.01, 77.66) --
	(238.46, 76.95) --
	(238.91, 76.26) --
	(239.36, 75.59) --
	(239.81, 74.94) --
	(240.26, 74.31) --
	(240.71, 73.71) --
	(241.16, 73.14) --
	(241.60, 72.58) --
	(242.05, 72.05) --
	(242.50, 71.54) --
	(242.95, 71.05) --
	(243.40, 70.59) --
	(243.85, 70.14) --
	(244.30, 69.70) --
	(244.75, 69.29) --
	(245.19, 68.89) --
	(245.64, 68.51) --
	(246.09, 68.14) --
	(246.54, 67.79) --
	(246.99, 67.45) --
	(247.44, 67.13) --
	(247.89, 66.82) --
	(248.33, 66.52) --
	(248.78, 66.23) --
	(249.23, 65.96) --
	(249.68, 65.69) --
	(250.13, 65.43) --
	(250.58, 65.17) --
	(251.03, 64.91) --
	(251.48, 64.64) --
	(251.92, 64.37) --
	(252.37, 64.09) --
	(252.82, 63.79) --
	(253.27, 63.48) --
	(253.72, 63.16) --
	(254.17, 62.83) --
	(254.62, 62.50) --
	(255.07, 62.16) --
	(255.51, 61.82) --
	(255.96, 61.49) --
	(256.41, 61.17) --
	(256.86, 60.86) --
	(257.31, 60.57) --
	(257.76, 60.30) --
	(258.21, 60.04) --
	(258.65, 59.81) --
	(259.10, 59.60) --
	(259.55, 59.41) --
	(260.00, 59.23) --
	(260.45, 59.08) --
	(260.90, 58.94) --
	(261.35, 58.81) --
	(261.80, 58.71) --
	(262.24, 58.61) --
	(262.69, 58.53) --
	(263.14, 58.47) --
	(263.59, 58.41) --
	(264.04, 58.37) --
	(264.49, 58.33) --
	(264.94, 58.31) --
	(265.39, 58.29) --
	(265.83, 58.28) --
	(266.28, 58.27) --
	(266.73, 58.27) --
	(267.18, 58.27) --
	(267.63, 58.26) --
	(268.08, 58.25) --
	(268.53, 58.24) --
	(268.97, 58.22) --
	(269.42, 58.20) --
	(269.87, 58.17) --
	(270.32, 58.14) --
	(270.77, 58.10) --
	(271.22, 58.06) --
	(271.67, 58.02) --
	(272.12, 57.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.70, 51.21) -- (219.70, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.32, 51.21) -- (187.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.84, 51.21) -- (224.84, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.85, 51.21) -- (196.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.46, 51.21) -- (194.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.92, 51.21) -- (214.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.07, 51.21) -- (207.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (241.70, 51.21) -- (241.70, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.09, 51.21) -- (227.09, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.97, 51.21) -- (218.97, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (248.52, 51.21) -- (248.52, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.24, 51.21) -- (226.24, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.82, 51.21) -- (174.82, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (248.64, 51.21) -- (248.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (240.16, 51.21) -- (240.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.88, 51.21) -- (201.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.87, 51.21) -- (214.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.85, 51.21) -- (222.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.06, 51.21) -- (228.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.98, 51.21) -- (188.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.40, 51.21) -- (180.40, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.59, 51.21) -- (198.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.46, 51.21) -- (216.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.40, 51.21) -- (203.40, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.18, 51.21) -- (200.18, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.89, 51.21) -- (212.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.89, 51.21) -- (207.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.13, 51.21) -- (201.13, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.84, 51.21) -- (193.84, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.60, 51.21) -- (196.60, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.81, 51.21) -- (170.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.34, 51.21) -- (176.34, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.35, 51.21) -- (173.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (235.06, 51.21) -- (235.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.15, 51.21) -- (231.15, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.75, 51.21) -- (216.75, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.40, 51.21) -- (221.40, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.22, 51.21) -- (236.22, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.50, 51.21) -- (178.50, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.03, 51.21) -- (234.03, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.35, 51.21) -- (185.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.39, 51.21) -- (231.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.66, 51.21) -- (156.66, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (150.19, 51.21) -- (150.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.35, 51.21) -- (205.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.95, 51.21) -- (192.95, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.90, 51.21) -- (204.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (249.29, 51.21) -- (249.29, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (243.13, 51.21) -- (243.13, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.52, 51.21) -- (182.52, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.70, 51.21) -- (187.70, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (255.25, 51.21) -- (255.25, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.99, 51.21) -- (185.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (128.62, 51.21) -- (128.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.08, 51.21) -- (226.08, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.26, 51.21) -- (199.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.74, 51.21) -- (204.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.55, 51.21) -- (197.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.96, 51.21) -- (195.96, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.69, 51.21) -- (176.69, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.45, 51.21) -- (207.45, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (133.55, 51.21) -- (133.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.99, 51.21) -- (208.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.19, 51.21) -- (220.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.32, 51.21) -- (163.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (248.88, 51.21) -- (248.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.96, 51.21) -- (225.96, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (159.12, 51.21) -- (159.12, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.20, 51.21) -- (236.20, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.14, 51.21) -- (201.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.24, 51.21) -- (194.24, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.67, 51.21) -- (208.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.17, 51.21) -- (223.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.17, 51.21) -- (164.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (239.05, 51.21) -- (239.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.04, 51.21) -- (222.04, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.74, 51.21) -- (194.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.99, 51.21) -- (216.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.77, 51.21) -- (232.77, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.47, 51.21) -- (205.47, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (246.15, 51.21) -- (246.15, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.35, 51.21) -- (225.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.02, 51.21) -- (216.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (145.87, 51.21) -- (145.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.34, 51.21) -- (202.34, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.47, 51.21) -- (206.47, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.24, 51.21) -- (212.24, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.79, 51.21) -- (190.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (257.14, 51.21) -- (257.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (227.63, 51.21) -- (227.63, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (250.78, 51.21) -- (250.78, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.38, 51.21) -- (195.38, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (111.85, 51.21) -- (111.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (242.13, 51.21) -- (242.13, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.22, 51.21) -- (156.22, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.42, 51.21) -- (234.42, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.05, 51.21) -- (201.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.59, 51.21) -- (221.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.49, 51.21) -- (202.49, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.89, 51.21) -- (173.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (249.32, 51.21) -- (249.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.23, 51.21) -- (214.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.59, 51.21) -- (218.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.75, 51.21) -- (164.75, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (234.79, 51.21) -- (234.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.40, 51.21) -- (210.40, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (171.10, 51.21) -- (171.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.26, 51.21) -- (210.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.64, 51.21) -- (188.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.60, 51.21) -- (194.60, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.35, 51.21) -- (215.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.14, 51.21) -- (197.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (226.17, 51.21) -- (226.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.51, 51.21) -- (232.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.28, 51.21) -- (205.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.40, 51.21) -- (195.40, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.93, 51.21) -- (178.93, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.69, 51.21) -- (216.69, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (231.25, 51.21) -- (231.25, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.74, 51.21) -- (222.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.98, 51.21) -- (211.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.80, 51.21) -- (174.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.07, 51.21) -- (204.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (150.46, 51.21) -- (150.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (221.80, 51.21) -- (221.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.98, 51.21) -- (174.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.65, 51.21) -- (195.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.13, 51.21) -- (225.13, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.42, 51.21) -- (223.42, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.80, 51.21) -- (191.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.01, 51.21) -- (187.01, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.54, 51.21) -- (198.54, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (167.08, 51.21) -- (167.08, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.56, 51.21) -- (200.56, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.19, 51.21) -- (204.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (229.57, 51.21) -- (229.57, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.64, 51.21) -- (202.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.49, 51.21) -- (185.49, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.71, 51.21) -- (198.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.43, 51.21) -- (228.43, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.89, 51.21) -- (206.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.68, 51.21) -- (156.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.98, 51.21) -- (184.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.88, 51.21) -- (195.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.28, 51.21) -- (175.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.03, 51.21) -- (215.03, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.63, 51.21) -- (205.63, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.86, 51.21) -- (177.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.68, 51.21) -- (201.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.14, 51.21) -- (206.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.56, 51.21) -- (224.56, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.74, 51.21) -- (196.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.53, 51.21) -- (189.53, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (122.72, 51.21) -- (122.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.96, 51.21) -- (196.96, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.65, 51.21) -- (181.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (240.50, 51.21) -- (240.50, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.62, 51.21) -- (212.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.71, 51.21) -- (186.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.76, 51.21) -- (196.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.83, 51.21) -- (198.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.09, 51.21) -- (225.09, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.42, 51.21) -- (183.42, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (238.16, 51.21) -- (238.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.10, 51.21) -- (207.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.76, 51.21) -- (176.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.37, 51.21) -- (211.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.68, 51.21) -- (215.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (159.39, 51.21) -- (159.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.00, 51.21) -- (222.00, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.81, 51.21) -- (213.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (237.31, 51.21) -- (237.31, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.64, 51.21) -- (196.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (230.28, 51.21) -- (230.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.79, 51.21) -- (204.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.10, 51.21) -- (182.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (223.96, 51.21) -- (223.96, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.53, 51.21) -- (209.53, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (237.84, 51.21) -- (237.84, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.41, 51.21) -- (200.41, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.61, 51.21) -- (199.61, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.99, 51.21) -- (164.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.02, 51.21) -- (191.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.96, 51.21) -- (188.96, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.07, 51.21) -- (208.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.27, 51.21) -- (215.27, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.30, 51.21) -- (224.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.83, 51.21) -- (233.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (247.53, 51.21) -- (247.53, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (153.58, 51.21) -- (153.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.41, 51.21) -- (201.41, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.96, 51.21) -- (193.96, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.01, 51.21) -- (190.01, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (236.94, 51.21) -- (236.94, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.21, 51.21) -- (186.21, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.23, 51.21) -- (224.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (228.51, 51.21) -- (228.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (232.28, 51.21) -- (232.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (146.26, 51.21) -- (146.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.09, 51.21) -- (208.09, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.59, 51.21) -- (222.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.07, 51.21) -- (206.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.52, 51.21) -- (199.52, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (225.11, 51.21) -- (225.11, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.08, 51.21) -- (209.08, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.85, 51.21) -- (213.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (233.33, 51.21) -- (233.33, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.15, 51.21) -- (217.15, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.91, 51.21) -- (174.91, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.71, 51.21) -- (220.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.99, 51.21) -- (195.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.70, 51.21) -- (175.70, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.08, 51.21) -- (195.08, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.25, 51.21) -- (183.25, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.78, 51.21) -- (182.78, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.15, 51.21) -- (207.15, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.10, 51.21) -- (199.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.69, 51.21) -- (210.69, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.92, 51.21) -- (211.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.79, 51.21) -- (208.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.07, 51.21) -- (205.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.56, 51.21) -- (212.56, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.87, 51.21) -- (173.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.05, 51.21) -- (219.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.92, 51.21) -- (198.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.60, 51.21) -- (185.60, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.88, 51.21) -- (209.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.81, 51.21) -- (212.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.05, 51.21) -- (215.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.89, 51.21) -- (187.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.58, 51.21) -- (172.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.30, 51.21) -- (188.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.65, 51.21) -- (212.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.70, 51.21) -- (187.70, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.63, 51.21) -- (196.63, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.35, 51.21) -- (207.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.19, 51.21) -- (204.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.72, 51.21) -- (189.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.18, 51.21) -- (186.18, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.71, 51.21) -- (193.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (167.99, 51.21) -- (167.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.34, 51.21) -- (176.34, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (167.56, 51.21) -- (167.56, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.99, 51.21) -- (209.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.50, 51.21) -- (203.50, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.46, 51.21) -- (207.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.41, 51.21) -- (208.41, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.56, 51.21) -- (215.56, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.50, 51.21) -- (178.50, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.87, 51.21) -- (211.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.59, 51.21) -- (183.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (220.28, 51.21) -- (220.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.66, 51.21) -- (156.66, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (150.19, 51.21) -- (150.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.26, 51.21) -- (193.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.07, 51.21) -- (184.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.44, 51.21) -- (190.44, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.44, 51.21) -- (211.44, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.78, 51.21) -- (212.78, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.05, 51.21) -- (179.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.70, 51.21) -- (187.70, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (222.53, 51.21) -- (222.53, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.99, 51.21) -- (185.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (128.62, 51.21) -- (128.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.55, 51.21) -- (218.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.71, 51.21) -- (191.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.42, 51.21) -- (201.42, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.06, 51.21) -- (190.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.71, 51.21) -- (195.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.69, 51.21) -- (176.69, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.98, 51.21) -- (182.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (133.55, 51.21) -- (133.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.02, 51.21) -- (202.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.86, 51.21) -- (212.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.05, 51.21) -- (219.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.89, 51.21) -- (195.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (159.12, 51.21) -- (159.12, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.11, 51.21) -- (219.11, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.86, 51.21) -- (194.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.24, 51.21) -- (194.24, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.75, 51.21) -- (201.75, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.05, 51.21) -- (214.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.63, 51.21) -- (216.63, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.05, 51.21) -- (211.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.07, 51.21) -- (184.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.35, 51.21) -- (203.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.33, 51.21) -- (215.33, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.69, 51.21) -- (197.69, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.12, 51.21) -- (208.12, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.92, 51.21) -- (197.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.42, 51.21) -- (215.42, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (145.87, 51.21) -- (145.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.41, 51.21) -- (196.41, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.21, 51.21) -- (186.21, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.83, 51.21) -- (186.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.34, 51.21) -- (182.34, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (224.64, 51.21) -- (224.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.40, 51.21) -- (214.40, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.52, 51.21) -- (217.52, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.20, 51.21) -- (193.20, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (111.85, 51.21) -- (111.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.29, 51.21) -- (218.29, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.22, 51.21) -- (156.22, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (219.68, 51.21) -- (219.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.05, 51.21) -- (200.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.03, 51.21) -- (193.03, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.81, 51.21) -- (190.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.89, 51.21) -- (173.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.49, 51.21) -- (218.49, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.77, 51.21) -- (211.77, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (213.92, 51.21) -- (213.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.75, 51.21) -- (164.75, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.77, 51.21) -- (217.77, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.19, 51.21) -- (200.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (171.10, 51.21) -- (171.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.89, 51.21) -- (206.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.64, 51.21) -- (188.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.60, 51.21) -- (194.60, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.92, 51.21) -- (197.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.38, 51.21) -- (187.38, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.48, 51.21) -- (207.48, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.51, 51.21) -- (216.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.17, 51.21) -- (201.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.15, 51.21) -- (192.15, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.93, 51.21) -- (178.93, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.81, 51.21) -- (210.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.77, 51.21) -- (215.77, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.12, 51.21) -- (212.12, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.48, 51.21) -- (183.48, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.16, 51.21) -- (174.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.74, 51.21) -- (201.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (150.46, 51.21) -- (150.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.79, 51.21) -- (206.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.72, 51.21) -- (170.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.07, 51.21) -- (195.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.75, 51.21) -- (216.75, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.05, 51.21) -- (211.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.80, 51.21) -- (191.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.36, 51.21) -- (183.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.54, 51.21) -- (198.54, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.16, 51.21) -- (164.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.76, 51.21) -- (182.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.31, 51.21) -- (198.31, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.31, 51.21) -- (217.31, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.64, 51.21) -- (202.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.49, 51.21) -- (185.49, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.35, 51.21) -- (181.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.05, 51.21) -- (214.05, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.31, 51.21) -- (199.31, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.68, 51.21) -- (156.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.98, 51.21) -- (184.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.49, 51.21) -- (185.49, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.28, 51.21) -- (175.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.01, 51.21) -- (203.01, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.53, 51.21) -- (193.53, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.86, 51.21) -- (177.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.97, 51.21) -- (198.97, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.51, 51.21) -- (198.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.24, 51.21) -- (202.24, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.06, 51.21) -- (175.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.10, 51.21) -- (189.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (122.72, 51.21) -- (122.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.96, 51.21) -- (196.96, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.61, 51.21) -- (218.61, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.07, 51.21) -- (184.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.17, 51.21) -- (182.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.76, 51.21) -- (196.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.88, 51.21) -- (189.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.91, 51.21) -- (195.91, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.80, 51.21) -- (172.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.07, 51.21) -- (211.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.71, 51.21) -- (203.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.60, 51.21) -- (163.60, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.07, 51.21) -- (195.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.17, 51.21) -- (204.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (159.39, 51.21) -- (159.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (211.57, 51.21) -- (211.57, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.29, 51.21) -- (204.29, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (218.04, 51.21) -- (218.04, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.07, 51.21) -- (184.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.54, 51.21) -- (209.54, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.07, 51.21) -- (184.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.10, 51.21) -- (182.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.84, 51.21) -- (202.84, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.09, 51.21) -- (207.09, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.42, 51.21) -- (212.42, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.89, 51.21) -- (187.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.80, 51.21) -- (184.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.99, 51.21) -- (164.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.68, 51.21) -- (182.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.67, 51.21) -- (185.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.66, 51.21) -- (195.66, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.85, 51.21) -- (199.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.47, 51.21) -- (214.47, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (217.41, 51.21) -- (217.41, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (214.52, 51.21) -- (214.52, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (153.58, 51.21) -- (153.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.00, 51.21) -- (201.00, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.38, 51.21) -- (187.38, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.01, 51.21) -- (190.01, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.76, 51.21) -- (209.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.21, 51.21) -- (186.21, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.22, 51.21) -- (212.22, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.33, 51.21) -- (215.33, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.92, 51.21) -- (206.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (146.26, 51.21) -- (146.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.47, 51.21) -- (200.47, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (215.97, 51.21) -- (215.97, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.77, 51.21) -- (202.77, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.85, 51.21) -- (197.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (216.13, 51.21) -- (216.13, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.81, 51.21) -- (204.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.17, 51.21) -- (205.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.52, 51.21) -- (206.52, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.37, 51.21) -- (201.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.91, 51.21) -- (174.91, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.14, 51.21) -- (209.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.22, 51.21) -- (184.22, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.74, 51.21) -- (166.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.07, 51.21) -- (184.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.35, 51.21) -- (173.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.95, 51.21) -- (196.95, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.88, 51.21) -- (187.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.85, 51.21) -- (198.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.82, 51.21) -- (205.82, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.76, 51.21) -- (198.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.63, 51.21) -- (193.63, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.89, 51.21) -- (200.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (165.17, 51.21) -- (165.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.21, 51.21) -- (207.21, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.88, 51.21) -- (187.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.17, 51.21) -- (175.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.49, 51.21) -- (200.49, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.25, 51.21) -- (202.25, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.53, 51.21) -- (203.53, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.83, 51.21) -- (180.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (165.36, 51.21) -- (165.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.86, 51.21) -- (176.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.87, 51.21) -- (202.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.36, 51.21) -- (177.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (189.73, 51.21) -- (189.73, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.69, 51.21) -- (199.69, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.98, 51.21) -- (197.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.16, 51.21) -- (179.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.04, 51.21) -- (176.04, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.92, 51.21) -- (183.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.88, 51.21) -- (175.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.76, 51.21) -- (160.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.38, 51.21) -- (198.38, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.43, 51.21) -- (192.43, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.87, 51.21) -- (202.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.80, 51.21) -- (200.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.58, 51.21) -- (204.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (171.73, 51.21) -- (171.73, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.89, 51.21) -- (200.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.29, 51.21) -- (179.29, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.76, 51.21) -- (209.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.66, 51.21) -- (156.66, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (150.19, 51.21) -- (150.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.79, 51.21) -- (181.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.16, 51.21) -- (179.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.72, 51.21) -- (199.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.32, 51.21) -- (201.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (167.32, 51.21) -- (167.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.41, 51.21) -- (187.41, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (210.69, 51.21) -- (210.69, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.99, 51.21) -- (185.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (128.62, 51.21) -- (128.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.78, 51.21) -- (206.78, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.68, 51.21) -- (181.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.26, 51.21) -- (194.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.07, 51.21) -- (182.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.58, 51.21) -- (188.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.22, 51.21) -- (174.22, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (133.55, 51.21) -- (133.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.01, 51.21) -- (193.01, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.76, 51.21) -- (201.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.21, 51.21) -- (207.21, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.07, 51.21) -- (184.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.75, 51.21) -- (207.75, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.18, 51.21) -- (186.18, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.50, 51.21) -- (186.50, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.73, 51.21) -- (191.73, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.87, 51.21) -- (204.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.85, 51.21) -- (204.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.35, 51.21) -- (202.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.20, 51.21) -- (192.20, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.64, 51.21) -- (204.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.44, 51.21) -- (188.44, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.46, 51.21) -- (199.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.08, 51.21) -- (186.08, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.28, 51.21) -- (209.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (145.87, 51.21) -- (145.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.56, 51.21) -- (185.56, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.27, 51.21) -- (176.27, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.04, 51.21) -- (176.04, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (212.80, 51.21) -- (212.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.03, 51.21) -- (204.03, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.89, 51.21) -- (205.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.32, 51.21) -- (184.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (111.85, 51.21) -- (111.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.65, 51.21) -- (206.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.22, 51.21) -- (156.22, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (209.47, 51.21) -- (209.47, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.26, 51.21) -- (198.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.79, 51.21) -- (181.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.16, 51.21) -- (179.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.89, 51.21) -- (173.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.65, 51.21) -- (206.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.13, 51.21) -- (204.13, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.68, 51.21) -- (204.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (162.60, 51.21) -- (162.60, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.07, 51.21) -- (206.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.45, 51.21) -- (198.45, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (168.65, 51.21) -- (168.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.92, 51.21) -- (197.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.19, 51.21) -- (184.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.25, 51.21) -- (193.25, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.08, 51.21) -- (186.08, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.21, 51.21) -- (180.21, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.91, 51.21) -- (195.91, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (208.14, 51.21) -- (208.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.46, 51.21) -- (192.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.60, 51.21) -- (187.60, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.28, 51.21) -- (169.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.09, 51.21) -- (199.09, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.32, 51.21) -- (204.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.56, 51.21) -- (200.56, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.63, 51.21) -- (173.63, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (168.19, 51.21) -- (168.19, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.25, 51.21) -- (191.25, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (150.46, 51.21) -- (150.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.26, 51.21) -- (200.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (161.57, 51.21) -- (161.57, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (186.43, 51.21) -- (186.43, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.96, 51.21) -- (205.96, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.21, 51.21) -- (201.21, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.80, 51.21) -- (191.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.94, 51.21) -- (197.94, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (155.40, 51.21) -- (155.40, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.81, 51.21) -- (188.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.47, 51.21) -- (205.47, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.38, 51.21) -- (197.38, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.67, 51.21) -- (184.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (171.46, 51.21) -- (171.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (202.84, 51.21) -- (202.84, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.80, 51.21) -- (190.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.68, 51.21) -- (156.68, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.98, 51.21) -- (184.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.33, 51.21) -- (175.33, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.28, 51.21) -- (175.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (193.93, 51.21) -- (193.93, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.78, 51.21) -- (183.78, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.51, 51.21) -- (174.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.97, 51.21) -- (188.97, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (187.88, 51.21) -- (187.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.00, 51.21) -- (191.00, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (165.64, 51.21) -- (165.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.16, 51.21) -- (179.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (122.72, 51.21) -- (122.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.12, 51.21) -- (194.12, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.89, 51.21) -- (206.89, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.73, 51.21) -- (175.73, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.81, 51.21) -- (192.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.57, 51.21) -- (180.57, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (184.07, 51.21) -- (184.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (167.70, 51.21) -- (167.70, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.61, 51.21) -- (199.61, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (200.61, 51.21) -- (200.61, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (154.45, 51.21) -- (154.45, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.02, 51.21) -- (188.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (198.64, 51.21) -- (198.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (149.25, 51.21) -- (149.25, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.24, 51.21) -- (201.24, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (192.90, 51.21) -- (192.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (207.61, 51.21) -- (207.61, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.92, 51.21) -- (197.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.10, 51.21) -- (182.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.00, 51.21) -- (191.00, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.10, 51.21) -- (197.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.45, 51.21) -- (201.45, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.06, 51.21) -- (176.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.02, 51.21) -- (179.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.02, 51.21) -- (176.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.03, 51.21) -- (179.03, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.55, 51.21) -- (185.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (188.14, 51.21) -- (188.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.71, 51.21) -- (204.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.88, 51.21) -- (205.88, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (203.48, 51.21) -- (203.48, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.31, 51.21) -- (195.31, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.04, 51.21) -- (176.04, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.58, 51.21) -- (185.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.92, 51.21) -- (197.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.86, 51.21) -- (185.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (201.13, 51.21) -- (201.13, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (205.64, 51.21) -- (205.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (196.23, 51.21) -- (196.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (146.26, 51.21) -- (146.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.79, 51.21) -- (191.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (204.73, 51.21) -- (204.73, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (191.93, 51.21) -- (191.93, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (195.27, 51.21) -- (195.27, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (206.65, 51.21) -- (206.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (197.53, 51.21) -- (197.53, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.24, 51.21) -- (194.24, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (194.92, 51.21) -- (194.92, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (190.54, 51.21) -- (190.54, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (168.40, 51.21) -- (168.40, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (199.12, 51.21) -- (199.12, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (157.33, 51.21) -- (157.33, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (139.83, 51.21) -- (139.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.58, 51.21) -- (156.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.46, 51.21) -- (169.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (171.36, 51.21) -- (171.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.26, 51.21) -- (180.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (171.36, 51.21) -- (171.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.14, 51.21) -- (166.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.86, 51.21) -- (173.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (139.83, 51.21) -- (139.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.72, 51.21) -- (179.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (150.27, 51.21) -- (150.27, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.71, 51.21) -- (176.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.35, 51.21) -- (175.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.04, 51.21) -- (176.04, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (154.30, 51.21) -- (154.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (139.83, 51.21) -- (139.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.29, 51.21) -- (177.29, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.62, 51.21) -- (174.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.58, 51.21) -- (156.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (132.90, 51.21) -- (132.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (158.27, 51.21) -- (158.27, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (138.18, 51.21) -- (138.18, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (171.36, 51.21) -- (171.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.14, 51.21) -- (166.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.98, 51.21) -- (177.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.35, 51.21) -- (175.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.36, 51.21) -- (177.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (149.28, 51.21) -- (149.28, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.86, 51.21) -- (173.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (168.93, 51.21) -- (168.93, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.27, 51.21) -- (182.27, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (132.90, 51.21) -- (132.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (139.83, 51.21) -- (139.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (154.30, 51.21) -- (154.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.62, 51.21) -- (174.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (139.83, 51.21) -- (139.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.43, 51.21) -- (166.43, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (183.20, 51.21) -- (183.20, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (165.63, 51.21) -- (165.63, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (121.06, 51.21) -- (121.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.72, 51.21) -- (179.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (154.30, 51.21) -- (154.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.46, 51.21) -- (169.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.53, 51.21) -- (156.53, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.87, 51.21) -- (166.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (121.06, 51.21) -- (121.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.14, 51.21) -- (166.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.62, 51.21) -- (174.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (121.06, 51.21) -- (121.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.72, 51.21) -- (179.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.58, 51.21) -- (156.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (121.06, 51.21) -- (121.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.26, 51.21) -- (180.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (162.02, 51.21) -- (162.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.14, 51.21) -- (166.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.10, 51.21) -- (178.10, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (121.06, 51.21) -- (121.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.36, 51.21) -- (177.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.65, 51.21) -- (176.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.14, 51.21) -- (166.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.36, 51.21) -- (177.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (162.02, 51.21) -- (162.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.17, 51.21) -- (181.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (158.59, 51.21) -- (158.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.79, 51.21) -- (181.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (132.90, 51.21) -- (132.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (158.59, 51.21) -- (158.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.58, 51.21) -- (151.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (185.31, 51.21) -- (185.31, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.75, 51.21) -- (177.75, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.58, 51.21) -- (178.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (157.76, 51.21) -- (157.76, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (111.85, 51.21) -- (111.85, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.16, 51.21) -- (179.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (143.80, 51.21) -- (143.80, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (182.27, 51.21) -- (182.27, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.62, 51.21) -- (174.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (154.30, 51.21) -- (154.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (160.39, 51.21) -- (160.39, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.16, 51.21) -- (179.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.72, 51.21) -- (179.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.66, 51.21) -- (177.66, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (141.86, 51.21) -- (141.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.58, 51.21) -- (178.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.35, 51.21) -- (175.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (146.47, 51.21) -- (146.47, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.43, 51.21) -- (170.43, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (158.59, 51.21) -- (158.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.17, 51.21) -- (170.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (158.59, 51.21) -- (158.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (158.13, 51.21) -- (158.13, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (168.42, 51.21) -- (168.42, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (181.30, 51.21) -- (181.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.84, 51.21) -- (166.84, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (147.81, 51.21) -- (147.81, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.36, 51.21) -- (177.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.07, 51.21) -- (173.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (143.79, 51.21) -- (143.79, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (165.98, 51.21) -- (165.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (132.90, 51.21) -- (132.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.17, 51.21) -- (176.17, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (138.12, 51.21) -- (138.12, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (162.02, 51.21) -- (162.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.58, 51.21) -- (178.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.62, 51.21) -- (174.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.46, 51.21) -- (169.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.62, 51.21) -- (174.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (139.52, 51.21) -- (139.52, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.07, 51.21) -- (164.07, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.98, 51.21) -- (177.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.43, 51.21) -- (170.43, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.35, 51.21) -- (175.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.57, 51.21) -- (164.57, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (141.65, 51.21) -- (141.65, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.46, 51.21) -- (169.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.58, 51.21) -- (156.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (162.98, 51.21) -- (162.98, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (162.02, 51.21) -- (162.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (139.83, 51.21) -- (139.83, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (121.06, 51.21) -- (121.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.71, 51.21) -- (170.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (132.90, 51.21) -- (132.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.72, 51.21) -- (179.72, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.69, 51.21) -- (173.69, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (154.30, 51.21) -- (154.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (156.58, 51.21) -- (156.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.23, 51.21) -- (172.23, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.71, 51.21) -- (176.71, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (132.90, 51.21) -- (132.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (167.32, 51.21) -- (167.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.86, 51.21) -- (173.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (124.29, 51.21) -- (124.29, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.45, 51.21) -- (174.45, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.14, 51.21) -- (166.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (180.26, 51.21) -- (180.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.43, 51.21) -- (170.43, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.74, 51.21) -- (144.74, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (162.02, 51.21) -- (162.02, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.82, 51.21) -- (170.82, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (175.35, 51.21) -- (175.35, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.64, 51.21) -- (148.64, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (151.67, 51.21) -- (151.67, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (132.90, 51.21) -- (132.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (152.29, 51.21) -- (152.29, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (154.30, 51.21) -- (154.30, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (158.59, 51.21) -- (158.59, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (161.62, 51.21) -- (161.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.90, 51.21) -- (177.90, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.58, 51.21) -- (178.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (176.04, 51.21) -- (176.04, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (121.06, 51.21) -- (121.06, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (171.61, 51.21) -- (171.61, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] ( 31.37, 51.21) -- ( 31.37, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (148.55, 51.21) -- (148.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (161.42, 51.21) -- (161.42, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (170.43, 51.21) -- (170.43, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (174.62, 51.21) -- (174.62, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (178.58, 51.21) -- (178.58, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (169.46, 51.21) -- (169.46, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (139.26, 51.21) -- (139.26, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (166.14, 51.21) -- (166.14, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (177.36, 51.21) -- (177.36, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (164.87, 51.21) -- (164.87, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (172.55, 51.21) -- (172.55, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (179.16, 51.21) -- (179.16, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.99, 51.21) -- (173.99, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (167.32, 51.21) -- (167.32, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (167.86, 51.21) -- (167.86, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (163.51, 51.21) -- (163.51, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (144.66, 51.21) -- (144.66, 54.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.1pt,line join=round,line cap=round] (173.86, 51.21) -- (173.86, 54.71);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37, 51.21) rectangle (283.58,193.24);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 54.91) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 80.74) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,106.56) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,132.38) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,158.21) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,184.03) {0.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 57.67) --
	( 31.37, 57.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 83.49) --
	( 31.37, 83.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,109.32) --
	( 31.37,109.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,135.14) --
	( 31.37,135.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,160.96) --
	( 31.37,160.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,186.79) --
	( 31.37,186.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 69.75, 48.46) --
	( 69.75, 51.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.99, 48.46) --
	(120.99, 51.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.23, 48.46) --
	(172.23, 51.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.48, 48.46) --
	(223.48, 51.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (274.72, 48.46) --
	(274.72, 51.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 69.75, 40.75) {0.002};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (120.99, 40.75) {0.050};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (172.23, 40.75) {1.000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (223.48, 40.75) {20.086};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (274.72, 40.75) {403.429};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (157.47, 30.25) {Cumulative exposure (mg/m$^3\cdot\,$years)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,122.23) {Density};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 71.86,  5.50) rectangle (243.09, 26.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 82.86, 11.00) rectangle ( 92.86, 21.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line cap=rect] ( 83.57, 11.71) rectangle ( 92.15, 20.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (119.85, 11.00) rectangle (129.85, 21.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 7pt off 3pt ,line cap=rect] (120.57, 11.71) rectangle (129.14, 20.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (156.85, 11.00) rectangle (166.85, 21.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect] (157.56, 11.71) rectangle (166.14, 20.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (189.35, 11.00) rectangle (199.35, 21.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line cap=rect] (190.06, 11.71) rectangle (198.64, 20.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 98.36, 12.90) {0.05};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (135.35, 12.90) {0.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (172.35, 12.90) {0.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,198.74);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (204.85, 12.90) {No limit};
\end{scope}
\end{tikzpicture}

\end{document}