### 2022-09-15 reports/paper/05-appendix.rmd

---
output:
  pdf_document:
    keep_tex: true
header-includes:
- \usepackage{longtable}
- \usepackage{setspace}\onehalfspacing
- \usepackage{verbatim}
- \usepackage{framed}
- \usepackage{pgf}
- \usepackage{pgfplots}
- \pgfplotsset{compat=1.17}
- \usepackage{tikz}
- \usetikzlibrary{graphs, arrows, automata, positioning, decorations}
- \tikzset{invisible/.style={opacity=0}, visible on/.style={alt={#1{}{invisible}}}, alt/.code args={<#1>#2#3}{\alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}},}
- \usepackage{textgreek}
- \usepackage{framed}
- \usepackage{xcolor}
- \renewenvironment{quote}{\begin{minipage}{\linewidth}\colorlet{shadecolor}{orange!15}\begin{snugshade}}{\end{snugshade}\end{minipage}}
- \usepackage{booktabs}
- \usepackage{adjustbox}
- \newcommand{\ind}{\perp\mkern-10mu\perp}
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/asa.csl')`"
---

```{=tex}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\arraystretch}{1}
```

```{r, include=FALSE}
knitr::opts_chunk$set(
	echo = F, results = 'asis',
	warning = F,
	tab.cap.tnd = 1,
	ft.arraystretch = 1.15)
if (!'tidyr' %in% loadedNamespaces()) {library(tidyverse)}
library(flextable)
```

# Appendix

![Oberved and post-intervention distributions of average annual exposure (mg/m^3^). The rugplot shows maximum cumulative exposure accrued by non-Hodgkin lymphoma cases.](`r here::here("resources/images", "shift-avg-soluble.png")`){width=6in}

![Oberved and post-intervention distributions of cumulative exposure to soluble metalworking fluids (mg/m$^3\cdot$years) at end of follow-up. The rugplot shows maximum cumulative exposure accrued by non-Hodgkin lymphoma cases.](`r here::here("resources/images", "shift-cum-soluble.png")`){width=6in}

```{r, tab.cap.pre="Table A", include=F}
str.flextab <- readRDS("str.flextab.rds")
str.flextab
```

```{r, include=F}
syn.flextab <- readRDS("syn.flextab.rds")
syn.flextab
```