### 2022-09-15 reports/slides/slides.Rmd

---
title: "Evaluating hypothetical limits on metalworking fluid exposure for reducing non-Hodgkin lymphoma incidence"
subtitle: "An application of the hazard-extended parametric g-formula"
# author: "Kevin Chen"
date: \today
# classoption: aspectratio=169
fontsize: 11pt
output:
  beamer_presentation:
    toc: false
    # latex_engine: lualatex
    includes:
      in_header: "~/HeadRs/BeamerHead.sty"
    keep_tex: true
header-includes:
- \setlength{\itemindent}{0em}
- \setlength{\leftmargini}{1em}
- \setlength{\leftmarginii}{1em}
- \setlength{\leftmarginiii}{1em}
- \usepackage{bbold}
# - \setlist*[itemize]{labelindent=\parindent,itemindent=0pt,leftmargin=*}
# - \usepackage{pgfpages}
# - \setbeamertemplate{note page}[plain]
# - \setbeameroption{show notes on second screen=right}
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
---

```{r, include=F}
library(tidyverse)
library(xtable)
```

# Non-Hodgkin Lymphoma

- 7th most common cancer in the US
- Incidence rose dramatically 1960 to 1990 around the world
	- In the US:  8.5 to 18.3 per 100,000 between 1973 and 2019
- Strongest risk factors [@Ekstrom-Smedby_2006]
	- Immunosuppression, both genetic and acquired
	- Infection by _H. pylori_, Epstein-Barr virus, Hepatitis C virus, HIV
- Known risk factors account for about 50% of observed incidence [@Hartge_1992]

\note{
In order of frequency (2020):
1. breast
2. lung/bronchus
3. prostate
4. colorectal
5. melanoma
6. bladder
7. NHL
8. kidney \& renal pelvis,
9. endometrial
10. leukemia
}

# NHL and peseticides

- Rise in NHL mirrors rise in chemicalization of agriculture, industry, and warfare [@Nelson_2005]
	- Agricultural exposure to pesticides very well studied
	- Meta-analyses yielded consistent associations with pesticides:\
	carbamate, organophosphorous, triazine, organochlorine [@Schinasi_2014]

\begin{center}
\includegraphics[width=2in]{./resources/pesticide_aircraft.jpg}
\end{center}

\note{
Several organophosphorous pesticides considered possibly carcinogenic by IARC
Several chlorinated solvents are IARC Group 1 carcinogens
}

# Soluble MWF

- Water-based soluble MWFs contain diverse additives:\
	- Triazines: biocides
	- Nitrites: rust inhibitors
	- Sulfonates: emulsifiers
	- Organochlorines: extreme pressure performance enhancers
- In UAW-GM, soluble MWF was used more frequently and in much larger quantities than straight and synthetic MWF
- Globally, soluble MWF is the most common

# MWF exposure in UAW-GM

![Median average annual exposure among exposed workers over time.](`r here::here("resources/images", "exposure.png")`){width=4in}

- 89% ever exposed to soluble
- 57% ever exposed to straight
- 35% ever exposed to synthetic

# MWF exposure and NHL in UAW-GM

![NHL and soluble MWF exposure in recent cancer incidence analysis (Hilary's paper, in press).](`r here::here("../gm-cancer-inc/reports/resources/Lag 21/subplots", "AJE-01232-2021 Colbeth Figure 2N.pdf")`){width=4in}

# Iterated conditional expectation (ICE) g-formula

ICE g-formula estimators rely on the tower property of expectation (from law of total probability)

- For example, $\mathbb E\, \mathbb E \left[Y \mid A = \check{a}, W \right] = \mathbb E \left[Y \mid A = \check{a}\right]$
- For discrete $W$, we have\
$\mathbb E \left[Y \mid A = \check{a}\right] = \sum_{w \in \mathcal W} \mathbb E\, \left[Y \mid A = \check{a}, W = w\right] \Prob{W = w}$\
This is the non-parametric $g$-computation result, which we can estimate parametrically by specifying a model:
\begin{center}$\mathbb E\left[Y \mid A = \check{a}, W = w\right] = \beta_0 + \beta_1 \check{a} + \beta_2 w$\end{center}

	- Avoids the need for propensity score
	- Relies heavily on correct model specification
	- Equivalent to model-based standardization

# ICE g-formula for multiple timepoints

\footnotesize\renewcommand{\arraystretch}{1.2}
\begin{tabular}{ll}
Put & $\psi =  \E[f_{Y_J}]{Y_J(1 - Y_{J - 1}) \mid \bar L_{J}, \bar A_{J} = \bar A^g_{J}, \bar Y_{J - 1} = C_J = 0}$ \\
Put & $\psi =  \E[f_{L_{J}}]{\,\psi\, \mid \bar L_{J - 1}, \bar A_{J - 1} = \bar A^g_{J - 1}, \bar Y_{J - 1} = C_{J} = 0}$ \\
Put & $\psi =  \E[f_{Y_{J - 1}}]{Y_{J - 1} (1 - Y_{J - 2}) + \,\psi\, \mid \bar L_{J - 1}, \bar A_{J - 1} = \bar A^g_{J - 1}, \bar Y_{J - 2} = C_{J - 1} = 0}$ \\
& $\hspace{4pt} \vdots$ \\
Put & $\psi =  \E[f_{Y_1}]{Y_1 + \,\psi\, \mid L_1, A_1 = A^g_1, \bar Y_{0} = C_0 = 0}$ \\
\\
$\leadsto$ &  $\E{Y^g_J} =  \E[f_{L_{1}}]{\,\psi\,}$
\end{tabular}

For _hazard-extended_ ICE g-formula,\
we use discrete hazard estimate for interval $(j - 1, j]$ in place of $Y_j$

# Intervention of interest

- Static stochastic interventions: limit soluble MWF to\
	- 0.5 mg/m^3^ (NIOSH REL)
	- 0.25 mg/m^3^
	- 0.05 mg/m^3^
- Dynamic stochastic interventions: reduce soluble MWF such that total MWF exposure limited to 0.5, 0.25, and 0.05 mg/m^3^
	- Translates to the following exposure limits on soluble MWF:
		- $\max\left(0,\ 0.5 - \text{str} - \text{syn}\right)$
		- $\max\left(0,\ 0.25 - \text{str} - \text{syn}\right)$
		- $\max\left(0,\ 0.05 - \text{str} - \text{syn}\right)$

# Illustration of interventions when applied

![Observed and post-intervention levels of exposure in three example person-years.](`r here::here("resources", "interventions.pdf")`){width=4.25in}

# Results

\scriptsize\setlength{\tabcolsep}{3pt}
```{r, echo=F, results="asis", warning=F, message=F}
sol.tab <- readRDS(here::here('resources', 'sol.tab.rds'))
names(sol.tab)[grep("%", names(sol.tab))] <- gsub('%', '\\\\%', names(sol.tab)[grep("%", names(sol.tab))])
names(sol.tab)[1] <- "Exposure limit on soluble MWF (mg/m$^3$)"
sol.tab %>% xtable(align = 'cp{0.27\\linewidth}C{0.18\\linewidth}C{0.1\\linewidth}ccc') %>%
	print(comment = F, include.rownames = F,
				sanitize.text.function = function(x) {x})
```

# Discussion

- Assumption of correct model specification is doing heavy duty
	- Trade-off between satisfying positivity and correct model specification
	- But approach is much more flexible than Cox PH
- Why bother with stochastic intervention?
	- In reality, companies may not reduce exposures already below REL
	- In estimation, we can achieve positivity more easily
- Why bother with complex _dynamic_ stochastic intervention
	- NIOSH REL is for total MWF exposure?
	- Dynamic interventions can achieve risk reductions by intervening on fewer person-years (analytically) and processes (in reality)

# Citations

\singlespacing\footnotesize