### 2022-09-15 resources/interventions.tex

% Created by tikzDevice version 0.12.3.1 on 2022-06-28 15:57:56
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\end{document}
re}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (469.75,216.81);
\begin{scope}
\path[clip] ( 25.15, 47.51) rectangle (167.85,198.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 25.15, 47.51) rectangle (167.85,198.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] ( 31.63, 54.37) rectangle ( 68.70,164.08);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] ( 77.97, 54.37) rectangle (115.03,122.94);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (124.30, 54.37) rectangle (161.36, 68.09);
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] ( 31.63,164.08) rectangle ( 68.70,191.51);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] ( 77.97,122.94) rectangle (115.03,191.51);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (124.30, 68.09) rectangle (161.36, 81.80);
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (  3.83, 54.37) rectangle (  3.83, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (  3.83, 54.37) rectangle (  3.83, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (  3.83, 54.37) rectangle (  3.83, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 25.15, 88.66) -- (167.85, 88.66);
\end{scope}
\begin{scope}
\path[clip] (173.35, 47.51) rectangle (316.05,198.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (173.35, 47.51) rectangle (316.05,198.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (179.84, 54.37) rectangle (216.90, 88.66);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (226.17, 54.37) rectangle (263.23, 88.66);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (272.50, 54.37) rectangle (309.57, 68.09);
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (179.84, 88.66) rectangle (216.90,116.08);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (226.17, 88.66) rectangle (263.23,157.22);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (272.50, 68.09) rectangle (309.57, 81.80);
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (152.04, 54.37) rectangle (152.04, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (152.04, 54.37) rectangle (152.04, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (152.04, 54.37) rectangle (152.04, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (173.35, 88.66) -- (316.05, 88.66);
\end{scope}
\begin{scope}
\path[clip] (321.55, 47.51) rectangle (464.26,198.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (321.55, 47.51) rectangle (464.25,198.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (328.04, 54.37) rectangle (365.10, 61.23);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (374.37, 54.37) rectangle (411.44, 54.37);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (420.70, 54.37) rectangle (457.77, 68.09);
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (328.04, 61.23) rectangle (365.10, 88.66);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (374.37, 54.37) rectangle (411.44,122.94);

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (420.70, 68.09) rectangle (457.77, 81.80);
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (300.24, 54.37) rectangle (300.24, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (300.24, 54.37) rectangle (300.24, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] (300.24, 54.37) rectangle (300.24, 54.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (321.55, 88.66) -- (464.26, 88.66);
\end{scope}
\begin{scope}
\path[clip] ( 25.15,198.36) rectangle (167.85,211.31);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round,fill=fillColor] ( 25.15,198.36) rectangle (167.85,211.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 27.65,201.74) {A) No intervention};
\end{scope}
\begin{scope}
\path[clip] (173.35,198.36) rectangle (316.05,211.31);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round,fill=fillColor] (173.35,198.36) rectangle (316.05,211.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (175.85,201.74) {B) Static intervention};
\end{scope}
\begin{scope}
\path[clip] (321.55,198.36) rectangle (464.26,211.31);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round,fill=fillColor] (321.55,198.36) rectangle (464.25,211.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (324.05,201.74) {C) Dynamic intervention};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 50.17, 37.06) {p-yr 1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 96.50, 37.06) {p-yr 2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (142.83, 37.06) {p-yr 3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (198.37, 37.06) {p-yr 1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (244.70, 37.06) {p-yr 2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (291.03, 37.06) {p-yr 3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (346.57, 37.06) {p-yr 1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (392.90, 37.06) {p-yr 2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (439.24, 37.06) {p-yr 3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15, 47.51) --
	( 25.15,198.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20, 51.62) {0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40, 54.37) --
	( 25.15, 54.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40, 88.66) --
	( 25.15, 88.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,122.94) --
	( 25.15,122.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,157.22) --
	( 25.15,157.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,191.51) --
	( 25.15,191.51);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,122.94) {Average annual exposure (mg/m$^3$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.28,  5.50) rectangle (189.07, 24.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 60.28, 11.00) rectangle ( 68.28, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line cap=rect,fill=fillColor] ( 60.99, 11.71) rectangle ( 67.56, 18.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 73.78, 11.90) {Hypothetical exposure limit};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (200.07,  5.50) rectangle (440.13, 24.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (211.07, 11.00) rectangle (219.07, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (211.79, 11.71) rectangle (218.36, 18.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (211.79, 11.71) rectangle (218.36, 18.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (334.17, 11.00) rectangle (342.17, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (334.88, 11.71) rectangle (341.46, 18.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor] (334.88, 11.71) rectangle (341.46, 18.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (224.57, 11.90) {Exposure to soluble MWF};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (469.75,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (347.67, 11.90) {Other MWF  exposure};
\end{scope}
\end{tikzpicture}

\end{document}