### 2022-09-30 reports/ACIC (2022)/poster/resources/rr.tex
Clean up and prepare for submission

% Created by tikzDevice version 0.12.3.1 on 2022-09-30 16:23:45
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\end{document}