### 2022-09-30 reports/paper/paper.Rmd

---
# fontsize: 11pt
# mainfont: arial
output:
  # pdf_document:
  #   fig_caption: true
  #   latex_engine: lualatex
  #   includes:
  #     in_header: "~/HeadRs/Stathead.sty"
  #   keep_tex: true
  bookdown::word_document2:
    number_sections: false
    global_numbering: true
    reference_docx: "~/HeadRs/rmd_templates/style-guide-arial-11.docx"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
notes-after-punctuation: true
---

\fancyhead[R]{}
\fancyhead[L]{}
\renewcommand{\headrulewidth}{0pt}
\singlespacing

\thispagestyle{empty}

\setlength{\columnseprule}{0pt}

```{r, include=FALSE}
knitr::opts_chunk$set(echo=F, results = 'asis')
library(bibtex)
biblio <- bibtex::read.bib('~/Box Sync/papers/EnvEpi.bib')
```

**Type of manuscript:** Original Research Article

**Manuscript title:** Evaluating hypothetical limits on metalworking fluid exposure for reducing non-Hodgkin lymphoma incidence: An application of the hazard-extended parametric g-formula

**Authors:** Kevin T. Chen^a,b^, Ellen A. Eisen^b^, Sally Picciotto^b^

^a^ University of California, Berkeley School of Public Health, Division of Epidemiology \& Biostatistics

^b^ University of California, Berkeley School of Public Health, Division of Environmental Health Sciences

**Corresponding author:**\
Kevin T. Chen\
2121 Berkeley Way\
Room 5302, Desk 5305-1\
Berkeley, CA 94720-7360\
kevchen@berkeley.edu\
(510) 387-3327

**Running head suggestion:** Limits on metalworking fluid for preventing NHL.

**Conflicts of interests:** Authors declare no conflicts of interest.

**Sources of financial support:** This work was supported by Training Grant T42OH008429 and Research Project Grant R01OH011092, both funded by the National Institute for Occupational Safety and Health (NIOSH) / Centers for Disease Control and Prevention (CDC).

**Data and computing code:** Code for reproducing analyses and reports are available on [Github](https://github.com/kvntchn/gm-nhl-ice). Data are available upon reasonable request.

**Acknowledgements:**

\newpage

# Abstract

(Limit: 250 words)

**Background:** Non-Hodgkin lymphoma (NHL) incidence has increased significantly since 1960 and was recently linked with workplace exposure in the United Auto Workers-General Motors (UAW-GM) Cohort Study. Here, we investigate hypothetical interventions on workplace exposure to oil- and water-based soluble metalworking fluids (MWF) in relation to NHL risk. Interventions on entire classes of occupational exposures rather than particular agents may result in stronger lasting protections for worker health.

**Methods:** We estimated counterfactual risk of NHL between 1985 and 2004 in the UAW-GM cohort of autoworkers exposed to straight, soluble, and synthetic MWF at three Michigan plants under hypothetical interventions. We estimated the effect of static and dynamic stochastic interventions on average annual exposure to soluble MWF. Unlike the static interventions, dynamic interventions take into account exposures to the other two classes of metalworking fluids, reflecting the fact that currently recommended limits apply to the total of all three exposures.

**Results:** During follow-up, 231 NHL cases occurred. Stronger hypothetical limits on average annual exposure to soluble MWF resulted in monotonically stronger reductions in NHL risk. Capping exposure to soluble MWF at 0.05 mg/m^3^ resulted in a risk ratio of 0.79 (95% CI: 0.61, 0.96). Capping total MWF exposure at 0.05 mg/m^3^ by intervening only on exposure to soluble MWF resulted in a risk ratio of 0.72 (95% CI: 0.44, 0.97).

**Conclusion:** We considered interventions on soluble MWF that address total MWF exposure levels and showed that stronger limits on average annual exposure to soluble MWF may confer stronger protection against NHL risk.

# What this study adds

(Limit: 100 words)

Little is known about possible links between occupational exposures and non-Hodgkin lymphoma. Existing studies are generally single time-point case-control studies with limited exposure assessment. We leveraged the large study population size and quantitative exposure assessment of the United Auto Workers-General Motors Cohort Study to estimate counterfactual risk of non-Hodgkin lymphoma under realistic interventions using a novel causal estimator capable of adjusting for time-varying confounding affected by past exposure.

\newpage

```{r child = '01-introduction.rmd'}
```

```{r child = '02-methods.rmd'}
```

```{r child = '03-results.rmd'}
```

```{r child = '04-discussion.rmd'}
```

\newpage

# Citations

<div id="refs"></div>