### 2023-03-02 reports/paper/02-methods.rmd

---
output:
  word_document:
    reference_docx: "~/HeadRs/rmd_templates/style-guide-arial-11.docx"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/asa.csl')`"
---

```{r, include=FALSE, eval=F}
knitr::opts_chunk$set(echo=F, results = 'asis')
library(bibtex)
biblio <- bibtex::read.bib('~/Box Sync/papers/EnvEpi.bib')
```

# Methods

We estimated NHL risk from 1985 to 2005 under hypothetical limits on average annual exposure to soluble MWF by applying the hazard-extended iterative conditional expectation (ICE) parametric g-formula estimator [@Wen_2020]. We leveraged time-varying quantitative MWF exposure data in tandem with employment records to adjust for the HWSE. First, we estimated the expected number of NHL cases that we would observe if there were no censoring by competing risks. Then, we contrasted this counterfactual risk to that when in addition, the hypothetical exposure limits of (1) the NIOSH Recommended Exposure Limit (REL) for total particulate mass (PM) from MWFs (0.5 mg/m^3^), (2) half the REL (0.25 mg/m^3^), and (3) a tenth of the REL (0.05 mg/m^3^) were enforced for soluble MWF over workers' entire working lifetimes [@niosh_1998]. Under these hypothetical interventions, exposure was set at the hypothetical limit if observed exposure was greater than that limit.  Otherwise, exposure not intervened upon. This kind of intervention is known as a threshold longitudinal modified treatment policy [@Taubman_2009;@Richardson_2013;@Haneuse_2013;@Young_2014]. Longitudinal modified treatment policies are intervention strategies that depend on the natural value of exposure at a particular time ie the value of exposure that would have been observed at that time if the intervention were discontinued immediately prior [@Richardson_2013;@Diaz_2021].

## Study population

The UAW-GM cohort includes all hourly workers at three automobile manufacturing plants in Michigan who had worked at least 3 years by 1985. Past papers provide detailed descriptions of the cohort [@Eisen_1992;@Eisen_2001]. The present study population (N = 34,738) was restricted to the autoworkers who were at work in 1941 or not yet hired, missing no more than half of their employment history, and still alive at the start of follow-up. Autoworkers in the study population were followed for NHL incidence from January 1, 1985 until NHL diagnosis, death, January 1, 2005 (10 years after the end of employment record availability), or age 108 years (the oldest observed age at death), whichever came earlier.

## Outcome and covariates

We identified incident cancers in the UAW-GM Cohort that occurred between 1985 and 2005 by linkage to the Michigan Cancer Registry (MCR). Workers at Plants 1 and 2, located in the greater Detroit metropolitan area, were also linked to the Detroit Regional Registry of the Surveillance, Epidemiology, and End Results (SEER) Program. Cancer types were distinguished using site and histology codes conforming to the International classification of Diseases for Oncology, 3^rd^ edition (ICD-O-3). Non-Hodgkin lymphoma was defined by cancers with any of the following ICD-O-3 Histology codes: 9590-9597, 9670-9671, 9673, 9675, 9678-9680, 9684, 9687-9691, 9695, 9698-9702, 9705, 9708-9709, 9712, 9714-9719, 9724-9729, 9735, 9737-9738, 9811-9818, 9823, 9827, 9837. Subtypes of NHL were also identified for sensitivity analyses. Details regarding cancer incidence follow-up are described elsewhere [@Colbeth_2022]. Vital status was ascertained from company records and by linkage to Social Security Administration, National Death Index, and state mortality files.

Covariates including year of hire, sex, race, and plant location were obtained from company records. Race was missing for about 16% of the cohort, most commonly among workers hired before 1960 in Plant 2. In analyses, missing race was considered a distinct category. All covariates were coded as categorical variables. Cut-points for continuous covariates were determined according to the quantiles among cases.

## Exposure

Quantitative measurement of time-varying MWF exposure is a distinguishing strength of the UAW-GM cohort relative to other studies of occupational exposure. Exposure assessment was based on direct sampling as well as company records. Company industrial hygienists collected several hundred personal and area samples for total particulate matter (mg/m^3^) composed of MWF over many decades. Research industrial hygienists collected additional air sampling data when the cohort study was launched in the mid 1980s. These data were combined with the historical data and company records to derive quantitative 8-hour time-weighted average exposure estimates to soluble, straight, and synthetic MWFs for each combination of job, department, and plant over time. Workers' time-weighted average annual exposure to each MWF type was determined by combining this job-exposure matrix with employment records, which recorded time-varying job type, department, and plant through 1994. For employment records that were at least half complete, gaps in the record were interpolated by carrying forward the last known job type. The exposure assessment is described in detail elsewhere [@Hallock_1994;@Woskie_1994;@Woskie_2003]. Previous analyses of NHL applied exposure lags of 1 to 20 years; we lagged cumulative MWF exposures by 10 years to account for disease latency and therefore ended follow-up on January 1, 2005 [@Smith_2007;@Karipidis_2007;@Zhang_2019]. In analyses, MWF exposure history was summarized as the cumulative sum of annual exposure intensities and coded as categorical variables with cut-points determined according the quantiles of cumulative exposure among cases.

## Statistical methods

The hazard-extended ICE parametric g-formula may be thought of as a series of model-based standardization steps, which we implemented using logistic regression. We split the 20-year follow-up period into eight time periods; the first two periods spanned four years each, and the remaining six periods spanned two years each. The first two periods are longer in length to account for the smaller number of cases in those years. Post-intervention estimates of the discrete hazard of NHL given all exposures and covariates were combined iteratively from the end of follow-up to the start. In each iteration, predicted discrete hazards were standardized over post-intervention exposure and covariate histories before combining with discrete hazards from the previous iteration. This iterative process results in sequentially standardized estimates of NHL risk over the entire follow-up period. Averaging over the baseline distribution of covariates yields the counterfactual risk estimate of NHL when the intervention of interest was enforced for the entire study population.

Post-intervention exposure and exposure history were summarized as cumulative exposure. We modeled discrete hazards by fitting a pooled logistic regression for NHL over at-risk person-periods given cumulative exposure to straight, soluble, and synthetic MWFs, employment status, cumulative time off, year of hire, sex (male/female), race (Black/white/unknown), and plant (Plant 1/Plant 2/Plant 3). Cumulative exposure to MWFs, employment status, and cumulative time off were lagged 10 years. All continuous variables were represented as categorical variables with cut points determined by the tertiles of nonzero values among NHL cases. An overview of the general steps of the estimation procedure are presented below.

0.  For the hypothetical exposure limit of interest, compute the cumulative exposure
    each worker would have accrued by the end of each follow-up period since hire.
1.  Fit a pooled logistic regression to the observed data for NHL on covariates and
    cumulative exposure over all at-risk person-periods, excluding those ending with a
    censoring event.
2.  Replace the cumulative exposure vector in the observed data with the
    post-intervention cumulative exposure vector. Using the model fitted in the 
    previous step, compute estimates of the post-intervention discrete hazard for
    each at-risk person-period including those that end with a censoring event.
3.  Set $k = K - 1$ where $K = 8$ is the number of follow-up periods:

	a) Among those who were event-free and uncensored through the $k$^th^ period,
		 fit a logistic regression on the predicted post-intervention discrete hazard
		 for the period spanning the $(k + 1)$ through $K$^th^ periods given observed
		 covariate and exposure history up through (and including) the $k$^th^ period.
	b) Replace the cumulative exposure vector in the observed data with the
		 post-intervention cumulative exposure vector. Using the model fitted in the
		 previous step, obtain predicted values for all those who were at-risk in the
		 $k$^th^ period.
	c) Compute the predicted post-intervention discrete hazard for the period spanning
		 the $k$ through $K$^th^ periods by multiplying the predicted values from the
		 previous step by 1 minus the discrete hazard estimate used to fit the model
		 in step 3.a and adding that same discrete hazard estimate to the product.
	d) If $k > 1$, set $k = k - 1$ and return to step 3.a.

4.  Compute the counterfactual risk by averaging the predicted post-intervention
		discrete hazards for the entire follow-up period for all units.

We estimated risk under the observed distribution of soluble MWF exposure (natural course) and under the six interventions. We contrasted the risk under intervention to that under the natural course by computing relative risks. Confidence intervals were computed using the nonparametric bootstrap with 1000 Monte Carlo samples from the population at risk at start of follow-up and centering on the estimate computed from observed data. All the necessary script used to reproduce the analyses are available on [GitHub](https://github.com/kvntchn/gm-nhl-ice).