### 2023-03-02 reports/paper/paper.Rmd

---
# fontsize: 11pt
# mainfont: arial
output:
  # pdf_document:
  #   fig_caption: true
  #   latex_engine: lualatex
  #   includes:
  #     in_header: "~/HeadRs/Stathead.sty"
  #   keep_tex: true
  bookdown::word_document2:
    number_sections: false
    global_numbering: true
    reference_docx: "~/HeadRs/rmd_templates/style-guide-arial-11.docx"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
notes-after-punctuation: true
---

\fancyhead[R]{}
\fancyhead[L]{}
\renewcommand{\headrulewidth}{0pt}
\singlespacing

\thispagestyle{empty}

\setlength{\columnseprule}{0pt}

```{r, include=FALSE}
knitr::opts_chunk$set(echo=F, results = 'asis')
library(bibtex)
biblio <- bibtex::read.bib('~/Box Sync/papers/EnvEpi.bib')
```

**Type of manuscript:** Original Research Article

**Manuscript title:** Hypothetical limits on workplace exposure for reducing non-Hodgkin lymphoma: An application of the hazard-extended iterative conditional expectation parametric g-formula

**Authors:** Kevin T. Chen^a,b,c^, Ellen A. Eisen^b^, Sally Picciotto^b^

^a^ University of California, Berkeley School of Public Health, Division of Epidemiology \& Biostatistics

^b^ University of California, Berkeley School of Public Health, Division of Environmental Health Sciences

^c^ University of California, Berkeley Department of Statistics

**Corresponding author:**\
Kevin T. Chen\
2121 Berkeley Way\
Room 5302, Desk 5305-1\
Berkeley, CA 94720-7360\
kevchen@berkeley.edu\
(510) 387-3327

**Running head suggestion:** Limits on metalworking fluid for preventing NHL.

**Conflicts of interests:** Authors declare no conflicts of interest.

**Sources of financial support:** This work was supported by Training Grant T42OH008429 and Research Project Grant R01OH011092, both funded by the National Institute for Occupational Safety and Health (NIOSH) / Centers for Disease Control and Prevention (CDC).

**Data and computing code:** Code for reproducing analyses and reports are available on [Github](https://github.com/kvntchn/gm-nhl-ice). Data are available upon reasonable request.

**Keywords:** Metalworking fluid, occupational exposure, non-Hodgkin lymphoma, cancer, causal inference, g-formula, healthy-worker effect, United States

\newpage

# Abstract

(Limit: 250 words)

**Background:** Non-Hodgkin lymphoma (NHL) incidence has increased significantly since 1960 and was recently linked to workplace exposure to soluble metalworking fluids (MWF) in a standard survival analysis of the United Auto Workers-General Motors (UAW-GM) cohort. Here, we investigate the causal effects of hypothetical limits on soluble MWF exposure in relation to NHL risk using an iterative conditional expectation (ICE) parametric g-formula estimator.

**Methods:** We estimated counterfactual risk of NHL between 1985 and 2005 in the UAW-GM cohort of autoworkers at three Michigan plants (n = 34,734) under hypothetical limits on average annual exposure to soluble MWF. We applied the hazard-extended ICE parametric g-formula, an estimator capable of adjusting for time-varying confounding affected by past exposure.

**Results:** During follow-up, 231 NHL cases occurred. Stronger hypothetical limits on soluble MWF exposure resulted in monotonic reductions in NHL risk estimates. Capping average annual exposure at 0.5 mg/m^3^, the NIOSH recommended exposure limit, would have prevented 44 (95% CI: -6, 91) cases. Capping at 0.25 mg/m^3^ would have prevented 52 (95% CI: -5, 106) cases. Finally, capping at 0.05 mg/m^3^ would have prevented 71 (95% CI: 12, 129) cases.

**Conclusion:** Stronger limits on soluble MWF exposure would have prevented more cases of NHL in the UAW-GM cohort. Our application demonstrates the utility of an ICE g-formula estimator for estimating the effect of realistic exposure interventions on a survival outcome. Greater availability of worked examples and software would facilitate wider application of ICE g-formula estimators.

\newpage

```{r child = '01-introduction.rmd'}
```

```{r child = '02-methods.rmd'}
```

```{r child = '03-results.rmd'}
```

```{r child = '04-discussion.rmd'}
```

\newpage

# Citations

<div id="refs"></div>