### 2023-03-02 resources/interventions.tex

% Created by tikzDevice version 0.12.3.1 on 2023-02-02 14:54:25
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage{upgreek}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}[Extension = .ttf,
			UprightFont = *,
			BoldFont = *bd,
			talicFont = *i,
			BoldItalicFont = *bi]

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\end{document}
ill opacity=0.00] (0,0) rectangle (361.35,216.81);
\begin{scope}
\path[clip] ( 25.15, 45.51) rectangle (187.75,198.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 25.15, 45.51) rectangle (187.75,198.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] ( 32.54, 52.46) rectangle ( 74.77,163.62);

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] ( 85.33, 52.46) rectangle (127.56,121.94);

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (138.12, 52.46) rectangle (180.36, 66.36);
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] ( 32.54,163.62) rectangle ( 74.77,191.41);

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] ( 85.33,121.94) rectangle (127.56,191.41);

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (138.12, 66.36) rectangle (180.36, 80.25);
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,fill=fillColor] (  0.86, 52.46) rectangle (  0.86, 52.46);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,fill=fillColor] (  0.86, 52.46) rectangle (  0.86, 52.46);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,fill=fillColor] (  0.86, 52.46) rectangle (  0.86, 52.46);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 25.15, 87.20) -- (187.75, 87.20);
\end{scope}
\begin{scope}
\path[clip] (193.25, 45.51) rectangle (355.85,198.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (193.25, 45.51) rectangle (355.85,198.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (200.64, 52.46) rectangle (242.87, 87.20);

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (253.43, 52.46) rectangle (295.67, 87.20);

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (306.22, 52.46) rectangle (348.46, 66.36);
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (200.64, 87.20) rectangle (242.87,114.99);

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (253.43, 87.20) rectangle (295.67,156.68);

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (306.22, 66.36) rectangle (348.46, 80.25);
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,fill=fillColor] (168.96, 52.46) rectangle (168.96, 52.46);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,fill=fillColor] (168.96, 52.46) rectangle (168.96, 52.46);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,fill=fillColor] (168.96, 52.46) rectangle (168.96, 52.46);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (193.25, 87.20) -- (355.85, 87.20);
\end{scope}
\begin{scope}
\path[clip] ( 25.15,198.36) rectangle (187.75,211.31);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round,fill=fillColor] ( 25.15,198.36) rectangle (187.75,211.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 27.65,201.74) {A) No intervention};
\end{scope}
\begin{scope}
\path[clip] (193.25,198.36) rectangle (355.85,211.31);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round,fill=fillColor] (193.25,198.36) rectangle (355.85,211.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (195.75,201.74) {B) Intervention};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 53.66, 35.06) {p-yr 1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (106.45, 35.06) {p-yr 2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (159.24, 35.06) {p-yr 3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (221.76, 35.06) {p-yr 1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (274.55, 35.06) {p-yr 2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (327.34, 35.06) {p-yr 3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 25.15, 45.51) --
	( 25.15,198.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 20.20, 49.71) {0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40, 52.46) --
	( 25.15, 52.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40, 87.20) --
	( 25.15, 87.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,121.94) --
	( 25.15,121.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,156.68) --
	( 25.15,156.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 22.40,191.41) --
	( 25.15,191.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,121.94) {Average annual exposure (mg/m$^3$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 75.97, 19.50) rectangle (305.02, 22.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 81.47, 19.50) rectangle ( 89.47, 27.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] ( 82.18, 20.21) rectangle ( 88.76, 26.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] ( 82.18, 20.21) rectangle ( 88.76, 26.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (204.57, 19.50) rectangle (212.57, 27.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (205.28, 20.21) rectangle (211.86, 26.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 0.6pt,fill=fillColor] (205.28, 20.21) rectangle (211.86, 26.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 94.97, 20.40) {Exposure to soluble MWF};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (218.07, 20.40) {Other MWF  exposure};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (126.10,  5.50) rectangle (254.90,  8.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (131.60,  5.50) rectangle (139.60, 13.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,fill=fillColor] (132.31,  6.21) rectangle (138.89, 12.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (145.10,  6.40) {Hypothetical exposure limit};
\end{scope}
\end{tikzpicture}

\end{document}