### 2023-04-08 resources/exposure.tex

% Created by tikzDevice version 0.12.3.1 on 2023-04-06 10:31:22
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage{upgreek}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,180.67);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,180.68);
\end{scope}
\begin{scope}
\path[clip] ( 35.37, 28.21) rectangle (162.62,162.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37, 28.21) rectangle (162.62,162.23);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 44.76) --
	(162.62, 44.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 69.57) --
	(162.62, 69.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 94.37) --
	(162.62, 94.37);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,119.17) --
	(162.62,119.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,143.98) --
	(162.62,143.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.54, 28.21) --
	( 63.54,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (100.86, 28.21) --
	(100.86,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (138.18, 28.21) --
	(138.18,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 32.36) --
	(162.62, 32.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 57.16) --
	(162.62, 57.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 81.97) --
	(162.62, 81.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,106.77) --
	(162.62,106.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,131.58) --
	(162.62,131.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,156.38) --
	(162.62,156.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.88, 28.21) --
	( 44.88,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 82.20, 28.21) --
	( 82.20,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (119.52, 28.21) --
	(119.52,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.83, 28.21) --
	(156.83,162.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.75,139.99) --
	( 48.61,123.33) --
	( 50.48,130.79) --
	( 52.35,140.45) --
	( 54.21,128.18) --
	( 56.08,130.82) --
	( 57.94,143.76) --
	( 59.81,143.17) --
	( 61.68,156.14) --
	( 63.54,142.11) --
	( 65.41,141.53) --
	( 67.27,118.13) --
	( 69.14,127.52) --
	( 71.00,133.98) --
	( 72.87,137.54) --
	( 74.74,132.30) --
	( 76.60,138.56) --
	( 78.47,120.48) --
	( 80.33,141.46) --
	( 82.20,145.32) --
	( 84.07,141.53) --
	( 85.93,149.94) --
	( 87.80,150.14) --
	( 89.66,146.20) --
	( 91.53,141.53) --
	( 93.39,141.53) --
	( 95.26,141.53) --
	( 97.13,141.53) --
	( 98.99,142.55) --
	(100.86, 71.73) --
	(102.72, 72.60) --
	(104.59, 69.29) --
	(106.46, 69.92) --
	(108.32, 62.24) --
	(110.19, 63.56) --
	(112.05, 66.32) --
	(113.92, 68.18) --
	(115.78, 66.95) --
	(117.65, 68.18) --
	(119.52, 57.01) --
	(121.38, 61.53) --
	(123.25, 55.24) --
	(125.11, 55.83) --
	(126.98, 65.33) --
	(128.85, 61.53) --
	(130.71, 57.67) --
	(132.58, 56.97) --
	(134.44, 54.48) --
	(136.31, 54.14) --
	(138.18, 54.08) --
	(140.04, 53.17) --
	(141.91, 52.60) --
	(143.77, 50.78) --
	(145.64, 44.72);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37, 28.21) rectangle (162.62,162.23);
\end{scope}
\begin{scope}
\path[clip] (168.12, 28.21) rectangle (295.37,162.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (168.12, 28.21) rectangle (295.37,162.23);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (168.12, 44.76) --
	(295.37, 44.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] (168.12, 69.57) --
	(295.37, 69.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] (168.12, 94.37) --
	(295.37, 94.37);

\path[draw=drawColor,line width= 0.3pt,line join=round] (168.12,119.17) --
	(295.37,119.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] (168.12,143.98) --
	(295.37,143.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] (196.29, 28.21) --
	(196.29,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (233.61, 28.21) --
	(233.61,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (270.93, 28.21) --
	(270.93,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.12, 32.36) --
	(295.37, 32.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.12, 57.16) --
	(295.37, 57.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.12, 81.97) --
	(295.37, 81.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.12,106.77) --
	(295.37,106.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.12,131.58) --
	(295.37,131.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.12,156.38) --
	(295.37,156.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (177.63, 28.21) --
	(177.63,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.95, 28.21) --
	(214.95,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (252.27, 28.21) --
	(252.27,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (289.58, 28.21) --
	(289.58,162.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (179.50, 64.49) --
	(181.37, 67.76) --
	(183.23, 54.42) --
	(185.10, 58.48) --
	(186.96, 50.99) --
	(188.83, 54.60) --
	(190.69, 56.67) --
	(192.56, 51.09) --
	(194.43, 53.62) --
	(196.29, 45.62) --
	(198.16, 43.01) --
	(200.02, 43.01) --
	(201.89, 43.37) --
	(203.76, 43.37) --
	(205.62, 43.37) --
	(207.49, 43.37) --
	(209.35, 43.37) --
	(211.22, 43.37) --
	(213.08, 43.37) --
	(214.95, 44.40) --
	(216.82, 43.37) --
	(218.68, 46.38) --
	(220.55, 45.62) --
	(222.41, 48.79) --
	(224.28, 54.13) --
	(226.15, 55.09) --
	(228.01, 55.09) --
	(229.88, 55.09) --
	(231.74, 55.09) --
	(233.61, 45.48) --
	(235.48, 46.84) --
	(237.34, 48.19) --
	(239.21, 45.83) --
	(241.07, 41.35) --
	(242.94, 41.35) --
	(244.80, 39.86) --
	(246.67, 39.08) --
	(248.54, 39.08) --
	(250.40, 39.08) --
	(252.27, 39.08) --
	(254.13, 39.08) --
	(256.00, 38.03) --
	(257.87, 38.83) --
	(259.73, 39.08) --
	(261.60, 39.08) --
	(263.46, 40.81) --
	(265.33, 42.05) --
	(267.19, 42.01) --
	(269.06, 39.65) --
	(270.93, 41.22) --
	(272.79, 41.91) --
	(274.66, 41.29) --
	(276.52, 41.67) --
	(278.39, 39.78);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (168.12, 28.21) rectangle (295.37,162.23);
\end{scope}
\begin{scope}
\path[clip] (300.87, 28.21) rectangle (428.12,162.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (300.87, 28.21) rectangle (428.12,162.23);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (300.87, 44.76) --
	(428.12, 44.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] (300.87, 69.57) --
	(428.12, 69.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] (300.87, 94.37) --
	(428.12, 94.37);

\path[draw=drawColor,line width= 0.3pt,line join=round] (300.87,119.17) --
	(428.12,119.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] (300.87,143.98) --
	(428.12,143.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] (329.04, 28.21) --
	(329.04,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (366.36, 28.21) --
	(366.36,162.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (403.68, 28.21) --
	(403.68,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (300.87, 32.36) --
	(428.12, 32.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (300.87, 57.16) --
	(428.12, 57.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (300.87, 81.97) --
	(428.12, 81.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (300.87,106.77) --
	(428.12,106.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (300.87,131.58) --
	(428.12,131.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (300.87,156.38) --
	(428.12,156.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.38, 28.21) --
	(310.38,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.70, 28.21) --
	(347.70,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (385.02, 28.21) --
	(385.02,162.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (422.34, 28.21) --
	(422.34,162.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.25, 72.54) --
	(314.12,127.07) --
	(315.98,127.07) --
	(317.85,113.74) --
	(319.71, 88.56) --
	(321.58,100.41) --
	(323.45, 76.72) --
	(325.31, 54.98) --
	(327.18, 53.98) --
	(329.04, 48.46) --
	(330.91, 70.36) --
	(332.77, 71.10) --
	(334.64, 44.28) --
	(336.51, 44.28) --
	(338.37, 45.88) --
	(340.24, 44.35) --
	(342.10, 48.14) --
	(343.97, 44.28) --
	(345.84, 44.28) --
	(347.70, 44.97) --
	(349.57, 44.28) --
	(351.43, 44.28) --
	(353.30, 44.40) --
	(355.17, 45.60) --
	(357.03, 47.66) --
	(358.90, 45.21) --
	(360.76, 48.23) --
	(362.63, 47.55) --
	(364.49, 50.24) --
	(366.36, 35.96) --
	(368.23, 36.25) --
	(370.09, 36.25) --
	(371.96, 35.89) --
	(373.82, 34.30) --
	(375.69, 34.30) --
	(377.56, 34.62) --
	(379.42, 34.66) --
	(381.29, 34.45) --
	(383.15, 35.26) --
	(385.02, 35.37) --
	(386.88, 35.70) --
	(388.75, 35.18) --
	(390.62, 35.67) --
	(392.48, 36.25) --
	(394.35, 36.25) --
	(396.21, 38.05) --
	(398.08, 38.07) --
	(399.95, 38.05) --
	(401.81, 37.68) --
	(403.68, 38.10) --
	(405.54, 37.67) --
	(407.41, 36.10) --
	(409.27, 39.85) --
	(411.14, 37.05);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.87, 28.21) rectangle (428.12,162.23);
\end{scope}
\begin{scope}
\path[clip] ( 35.37,162.23) rectangle (162.62,175.17);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 35.37,162.23) rectangle (162.62,175.17);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 98.99,165.60) {Soluble};
\end{scope}
\begin{scope}
\path[clip] (168.12,162.23) rectangle (295.37,175.17);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (168.12,162.23) rectangle (295.37,175.17);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (231.74,165.60) {Straight};
\end{scope}
\begin{scope}
\path[clip] (300.87,162.23) rectangle (428.12,175.17);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (300.87,162.23) rectangle (428.12,175.17);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (364.49,165.60) {Synthetic};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.88, 25.46) --
	( 44.88, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 82.20, 25.46) --
	( 82.20, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (119.52, 25.46) --
	(119.52, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.83, 25.46) --
	(156.83, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 44.88, 17.75) {1940};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 82.20, 17.75) {1960};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (119.52, 17.75) {1980};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (156.83, 17.75) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (177.63, 25.46) --
	(177.63, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.95, 25.46) --
	(214.95, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (252.27, 25.46) --
	(252.27, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (289.58, 25.46) --
	(289.58, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (177.63, 17.75) {1940};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (214.95, 17.75) {1960};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (252.27, 17.75) {1980};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (289.58, 17.75) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.38, 25.46) --
	(310.38, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.70, 25.46) --
	(347.70, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (385.02, 25.46) --
	(385.02, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (422.34, 25.46) --
	(422.34, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (310.38, 17.75) {1940};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (347.70, 17.75) {1960};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (385.02, 17.75) {1980};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (422.34, 17.75) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 29.61) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 54.41) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 79.21) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,104.02) {0.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,128.82) {1.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,153.63) {1.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 32.36) --
	( 35.37, 32.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 57.16) --
	( 35.37, 57.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 81.97) --
	( 35.37, 81.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,106.77) --
	( 35.37,106.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,131.58) --
	( 35.37,131.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,156.38) --
	( 35.37,156.38);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (231.74,  7.25) {Calendar year};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70, 95.22) {Median exposure (mg/m$^3$) among exposed};
\end{scope}
\end{tikzpicture}

\end{document}