### 2023-04-08 resources/exposure_long.tex

% Created by tikzDevice version 0.12.3.1 on 2023-04-06 10:31:57
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage{upgreek}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (216.81,397.48);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (216.81,397.48);
\end{scope}
\begin{scope}
\path[clip] ( 35.37,274.39) rectangle (211.31,379.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37,274.39) rectangle (211.31,379.04);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,287.32) --
	(211.31,287.32);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,306.68) --
	(211.31,306.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,326.05) --
	(211.31,326.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,345.42) --
	(211.31,345.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,364.79) --
	(211.31,364.79);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 66.99,274.39) --
	( 66.99,379.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (103.34,274.39) --
	(103.34,379.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.70,274.39) --
	(139.70,379.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (176.05,274.39) --
	(176.05,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,277.63) --
	(211.31,277.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,297.00) --
	(211.31,297.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,316.37) --
	(211.31,316.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,335.74) --
	(211.31,335.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,355.10) --
	(211.31,355.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,374.47) --
	(211.31,374.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.82,274.39) --
	( 48.82,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.17,274.39) --
	( 85.17,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.52,274.39) --
	(121.52,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.87,274.39) --
	(157.87,379.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (194.22,274.39) --
	(194.22,379.04);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 86.99,361.67) --
	( 88.80,348.66) --
	( 90.62,354.49) --
	( 92.44,362.03) --
	( 94.26,352.45) --
	( 96.07,354.51) --
	( 97.89,364.62) --
	( 99.71,364.16) --
	(101.53,374.28) --
	(103.34,363.33) --
	(105.16,362.87) --
	(106.98,344.61) --
	(108.80,351.93) --
	(110.62,356.98) --
	(112.43,359.76) --
	(114.25,355.67) --
	(116.07,360.55) --
	(117.89,346.44) --
	(119.70,362.82) --
	(121.52,365.84) --
	(123.34,362.87) --
	(125.16,369.44) --
	(126.97,369.60) --
	(128.79,366.53) --
	(130.61,362.87) --
	(132.43,362.87) --
	(134.24,362.87) --
	(136.06,362.87) --
	(137.88,363.68) --
	(139.70,308.37) --
	(141.51,309.05) --
	(143.33,306.47) --
	(145.15,306.96) --
	(146.97,300.96) --
	(148.78,301.99) --
	(150.60,304.15) --
	(152.42,305.60) --
	(154.24,304.64) --
	(156.06,305.60) --
	(157.87,296.88) --
	(159.69,300.41) --
	(161.51,295.50) --
	(163.33,295.96) --
	(165.14,303.38) --
	(166.96,300.41) --
	(168.78,297.39) --
	(170.60,296.85) --
	(172.41,294.91) --
	(174.23,294.64) --
	(176.05,294.59) --
	(177.87,293.88) --
	(179.68,293.44) --
	(181.50,292.01) --
	(183.32,287.29);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37,274.39) rectangle (211.31,379.04);
\end{scope}
\begin{scope}
\path[clip] ( 35.37,151.30) rectangle (211.31,255.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37,151.30) rectangle (211.31,255.95);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,164.23) --
	(211.31,164.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,183.59) --
	(211.31,183.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,202.96) --
	(211.31,202.96);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,222.33) --
	(211.31,222.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,241.70) --
	(211.31,241.70);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 66.99,151.30) --
	( 66.99,255.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (103.34,151.30) --
	(103.34,255.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.70,151.30) --
	(139.70,255.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (176.05,151.30) --
	(176.05,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,154.54) --
	(211.31,154.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,173.91) --
	(211.31,173.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,193.28) --
	(211.31,193.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,212.65) --
	(211.31,212.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,232.01) --
	(211.31,232.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,251.38) --
	(211.31,251.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.82,151.30) --
	( 48.82,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.17,151.30) --
	( 85.17,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.52,151.30) --
	(121.52,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.87,151.30) --
	(157.87,255.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (194.22,151.30) --
	(194.22,255.95);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 86.99,179.63) --
	( 88.80,182.18) --
	( 90.62,171.77) --
	( 92.44,174.94) --
	( 94.26,169.09) --
	( 96.07,171.91) --
	( 97.89,173.52) --
	( 99.71,169.17) --
	(101.53,171.14) --
	(103.34,164.90) --
	(105.16,162.86) --
	(106.98,162.86) --
	(108.80,163.14) --
	(110.62,163.14) --
	(112.43,163.14) --
	(114.25,163.14) --
	(116.07,163.14) --
	(117.89,163.14) --
	(119.70,163.14) --
	(121.52,163.95) --
	(123.34,163.14) --
	(125.16,165.49) --
	(126.97,164.90) --
	(128.79,167.37) --
	(130.61,171.54) --
	(132.43,172.29) --
	(134.24,172.29) --
	(136.06,172.29) --
	(137.88,172.29) --
	(139.70,164.79) --
	(141.51,165.85) --
	(143.33,166.90) --
	(145.15,165.06) --
	(146.97,161.56) --
	(148.78,161.56) --
	(150.60,160.39) --
	(152.42,159.79) --
	(154.24,159.79) --
	(156.06,159.79) --
	(157.87,159.79) --
	(159.69,159.79) --
	(161.51,158.97) --
	(163.33,159.59) --
	(165.14,159.79) --
	(166.96,159.79) --
	(168.78,161.14) --
	(170.60,162.11) --
	(172.41,162.08) --
	(174.23,160.23) --
	(176.05,161.46) --
	(177.87,162.00) --
	(179.68,161.52) --
	(181.50,161.81) --
	(183.32,160.34);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37,151.30) rectangle (211.31,255.95);
\end{scope}
\begin{scope}
\path[clip] ( 35.37, 28.21) rectangle (211.31,132.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37, 28.21) rectangle (211.31,132.86);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 41.14) --
	(211.31, 41.14);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 60.50) --
	(211.31, 60.50);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 79.87) --
	(211.31, 79.87);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 99.24) --
	(211.31, 99.24);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,118.61) --
	(211.31,118.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 66.99, 28.21) --
	( 66.99,132.86);

\path[draw=drawColor,line width= 0.3pt,line join=round] (103.34, 28.21) --
	(103.34,132.86);

\path[draw=drawColor,line width= 0.3pt,line join=round] (139.70, 28.21) --
	(139.70,132.86);

\path[draw=drawColor,line width= 0.3pt,line join=round] (176.05, 28.21) --
	(176.05,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 31.45) --
	(211.31, 31.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 50.82) --
	(211.31, 50.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 70.19) --
	(211.31, 70.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 89.56) --
	(211.31, 89.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,108.92) --
	(211.31,108.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,128.29) --
	(211.31,128.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.82, 28.21) --
	( 48.82,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.17, 28.21) --
	( 85.17,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.52, 28.21) --
	(121.52,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.87, 28.21) --
	(157.87,132.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (194.22, 28.21) --
	(194.22,132.86);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 86.99, 62.82) --
	( 88.80,105.41) --
	( 90.62,105.41) --
	( 92.44, 95.00) --
	( 94.26, 75.34) --
	( 96.07, 84.59) --
	( 97.89, 66.09) --
	( 99.71, 49.12) --
	(101.53, 48.33) --
	(103.34, 44.02) --
	(105.16, 61.12) --
	(106.98, 61.70) --
	(108.80, 40.76) --
	(110.62, 40.76) --
	(112.43, 42.01) --
	(114.25, 40.81) --
	(116.07, 43.77) --
	(117.89, 40.76) --
	(119.70, 40.76) --
	(121.52, 41.30) --
	(123.34, 40.76) --
	(125.16, 40.76) --
	(126.97, 40.85) --
	(128.79, 41.79) --
	(130.61, 43.40) --
	(132.43, 41.49) --
	(134.24, 43.85) --
	(136.06, 43.31) --
	(137.88, 45.41) --
	(139.70, 34.26) --
	(141.51, 34.49) --
	(143.33, 34.49) --
	(145.15, 34.20) --
	(146.97, 32.97) --
	(148.78, 32.97) --
	(150.60, 33.21) --
	(152.42, 33.25) --
	(154.24, 33.08) --
	(156.06, 33.71) --
	(157.87, 33.80) --
	(159.69, 34.06) --
	(161.51, 33.66) --
	(163.33, 34.04) --
	(165.14, 34.49) --
	(166.96, 34.49) --
	(168.78, 35.90) --
	(170.60, 35.91) --
	(172.41, 35.90) --
	(174.23, 35.61) --
	(176.05, 35.93) --
	(177.87, 35.60) --
	(179.68, 34.37) --
	(181.50, 37.30) --
	(183.32, 35.12);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37, 28.21) rectangle (211.31,132.86);
\end{scope}
\begin{scope}
\path[clip] ( 35.37,132.86) rectangle (211.31,145.80);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 35.37,132.86) rectangle (211.31,145.80);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (123.34,136.23) {Synthetic};
\end{scope}
\begin{scope}
\path[clip] ( 35.37,255.95) rectangle (211.31,268.89);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 35.37,255.95) rectangle (211.31,268.89);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (123.34,259.32) {Straight};
\end{scope}
\begin{scope}
\path[clip] ( 35.37,379.04) rectangle (211.31,391.99);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 35.37,379.04) rectangle (211.31,391.98);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (123.34,382.41) {Soluble};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.82, 25.46) --
	( 48.82, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.17, 25.46) --
	( 85.17, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.52, 25.46) --
	(121.52, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.87, 25.46) --
	(157.87, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (194.22, 25.46) --
	(194.22, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 48.82, 17.75) {1920};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 85.17, 17.75) {1940};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (121.52, 17.75) {1960};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (157.87, 17.75) {1980};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (194.22, 17.75) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,274.88) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,294.25) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,313.61) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,332.98) {0.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,352.35) {1.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,371.72) {1.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,277.63) --
	( 35.37,277.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,297.00) --
	( 35.37,297.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,316.37) --
	( 35.37,316.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,335.74) --
	( 35.37,335.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,355.10) --
	( 35.37,355.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,374.47) --
	( 35.37,374.47);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,151.79) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,171.16) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,190.52) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,209.89) {0.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,229.26) {1.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,248.63) {1.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,154.54) --
	( 35.37,154.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,173.91) --
	( 35.37,173.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,193.28) --
	( 35.37,193.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,212.65) --
	( 35.37,212.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,232.01) --
	( 35.37,232.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,251.38) --
	( 35.37,251.38);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 28.70) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 48.06) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 67.43) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 86.80) {0.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,106.17) {1.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,125.54) {1.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 31.45) --
	( 35.37, 31.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 50.82) --
	( 35.37, 50.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 70.19) --
	( 35.37, 70.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 89.56) --
	( 35.37, 89.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,108.92) --
	( 35.37,108.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,128.29) --
	( 35.37,128.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (123.34,  7.25) {Calendar year};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,397.48);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,203.62) {Median exposure (mg/m$^3$) among exposed};
\end{scope}
\end{tikzpicture}

\end{document}