### 2023-06-12 reports/SER (2023)/nhl-ice-poster.tex
SER postr

% !TeX program = lualatex
%----------------------------------------------------------------------------------------
%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[20pt,final]{beamer}

\usepackage[scale=1]{beamerposter} % Use the beamerposter package for laying out the poster
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{titlelike}         {bg=yaleblue,fg=white}
\setbeamercolor{frametitle}        {bg=yaleblue!10,fg=yaleblue}
\setbeamercolor{block title}{fg=yaleblue,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
% \setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
% \setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
\setbeamercolor{block alerted title}{fg=white,bg=calyellow}
\setbeamercolor{block alerted body}{fg=black,bg=white}

\setbeamercolor{enumerate item}{fg=dblue!70}
\setbeamercolor{itemize item}{fg=dblue!70}
\setbeamercolor{itemize subitem}{fg=dblue!70}

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{42in} % A0 width: 46.8in
\setlength{\paperheight}{33.5in} % A0 height: 33.1in
\setlength{\sepwid}{0.02\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.3\paperwidth} % Width of one column
\setlength{\twocolwid}{2\onecolwid} % Width of two columns
\addtolength{\twocolwid}{\sepwid}
\setlength{\threecolwid}{3\onecolwid} % Width of three columns
\addtolength{\threecolwid}{2\sepwid}
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
\setlength\labelsep{\dimexpr\labelsep + 0.5em\relax}
\setlength{\parskip}{1em}
\setlength\columnsep{\sepwid}

\setbeamertemplate{itemize/enumerate body begin}{\normalsize}
\setbeamertemplate{itemize/enumerate subbody begin}{\normalsize}
\setbeamertemplate{itemize/enumerate subsubbody begin}{\normalsize}

% declare `cmex` to be arbitrary scalable
\DeclareFontShape{OMX}{cmex}{m}{n}{
  <-7.5> cmex7
  <7.5-8.5> cmex8
  <8.5-9.5> cmex9
  <9.5-> cmex10
}{}

\SetSymbolFont{largesymbols}{normal}{OMX}{cmex}{m}{n}
\SetSymbolFont{largesymbols}{bold}  {OMX}{cmex}{m}{n}

%-----------------------------------------------------------
% \renewcommand{\figurename}{Fig.}

\usepackage{multicol}
\setbeamerfont{caption}{size=\normalsize}
\usepackage{caption}
\captionsetup{labelformat=simple, labelsep=colon}

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

\usepackage{bbm} % more BB
\input{\string~/HeadRs/common_supplement.tex}

\usepackage[sorting=none]{biblatex}
\renewcommand*{\bibfont}{\footnotesize}
\addbibresource{ser_poster.bib}

% ---------------------------------------------------------------------------
\setbeamertemplate{headline}{
 \leavevmode
  \begin{columns}
   \begin{column}{\linewidth} % line
    \vskip1cm
    \centering
    \usebeamercolor{title in headline}{\color{jblue}
        \Huge % CHANGE this line for the size of the title
        {\textbf{\inserttitle}}\\[0.5ex]}
    \usebeamercolor{author in headline}{\color{fg}
        \Large % CHANGE this line for the size of the author
        {\insertauthor}\\[1ex]}
    \usebeamercolor{institute in headline}{\color{fg}
        \large % CHANGE this line for the size of the institut
        {\insertinstitute}\\[1ex]}
    \vskip1cm
   \end{column}
   \vspace{1cm}
  \end{columns}
 \vspace{0.5in} % CHANGE this line for the space between your title and the horizontal rule
 \begin{beamercolorbox}[wd=47in,colsep=0.1cm]{cboxb}\end{beamercolorbox}
 \vspace{0.1in} % CHANGE this line for the space between your horizontal rule and your main body
}

%----------------------------------------------------------------------------------------
%  TITLE SECTION
%----------------------------------------------------------------------------------------
\title{\huge Risk of non-Hodgkin lymphoma under hypothetical interventions with guaranteed positivity} % Poster title
\author{Kevin T. Chen\textsuperscript{\small 1}, Sally Picciotto\textsuperscript{\small 2}, Ellen A. Eisen\textsuperscript{\small 2}} % Author(s)
\institute{University of California, Berkeley -- School of Public Health} % Institution(s)
%----------------------------------------------------------------------------------------

\begin{document}

% add logo
\addtobeamertemplate{headline}{}
{\begin{tikzpicture}[remember picture, overlay]
    \node [anchor=north west, inner sep=1.5cm]  at (current page.north west)
    {\includegraphics[height=9cm]{ucberkeleyseal_139_540.eps}};
  \end{tikzpicture}}
\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

  \vspace{0.01\paperheight}

  \begin{columns}[t,totalwidth=\onecolwid]

    \begin{column}{\sepwid}\end{column} % Empty spacer column

    \begin{column}{\onecolwid} % The first column
      \vbox to 0.78\paperheight{

        %----------------------------------------------------------------------------------------
        %  OBJECTIVES
        %----------------------------------------------------------------------------------------

        \begin{block}{Backgound}

          \begin{itemize}
            \item \textbf{Non-Hodgkin Lymphoma} (NHL) incidence was associated with
                  exposure to
                  \textbf{soluble metalworking fluid} (MWF) in a
                  Cox analysis of the
                  \textbf{United Auto Workers-General Motors} (UAW-GM) \textbf{cohort}
                  \nocite{Colbeth_2022}
            \item Unlike traditional regression analysis, \emph{causal inference} methods
                  \begin{itemize}
                    \item Can adjust for time-varying confounding affected by past exposure \nocite{Arrighi_1994}
                    \item Estimate population effects of hypothetical interventions
                  \end{itemize}
            \item Causal inference in statistics requires \textbf{positivity}, which is
                  \textbf{not always assessed or addressed} \nocite{Petersen_2012}
            \item Here, we specified \textbf{\emph{supportable} interventions} on soluble MWF exposure
                  in the UAW-GM cohort that \textbf{guarantee positivity}
            \item We estimated the effect of supportable interventions on
                  NHL risk in the UAW-GM cohort (1985-2015) using the
                  \textbf{hazard-extended iterative conditional expectation (ICE) parametric g-formula}
                  \nocite{Wen_2020,Young_2014}
            \item Unlike the classic parametric g-formula, ICE g-formula estimators do not require
                  parametric specification of the full joint distribution of the confounders,
                  exposure, and outcome
          \end{itemize}

        \end{block}

        \begin{block}{Target and supportable exposure limits}

          \begin{itemize}
            \item Suppose we want to know the effect of capping exposures at a \textbf{target exposure limit},
                  but estimation may not be supported by observed data
            \item Instead, we define \textbf{supportable exposure limits}
                  for every time $k$ and unique combination of confounder and exposure histories
                  $(\bar l_{k}, \bar a_{k - 1})$
                  \begin{itemize}
                    \item Limit at greatest observed exposure $\le$ \textbf{target limit}, if exists
                    \item No limit, if all observed exposures $>$ \textbf{target limit}
                  \end{itemize}
            \item Propensity scores for exposure at the supportable exposure limits
                  are guaranteed to be \textbf{strictly positive}
                  % \item Supportable exposure limits may be equal to, less than, or greater than the target exposure limits
            \item \textbf{Supportable intervention rule} for every
                  $(\bar l_{k}, \bar a_{k - 1})$, reduces exposures $a_k$ above the
                  \textbf{supportable exposure limit} to that limit, but allows exposures
                  below to vary naturally (Figures 1 and 2)
            \item Applying the supportable intervention rule to the observed data
                  induces the intervention distribution, which defines the
                  \textbf{stochastic dynamic intervention} with \textbf{guaranteed positvity}
          \end{itemize}

          \vspace{0.1\baselineskip}
          \begin{figure}
            \caption{Marginal distribution of nonzero exposure at $k = 2$
              before and after applying the supportable intervention rule}
            \begin{tikzpicture}[x=2.5pt,y=2.5pt]
              \input{resources/intervention-density.tex}
            \end{tikzpicture}
          \end{figure}

        \end{block}

      }\end{column} % End of the first column

    \begin{column}{\sepwid}\end{column} % Empty spacer column

    \begin{column}{\onecolwid} % Begin a column which is two columns wide (column 2)
      \vbox to 0.78\paperheight{

        \vspace{-0.5\baselineskip}
        \begin{figure}
          \caption{Distribution of nonzero exposure for three distinct confounder and
          exposure histories at $k = 2$ before and after applying the supportable intervention rule}
          \begin{tikzpicture}[x=2.5pt,y=2.5pt]
            \input{resources/intervention-density-witnin-combos.tex}
          \end{tikzpicture}
        \end{figure}

        \vspace{0.25\baselineskip}
        \begin{table}
          \caption{Demographic characteristics of the full cohort and the NHL cases}
          \small
          \begin{tabular}{p{0.3\linewidth}rlcrl}
            \toprule
                                & \multicolumn{2}{c}{Study population\phantom{0000}} &
                                & \multicolumn{2}{c}{NHL cases\phantom{00000}}                                                 \\
            \midrule
            $N$ (person-years)  & 33,134                                             & (794,733)     &  & 339  & (5,809)       \\
            Race                                                                                                               \\
            \hspace{1em}White   & 21,315                                             & (64\%)        &  & 250  & (74\%)        \\
            \hspace{1em}Black   & 6,250                                              & (19\%)        &  & 40   & (12\%)        \\
            \hspace{1em}Unknown & 5,569                                              & (17\%)        &  & 49   & (14\%)        \\
            Sex                                                                                                                \\
            \hspace{1em}Male    & 30,249                                             & (87\%)        &  & 206  & (89\%)        \\
            \hspace{1em}Female  & 4,499                                              & (13\%)        &  & 25   & (11\%)        \\
            % Plant                                                                                                                 \\
            % \hspace{1em}Plant 1    & 7,273                                              & (22\%)        &  & 70   & (21\%)        \\
            % \hspace{1em}Plant 2    & 14,251                                             & (43\%)        &  & 137  & (40\%)        \\
            % \hspace{1em}Plant 3    & 11,610                                             & (35\%)        &  & 132  & (39\%)        \\
            % Ever exposed to MWFs \\
            % \hspace{1em}Soluble & 29,010  & (88\%)    & 299  & (88\%) \\
            % \hspace{1em}Straight & 18,710  & (56\%)    & 197  & (58\%) \\
            % \hspace{1em}Synthetic & 11,824  & (36\%)  & 111  & (33\%) \\
            % Deceased by 2015         & 14,434 & (44\%)        &   & 53   & (16\%)            \\
            % % \hline
            % % Years of follow-up & 20 & (16.46, 20) &  & 13 & (6.83, 17) \\
            % Year of birth            & 1941   & (1927, 1950)  &   & 1935 & (1926, 1945)      \\
            % Year of hire             & 1967   & (1953, 1976)  &   & 1964 & (1953, 1971)      \\
            % Age at hire              & 23.6   & (20.0, 30.3)  &   & 25.3 & (20.2, 32.9)      \\
            % Year of leaving work$^*$ & 1981   & (1970, 1989)  &   & 1981 & (1971, 1989)      \\
            % Age at leaving work$^*$  & 45.2   & (31.8, 57.3)  &   & 53.0 & (36.4, 60.4)      \\
            Years at work       & 15.2                                               & (7.0, 26.6)   &  & 21.0 & (7.8, 29.9)   \\
            % Age at death$^\sharp$    & 73.4   & (64.4, 81.3)  &   & 73.0 & (66.3, 80.8)      \\
            Cumulative exposure$^\flat$
                                & 4.33                                               & (1.71, 10.69) &  & 5.43 & (2.19, 14.33) \\
            \bottomrule
            \multicolumn{6}{l}{\small Statistics shown are count (percent) or median (first and third quartiles).}             \\
            % \multicolumn{6}{l}{\small $^\sharp$ Among those who left work by 1995, when employment records ended.}                \\
            \multicolumn{6}{l}{\small $^\flat$ Among those who were ever-exposed; units in mg/m$^3\,\cdot\,$year.}             \\
            % \multicolumn{6}{l}{\small Abbreviations: NHL: non-Hodgkin lymphoma.}                                                  \\
            \bottomrule
          \end{tabular}
        \end{table}

        \vspace{1.5\baselineskip}
        \begin{block}{\normalsize Affiliations and acknowledgements \vspace*{0.2\baselineskip}}
          \vspace{-0.7\baselineskip}\small
          \begin{center}{
              \textsuperscript{1} Division of Epidemiology \& Biostatistics}
            \hspace{2em}
            \textsuperscript{2} Division of Environmental Health Sciences\\[6pt]
            Funding: CDC, NIOSH (R01OH011092) and NIOSH Training Grant (T42OH008429)
          \end{center}
        \end{block}

      }\end{column} % End of column 2

    \begin{column}{\sepwid}\end{column}

    \begin{column}{\onecolwid} % Begin a column which is two columns wide (column 2)
      \vbox to 0.78\paperheight{

        \begin{alertblock}{Main results}
          \vspace{-0.6\baselineskip}
          \begin{figure}
            \caption{Counterfactual number of cases averted under supportable intervention rules
              based on five different target exposure limits and no censoring,
              with 95\% bootstrap confidence intervals.}
            \begin{tikzpicture}[x=2.5pt,y=2.5pt]
              \input{resources/rd.tex}
            \end{tikzpicture}
          \end{figure}

          \begin{itemize}
            \item There would have been \textbf{502 NHL cases} if there was no censoring
            \item Stronger target exposure limits \textbf{monotonically reduced} NHL risk
            \item Setting the target exposure limit at the
                  \textbf{NIOSH recommended exposure limit}
                  0.5 mg/m\textsuperscript{3}
                  for total particulate mass derived from MWF
                  \textbf{would have averted 112 (95\% CI: 57, 181) NHL cases}
          \end{itemize}
        \end{alertblock}

        \begin{block}{Conclusions and discussion}
          \begin{itemize}
            \item \textbf{Stronger limits on exposure to soluble MWF provide stronger protections against NHL}
                  % \item Causal inference relies on assumptions\vspace*{-0.2\baselineskip}
                  %       \begin{multicols}{2}\begin{itemize}
                  %           \item Exchangeability
                  %           \item Positivity
                  %           \item Consistency
                  %           \item Correct model specification
                  %         \end{itemize}\end{multicols}
            \item During the anticipated rebound in domestic manufacturing,
                  protecting worker health should be a priority
            \item We evaluated supportable interventions with \textbf{guaranteed positivity} and therefore
                  \textbf{avoid extrapolation}
            \item We expect uniformly-enforced target exposure limits to have even
                  \textbf{stronger protective effects}
            \item The \textbf{classic parametric g-formula} estimator can also estimate
                  effects of supportable intervention rules, but
                  \textbf{requires many more parametric assumptions}
                  than ICE g-formula estimators \nocite{Taubman_2009}
          \end{itemize}
        \end{block}

        \vspace{3\baselineskip}

        \begin{block}{\normalsize References \vspace*{0.2\baselineskip}}
          \printbibliography
        \end{block}
        %     }\end{column} % End of column 2
        % \begin{column}{\sepwid}\end{column}
        % \begin{column}{\onecolwid}\vbox to 0.78\paperheight{%

        % \begin{block}{Overview of analysis steps}
        %   \begin{enumerate}\itemsep=6pt
        %     \item Fit a model for $h_{k, k - 1} = \mathbb E\left[Y_k \mid \bar l_{k - 1},
        %     \bar a_{k - 1}, Y_{k - 1} = C_k = 0\right]$.
        %     \item Predict $h^\text{int}_{k, k - 1}(a_{k - 1})$ for all possible $a_{k -1}$
        %     using the model from \mynum{1}.
        %     \item Estimate the counterfactual hazard $h^\text{int}_{K, K - 1}$
        %     $$\hat h^\text{int}_{K, K - 1} = \sum_{a_{K - 1}}
        %     \hat h^\text{int}_{k, k - 1}(a_{k - 1})
        %     \times f^\text{int}(a_{k - 1} \mid \bar l_{k - 1}, \bar a_{k - 2},
        %     Y_{k - 1} = C_{k - 1} = 0)$$
        %     \item Set $k = K - 2$. Iteratively:
        %           \vspace{2pt}\begin{enumerate}\itemsep=8pt
        %             \item Regress $\hat h^\text{int}_{K, k + 1}$ on $\bar L_k$
        %             and $\bar A_k$ among those with $Y_{k + 1} = C_{k + 1} = 0$.
        %             \item Predict $h^\text{int}_{K, k + 1} (a_k)$ for all possible $a_{k}$
        %             using the model from \mynum{a}.
        %             \item Estimate the counterfactual hazard for interval $[k, K]$
        %             \begin{align*}
        %               \hat h^\text{int}_{K, k} = \sum_{a_k}
        %               & \left(
        %                 \hat h^\text{int}_{K, k + 1} (a_k)
        %                 \times (1 - \hat h^\text{int}_{k + 1, k} (a_k)) +
        %                 \hat h^\text{int}_{k + 1, k} (a_k)
        %               \right) \\
        %               & \hspace{2em} \times f^\text{int}(a_{k} \mid
        %               \bar l_{k}, \bar a_{k - 1}, Y_{k} = C_{k} = 0)
        %             \end{align*}
        %             \item If $k > 1$, set $k \leftarrow k - 1$ and return to \mynum{a}.
        %           \end{enumerate}\vspace{2pt}
        %     \item Estimate $\mathbb E\left[ Y_K^\text{int} \right]$ using
        %     $n^{-1}\sum_i^n \hat h^\text{int}_{K, 0}$.
        %   \end{enumerate}

        % \end{block}

      }\end{column} % End of column 3

    \begin{column}{\sepwid}\end{column}

  \end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}