### 2023-06-12 reports/SER (2023)/resources/intervention-density-combined.tex

% Created by tikzDevice version 0.12.3.1 on 2023-06-05 10:01:50
% !TEX encoding = UTF-8 Unicode
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (903.38,180.67);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (903.38,180.68);
\end{scope}
\begin{scope}
\path[clip] ( 33.90, 45.85) rectangle (245.64,164.10);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 33.90, 45.85) rectangle (245.64,164.10);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 43.52, 51.22) rectangle ( 50.92, 54.00);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 50.92, 51.22) rectangle ( 58.33, 54.83);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 58.33, 51.22) rectangle ( 65.73, 58.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 65.73, 51.22) rectangle ( 73.14, 54.23);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 73.14, 51.22) rectangle ( 80.54, 55.10);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 80.54, 51.22) rectangle ( 87.94, 54.07);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 87.94, 51.22) rectangle ( 95.35, 54.09);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 95.35, 51.22) rectangle (102.75, 63.00);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (102.75, 51.22) rectangle (110.15, 58.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (110.15, 51.22) rectangle (117.56, 54.04);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (117.56, 51.22) rectangle (124.96, 55.71);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (124.96, 51.22) rectangle (132.36, 56.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (132.36, 51.22) rectangle (139.77, 54.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (139.77, 51.22) rectangle (147.17, 53.46);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (147.17, 51.22) rectangle (154.58, 59.90);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (154.58, 51.22) rectangle (161.98, 55.03);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (161.98, 51.22) rectangle (169.38, 53.02);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (169.38, 51.22) rectangle (176.79, 55.55);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (176.79, 51.22) rectangle (184.19, 55.39);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (184.19, 51.22) rectangle (191.59, 55.55);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (191.59, 51.22) rectangle (199.00, 53.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (199.00, 51.22) rectangle (206.40, 52.56);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (206.40, 51.22) rectangle (213.80, 54.54);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (213.80, 51.22) rectangle (221.21, 52.94);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (221.21, 51.22) rectangle (228.61, 60.40);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (228.61, 51.22) rectangle (236.02, 51.93);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 43.52, 51.22) rectangle ( 50.92, 63.05);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 50.92, 51.22) rectangle ( 58.33, 54.57);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 58.33, 51.22) rectangle ( 65.73, 59.70);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 65.73, 51.22) rectangle ( 73.14, 54.88);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 73.14, 51.22) rectangle ( 80.54, 55.88);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 80.54, 51.22) rectangle ( 87.94, 54.55);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 87.94, 51.22) rectangle ( 95.35, 55.13);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 95.35, 51.22) rectangle (102.75, 67.26);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (102.75, 51.22) rectangle (110.15, 69.47);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (110.15, 51.22) rectangle (117.56, 66.25);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (117.56, 51.22) rectangle (124.96, 52.97);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (124.96, 51.22) rectangle (132.36, 52.74);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (132.36, 51.22) rectangle (139.77, 52.09);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (139.77, 51.22) rectangle (147.17, 51.69);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (147.17, 51.22) rectangle (154.58, 54.31);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (154.58, 51.22) rectangle (161.98, 52.07);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (161.98, 51.22) rectangle (169.38, 51.69);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (169.38, 51.22) rectangle (176.79, 52.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (176.79, 51.22) rectangle (184.19, 52.14);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (184.19, 51.22) rectangle (191.59, 52.85);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (191.59, 51.22) rectangle (199.00, 51.97);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (199.00, 51.22) rectangle (206.40, 51.66);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (206.40, 51.22) rectangle (213.80, 52.15);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (213.80, 51.22) rectangle (221.21, 51.69);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (221.21, 51.22) rectangle (228.61, 54.97);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (228.61, 51.22) rectangle (236.02, 51.37);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (117.56, 45.85) -- (117.56,164.10);

\path[draw=drawColor,line width= 0.9pt,line join=round,line cap=round] ( 33.90, 45.85) rectangle (245.64,164.10);
\end{scope}
\begin{scope}
\path[clip] (251.14, 45.85) rectangle (462.89,164.10);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.14, 45.85) rectangle (462.89,164.10);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (260.77, 51.22) rectangle (268.17, 87.06);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (268.17, 51.22) rectangle (275.57, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (275.57, 51.22) rectangle (282.98, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (282.98, 51.22) rectangle (290.38, 63.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (290.38, 51.22) rectangle (297.78, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (297.78, 51.22) rectangle (305.19, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (305.19, 51.22) rectangle (312.59, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (312.59, 51.22) rectangle (319.99, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (319.99, 51.22) rectangle (327.40, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (327.40, 51.22) rectangle (334.80, 63.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (334.80, 51.22) rectangle (342.21, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (342.21, 51.22) rectangle (349.61, 63.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (349.61, 51.22) rectangle (357.01, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (357.01, 51.22) rectangle (364.42, 63.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (364.42, 51.22) rectangle (371.82, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (371.82, 51.22) rectangle (379.22, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (379.22, 51.22) rectangle (386.63, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (386.63, 51.22) rectangle (394.03, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (394.03, 51.22) rectangle (401.43, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (401.43, 51.22) rectangle (408.84, 63.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (408.84, 51.22) rectangle (416.24, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (416.24, 51.22) rectangle (423.65, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (423.65, 51.22) rectangle (431.05, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (431.05, 51.22) rectangle (438.45, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (438.45, 51.22) rectangle (445.86, 63.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (445.86, 51.22) rectangle (453.26, 51.22);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (260.77, 51.22) rectangle (268.17, 87.06);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (268.17, 51.22) rectangle (275.57, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (275.57, 51.22) rectangle (282.98, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (282.98, 51.22) rectangle (290.38, 63.17);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (290.38, 51.22) rectangle (297.78, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (297.78, 51.22) rectangle (305.19, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (305.19, 51.22) rectangle (312.59, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (312.59, 51.22) rectangle (319.99, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (319.99, 51.22) rectangle (327.40, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (327.40, 51.22) rectangle (334.80,110.95);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (334.80, 51.22) rectangle (342.21, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (342.21, 51.22) rectangle (349.61, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (349.61, 51.22) rectangle (357.01, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (357.01, 51.22) rectangle (364.42, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (364.42, 51.22) rectangle (371.82, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (371.82, 51.22) rectangle (379.22, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (379.22, 51.22) rectangle (386.63, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (386.63, 51.22) rectangle (394.03, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (394.03, 51.22) rectangle (401.43, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (401.43, 51.22) rectangle (408.84, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (408.84, 51.22) rectangle (416.24, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (416.24, 51.22) rectangle (423.65, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (423.65, 51.22) rectangle (431.05, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (431.05, 51.22) rectangle (438.45, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (438.45, 51.22) rectangle (445.86, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (445.86, 51.22) rectangle (453.26, 51.22);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (334.80, 45.85) -- (334.80,164.10);

\path[draw=drawColor,line width= 0.9pt,line join=round,line cap=round] (251.14, 45.85) rectangle (462.89,164.10);
\end{scope}
\begin{scope}
\path[clip] (468.39, 45.85) rectangle (680.13,164.10);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (468.39, 45.85) rectangle (680.13,164.10);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (478.01, 51.22) rectangle (485.41, 78.10);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (485.41, 51.22) rectangle (492.82, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (492.82, 51.22) rectangle (500.22, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (500.22, 51.22) rectangle (507.62, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (507.62, 51.22) rectangle (515.03, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (515.03, 51.22) rectangle (522.43, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (522.43, 51.22) rectangle (529.84, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (529.84, 51.22) rectangle (537.24, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (537.24, 51.22) rectangle (544.64, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (544.64, 51.22) rectangle (552.05, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (552.05, 51.22) rectangle (559.45,104.97);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (559.45, 51.22) rectangle (566.85, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (566.85, 51.22) rectangle (574.26, 78.10);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (574.26, 51.22) rectangle (581.66, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (581.66, 51.22) rectangle (589.07, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (589.07, 51.22) rectangle (596.47, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (596.47, 51.22) rectangle (603.87, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (603.87, 51.22) rectangle (611.28, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (611.28, 51.22) rectangle (618.68, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (618.68, 51.22) rectangle (626.08, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (626.08, 51.22) rectangle (633.49, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (633.49, 51.22) rectangle (640.89, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (640.89, 51.22) rectangle (648.29, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (648.29, 51.22) rectangle (655.70, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (655.70, 51.22) rectangle (663.10, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (663.10, 51.22) rectangle (670.51, 51.22);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (478.01, 51.22) rectangle (485.41,158.73);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (485.41, 51.22) rectangle (492.82, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (492.82, 51.22) rectangle (500.22, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (500.22, 51.22) rectangle (507.62, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (507.62, 51.22) rectangle (515.03, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (515.03, 51.22) rectangle (522.43, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (522.43, 51.22) rectangle (529.84, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (529.84, 51.22) rectangle (537.24, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (537.24, 51.22) rectangle (544.64, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (544.64, 51.22) rectangle (552.05, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (552.05, 51.22) rectangle (559.45, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (559.45, 51.22) rectangle (566.85, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (566.85, 51.22) rectangle (574.26, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (574.26, 51.22) rectangle (581.66, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (581.66, 51.22) rectangle (589.07, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (589.07, 51.22) rectangle (596.47, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (596.47, 51.22) rectangle (603.87, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (603.87, 51.22) rectangle (611.28, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (611.28, 51.22) rectangle (618.68, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (618.68, 51.22) rectangle (626.08, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (626.08, 51.22) rectangle (633.49, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (633.49, 51.22) rectangle (640.89, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (640.89, 51.22) rectangle (648.29, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (648.29, 51.22) rectangle (655.70, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (655.70, 51.22) rectangle (663.10, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (663.10, 51.22) rectangle (670.51, 51.22);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (552.05, 45.85) -- (552.05,164.10);

\path[draw=drawColor,line width= 0.9pt,line join=round,line cap=round] (468.39, 45.85) rectangle (680.13,164.10);
\end{scope}
\begin{scope}
\path[clip] (685.63, 45.85) rectangle (897.38,164.10);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (685.63, 45.85) rectangle (897.38,164.10);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (695.25, 51.22) rectangle (702.66, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (702.66, 51.22) rectangle (710.06, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (710.06, 51.22) rectangle (717.47, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (717.47, 51.22) rectangle (724.87, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (724.87, 51.22) rectangle (732.27, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (732.27, 51.22) rectangle (739.68, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (739.68, 51.22) rectangle (747.08, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (747.08, 51.22) rectangle (754.48, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (754.48, 51.22) rectangle (761.89, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (761.89, 51.22) rectangle (769.29, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (769.29, 51.22) rectangle (776.70, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (776.70, 51.22) rectangle (784.10, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (784.10, 51.22) rectangle (791.50, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (791.50, 51.22) rectangle (798.91, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (798.91, 51.22) rectangle (806.31, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (806.31, 51.22) rectangle (813.71, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (813.71, 51.22) rectangle (821.12, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (821.12, 51.22) rectangle (828.52, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (828.52, 51.22) rectangle (835.92, 57.94);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (835.92, 51.22) rectangle (843.33, 57.94);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (843.33, 51.22) rectangle (850.73, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (850.73, 51.22) rectangle (858.14, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (858.14, 51.22) rectangle (865.54, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (865.54, 51.22) rectangle (872.94, 57.94);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (872.94, 51.22) rectangle (880.35,138.57);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (880.35, 51.22) rectangle (887.75, 51.22);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (695.25, 51.22) rectangle (702.66, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (702.66, 51.22) rectangle (710.06, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (710.06, 51.22) rectangle (717.47, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (717.47, 51.22) rectangle (724.87, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (724.87, 51.22) rectangle (732.27, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (732.27, 51.22) rectangle (739.68, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (739.68, 51.22) rectangle (747.08, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (747.08, 51.22) rectangle (754.48, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (754.48, 51.22) rectangle (761.89, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (761.89, 51.22) rectangle (769.29, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (769.29, 51.22) rectangle (776.70, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (776.70, 51.22) rectangle (784.10, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (784.10, 51.22) rectangle (791.50, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (791.50, 51.22) rectangle (798.91, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (798.91, 51.22) rectangle (806.31, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (806.31, 51.22) rectangle (813.71, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (813.71, 51.22) rectangle (821.12, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (821.12, 51.22) rectangle (828.52, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (828.52, 51.22) rectangle (835.92, 57.94);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (835.92, 51.22) rectangle (843.33, 57.94);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (843.33, 51.22) rectangle (850.73, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (850.73, 51.22) rectangle (858.14, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (858.14, 51.22) rectangle (865.54, 51.22);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (865.54, 51.22) rectangle (872.94, 57.94);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (872.94, 51.22) rectangle (880.35,138.57);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (880.35, 51.22) rectangle (887.75, 51.22);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (769.29, 45.85) -- (769.29,164.10);

\path[draw=drawColor,line width= 0.9pt,line join=round,line cap=round] (685.63, 45.85) rectangle (897.38,164.10);
\end{scope}
\begin{scope}
\path[clip] ( 33.90,164.10) rectangle (245.64,180.67);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 33.90,164.10) rectangle (245.64,180.68);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (139.77,169.36) {Marginal distribution};
\end{scope}
\begin{scope}
\path[clip] (251.14,164.10) rectangle (462.89,180.67);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (251.14,164.10) rectangle (462.89,180.68);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (357.01,169.36) {Supported limit is at target limit};
\end{scope}
\begin{scope}
\path[clip] (468.39,164.10) rectangle (680.13,180.67);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (468.39,164.10) rectangle (680.13,180.68);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (574.26,169.36) {Supported limit is below target limit};
\end{scope}
\begin{scope}
\path[clip] (685.63,164.10) rectangle (897.38,180.67);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (685.63,164.10) rectangle (897.38,180.68);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (791.50,169.36) {No limit (none observed below target limit)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.52, 43.10) --
	( 43.52, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.75, 43.10) --
	(102.75, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.98, 43.10) --
	(161.98, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.21, 43.10) --
	(221.21, 45.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 43.52, 34.70) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (102.75, 34.70) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (161.98, 34.70) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (221.21, 34.70) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (260.77, 43.10) --
	(260.77, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (319.99, 43.10) --
	(319.99, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.22, 43.10) --
	(379.22, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (438.45, 43.10) --
	(438.45, 45.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (260.77, 34.70) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (319.99, 34.70) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (379.22, 34.70) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (438.45, 34.70) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.01, 43.10) --
	(478.01, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (537.24, 43.10) --
	(537.24, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (596.47, 43.10) --
	(596.47, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (655.70, 43.10) --
	(655.70, 45.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (478.01, 34.70) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (537.24, 34.70) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (596.47, 34.70) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (655.70, 34.70) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (695.25, 43.10) --
	(695.25, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (754.48, 43.10) --
	(754.48, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (813.71, 43.10) --
	(813.71, 45.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (872.94, 43.10) --
	(872.94, 45.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (695.25, 34.70) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (754.48, 34.70) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (813.71, 34.70) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (872.94, 34.70) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95, 48.12) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95, 75.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,101.88) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,128.75) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,155.63) {40};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15, 51.22) --
	( 33.90, 51.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15, 78.10) --
	( 33.90, 78.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15,104.97) --
	( 33.90,104.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15,131.85) --
	( 33.90,131.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15,158.73) --
	( 33.90,158.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (465.64, 20.75) {Exposure to soluble MWF (mg/m\textsuperscript{3})};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 12.20,104.97) {Density};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (290.12,  0.00) rectangle (478.56,  8.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (290.12,  5.90) {Distribution:};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (345.06,  5.00) rectangle (353.06, 13.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.8pt,fill=fillColor,fill opacity=0.20] (346.06,  6.00) rectangle (352.06, 12.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (398.62,  5.00) rectangle (406.62, 13.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.8pt,fill=fillColor,fill opacity=0.20] (399.62,  6.00) rectangle (405.62, 12.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (357.56,  5.90) {Observed};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (411.12,  5.90) {Post-intervention};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (489.56,  0.00) rectangle (641.15,  8.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (489.56,  5.90) {Target exposure limit:};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (581.39,  5.00) rectangle (589.39, 13.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (585.39,  5.00) -- (585.39, 13.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (903.38,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (593.89,  5.90) {0.25 mg/m\textsuperscript{3}};
\end{scope}