### 2023-06-12 reports/SER (2023)/resources/intervention-density-combo2.tex

% Created by tikzDevice version 0.12.3.1 on 2023-06-02 02:14:31
% !TEX encoding = UTF-8 Unicode
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (271.01,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (271.01,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 33.90, 40.85) rectangle (265.01,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 33.90, 40.85) rectangle (265.01,216.81);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 44.40, 48.84) rectangle ( 52.48, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 52.48, 48.84) rectangle ( 60.56, 94.55);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 60.56, 48.84) rectangle ( 68.64,105.98);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 68.64, 48.84) rectangle ( 76.73, 71.70);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 76.73, 48.84) rectangle ( 84.81, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 84.81, 48.84) rectangle ( 92.89, 71.70);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 92.89, 48.84) rectangle (100.97, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (100.97, 48.84) rectangle (109.05, 94.55);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (109.05, 48.84) rectangle (117.13, 71.70);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (117.13, 48.84) rectangle (125.21, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (125.21, 48.84) rectangle (133.29, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (133.29, 48.84) rectangle (141.37, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (141.37, 48.84) rectangle (149.45, 83.12);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (149.45, 48.84) rectangle (157.54, 71.70);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (157.54, 48.84) rectangle (165.62, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (165.62, 48.84) rectangle (173.70, 60.27);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (173.70, 48.84) rectangle (181.78, 60.27);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (181.78, 48.84) rectangle (189.86, 60.27);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (189.86, 48.84) rectangle (197.94, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (197.94, 48.84) rectangle (206.02, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (206.02, 48.84) rectangle (214.10, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (214.10, 48.84) rectangle (222.18, 60.27);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (222.18, 48.84) rectangle (230.26, 60.27);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (230.26, 48.84) rectangle (238.35, 71.70);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (238.35, 48.84) rectangle (246.43, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (246.43, 48.84) rectangle (254.51, 48.84);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 44.40, 48.84) rectangle ( 52.48, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 52.48, 48.84) rectangle ( 60.56, 94.55);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 60.56, 48.84) rectangle ( 68.64,105.98);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 68.64, 48.84) rectangle ( 76.73, 71.70);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 76.73, 48.84) rectangle ( 84.81, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 84.81, 48.84) rectangle ( 92.89, 71.70);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 92.89, 48.84) rectangle (100.97, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (100.97, 48.84) rectangle (109.05, 94.55);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (109.05, 48.84) rectangle (117.13,208.81);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (117.13, 48.84) rectangle (125.21, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (125.21, 48.84) rectangle (133.29, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (133.29, 48.84) rectangle (141.37, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (141.37, 48.84) rectangle (149.45, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (149.45, 48.84) rectangle (157.54, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (157.54, 48.84) rectangle (165.62, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (165.62, 48.84) rectangle (173.70, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (173.70, 48.84) rectangle (181.78, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (181.78, 48.84) rectangle (189.86, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (189.86, 48.84) rectangle (197.94, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (197.94, 48.84) rectangle (206.02, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (206.02, 48.84) rectangle (214.10, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (214.10, 48.84) rectangle (222.18, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (222.18, 48.84) rectangle (230.26, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (230.26, 48.84) rectangle (238.35, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (238.35, 48.84) rectangle (246.43, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (246.43, 48.84) rectangle (254.51, 48.84);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.9pt,line join=round,line cap=round] ( 33.90, 40.85) rectangle (265.01,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95, 45.75) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95, 90.02) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,134.30) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,178.58) {15};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15, 48.84) --
	( 33.90, 48.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15, 93.12) --
	( 33.90, 93.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15,137.40) --
	( 33.90,137.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.15,181.67) --
	( 33.90,181.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.40, 38.10) --
	( 44.40, 40.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.05, 38.10) --
	(109.05, 40.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (173.70, 38.10) --
	(173.70, 40.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.35, 38.10) --
	(238.35, 40.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 44.40, 29.70) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (109.05, 29.70) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (173.70, 29.70) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (238.35, 29.70) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (149.45, 15.75) {Exposure to soluble MWF (mg/m\textsuperscript{3})};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 12.20,128.83) {Density};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 55.24,  0.00) rectangle (243.67,  3.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 55.24,  0.90) {Distribution:};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (110.17,  0.00) rectangle (118.17,  8.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.8pt,fill=fillColor,fill opacity=0.20] (111.17,  1.00) rectangle (117.17,  7.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (163.74,  0.00) rectangle (171.74,  8.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.8pt,fill=fillColor,fill opacity=0.20] (164.74,  1.00) rectangle (170.74,  7.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (122.67,  0.90) {Observed};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (176.24,  0.90) {Post intervention};
\end{scope}