### 2023-06-12 reports/SER (2023)/resources/intervention-density-combo3.tex

% Created by tikzDevice version 0.12.3.1 on 2023-06-02 02:14:31
% !TEX encoding = UTF-8 Unicode
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (271.01,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (271.01,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 40.89, 40.85) rectangle (265.01,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.89, 40.85) rectangle (265.01,216.81);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 51.08, 48.84) rectangle ( 56.31, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 56.31, 48.84) rectangle ( 61.53, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 61.53, 48.84) rectangle ( 66.75, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 66.75, 48.84) rectangle ( 71.98, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 71.98, 48.84) rectangle ( 77.20, 54.00);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 77.20, 48.84) rectangle ( 82.43, 51.42);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 82.43, 48.84) rectangle ( 87.65, 82.39);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 87.65, 48.84) rectangle ( 92.87, 54.00);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 92.87, 48.84) rectangle ( 98.10, 97.87);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 98.10, 48.84) rectangle (103.32, 66.91);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (103.32, 48.84) rectangle (108.55, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (108.55, 48.84) rectangle (113.77, 64.33);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (113.77, 48.84) rectangle (119.00, 56.58);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (119.00, 48.84) rectangle (124.22, 69.49);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (124.22, 48.84) rectangle (129.44, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (129.44, 48.84) rectangle (134.67, 51.42);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (134.67, 48.84) rectangle (139.89, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (139.89, 48.84) rectangle (145.12, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (145.12, 48.84) rectangle (150.34, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (150.34, 48.84) rectangle (155.57, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (155.57, 48.84) rectangle (160.79, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (160.79, 48.84) rectangle (166.01, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (166.01, 48.84) rectangle (171.24, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (171.24, 48.84) rectangle (176.46, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (176.46, 48.84) rectangle (181.69, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (181.69, 48.84) rectangle (186.91, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (186.91, 48.84) rectangle (192.13, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (192.13, 48.84) rectangle (197.36, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (197.36, 48.84) rectangle (202.58, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (202.58, 48.84) rectangle (207.81, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (207.81, 48.84) rectangle (213.03, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (213.03, 48.84) rectangle (218.26, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (218.26, 48.84) rectangle (223.48, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (223.48, 48.84) rectangle (228.70, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (228.70, 48.84) rectangle (233.93, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (233.93, 48.84) rectangle (239.15, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (239.15, 48.84) rectangle (244.38, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (244.38, 48.84) rectangle (249.60, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (249.60, 48.84) rectangle (254.83, 48.84);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 51.08, 48.84) rectangle ( 56.31, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 56.31, 48.84) rectangle ( 61.53, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 61.53, 48.84) rectangle ( 66.75, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 66.75, 48.84) rectangle ( 71.98, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 71.98, 48.84) rectangle ( 77.20,208.81);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 77.20, 48.84) rectangle ( 82.43, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 82.43, 48.84) rectangle ( 87.65, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 87.65, 48.84) rectangle ( 92.87, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 92.87, 48.84) rectangle ( 98.10, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] ( 98.10, 48.84) rectangle (103.32, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (103.32, 48.84) rectangle (108.55, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (108.55, 48.84) rectangle (113.77, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (113.77, 48.84) rectangle (119.00, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (119.00, 48.84) rectangle (124.22, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (124.22, 48.84) rectangle (129.44, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (129.44, 48.84) rectangle (134.67, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (134.67, 48.84) rectangle (139.89, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (139.89, 48.84) rectangle (145.12, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (145.12, 48.84) rectangle (150.34, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (150.34, 48.84) rectangle (155.57, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (155.57, 48.84) rectangle (160.79, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (160.79, 48.84) rectangle (166.01, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (166.01, 48.84) rectangle (171.24, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (171.24, 48.84) rectangle (176.46, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (176.46, 48.84) rectangle (181.69, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (181.69, 48.84) rectangle (186.91, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (186.91, 48.84) rectangle (192.13, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (192.13, 48.84) rectangle (197.36, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (197.36, 48.84) rectangle (202.58, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (202.58, 48.84) rectangle (207.81, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (207.81, 48.84) rectangle (213.03, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (213.03, 48.84) rectangle (218.26, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (218.26, 48.84) rectangle (223.48, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (223.48, 48.84) rectangle (228.70, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (228.70, 48.84) rectangle (233.93, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (233.93, 48.84) rectangle (239.15, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (239.15, 48.84) rectangle (244.38, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (244.38, 48.84) rectangle (249.60, 48.84);

\path[draw=drawColor,line width= 1.6pt,fill=fillColor,fill opacity=0.20] (249.60, 48.84) rectangle (254.83, 48.84);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.9pt,line join=round,line cap=round] ( 40.89, 40.85) rectangle (265.01,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94, 45.75) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94, 85.74) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94,125.73) {5.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94,165.72) {7.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94,205.71) {10.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14, 48.84) --
	( 40.89, 48.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14, 88.84) --
	( 40.89, 88.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14,128.83) --
	( 40.89,128.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14,168.82) --
	( 40.89,168.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14,208.81) --
	( 40.89,208.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.08, 38.10) --
	( 51.08, 40.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.32, 38.10) --
	(103.32, 40.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.57, 38.10) --
	(155.57, 40.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.81, 38.10) --
	(207.81, 40.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (260.05, 38.10) --
	(260.05, 40.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 51.08, 29.70) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (103.32, 29.70) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (155.57, 29.70) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (207.81, 29.70) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (260.05, 29.70) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (152.95, 15.75) {Exposure to soluble MWF (mg/m\textsuperscript{3})};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 12.20,128.83) {Density};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 58.74,  0.00) rectangle (247.17,  3.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 58.74,  0.90) {Distribution:};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (113.67,  0.00) rectangle (121.67,  8.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{186,129,35}
\definecolor{fillColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.8pt,fill=fillColor,fill opacity=0.20] (114.67,  1.00) rectangle (120.67,  7.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (167.24,  0.00) rectangle (175.24,  8.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,60,97}
\definecolor{fillColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.8pt,fill=fillColor,fill opacity=0.20] (168.24,  1.00) rectangle (174.24,  7.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (126.17,  0.90) {Observed};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (179.74,  0.90) {Post intervention};
\end{scope}