### 2023-06-12 reports/SER (2023)/resources/intervention-density-witnin-combos.tex

% Created by tikzDevice version 0.12.3.1 on 2023-06-12 13:31:15
% !TEX encoding = UTF-8 Unicode
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (271.01,437.23);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (271.01,437.23);
\end{scope}
\begin{scope}
\path[clip] ( 33.90,319.44) rectangle (265.01,420.66);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 33.90,319.44) rectangle (265.01,420.66);
\definecolor{fillColor}{RGB}{163,207,235}

\path[fill=fillColor] ( 44.40,324.04) rectangle ( 52.48,354.71);

\path[fill=fillColor] ( 52.48,324.04) rectangle ( 60.56,324.04);

\path[fill=fillColor] ( 60.56,324.04) rectangle ( 68.64,324.04);

\path[fill=fillColor] ( 68.64,324.04) rectangle ( 76.73,334.26);

\path[fill=fillColor] ( 76.73,324.04) rectangle ( 84.81,324.04);

\path[fill=fillColor] ( 84.81,324.04) rectangle ( 92.89,324.04);

\path[fill=fillColor] ( 92.89,324.04) rectangle (100.97,324.04);

\path[fill=fillColor] (100.97,324.04) rectangle (109.05,324.04);

\path[fill=fillColor] (109.05,324.04) rectangle (117.13,324.04);

\path[fill=fillColor] (117.13,324.04) rectangle (125.21,375.16);

\path[fill=fillColor] (125.21,324.04) rectangle (133.29,324.04);

\path[fill=fillColor] (133.29,324.04) rectangle (141.37,324.04);

\path[fill=fillColor] (141.37,324.04) rectangle (149.45,324.04);

\path[fill=fillColor] (149.45,324.04) rectangle (157.54,324.04);

\path[fill=fillColor] (157.54,324.04) rectangle (165.62,324.04);

\path[fill=fillColor] (165.62,324.04) rectangle (173.70,324.04);

\path[fill=fillColor] (173.70,324.04) rectangle (181.78,324.04);

\path[fill=fillColor] (181.78,324.04) rectangle (189.86,324.04);

\path[fill=fillColor] (189.86,324.04) rectangle (197.94,324.04);

\path[fill=fillColor] (197.94,324.04) rectangle (206.02,324.04);

\path[fill=fillColor] (206.02,324.04) rectangle (214.10,324.04);

\path[fill=fillColor] (214.10,324.04) rectangle (222.18,324.04);

\path[fill=fillColor] (222.18,324.04) rectangle (230.26,324.04);

\path[fill=fillColor] (230.26,324.04) rectangle (238.35,324.04);

\path[fill=fillColor] (238.35,324.04) rectangle (246.43,324.04);

\path[fill=fillColor] (246.43,324.04) rectangle (254.51,324.04);
\definecolor{drawColor}{RGB}{186,129,35}

\path[draw=drawColor,line width= 3.2pt] ( 44.40,324.04) rectangle ( 52.48,354.71);

\path[draw=drawColor,line width= 3.2pt] ( 52.48,324.04) rectangle ( 60.56,324.04);

\path[draw=drawColor,line width= 3.2pt] ( 60.56,324.04) rectangle ( 68.64,324.04);

\path[draw=drawColor,line width= 3.2pt] ( 68.64,324.04) rectangle ( 76.73,334.26);

\path[draw=drawColor,line width= 3.2pt] ( 76.73,324.04) rectangle ( 84.81,324.04);

\path[draw=drawColor,line width= 3.2pt] ( 84.81,324.04) rectangle ( 92.89,324.04);

\path[draw=drawColor,line width= 3.2pt] ( 92.89,324.04) rectangle (100.97,324.04);

\path[draw=drawColor,line width= 3.2pt] (100.97,324.04) rectangle (109.05,324.04);

\path[draw=drawColor,line width= 3.2pt] (109.05,324.04) rectangle (117.13,324.04);

\path[draw=drawColor,line width= 3.2pt] (117.13,324.04) rectangle (125.21,334.26);

\path[draw=drawColor,line width= 3.2pt] (125.21,324.04) rectangle (133.29,324.04);

\path[draw=drawColor,line width= 3.2pt] (133.29,324.04) rectangle (141.37,334.26);

\path[draw=drawColor,line width= 3.2pt] (141.37,324.04) rectangle (149.45,324.04);

\path[draw=drawColor,line width= 3.2pt] (149.45,324.04) rectangle (157.54,334.26);

\path[draw=drawColor,line width= 3.2pt] (157.54,324.04) rectangle (165.62,324.04);

\path[draw=drawColor,line width= 3.2pt] (165.62,324.04) rectangle (173.70,324.04);

\path[draw=drawColor,line width= 3.2pt] (173.70,324.04) rectangle (181.78,324.04);

\path[draw=drawColor,line width= 3.2pt] (181.78,324.04) rectangle (189.86,324.04);

\path[draw=drawColor,line width= 3.2pt] (189.86,324.04) rectangle (197.94,324.04);

\path[draw=drawColor,line width= 3.2pt] (197.94,324.04) rectangle (206.02,334.26);

\path[draw=drawColor,line width= 3.2pt] (206.02,324.04) rectangle (214.10,324.04);

\path[draw=drawColor,line width= 3.2pt] (214.10,324.04) rectangle (222.18,324.04);

\path[draw=drawColor,line width= 3.2pt] (222.18,324.04) rectangle (230.26,324.04);

\path[draw=drawColor,line width= 3.2pt] (230.26,324.04) rectangle (238.35,324.04);

\path[draw=drawColor,line width= 3.2pt] (238.35,324.04) rectangle (246.43,334.26);

\path[draw=drawColor,line width= 3.2pt] (246.43,324.04) rectangle (254.51,324.04);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.6pt,dash pattern=on 4pt off 4pt ,line join=round] (125.21,319.44) -- (125.21,420.66);

\path[draw=drawColor,line width= 3.2pt,line join=round,line cap=round] ( 33.90,319.44) rectangle (265.01,420.66);
\end{scope}
\begin{scope}
\path[clip] ( 33.90,196.14) rectangle (265.01,297.37);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 33.90,196.14) rectangle (265.01,297.37);
\definecolor{fillColor}{RGB}{163,207,235}

\path[fill=fillColor] ( 44.40,200.74) rectangle ( 52.48,292.77);

\path[fill=fillColor] ( 52.48,200.74) rectangle ( 60.56,200.74);

\path[fill=fillColor] ( 60.56,200.74) rectangle ( 68.64,200.74);

\path[fill=fillColor] ( 68.64,200.74) rectangle ( 76.73,200.74);

\path[fill=fillColor] ( 76.73,200.74) rectangle ( 84.81,200.74);

\path[fill=fillColor] ( 84.81,200.74) rectangle ( 92.89,200.74);

\path[fill=fillColor] ( 92.89,200.74) rectangle (100.97,200.74);

\path[fill=fillColor] (100.97,200.74) rectangle (109.05,200.74);

\path[fill=fillColor] (109.05,200.74) rectangle (117.13,200.74);

\path[fill=fillColor] (117.13,200.74) rectangle (125.21,200.74);

\path[fill=fillColor] (125.21,200.74) rectangle (133.29,200.74);

\path[fill=fillColor] (133.29,200.74) rectangle (141.37,200.74);

\path[fill=fillColor] (141.37,200.74) rectangle (149.45,200.74);

\path[fill=fillColor] (149.45,200.74) rectangle (157.54,200.74);

\path[fill=fillColor] (157.54,200.74) rectangle (165.62,200.74);

\path[fill=fillColor] (165.62,200.74) rectangle (173.70,200.74);

\path[fill=fillColor] (173.70,200.74) rectangle (181.78,200.74);

\path[fill=fillColor] (181.78,200.74) rectangle (189.86,200.74);

\path[fill=fillColor] (189.86,200.74) rectangle (197.94,200.74);

\path[fill=fillColor] (197.94,200.74) rectangle (206.02,200.74);

\path[fill=fillColor] (206.02,200.74) rectangle (214.10,200.74);

\path[fill=fillColor] (214.10,200.74) rectangle (222.18,200.74);

\path[fill=fillColor] (222.18,200.74) rectangle (230.26,200.74);

\path[fill=fillColor] (230.26,200.74) rectangle (238.35,200.74);

\path[fill=fillColor] (238.35,200.74) rectangle (246.43,200.74);

\path[fill=fillColor] (246.43,200.74) rectangle (254.51,200.74);
\definecolor{drawColor}{RGB}{186,129,35}

\path[draw=drawColor,line width= 3.2pt] ( 44.40,200.74) rectangle ( 52.48,223.75);

\path[draw=drawColor,line width= 3.2pt] ( 52.48,200.74) rectangle ( 60.56,200.74);

\path[draw=drawColor,line width= 3.2pt] ( 60.56,200.74) rectangle ( 68.64,200.74);

\path[draw=drawColor,line width= 3.2pt] ( 68.64,200.74) rectangle ( 76.73,200.74);

\path[draw=drawColor,line width= 3.2pt] ( 76.73,200.74) rectangle ( 84.81,200.74);

\path[draw=drawColor,line width= 3.2pt] ( 84.81,200.74) rectangle ( 92.89,200.74);

\path[draw=drawColor,line width= 3.2pt] ( 92.89,200.74) rectangle (100.97,200.74);

\path[draw=drawColor,line width= 3.2pt] (100.97,200.74) rectangle (109.05,200.74);

\path[draw=drawColor,line width= 3.2pt] (109.05,200.74) rectangle (117.13,200.74);

\path[draw=drawColor,line width= 3.2pt] (117.13,200.74) rectangle (125.21,200.74);

\path[draw=drawColor,line width= 3.2pt] (125.21,200.74) rectangle (133.29,246.75);

\path[draw=drawColor,line width= 3.2pt] (133.29,200.74) rectangle (141.37,200.74);

\path[draw=drawColor,line width= 3.2pt] (141.37,200.74) rectangle (149.45,223.75);

\path[draw=drawColor,line width= 3.2pt] (149.45,200.74) rectangle (157.54,200.74);

\path[draw=drawColor,line width= 3.2pt] (157.54,200.74) rectangle (165.62,200.74);

\path[draw=drawColor,line width= 3.2pt] (165.62,200.74) rectangle (173.70,200.74);

\path[draw=drawColor,line width= 3.2pt] (173.70,200.74) rectangle (181.78,200.74);

\path[draw=drawColor,line width= 3.2pt] (181.78,200.74) rectangle (189.86,200.74);

\path[draw=drawColor,line width= 3.2pt] (189.86,200.74) rectangle (197.94,200.74);

\path[draw=drawColor,line width= 3.2pt] (197.94,200.74) rectangle (206.02,200.74);

\path[draw=drawColor,line width= 3.2pt] (206.02,200.74) rectangle (214.10,200.74);

\path[draw=drawColor,line width= 3.2pt] (214.10,200.74) rectangle (222.18,200.74);

\path[draw=drawColor,line width= 3.2pt] (222.18,200.74) rectangle (230.26,200.74);

\path[draw=drawColor,line width= 3.2pt] (230.26,200.74) rectangle (238.35,200.74);

\path[draw=drawColor,line width= 3.2pt] (238.35,200.74) rectangle (246.43,200.74);

\path[draw=drawColor,line width= 3.2pt] (246.43,200.74) rectangle (254.51,200.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.6pt,dash pattern=on 4pt off 4pt ,line join=round] (125.21,196.14) -- (125.21,297.37);

\path[draw=drawColor,line width= 3.2pt,line join=round,line cap=round] ( 33.90,196.14) rectangle (265.01,297.37);
\end{scope}
\begin{scope}
\path[clip] ( 33.90, 72.85) rectangle (265.01,174.07);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 33.90, 72.85) rectangle (265.01,174.07);
\definecolor{fillColor}{RGB}{163,207,235}

\path[fill=fillColor] ( 44.40, 77.45) rectangle ( 52.48, 77.45);

\path[fill=fillColor] ( 52.48, 77.45) rectangle ( 60.56, 77.45);

\path[fill=fillColor] ( 60.56, 77.45) rectangle ( 68.64, 77.45);

\path[fill=fillColor] ( 68.64, 77.45) rectangle ( 76.73, 77.45);

\path[fill=fillColor] ( 76.73, 77.45) rectangle ( 84.81, 77.45);

\path[fill=fillColor] ( 84.81, 77.45) rectangle ( 92.89, 77.45);

\path[fill=fillColor] ( 92.89, 77.45) rectangle (100.97, 77.45);

\path[fill=fillColor] (100.97, 77.45) rectangle (109.05, 77.45);

\path[fill=fillColor] (109.05, 77.45) rectangle (117.13, 77.45);

\path[fill=fillColor] (117.13, 77.45) rectangle (125.21, 77.45);

\path[fill=fillColor] (125.21, 77.45) rectangle (133.29, 77.45);

\path[fill=fillColor] (133.29, 77.45) rectangle (141.37, 77.45);

\path[fill=fillColor] (141.37, 77.45) rectangle (149.45, 77.45);

\path[fill=fillColor] (149.45, 77.45) rectangle (157.54, 77.45);

\path[fill=fillColor] (157.54, 77.45) rectangle (165.62, 77.45);

\path[fill=fillColor] (165.62, 77.45) rectangle (173.70, 77.45);

\path[fill=fillColor] (173.70, 77.45) rectangle (181.78, 77.45);

\path[fill=fillColor] (181.78, 77.45) rectangle (189.86, 77.45);

\path[fill=fillColor] (189.86, 77.45) rectangle (197.94, 83.20);

\path[fill=fillColor] (197.94, 77.45) rectangle (206.02, 83.20);

\path[fill=fillColor] (206.02, 77.45) rectangle (214.10, 77.45);

\path[fill=fillColor] (214.10, 77.45) rectangle (222.18, 77.45);

\path[fill=fillColor] (222.18, 77.45) rectangle (230.26, 77.45);

\path[fill=fillColor] (230.26, 77.45) rectangle (238.35, 83.20);

\path[fill=fillColor] (238.35, 77.45) rectangle (246.43,152.22);

\path[fill=fillColor] (246.43, 77.45) rectangle (254.51, 77.45);
\definecolor{drawColor}{RGB}{186,129,35}

\path[draw=drawColor,line width= 3.2pt] ( 44.40, 77.45) rectangle ( 52.48, 77.45);

\path[draw=drawColor,line width= 3.2pt] ( 52.48, 77.45) rectangle ( 60.56, 77.45);

\path[draw=drawColor,line width= 3.2pt] ( 60.56, 77.45) rectangle ( 68.64, 77.45);

\path[draw=drawColor,line width= 3.2pt] ( 68.64, 77.45) rectangle ( 76.73, 77.45);

\path[draw=drawColor,line width= 3.2pt] ( 76.73, 77.45) rectangle ( 84.81, 77.45);

\path[draw=drawColor,line width= 3.2pt] ( 84.81, 77.45) rectangle ( 92.89, 77.45);

\path[draw=drawColor,line width= 3.2pt] ( 92.89, 77.45) rectangle (100.97, 77.45);

\path[draw=drawColor,line width= 3.2pt] (100.97, 77.45) rectangle (109.05, 77.45);

\path[draw=drawColor,line width= 3.2pt] (109.05, 77.45) rectangle (117.13, 77.45);

\path[draw=drawColor,line width= 3.2pt] (117.13, 77.45) rectangle (125.21, 77.45);

\path[draw=drawColor,line width= 3.2pt] (125.21, 77.45) rectangle (133.29, 77.45);

\path[draw=drawColor,line width= 3.2pt] (133.29, 77.45) rectangle (141.37, 77.45);

\path[draw=drawColor,line width= 3.2pt] (141.37, 77.45) rectangle (149.45, 77.45);

\path[draw=drawColor,line width= 3.2pt] (149.45, 77.45) rectangle (157.54, 77.45);

\path[draw=drawColor,line width= 3.2pt] (157.54, 77.45) rectangle (165.62, 77.45);

\path[draw=drawColor,line width= 3.2pt] (165.62, 77.45) rectangle (173.70, 77.45);

\path[draw=drawColor,line width= 3.2pt] (173.70, 77.45) rectangle (181.78, 77.45);

\path[draw=drawColor,line width= 3.2pt] (181.78, 77.45) rectangle (189.86, 77.45);

\path[draw=drawColor,line width= 3.2pt] (189.86, 77.45) rectangle (197.94, 83.20);

\path[draw=drawColor,line width= 3.2pt] (197.94, 77.45) rectangle (206.02, 83.20);

\path[draw=drawColor,line width= 3.2pt] (206.02, 77.45) rectangle (214.10, 77.45);

\path[draw=drawColor,line width= 3.2pt] (214.10, 77.45) rectangle (222.18, 77.45);

\path[draw=drawColor,line width= 3.2pt] (222.18, 77.45) rectangle (230.26, 77.45);

\path[draw=drawColor,line width= 3.2pt] (230.26, 77.45) rectangle (238.35, 83.20);

\path[draw=drawColor,line width= 3.2pt] (238.35, 77.45) rectangle (246.43,152.22);

\path[draw=drawColor,line width= 3.2pt] (246.43, 77.45) rectangle (254.51, 77.45);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.6pt,dash pattern=on 4pt off 4pt ,line join=round] (125.21, 72.85) -- (125.21,174.07);

\path[draw=drawColor,line width= 3.2pt,line join=round,line cap=round] ( 33.90, 72.85) rectangle (265.01,174.07);
\end{scope}
\begin{scope}
\path[clip] ( 33.90,174.07) rectangle (265.01,190.64);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 3.2pt,line join=round,line cap=round,fill=fillColor] ( 33.90,174.07) rectangle (265.01,190.64);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (149.45,179.33) {No limit (all observed exposures $>$ target limit)};
\end{scope}
\begin{scope}
\path[clip] ( 33.90,297.37) rectangle (265.01,313.94);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 3.2pt,line join=round,line cap=round,fill=fillColor] ( 33.90,297.37) rectangle (265.01,313.94);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (149.45,302.62) {Supported exposure limit $<$ target limit};
\end{scope}
\begin{scope}
\path[clip] ( 33.90,420.66) rectangle (265.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{211,211,211}

\path[draw=drawColor,line width= 3.2pt,line join=round,line cap=round,fill=fillColor] ( 33.90,420.66) rectangle (265.01,437.23);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (149.45,425.92) {Supported exposure limit $=$ target limit};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 44.40, 70.10) --
	( 44.40, 72.85);

\path[draw=drawColor,line width= 3.2pt,line join=round] (109.05, 70.10) --
	(109.05, 72.85);

\path[draw=drawColor,line width= 3.2pt,line join=round] (173.70, 70.10) --
	(173.70, 72.85);

\path[draw=drawColor,line width= 3.2pt,line join=round] (238.35, 70.10) --
	(238.35, 72.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 44.40, 61.70) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (109.05, 61.70) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (173.70, 61.70) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (238.35, 61.70) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,320.94) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,343.95) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,366.95) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,389.96) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,412.96) {40};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,324.04) --
	( 33.90,324.04);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,347.04) --
	( 33.90,347.04);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,370.05) --
	( 33.90,370.05);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,393.06) --
	( 33.90,393.06);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,416.06) --
	( 33.90,416.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,197.64) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,220.65) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,243.65) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,266.66) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,289.67) {40};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,200.74) --
	( 33.90,200.74);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,223.75) --
	( 33.90,223.75);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,246.75) --
	( 33.90,246.75);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,269.76) --
	( 33.90,269.76);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,292.77) --
	( 33.90,292.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95, 74.35) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95, 97.35) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,120.36) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,143.36) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 28.95,166.37) {40};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15, 77.45) --
	( 33.90, 77.45);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,100.45) --
	( 33.90,100.45);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,123.46) --
	( 33.90,123.46);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,146.46) --
	( 33.90,146.46);

\path[draw=drawColor,line width= 3.2pt,line join=round] ( 31.15,169.47) --
	( 33.90,169.47);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (149.45, 47.75) {Exposure to soluble MWF (mg/m\textsuperscript{3})};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 12.20,246.75) {Density};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 42.74, 23.00) rectangle (256.17, 35.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 42.74, 30.90) {Distribution:\ \ \ };
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (104.17, 28.00) rectangle (116.17, 40.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{163,207,235}
\definecolor{fillColor}{RGB}{163,207,235}

\path[draw=drawColor,line width= 1.2pt,fill=fillColor] (105.67, 29.50) rectangle (114.67, 38.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (202.60, 28.00) rectangle (214.60, 40.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{186,129,35}

\path[draw=drawColor,line width= 1.2pt] (204.10, 29.50) rectangle (213.10, 38.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (121.17, 30.90) {Post-intervention\ \ \ \ };
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (219.60, 30.90) {Observed};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 71.16,  0.00) rectangle (227.75, 12.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 71.16,  7.90) {Target exposure limit:};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (163.49,  5.00) rectangle (175.49, 17.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.6pt,dash pattern=on 4pt off 4pt ,line join=round] (169.49,  5.00) -- (169.49, 17.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,437.23);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (180.49,  7.90) {0.25 mg/m\textsuperscript{3}};
\end{scope}