### 2023-06-12 reports/SER (2023)/resources/rd.tex

% Created by tikzDevice version 0.12.3.1 on 2023-06-01 11:24:54
% !TEX encoding = UTF-8 Unicode
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (271.01,162.61);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,162.61);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (271.01,162.61);
\end{scope}
\begin{scope}
\path[clip] ( 40.89, 26.85) rectangle (265.01,162.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.89, 26.85) rectangle (265.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round,fill=fillColor] (199.88,139.19) --
	(248.03,139.19) --
	(247.96,139.19) --
	(248.25,139.20) --
	(248.53,139.26) --
	(248.81,139.36) --
	(249.06,139.51) --
	(249.28,139.69) --
	(249.48,139.91) --
	(249.63,140.15) --
	(249.75,140.42) --
	(249.82,140.70) --
	(249.84,140.99) --
	(249.84,140.99) --
	(249.84,152.89) --
	(249.84,152.89) --
	(249.82,153.18) --
	(249.75,153.46) --
	(249.63,153.73) --
	(249.48,153.98) --
	(249.28,154.19) --
	(249.06,154.38) --
	(248.81,154.52) --
	(248.53,154.63) --
	(248.25,154.68) --
	(248.03,154.70) --
	(199.88,154.70) --
	(200.10,154.68) --
	(199.81,154.70) --
	(199.52,154.66) --
	(199.24,154.58) --
	(198.97,154.46) --
	(198.74,154.29) --
	(198.53,154.09) --
	(198.35,153.86) --
	(198.22,153.60) --
	(198.12,153.32) --
	(198.08,153.04) --
	(198.07,152.89) --
	(198.07,140.99) --
	(198.08,141.14) --
	(198.08,140.85) --
	(198.12,140.56) --
	(198.22,140.29) --
	(198.35,140.03) --
	(198.53,139.80) --
	(198.74,139.60) --
	(198.97,139.43) --
	(199.24,139.31) --
	(199.52,139.22) --
	(199.81,139.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] ( 40.89, 26.85) rectangle (265.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (223.95,144.19) {98 (45, 161)};
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round,fill=fillColor] (197.88,113.08) --
	(250.03,113.08) --
	(249.96,113.08) --
	(250.25,113.09) --
	(250.53,113.15) --
	(250.81,113.25) --
	(251.06,113.40) --
	(251.28,113.58) --
	(251.48,113.80) --
	(251.63,114.05) --
	(251.75,114.31) --
	(251.81,114.60) --
	(251.84,114.89) --
	(251.84,114.89) --
	(251.84,126.78) --
	(251.84,126.78) --
	(251.81,127.07) --
	(251.75,127.36) --
	(251.63,127.62) --
	(251.48,127.87) --
	(251.28,128.09) --
	(251.06,128.27) --
	(250.81,128.42) --
	(250.53,128.52) --
	(250.25,128.58) --
	(250.03,128.59) --
	(197.88,128.59) --
	(198.10,128.58) --
	(197.81,128.59) --
	(197.52,128.55) --
	(197.24,128.47) --
	(196.97,128.35) --
	(196.74,128.18) --
	(196.53,127.98) --
	(196.35,127.75) --
	(196.22,127.49) --
	(196.12,127.22) --
	(196.08,126.93) --
	(196.07,126.78) --
	(196.07,114.89) --
	(196.08,115.03) --
	(196.08,114.74) --
	(196.12,114.45) --
	(196.22,114.18) --
	(196.35,113.92) --
	(196.53,113.69) --
	(196.74,113.49) --
	(196.97,113.32) --
	(197.24,113.20) --
	(197.52,113.12) --
	(197.81,113.08) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] ( 40.89, 26.85) rectangle (265.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (223.95,118.08) {106 (51, 173)};
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round,fill=fillColor] (197.88, 86.97) --
	(250.03, 86.97) --
	(249.96, 86.97) --
	(250.25, 86.99) --
	(250.53, 87.04) --
	(250.81, 87.15) --
	(251.06, 87.29) --
	(251.28, 87.48) --
	(251.48, 87.69) --
	(251.63, 87.94) --
	(251.75, 88.21) --
	(251.81, 88.49) --
	(251.84, 88.78) --
	(251.84, 88.78) --
	(251.84,100.67) --
	(251.84,100.67) --
	(251.81,100.96) --
	(251.75,101.25) --
	(251.63,101.51) --
	(251.48,101.76) --
	(251.28,101.98) --
	(251.06,102.16) --
	(250.81,102.31) --
	(250.53,102.41) --
	(250.25,102.47) --
	(250.03,102.48) --
	(197.88,102.48) --
	(198.10,102.47) --
	(197.81,102.48) --
	(197.52,102.45) --
	(197.24,102.36) --
	(196.97,102.24) --
	(196.74,102.07) --
	(196.53,101.87) --
	(196.35,101.64) --
	(196.22,101.38) --
	(196.12,101.11) --
	(196.08,100.82) --
	(196.07,100.67) --
	(196.07, 88.78) --
	(196.08, 88.92) --
	(196.08, 88.63) --
	(196.12, 88.35) --
	(196.22, 88.07) --
	(196.35, 87.81) --
	(196.53, 87.58) --
	(196.74, 87.38) --
	(196.97, 87.21) --
	(197.24, 87.09) --
	(197.52, 87.01) --
	(197.81, 86.97) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] ( 40.89, 26.85) rectangle (265.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (223.95, 91.97) {112 (57, 181)};
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round,fill=fillColor] (197.88, 60.86) --
	(250.03, 60.86) --
	(249.96, 60.87) --
	(250.25, 60.88) --
	(250.53, 60.94) --
	(250.81, 61.04) --
	(251.06, 61.18) --
	(251.28, 61.37) --
	(251.48, 61.59) --
	(251.63, 61.83) --
	(251.75, 62.10) --
	(251.81, 62.38) --
	(251.84, 62.67) --
	(251.84, 62.67) --
	(251.84, 74.57) --
	(251.84, 74.57) --
	(251.81, 74.86) --
	(251.75, 75.14) --
	(251.63, 75.41) --
	(251.48, 75.65) --
	(251.28, 75.87) --
	(251.06, 76.05) --
	(250.81, 76.20) --
	(250.53, 76.30) --
	(250.25, 76.36) --
	(250.03, 76.37) --
	(197.88, 76.37) --
	(198.10, 76.36) --
	(197.81, 76.37) --
	(197.52, 76.34) --
	(197.24, 76.26) --
	(196.97, 76.13) --
	(196.74, 75.97) --
	(196.53, 75.77) --
	(196.35, 75.53) --
	(196.22, 75.28) --
	(196.12, 75.00) --
	(196.08, 74.71) --
	(196.07, 74.57) --
	(196.07, 62.67) --
	(196.08, 62.82) --
	(196.08, 62.53) --
	(196.12, 62.24) --
	(196.22, 61.96) --
	(196.35, 61.70) --
	(196.53, 61.47) --
	(196.74, 61.27) --
	(196.97, 61.11) --
	(197.24, 60.98) --
	(197.52, 60.90) --
	(197.81, 60.87) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] ( 40.89, 26.85) rectangle (265.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (223.95, 65.86) {118 (59, 192)};
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round,fill=fillColor] (197.88, 34.76) --
	(250.03, 34.76) --
	(249.96, 34.76) --
	(250.25, 34.77) --
	(250.53, 34.83) --
	(250.81, 34.93) --
	(251.06, 35.08) --
	(251.28, 35.26) --
	(251.48, 35.48) --
	(251.63, 35.72) --
	(251.75, 35.99) --
	(251.81, 36.27) --
	(251.84, 36.56) --
	(251.84, 36.56) --
	(251.84, 48.46) --
	(251.84, 48.46) --
	(251.81, 48.75) --
	(251.75, 49.03) --
	(251.63, 49.30) --
	(251.48, 49.54) --
	(251.28, 49.76) --
	(251.06, 49.95) --
	(250.81, 50.09) --
	(250.53, 50.19) --
	(250.25, 50.25) --
	(250.03, 50.27) --
	(197.88, 50.27) --
	(198.10, 50.25) --
	(197.81, 50.26) --
	(197.52, 50.23) --
	(197.24, 50.15) --
	(196.97, 50.02) --
	(196.74, 49.86) --
	(196.53, 49.66) --
	(196.35, 49.42) --
	(196.22, 49.17) --
	(196.12, 48.89) --
	(196.08, 48.60) --
	(196.07, 48.46) --
	(196.07, 36.56) --
	(196.08, 36.71) --
	(196.08, 36.42) --
	(196.12, 36.13) --
	(196.22, 35.85) --
	(196.35, 35.60) --
	(196.53, 35.36) --
	(196.74, 35.16) --
	(196.97, 35.00) --
	(197.24, 34.87) --
	(197.52, 34.79) --
	(197.81, 34.76) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] ( 40.89, 26.85) rectangle (265.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (223.95, 39.76) {129 (66, 214)};
\definecolor{drawColor}{RGB}{0,60,97}

\path[draw=drawColor,line width= 1.6pt,line join=round] (156.42,144.98) --
	(156.42,148.90);

\path[draw=drawColor,line width= 1.6pt,line join=round] (156.42,146.94) --
	( 84.79,146.94);

\path[draw=drawColor,line width= 1.6pt,line join=round] ( 84.79,144.98) --
	( 84.79,148.90);

\path[draw=drawColor,line width= 1.6pt,line join=round] (164.09,118.88) --
	(164.09,122.79);

\path[draw=drawColor,line width= 1.6pt,line join=round] (164.09,120.83) --
	( 88.92,120.83);

\path[draw=drawColor,line width= 1.6pt,line join=round] ( 88.92,118.88) --
	( 88.92,122.79);

\path[draw=drawColor,line width= 1.6pt,line join=round] (169.20, 92.77) --
	(169.20, 96.68);

\path[draw=drawColor,line width= 1.6pt,line join=round] (169.20, 94.73) --
	( 92.17, 94.73);

\path[draw=drawColor,line width= 1.6pt,line join=round] ( 92.17, 92.77) --
	( 92.17, 96.68);

\path[draw=drawColor,line width= 1.6pt,line join=round] (175.75, 66.66) --
	(175.75, 70.58);

\path[draw=drawColor,line width= 1.6pt,line join=round] (175.75, 68.62) --
	( 93.91, 68.62);

\path[draw=drawColor,line width= 1.6pt,line join=round] ( 93.91, 66.66) --
	( 93.91, 70.58);

\path[draw=drawColor,line width= 1.6pt,line join=round] (189.54, 40.55) --
	(189.54, 44.47);

\path[draw=drawColor,line width= 1.6pt,line join=round] (189.54, 42.51) --
	( 98.17, 42.51);

\path[draw=drawColor,line width= 1.6pt,line join=round] ( 98.17, 40.55) --
	( 98.17, 44.47);
\definecolor{fillColor}{RGB}{0,60,97}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (117.65,146.94) circle (  1.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (122.54,120.83) circle (  1.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (126.15, 94.73) circle (  1.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (129.92, 68.62) circle (  1.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (136.90, 42.51) circle (  1.86);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,draw opacity=0.70,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 57.26, 26.85) -- ( 57.26,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.9pt,line join=round,line cap=round] ( 40.89, 26.85) rectangle (265.01,162.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94, 39.41) {0.05};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94, 65.52) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94, 91.63) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94,117.74) {1.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 35.94,143.84) {2.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,162.61);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14, 42.51) --
	( 40.89, 42.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14, 68.62) --
	( 40.89, 68.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14, 94.73) --
	( 40.89, 94.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14,120.83) --
	( 40.89,120.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.14,146.94) --
	( 40.89,146.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,162.61);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 57.26, 24.10) --
	( 57.26, 26.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (119.00, 24.10) --
	(119.00, 26.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (180.74, 24.10) --
	(180.74, 26.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (242.48, 24.10) --
	(242.48, 26.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 57.26, 15.70) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (119.00, 15.70) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (180.74, 15.70) {200};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (242.48, 15.70) {300};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (152.95,  1.75) {Cases averted};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (271.01,162.61);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 12.20, 94.73) {Target limit (mg/m$^3$)};
\end{scope}