### 2023-06-12 reports/paper/paper.qmd

---
# fontsize: 11pt
# mainfont: arial
format:
 # html: default
 docx:
    number-sections: false
    global-numbering: true
    reference-doc: "../../../../HeadRs/rmd_templates/style-guide-arial-11.docx"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
notes-after-punctuation: true
execute:
  freeze: auto
editor_options: 
  chunk_output_type: inline
editor: 
  markdown: 
    wrap: 72
---

```{r, include=FALSE}
library(here)
knitr::opts_chunk$set(echo=F, results = 'asis')
library(bibtex)
biblio <- bibtex::read.bib('~/Box Sync/papers/EnvEpi.bib')
library(flextable)
library(data.table)
library(ggplot2); library(magrittr)
```

**Type of manuscript:** Original Research Article

**Manuscript title:** Hypothetical limits on workplace exposure for
reducing non-Hodgkin lymphoma: An application of the hazard-extended
iterative conditional expectation parametric g-formula

**Authors:** Kevin T. Chen^a^, Sally Picciotto^b^, Ellen A. Eisen^b^

^a^ University of California, Berkeley School of Public Health, Division
of Epidemiology & Biostatistics

**Corresponding author:**\
Kevin T. Chen\
2121 Berkeley Way\
Room 5302, Desk 5305-1\
Berkeley, CA 94720-7360\
kevchen\@berkeley.edu\
(510) 387-3327

**Running head suggestion:** Limits on metalworking fluid exposure to
prevent NHL.

**Conflicts of interests:** Authors declare no conflicts of interest.

**Sources of financial support:** This work was supported by Training
Grant T42OH008429 and Research Project Grant R01OH011092, both funded by
the National Institute for Occupational Safety and Health (NIOSH) /
Centers for Disease Control and Prevention (CDC).

**Data and computing code:** Code for reproducing analyses and reports
are available on [Github](https://github.com/kvntchn/gm-nhl-ice). Data
are available upon reasonable request.

**Keywords:** Metalworking fluid, occupational exposure, non-Hodgkin
lymphoma, cancer, causal inference, g-formula, healthy-worker effect,
United States

\newpage

# Abstract

(Limit: 250 words)

**Background:** Non-Hodgkin lymphoma (NHL) incidence has increased
substantially in the US since 1960 and was recently linked to workplace
exposure to soluble metalworking fluids (MWF) in a standard survival
analysis of the United Auto Workers-General Motors (UAW-GM) cohort. To
further explore this association, we estimate the effects of
hypothetical limits on soluble MWF exposure in relation to NHL risk in
the same cohort of autoworkers.

**Methods:** We estimated counterfactual risk of NHL between 1985 and
2015 in the UAW-GM cohort of autoworkers at three Michigan plants (n =
33,134) under hypothetical target exposure limits for soluble MWF. We
applied the hazard-extended iterative conditional expectation parametric
g-formula, an estimator capable of adjusting for time-varying
confounding affected by past exposure.

**Results:** 339 NHL cases occurred over the 30-year follow-up period.
Stronger target limits on soluble MWF exposure resulted in monotonic
reductions in NHL risk. Setting the target exposure limit at 2 mg/m^3^
would have prevented 98 (95% CI: 45, 161) cases of NHL. A target
exposure limit of 0.5 mg/m^3^, the NIOSH recommended exposure limit,
would have prevented 112 (95% CI: 57, 181) cases. Finally, a target
exposure limit of 0.05 mg/m^3^ would have prevented 129 (95% CI: 66,
214) cases.

**Conclusion:** Limits on soluble MWF exposure would have prevented
cases of NHL in the UAW-GM cohort. Although MWF use is expected to
decline as the automotive industry turns to electric vehicles, a boom in
domestic vehicle manufacturing could nonetheless lead to global
increases in worker exposure to MWF in the US. Maintaining or
strengthening exposure limits for MWF may protect workers from NHL.

\newpage

# Introduction

Non-Hodgkin Lymphoma (NHL) incidence in the United States doubled
between 1973 and 1994, before plateauing at around 19 per 100,000
persons per year, making it the seventh most common cancer in the
country [@seer_1994; @Ekstrom-Smedby_2006]. The strongest known risk
factor of NHL is immunosuppression, both congenital and acquired
[@Filipovich_1992; @Chiu_2015]. Modulation of immune function
contributes to the pathogenesis of NHL directly by perturbing lymphocyte
function or indirectly by reducing immunosurveillance against neoplasms.
Immunodeficiency may also interact with infection with Epstein-Barr
Virus, resulting in uncontrolled B-cell proliferation [@Oertel_2002].
Acquired immunodeficiency through infection with human immunodeficiency
virus (HIV) can lead to genetic lesions and dysregulation while
increasing vulnerability to infection by other oncogenic agents
[@Grulich_1999; @Alexander_2007]. However, since immunodeficiency and
infection with HIV are rare, they cannot fully explain the historic rise
or present burden of NHL [@Shiels_2013]. The rise in NHL incidence
coincided with a period of rapid and extensive chemicalization in
industry, agriculture and warfare; environmental and occupational
exposures may play an important explanatory role in the epidemiology of
NHL [@Nelson_2005; @Romero_2021].

Pesticide exposure among workers in agricultural settings was a common
target of NHL research in recent decades. A meta-analysis of 44 articles
published between 1980 and 2014 found statistically significant
associations between NHL and exposure to several classes of pesticides
including carbamate, organophosporus, triazine, and organochlorine
[@Schinasi_2014]. Occupational exposures associated with NHL are not
limited to the agricultural sector, however. Occupational groups
associated with NHL risk also include metal processors, health workers,
salespeople, machinists, and electricians [@Fritschi_2005a;
@Mester_2006; @Ekstrom-Smedby_2006]. Workers in these occupational
groups often come into contact with industrial chemicals such as
gasoline, solvents, coolants, and lubricants such as metalworking fluids
(MWF).

Metalworking fluids are complex mixtures of oil, water, and chemical
additives that optimize metal machining operations. During shaping,
grinding, and cutting operations, MWFs are misted, poured, or blasted at
high pressure onto work surfaces to remove debris, cool metal, improve
efficiency, and prevent deterioration of tools. Although MWFs are
essential to manufacturing processes, they also present a potential
health hazard to exposed workers through inhalation or ingestion of MWF
particulate mass.

There are three general types of MWF: straight, soluble, and synthetic.
The type of MWF applied has changed over the decades in response to
engineering, economic, and health considerations. Straight MWF, the
oldest class of MWF, is comprised mostly of mineral oils, which were
classified as carcinogenic by the International Agency for Research on
Cancer (IARC) in 1973, citing studies of occupational exposures among
workers in oil, textile, and metal industries [@IARC_1973]. In the same
decade, additional factors including the high cost of oil and the
introduction of performance-enhancing water-soluble chemical additives
spurred the gradual replacement of straight MWF with water-based soluble
and synthetic MWF for many, but not all metalworking operations
[@Childers_2005]. Although water-based MWFs contain little to no mineral
oil, they contain a richer cocktail of additives including chromates,
cycloalkanes, phenols, organochlorines, nitrites, sulfonates, and
triazines: classes of chemicals with known or plausible carcinogenicity
[@IARC_1987; @IARC_2014; @IARC_2017]. As the US prepares for a boom in
domestic automotive manufacturing, occupational exposure to MWF will
remain an important concern for worker health [@Armbrust_2021;
@Carr_2022; @Keilman_2023].

One challenge in estimating the causal effects of occupational exposures
on worker health is the Healthy Worker Survivor Effect (HWSE), the
process by which healthier individuals remain at work where they
accumulate more exposure while those more susceptible to the deleterious
health effects leave work [@Arrighi_1994]. A well-known analytical
method for causal inference in observational studies where the HWSE or
other forms of time-varying confounding affected by past exposure may be
operating is the parametric g-formula [@Robins_1986; @Taubman_2009;
@Richardson_2013]. A central requirement necessary for causal
interpretation of any analysis is positivity (overlap) ie adequate
variation in the exposure of interest within strata formed by confounder
and exposure history [@Petersen_2012]. Here, we specify hypothetical
interventions on exposure to soluble MWF in the United Auto
Workers-General Motors (UAW-GM) occupational cohort whose effects can be
estimated with guaranteed positivity. In particular, we estimate the
cumulative incidence of NHL between 1985 and 2015 under well-supported
(supportable) interventions by applying the the hazard-extended
iterative conditional expectation (ICE) parametric g-formula
[@Arrighi_1994; @Wen_2021].

<!-- The parametric g-formula has become increasingly popular among epidemiologists in the last decade, and there are several exemplary tutorial papers and illustrative examples [@Taubman_2009;@Breskin_2020]. This analytic approach is attractive in part because it allows investigators to estimate the population-level effect of realistic interventions on exposures that may depend on individuals' covariate history as well as their natural value of exposure. In addition to yielding estimates with a population-level interpretation, the parametric g-formula is also capable of adjusting for time-varying confounding possibly affected by past exposure in longitudinal settings, which is not possible with standard regression methods. In their seminal 2009 paper, Taubman et alia describe the parametric g-formula and illustrate its application in the estimation of the counterfactual risks under several interventions in the Nurses' Health Study [@Taubman_2009]. They introduced a non-iterative conditional expectation parametric g-formula that requires parametric specification of the full joint distribution of confounders, exposures, and outcomes over time. -->

<!-- Full parameterization is an attractive choice for investigators seeking to embed real causal knowledge into the estimation problem. For those who are not completely comfortable with the full set of parametric assumptions, the iterative conditional expectation (ICE) parametric g-formula may be a good alternative. The ICE g-formula was introduced as early as 2004 [@Robins_2004], but its parametric implementations remain far less common in practice than its non-ICE counterpart. Like the non-ICE parametric g-formula, the ICE expression enables adjustment for time-varying confounding possibly affected by past exposure. However, the ICE parametric g-formula requires the parametric specification of models for the outcome only, and not the confounders or exposure. Hence, it requires fewer parametric assumptions than the non-ICE parametric g-formula. Furthermore, the elegance of the ICE parametric g-formula allows analysts to implement the method using standard statistical software, which offers the prospect of greater interpretability and transparency than using "black box" implementations. -->

<!-- In recent years, there has been a surge in optimism for growth in American manufacturing [@Carr_2022;@Armbrust_2021]. While it may be easy to imagine a factory floor crowded with workers and heavy machinery, an essential part of the industrial production of physical goods is easy to overlook: lubricants and coolants. In industries that require metalworking, from cutting wafers for semiconductor production to stamping metals for electric vehicle battery casings, metalworking fluids (MWF) are essential. During use in drawing, stamping, cutting, and grinding operations, MWF are sprayed and poured onto tools and parts to provide lubrication and cooling. Parts may even be submerged in MWF in precision manufacturing applications [@Arrabiyeh_2020]. -->

# Methods

We estimated NHL cumulative incidence from 1985 to 2015 under
supportable interventions on average daily exposure to soluble MWF by
applying the hazard-extended ICE parametric g-formula estimator
[@Wen_2020]. First, we estimated the expected number of NHL cases that
we would observe if there were no censoring by competing risks. Then, we
contrasted this case count to that under supportable interventions based
on five hypothetical target exposure limits and no competing risks. The
five target exposure limits were (1) 2.0, (2) 1.0, (3) 0.5, (4) 0.25,
and (5) 0.05 mg/m^3^. The National Institute for Occupational Safety and
Health (NIOSH) recommended exposure limit (REL) for total particulate
mass (PM) composed of MWF is 0.5 mg/m^3^ [@niosh_1998]. For each target
exposure limit, period of follow-up, and unique combination of
confounder and exposure histories, we also defined a supportable
exposure limit, which was the maximum observed value of exposure at or
below the target exposure limit or no limit when all of the observed
values were above the target limit. For each target exposure limit, we
specified an intervention rule that reduced exposures above the
supportable exposure limit to that limit, but allowed exposures at or
below the limit to vary according to their natural distribution.
Applying this intervention rule to the observed distribution of exposure
produced the intervention distribution that defined the stochastic
dynamic intervention with guaranteed positivity.

## Study population

We used data from the UAW-GM cohort, which included all hourly workers
at three automobile manufacturing plants in Michigan who had worked at
least three years by 1985. Past papers provide detailed descriptions of
the cohort [@Eisen_1992; @Eisen_2001]. The large size of the study
population and rich time-varying, quantitative MWF exposure data provide
an opportunity to study this relatively rare cancer and evaluate
realistic interventions on MWF exposure in a longitudinal cohort
setting. The present study population (N = 33,134) was restricted to the
autoworkers who were at work in 1941 or not yet hired, missing no more
than half of their employment history, and still alive at the start of
follow-up. Autoworkers in the study population were followed for NHL
incidence from January 1, 1985 until NHL diagnosis, death, December 31,
2014 (30 years after the end of employment record availability), or the
oldest observed age at death (108 years), whichever came earlier.

## Outcome and covariates

We identified incident cancers in the UAW-GM cohort between 1985 and
2014 by linkage to the Michigan Cancer Registry (MCR). Workers at Plants
1 and 2, located in the greater Detroit metropolitan area, were also
linked to the Detroit Regional Registry of the Surveillance,
Epidemiology, and End Results (SEER) Program. Cancer types were
distinguished using site and histology codes conforming to the
International classification of Diseases for Oncology, 3^rd^ edition
(ICD-O-3). Non-Hodgkin lymphoma was defined by cancers with any of the
following ICD-O-3 histology codes: 9590-9597, 9670-9671, 9673, 9675,
9678-9680, 9684, 9687-9691, 9695, 9698-9702, 9705, 9708-9709, 9712,
9714-9719, 9724-9729, 9735, 9737-9738, 9811-9818, 9823, 9827, 9837.
Details regarding cancer incidence follow-up are described elsewhere
[@Colbeth_2023]. Vital status was ascertained from company records and
by linkage to Social Security Administration, National Death Index, and
state mortality files.

Potential confounders including year of hire, sex, race, time off,
employment status, and plant location were obtained from company
records. Race was missing for about 16% of the cohort, most commonly
among workers hired before 1960 in Plant 2. In analyses, missing race
was considered a distinct category. All potential confounders were coded
as categorical variables. Cut-points for categorizing continuous
covariates were determined according to the quantiles among cases.

## Exposure

Quantitative measurement of time-varying MWF exposure is a
distinguishing strength of the UAW-GM cohort study relative to other
occupational cohort studies. Exposure assessment was based on direct air
sampling as well as company records. Company industrial hygienists
collected several hundred personal and area samples for total
particulate matter (mg/m^3^) composed of MWF over many decades. Research
industrial hygienists collected additional air sampling data when the
cohort study was launched in the mid 1980s. These additional data
combined with historical data and company records constituted a
job-exposure matrix of quantitative 8-hour time-weighted average daily
exposure estimates to soluble, straight, and synthetic MWF for each
combination of job, department, and plant over time. Workers' average
daily exposure to each MWF type was determined by combining this
job-exposure matrix with employment records, which recorded time-varying
job type, department, and plant for each employee from hire to
termination or 1995, whichever came sooner. For employment records that
were at least half complete, gaps in the record were interpolated by
carrying forward the last known job type. The exposure assessment is
described in detail elsewhere [@Hallock_1994; @Woskie_1994;
@Woskie_2003].

Previous analyses of NHL applied exposure lags of 1 to 20 years to
account for disease latency; we lagged cumulative MWF exposures by 20
years and therefore ended follow-up on January 1, 2015 [@Smith_2007;
@Karipidis_2007; @Zhang_2019]. In analyses, MWF exposure history at
start of follow-up was summarized as the cumulative sum of annual
exposure intensities. Exposure was coded as categorical variables with
cut-points at zero and the quintiles of nonzero exposure among cases. We
estimated the effects of interventions on soluble MWF, the type of MWF
used most widely and in the greatest quantities. We treated exposure to
straight and synthetic MWF as potential confounders.

## Hazard-extended ICE parametric g-formula

We split the 30-year follow-up period into ten time periods based on the
deciles of the dates of diagnosis of NHL. The number of years per period
ranged from two to four years. The hazard-extended ICE parametric
g-formula involved two stages. In the first, we estimated counterfactual
discrete hazards over the person-periods. In the second, we pooled those
estimates to estimate the counterfactual risk over the entire follow-up
period. During pooling, we iteratively combined estimates of the
counterfactual hazard starting with the last two periods, then the last
three, then the last four, and so forth. In each iteration, we performed
model-based standardization over exposure and covariate histories before
combining the counterfactual discrete hazard with the risk over
subsequent periods. This iterative process resulted in a sequentially
standardized estimate of the counterfactual cumulative incidence of NHL
when the intervention of interest is enforced over all follow-up
periods. We took exposure to straight and synthetic MWF, employment
status, cumulative time off, age, duration of employment, sex
(male/female), race (Black/white/unknown), and plant (Plant 1/Plant
2/Plant 3) to be potential confounders. Exposure to MWF, employment
status, time off, and duration of employment were lagged 20 years. An
overview of the general steps of the estimation procedure are presented
below. Note that we refer to discrete hazards simply as hazards.

1.  Fit a pooled logistic regression of NHL on covariates and exposure
    over all at-risk person-periods.

2.  Predict the hazard for each person-period for each possible level of
    exposure using the model from step 1.

3.  Within strata formed by unique combinations of covariates, obtain
    the intervention distribution of exposure by applying the
    intervention rule to the observed exposure distribution.

4.  Within strata, estimate the counterfactual hazard for each
    person-period by taking a weighted sum of the predicted hazards from
    step 2.

5.  Starting at the penultimate period of follow-up, estimate the pooled
    counterfactual hazard:

    a)  Regress the counterfactual hazard pooled over all subsequent
        periods on exposure and covariate history at the current period.
    b)  Predict the pooled hazard for each person at risk at the start
        of the period and each possible level of exposure using the
        model from (a). Retrieve the predicted hazards from step 2. For
        each person and level of exposure, multiply the predicted pooled
        hazard by one minus the corresponding predicted hazard and add
        the predicted hazard to the product.
    c)  Within strata formed by exposure and covariate history, obtain
        the intervention distribution of exposure by applying the
        intervention rule to the observed exposure distribution.
    d)  Within strata, estimate the counterfactual hazard pooled over
        the present and subsequent periods for each person by taking a
        weighted sum of the predicted pooled hazards from (b).
    e)  If the present period is not first period, set the reference
        period to the preceding period and return to (a).

6.  Compute the counterfactual risk by averaging the counterfactual
    hazard over the entire follow-up period across all persons.

To account for censoring, fit the models in step 1 and step 5a among
those who were not censored and obtain predicted hazards for all at-risk
person-periods, including those that were censored.

We estimated cumulative incidence under the observed distribution of
soluble MWF exposure (natural course) and under the five supportable
exposure limits. We contrasted the cumulative incidence under
supportable limits to that under the natural course by computing the
number of cases averted and the cumulative incidence ratios. Confidence
intervals were computed using the basic nonparametric bootstrap with
1000 Monte Carlo samples from the population at the start of follow-up.
All the necessary script used to reproduce the analyses are available on
[GitHub](https://github.com/kvntchn/gm-nhl-ice).

# Results

```{r}
source(here::here("../gm-wrangling/wrangling/table1.R"))
clean.tab1 <- function(tab1) {
  # hspace.which <- grepl("\\\\hspace\\{.*\\}", rownames(tab1))
  rownames(tab1) <- gsub("\\\\hspace\\{.*\\}|\\\\hline |\\$.*\\$| among deceased", "", rownames(tab1))

  tab1[!grepl("median", tab1[,4]) & !is.na(tab1[,1]), 2]  <-  paste0(
    "(",
    tab1[!grepl("median", tab1[,4]) & !is.na(tab1[,1]), 2],
    ")"
  )

  tab1[grepl("median", tab1[,4]) & !is.na(tab1[,1]), 2] <- paste0(
    "(", tab1[grepl("median", tab1[,4]) & !is.na(tab1[,1]), 2], ", ",
    tab1[grepl("median", tab1[,4]) & !is.na(tab1[,1]), 3],
    ")"
  )

  return(tab1[,1:2])
}
full.tab1 <- readRDS(here::here("resources", "tab1.rds"))
nhl.tab1 <- readRDS(here::here("resources", "nhl.tab1.rds"))

tab1.clean <- clean.tab1(full.tab1)
nhl.tab1.clean <- clean.tab1(nhl.tab1)

tab1 <- cbind(hspace.logical = grepl("\\\\hspace\\{.*\\}", rownames(full.tab1)),
              variable = rownames(tab1.clean), tab1.clean, NA, nhl.tab1.clean)

tab1 <- tab1[!grepl("Years of follow-up", rownames(tab1)),]
tab1$variable[1] <- "N (person-years)"
```

Table 1 presents summary statistics of exposure and covariates for the
full study population and for those diagnosed with NHL between 1985 and
2005. The cohort is predominantly white (66%) and male (87%). The median
year of hire among those diagnosed with NHL was 1959 whereas the median
year of hire in the full study population was almost a decade later. Age
at hire was approximately the same among those with NHL and the full
study population. Median lagged cumulative exposure to all three MWF
types was higher among NHL cases. Soluble MWF were the most widely used
MWF type, with approximately 90% of workers ever exposed. Median
cumulative exposure among the exposed was 6.5 times higher for soluble
than for straight MWF. Figure 1 shows median annual average daily
exposure to the three MWF types among exposed workers over calendar
time. Exposure to MWF generally followed a downward trend over time.

```{r}
# tab1 <- tab1[tab1.which,]
tab1[c(1:13, 15, 14, 16:27, 29, 28, 30),-grep("hspace.logical", colnames(tab1))] |> 
  data.frame() |> flextable() |>
  delete_part(part = "header") |>
  add_header_row(values = c("", "Study population", "", "NHL cases"),
                 colwidths = c(1, 2, 1, 2)) |>
  hline_top(part = "header") |>
  hline_top() |> hline_bottom() |>
  hline(which((grepl("[0-9]\\)", tab1$spread))[-1])[1]) |> width(4, 0.1) |>
  width(1, 2.2) |> width(c(3, 6), 1.1) |> width(c(2, 5), 0.65) |>
  flextable::compose(
    i = grep("Cumulative exposure", rownames(tab1)), j = 1,
    value = as_paragraph("Cumulative exposure to MWF (mg/m",
                         as_sup("3"), "-years)")) |>
  flextable::align(j = 2:6, align = "center", part = "all") |>
  footnote(i = grep("Plant$", rownames(tab1)), j = 1, value = as_paragraph(
    " Plant of longest employment duration among those who worked at multiple plants"),
    ref_symbols = "a", inline = T) |>
  footnote(i = grep("Ever exposed|time off", rownames(tab1)), j = 1, value = as_paragraph(
    " Lagged 10 years"),
    ref_symbols = "b", inline = T) |>
  footnote(i = grep("work", rownames(tab1)), j = 1, value = as_paragraph(
    " Among those who left work by December 31, 1994"),
    ref_symbols = "c", inline = T) |>
  # footnote(i = grep("death", rownames(tab1)), j = 1, value = as_paragraph(
  #   " Among deceased"),
  #   ref_symbols = "d", inline = T) |>
  footnote(i = grep("Cumulative exposure", rownames(tab1)), j = 1, value = as_paragraph(
    " Among ever-exposed individuals, lagged 10 years."),
    ref_symbols = "d", inline = T) |>
  add_footer_row(values = "NHL: non-Hodgkin lymphoma.", colwidths = 6) |>
  padding(padding = 1.5, part = "all") |>
  padding(
    i = which(tab1$hspace.logical),
    j = 1,
    padding.left = 20
  ) |>
  merge_at(i = grep("Cumulative exposure", rownames(tab1))) |>
  hline_bottom(part = "footer") |> font(fontname = 'arial') |>
  set_caption(caption = "Summary of population characteristics. Statistics shown above the horizontal line are count (%). Those shown below are median (quartile 1, quartile 3).")
```

\

![Figure 1. Median annual average daily exposure to soluble, straight,
and synthetic metalworking fluids among exposed workers over
time.](../../resources/images/exposure.png){width="6.5in"}

```{r}
load(here::here("resources", paste0("estimates.rdata")))
estimates.bs <- readRDS(here::here("resources", paste0("estimates.bs.rds")))
```

```{r}
# Risk ratios
do.call(cbind, lapply(estimates.bs[,-(1:2)], function(x) {
  x/estimates.bs[,2]
})) %>% melt(measure.vars = names(.)) -> rr.bs

# Risk differences
do.call(cbind, lapply(estimates.bs[,-(1:2)], function(x) {
  x - estimates.bs[,2]
})) %>% melt(measure.vars = names(.)) -> rd.bs

estimates.bs[,-1] %>% melt(measure.vars = names(.)) -> risk.bs

# Collect point estimates
point_estimates <- data.table(
  variable = c(
    "200", "100", "50", "25", "05", "Any"),
  estimate =  c(sapply(estimates[2:6], mean),
                mean(estimates[[1]]))
)
```

```{r}
# Calculate risk ratios for each bootstrap sample
rr <- merge(
  rr.bs,
  point_estimates[,.(
    variable = paste0(variable, ".Any"),
    estimate = estimate/estimate[variable == "Any"]
  )],
  all.x = F,
  by = "variable")

# Get lower and upper 2.5th %tiles
rr <- rr[,.(
  rr = estimate[1],
  rr_lower = 2 * estimate[1] - quantile(value, 1 - 0.025),
  rr_upper = 2 * estimate[1] - quantile(value, 0.025)
  ),
variable]
```

```{r}
# Calculate risk differences for each bootstrap sample
rd <- merge(
  rd.bs,
  point_estimates[,.(
    variable = paste0(variable, ".Any"),
    estimate = estimate - estimate[variable == "Any"]
  )],
  all.x = F,
  by = "variable")

# Get lower and upper 2.5th %tiles
N <- 33134
rd <- rd[,.(
  rd = - estimate[1],
  rd_lower = - (2 * estimate[1] - quantile(value, 0.025)) * N,
  rd_upper = - (2 * estimate[1] - quantile(value, 1 - 0.025)) * N
  # wald.lower = exp(log(estimate[1]) + qnorm(0.025) * sd(log(value))),
  # wald.upper = exp(log(estimate[1]) - qnorm(0.025) * sd(log(value)))
),
variable]
```

```{r}
# Combine differences and ratios
contrasts.tab <- merge(rr, rd, by = 'variable')
```

```{r}
# Collect risks
risks <- merge(
  risk.bs,
  point_estimates,
  all.x = F,
  by = "variable")

# Get lower and upper 2.5th %tiles
risks <- risks[,.(estimate = estimate[1],
      lower = 2 * mean(value) - quantile(value, 1 - 0.025),
      upper = 2 * mean(value) - quantile(value, 0.025)
      # wald.lower = exp(log(estimate[1]) + qnorm(0.025) * sd(log(value))),
      # wald.upper = exp(log(estimate[1]) - qnorm(0.025) * sd(log(value)))
      ),
   .(variable.risks = variable)]
```

```{r}
intervention.labels <- data.table(
  variable = c(
    "200", "100", "50", "25", "05", "Any"
  ),
  label = c(
    "2.0", "1.0", "0.5", "0.25", "0.05", "None"
  )
)

# Add intervention labels
contrasts.tab <- merge(
  contrasts.tab, intervention.labels[,.(variable = paste0(variable, ".Any"), label)],
  all = T,
  by = "variable"
)

```

```{r}
clean_percent_py <- function(variable, digits = 1, scale = 100) {
  sprintf(paste0("%.", digits, "f"), round(
    intervened_py[match(
       sapply(variable, gsub, pattern = ".Any", replacement = ""),
       variable.dat
  ), percent_py_intervened] * scale,
  digits))
}

clean_risk <- function(variable, digits = 2, scale = 1000) {
  sprintf(paste0("%.", digits, "f"), round(risks[match(
       sapply(variable, gsub, pattern = ".Any", replacement = ""),
       variable.risks
  ), estimate] * scale, digits))
}

clean_risk.ci <- function(variable, digits = 2, scale = 1000) {
       x <- risks[match(
       sapply(variable, gsub, pattern = ".Any", replacement = ""),
       variable.risks
       ),.(lower, upper)]
       ifelse(!is.na(x$lower), paste0(
         "(",
         sprintf(paste0("%.", digits, "f"), round(x$lower * scale, digits)),
         ", ", 
         sprintf(paste0("%.", digits, "f"), round(x$upper * scale, digits)),
         ")"), NA)
}

clean_rr.ci <- function(estimate, lower, upper, digits = 2) {
  ifelse(!is.na(lower), paste0(
    "(",
    sprintf(paste0("%.", digits, "f"), round(lower, digits)),
    ", ",
    sprintf(paste0("%.", digits, "f"), round(upper, digits)),
    ")"), NA)
}

mwf.flextable <- function(x, n_col = 7) {
  flextable::compose(x, j = c(3, n_col), value = as_paragraph(rep("(95% CI)", 2)),
    part = "header"
  ) |>
  border_remove() |>
  hline_top(part = "header") |>
  hline_top() |> hline_bottom() |>
  flextable::align(j = 2:n_col, align = "center", part = "all") |>
  padding(padding = 1.5, part = "all")
}
```

The observed number of NHL cases over the 30-year follow-up period was
339. Table 2 presents the hazard-extended ICE parametric g-formula
estimates of the counterfactual number of cases, number of cases
averted, and cumulative incidence ratios contrasting hypothetical
interventions on exposure to soluble MWF based on different target
exposure limits and no competing risks. Under an intervention
eliminating competing risks, the estimated number of cases was 502 (95%
CI: 439, 555). Interventions based on stronger limits on average daily
exposure to soluble MWF resulted in monotonically stronger reductions in
the estimated cumulative incidence of NHL. Setting the target exposure
limit at the NIOSH REL 0.5 mg/m^3^ would have averted 112 (95% CI: 57,
181) NHL cases.

```{r, eval=F}
tikzDevice::tikz(here::here("reports/private/SER (2023)/resources/rd.tex"),
                 width = 3.75, height = 2.25,
								 documentDeclaration = "\\documentclass{beamer}",
                 bareBones = T)
contrasts.tab[,.(
     label,
     # `Person-years intervened (%)` = clean_percent_py(variable),
     `Cases averted`    = rd * N,
     rd_lower,
     rd_upper,
     result = paste0(
     	round(rd * N),
     	" (", round(rd_lower), ", ", round(rd_upper), ")")
   )][c(6, 3, 2, 5, 4, 1),] -> gg.tab 
gg.tab[label != "None"] |> ggplot(
    aes(x = factor(
    	label,
    	levels = c("0.05", "0.25", "0.5", "1.0", "2.0",  "None"),
    	labels = c("0.05", "0.25", "0.50", "1.00", "2.00",  "None")),
        y = `Cases averted`, ymin = rd_lower, ymax = rd_upper)
  ) +
	geom_label(aes(y = 270, label = result), size = 8/.pt, 
						 label.padding = unit(5, "pt")) +
  geom_errorbar(linewidth = 4 / .pt, width = 0.15, color = '#003c61') +
  geom_point(size = 4 / .pt, color = '#003c61') + 
  geom_hline(yintercept = 0, lty = 2, , alpha = 0.7) +
  labs(
    y = "Cases averted",
    x = "Target limit (mg/m$^3$)") +
  mytheme +
  coord_flip(ylim = c(-10, 320)) +
  theme(panel.grid = element_blank(),
  			axis.text = element_text(size = 9, colour = 'black'),
  			panel.border = element_rect(colour = "black", size = 0.75),
  			axis.title.y = element_text(size = 9, margin = margin(r = 6)),
  			axis.title.x = element_text(size = 9, margin = margin(t = 6)),
  			plot.margin = margin(l = 6, r = 6)
  )
dev.off()
```

```{r}
contrasts.tab[,.(
     label,
     # `Person-years intervened (%)` = clean_percent_py(variable),
     `Cases`  = clean_risk(variable, 0, N),
     `(95% CI)`      = clean_risk.ci(variable, 0, N),
     `Cases averted`    = round(rd * N),
     `(95% CI1)`        = paste0(
       "(",
       round(rd_lower), ", ",
       round(rd_upper), ")"),
     CIR              = sprintf("%.2f", round(rr, 2)),
     `(95% CI2)`   = clean_rr.ci(rr, rr_lower, rr_upper)
   )][c(6, 3, 2, 5, 4, 1),] -> sol.tab
sol.tab[label == "None", 4:7] <- NA

# save(sol.tab, mwf.flextable, file = here::here('resources', 'sol.tab.rdata'))

sol.tab |> flextable() |>
  flextable::compose(
    j = 1,
    value = as_paragraph("Exposure limit for soluble MWF (mg/m", as_sup("3"), ")"),
    part = "header"
  ) |>
  set_header_labels(
    `(95% CI1)` = "(95% CI)",
    `(95% CI2)` = "(95% CI)"
  ) |> mwf.flextable() |>
  width(width = c(1.1, 0.5, 0.9, 0.55, 0.9, 0.5, 0.9)) |>
  add_footer_row(values = paste(
    "CIR: cumulative incidence ratio. MWF: metalworking fluid",
    sep = "\n"), colwidths = 7) |>
  hline_bottom(part = "footer") |>
  set_caption(caption = "Counterfactual number of cases, number of cases averted, and cumulative incidence ratio estimates contrasting hypothetical limits on soluble MWF exposure to no intervention on exposure.")

```

# Discussion

Although NIOSH concluded that there exists substantial evidence linking
MWF exposure to several different cancers, their REL of 0.5 mg/m^3^ for
total particulate matter derived from any type of MWF targeted risk of
respiratory disorders rather than cancer [@niosh_1998; @Mirer_2003].
Exposure limits stronger than the NIOSH REL may provide additional
health protections not previously considered by policy makers. Using the
hazard-extended ICE parametric g-formula, We estimated the
counterfactual expected number of NHL cases from 1985 to 2015 in the
UAW-GM cohort if we enforced exposure limits based on five different
target exposure limits and found a monotonic exposure-dependent
relationship where stronger target exposure limits yielded lower NHL
case count estimates.

The g-formula is a well-known approach in causal inference used for the
identification and estimation of causal effects in the presence of
time-varying confounding affected by past exposure [@Robins_1986].
Standard representations of the g-formula include (1) a non-iterated
expectation over the joint density of covariates, (2) the ICE over time,
and (3) an inverse probability weighted expectation. The parametric
g-formula is a plug in estimator for the g-formula under its first,
non-iterative, representation. It involves the parametric modeling of
the full joint distribution of the outcome, exposure, and covariates
under study for each time point [@Taubman_2009; @Young_2014].
Counterfactual quantities under hypothetical interventions of interest
are computed from Monte Carlo samples from distributions implied by the
fitted parametric models. In longitudinal settings, this approach often
requires specifying and fitting large number of models in order to
satisfy the exchangeability assumptions necessary for causal
identification.
<!-- Since applications of the parametric g-formula are often context-laden, there are no unified guidelines for model specification. -->

Parametric estimators using the ICE representation of the g-formula
require modeling only the conditional outcome distributions, so they
require fewer parametric assumptions than non-iterative parametric
g-formula estimators. Counterfactual outcome estimates over the
follow-up period are computed from interval-specific conditional
estimates by applying the tower rule of expectation. Under the
assumptions of conditional exchangeability at all time points,
positivity, counterfactual consistency, and correct model specification,
the hazard-extended parametric g-formula yields unbiased estimates of
counterfactual risk with greater statistical efficiency than both
propensity score-based estimators and the classical ICE g-formula.

Conditional exchangeability means that for all time points, there is no
confounding of the relationship between exposure/censoring and both
future exposure/censoring and NHL status given the observed past,
including past exposure and covariates [@Richardson_2013; @Diaz_2021]. A
major threat to conditional exchangeability in longitudinal occupational
studies is the HWSE. We limit potential bias due to the HWSE by
conditioning on cumulative exposure, employment status, and cumulative
time off history at each time point. Cumulative time off and employment
status are reasonable mediators of the causal paths linking past health
to future exposure and health, but adjustment for these variables may
not be sufficient for eliminating bias due to the HWSE. Declines in a
worker's health may lead to reductions in work-related exposure without
affecting employment status or time off work [@Garcia_2017]. We expect
the absence of time-varying measures of worker health over the life
course to result in bias toward the null.

Positivity refers to the need for adequate variation in future exposure
among strata formed by observed covariate and intervention-compliant
exposure histories. Even under conditional exchangeability, where
exposures within these strata may be considered the result of
experimental assignment, expected counterfactual outcomes under
different exposures may not be estimable if there is excessive sparsity
in the observed distribution of exposures [@Maldonado_2002]. We
investigated static and dynamic stochastic interventions on soluble MWF
exposure that intervened only when average daily exposure exceeded the
hypothetical limit under consideration. Hence, our parameters of
interest achieve positivity more easily than those for static
deterministic interventions e.g. setting all to a single level of
exposure. Nonetheless, violations in positivity were still of concern
due to the high dimensionality of covariates, as is common in
longitudinal settings. We addressed sparsity by summarizing the 20 years
of follow-up over a coarser timescale with only 8 follow-up periods and
by representing covariates using fewer categories. Coarsening limits the
comprehensiveness of confounding control, but improves positivity. In
practice, causal inference using observational data must always balance
positivity, covariate adjustment, and model specification
[@Petersen_2012].
<!-- When estimating causal effects using estimators that do not require exposure modeling, as is the case with ICE g-formula methods, concerns over potential bias due to practical violations in positivity may be relaxed if correct model specification, discussed below, is attained. -->

The consistency assumption, also known as the
no-multiple-versions-of-treatment or stable unit treatment value
assumption, is that counterfactual outcomes under each possible exposure
value take on a unique value [@Cole_2009; @VanderWeele_2013]. This
assumption would be violated if there were multiple versions of
treatment causally associated with different outcomes. This basic notion
of consistency is violated in our analysis because our exposure of
interest is a complex mixture of diverse components with substantial
variation over time due to changes in formulation as well as the natural
physical, chemical, and biological changes in the MWF over the course of
its use and reuse [@Howell_2006]. However, causal effect estimates under
violations in the consistency assumption are still valid and unbiased if
there is adequate adjustment for confounders of the exposure-version
relationship [@VanderWeele_2013]. This may be thought of as conditional
consistency within strata, in which there is only one version of
treatment. Our analysis indexed time periods over calendar time and
adjusted for age, year of hire, and plant. In this way, we limited
potential for bias due to variation in MWF composition.

Correct model specification is a standard assumption in all parametric
and semi-parametric analyses. The estimator we applied offers greater
statistical efficiency than the classical ICE parametric g-formula
estimator because it leverages greater parametric smoothing. In causal
analyses of longitudinal cohort studies, both the hazard-extended and
classical ICE parametric g-formula estimators are less common than the
NICE parametric g-formula [@Keil_2014]. However, a major limitation of
the NICE g-formula is the g-null paradox: the guaranteed
misspecification of parametric models resulting in the false rejection
of the null hypothesis when the null is true and when there is
time-varying confounding affected by past exposure [@Naimi_2015;
@McGrath_2022]. As with all ICE g-formula estimators, the estimator we
applied is not subject to the g-null paradox. Furthermore, simulation
studies show that the variance of the hazard-extended ICE parametric
g-formula is similar to that of the NICE parametric g-formula, so we
expect the former to be no less conservative than the latter
[@Wen_2020].

<!-- In the last several decades, there have been substantial changes in the formulation of MWF, including the exclusion of nitrogenous compounds and the use of highly-refined mineral oils. The recent establishment of the Ecolabel for lubricants in the European Union reflects the shift toward prioritizing sustainability and health in the design and production of MWF and other lubricants [@EU_2018;@Skerlos_2008]. However, water-based MWF that satisfy Ecolabel criteria are nonetheless susceptible to the accumulation of dirt, metals, microbes, products of microbial metabolism, and unemulsified oil. These contaminants and their control strategies may also induce risks to human health. Furthermore, adoption of this new class of MWF remains limited, as is evidence of their safety. Considering the evolving complexity in MWF composition, epidemiologic investigations remain an invaluable approach for developing evidence-based occupational health policy and estimating their causal effect. -->

Much of the existing epidemiologic literature linking occupational and
environmental exposures to NHL report findings from case-control studies
where exposures are measured crudely as binary indicators of exposure or
membership in a particular occupational group [@Cano_2001;
@Costantini_2001; @Karunanayake_2008; @t-Mannetje_2016]. Associations
between occupations and NHL risk vary considerably, but one study of
working men in Kansas and Nebraska found strong associations between NHL
risk and occupations involving metalworking and motor vehicles
[@Zheng_2002]. Both of these occupations may entail exposure to soluble
MWF, which contain a number of additives of concern for human health and
for NHL risk in particular. Organic compounds containing phosphorous,
chlorine, sulfur, nitrogen, and boron are commonly added to soluble MWF
to control microbial growth, improve performance under high
heat/pressure, and inhibit corrosion [@Evans_2020]. Organophosphorus
compounds include organophosphate pesticides, which have been linked to
cancer risk in epidemiologic and animal studies. Some were classified as
possibly carcinogenic by the IARC [@IARC_2017]. Studies of occupational
exposure to chlorinated solvents and pesticides have also been linked to
NHL risk [@Cocco_2008; @Purdue_2011; @Cocco_2013; @Vlaanderen_2013;
@Callahan_2018]. In 2014, the IARC classified trichloroethylene,
tetrachloroethylene, and other chlorinated agents as Group 1 carcinogens
[@IARC_2014]. Chlorinated solvents are commonly used as degreasers in
industrial settings, but their use in the plants under study here was
rare [@Shrestha_2016]. The structural characteristics shared by MWF
additives and known/suspected carcinogens suggest potential similarities
in their behavior in biological systems.

This study investigated the effect of hypothetical limits on MWF
exposure by comparing the standardized distributions of NHL under
various distributions of cumulative exposure induced by applying upper
bounds to average daily exposure to soluble MWF. We selected these
hypothetical limits based on the NIOSH REL of 0.5 mg/m^3^ [@niosh_1998].
In the real world, there is no enforcement of the REL, but we
nonetheless observed average daily exposures below the REL in these GM
plants for many years. If the REL were enforced in the real world, we
would not expect reductions in exposure for these low-exposure
person-years. Hence, contrasting the counterfactual scenario where all
workers experienced average daily exposure at the REL to one where all
workers experienced average daily exposure at some higher level would
result in an overestimate of the expected real-world benefit of REL
enforcement.

<!-- The interventions we considered are stochastic interventions resulting in post-intervention distributions of exposure that better reflect what we would expect if exposure limits were enforced in the real world. The interventions which ignore exposure to straight and synthetic MWF are static. The interventions which reduce exposure to soluble MWF by a quantity determined by exposure to other MWF are dynamic. The dynamic stochastic interventions respect the fact that the REL is for total MWF, not just soluble MWF. If an individual experienced a large amount of exposure to straight and synthetic MWF in one year, then their exposure to soluble MWF should be bounded at a lower level so that the sum of their exposure to soluble, straight, and synthetic MWF is no greater than the hypothetical limit.  -->

# Conclusions

Associations between several occupations and risk of NHL have been
reported previously, but none evaluated the potential effect of
hypothetical limits on occupational exposures [@Ekstrom-Smedby_2006;
@Chiu_2015; @t-Mannetje_2016; @Callahan_2018]. We found evidence that
exposure to soluble MWF was associated with NHL incidence after
adjustment for time-varying confounding affected by prior exposure using
the hazard-extended ICE parametric g-formula. Reducing cumulative
exposure to soluble MWF by enforcing hypothetical, but realistic,
interventions on average daily exposure would reduce NHL incidence.
\newpage

# Citations

::: {#refs}
:::