### 2025-04-15 reports/paper/resources/intervention-density-witnin-combos-b.tex

% Created by tikzDevice version 0.12.6 on 2025-04-14 23:49:03
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage{upgreek}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (252.94,180.67);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (252.94,180.68);
\end{scope}
\begin{scope}
\path[clip] ( 39.16, 31.76) rectangle (246.94,165.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 39.16, 31.76) rectangle (246.95,165.67);
\definecolor{fillColor}{RGB}{211,211,211}

\path[fill=fillColor] ( 48.60, 37.85) rectangle ( 55.87, 37.85);

\path[fill=fillColor] ( 55.87, 37.85) rectangle ( 63.13, 37.85);

\path[fill=fillColor] ( 63.13, 37.85) rectangle ( 70.40, 37.85);

\path[fill=fillColor] ( 70.40, 37.85) rectangle ( 77.66, 37.85);

\path[fill=fillColor] ( 77.66, 37.85) rectangle ( 84.93, 37.85);

\path[fill=fillColor] ( 84.93, 37.85) rectangle ( 92.19, 37.85);

\path[fill=fillColor] ( 92.19, 37.85) rectangle ( 99.46, 37.85);

\path[fill=fillColor] ( 99.46, 37.85) rectangle (106.72, 37.85);

\path[fill=fillColor] (106.72, 37.85) rectangle (113.99, 37.85);

\path[fill=fillColor] (113.99, 37.85) rectangle (121.26,159.59);

\path[fill=fillColor] (121.26, 37.85) rectangle (128.52, 37.85);

\path[fill=fillColor] (128.52, 37.85) rectangle (135.79, 37.85);

\path[fill=fillColor] (135.79, 37.85) rectangle (143.05, 37.85);

\path[fill=fillColor] (143.05, 37.85) rectangle (150.32, 37.85);

\path[fill=fillColor] (150.32, 37.85) rectangle (157.58, 37.85);

\path[fill=fillColor] (157.58, 37.85) rectangle (164.85, 37.85);

\path[fill=fillColor] (164.85, 37.85) rectangle (172.11, 37.85);

\path[fill=fillColor] (172.11, 37.85) rectangle (179.38, 37.85);

\path[fill=fillColor] (179.38, 37.85) rectangle (186.64, 37.85);

\path[fill=fillColor] (186.64, 37.85) rectangle (193.91, 37.85);

\path[fill=fillColor] (193.91, 37.85) rectangle (201.17, 37.85);

\path[fill=fillColor] (201.17, 37.85) rectangle (208.44, 37.85);

\path[fill=fillColor] (208.44, 37.85) rectangle (215.70, 37.85);

\path[fill=fillColor] (215.70, 37.85) rectangle (222.97, 37.85);

\path[fill=fillColor] (222.97, 37.85) rectangle (230.23, 37.85);

\path[fill=fillColor] (230.23, 37.85) rectangle (237.50, 37.85);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt] ( 48.60, 37.85) rectangle ( 55.87, 37.85);

\path[draw=drawColor,line width= 0.4pt] ( 55.87, 37.85) rectangle ( 63.13, 37.85);

\path[draw=drawColor,line width= 0.4pt] ( 63.13, 37.85) rectangle ( 70.40, 37.85);

\path[draw=drawColor,line width= 0.4pt] ( 70.40, 37.85) rectangle ( 77.66, 37.85);

\path[draw=drawColor,line width= 0.4pt] ( 77.66, 37.85) rectangle ( 84.93, 37.85);

\path[draw=drawColor,line width= 0.4pt] ( 84.93, 37.85) rectangle ( 92.19, 37.85);

\path[draw=drawColor,line width= 0.4pt] ( 92.19, 37.85) rectangle ( 99.46, 37.85);

\path[draw=drawColor,line width= 0.4pt] ( 99.46, 37.85) rectangle (106.72, 37.85);

\path[draw=drawColor,line width= 0.4pt] (106.72, 37.85) rectangle (113.99, 37.85);

\path[draw=drawColor,line width= 0.4pt] (113.99, 37.85) rectangle (121.26,159.59);

\path[draw=drawColor,line width= 0.4pt] (121.26, 37.85) rectangle (128.52, 37.85);

\path[draw=drawColor,line width= 0.4pt] (128.52, 37.85) rectangle (135.79, 37.85);

\path[draw=drawColor,line width= 0.4pt] (135.79, 37.85) rectangle (143.05, 37.85);

\path[draw=drawColor,line width= 0.4pt] (143.05, 37.85) rectangle (150.32, 37.85);

\path[draw=drawColor,line width= 0.4pt] (150.32, 37.85) rectangle (157.58, 37.85);

\path[draw=drawColor,line width= 0.4pt] (157.58, 37.85) rectangle (164.85, 37.85);

\path[draw=drawColor,line width= 0.4pt] (164.85, 37.85) rectangle (172.11, 37.85);

\path[draw=drawColor,line width= 0.4pt] (172.11, 37.85) rectangle (179.38, 37.85);

\path[draw=drawColor,line width= 0.4pt] (179.38, 37.85) rectangle (186.64, 37.85);

\path[draw=drawColor,line width= 0.4pt] (186.64, 37.85) rectangle (193.91, 37.85);

\path[draw=drawColor,line width= 0.4pt] (193.91, 37.85) rectangle (201.17, 37.85);

\path[draw=drawColor,line width= 0.4pt] (201.17, 37.85) rectangle (208.44, 37.85);

\path[draw=drawColor,line width= 0.4pt] (208.44, 37.85) rectangle (215.70, 37.85);

\path[draw=drawColor,line width= 0.4pt] (215.70, 37.85) rectangle (222.97, 37.85);

\path[draw=drawColor,line width= 0.4pt] (222.97, 37.85) rectangle (230.23, 37.85);

\path[draw=drawColor,line width= 0.4pt] (230.23, 37.85) rectangle (237.50, 37.85);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 7pt off 3pt ,line join=round] (121.26, 31.76) -- (121.26,165.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.16, 31.76) rectangle (246.95,165.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 34.21, 35.09) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 34.21, 65.53) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 34.21, 95.96) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 34.21,126.40) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 34.21,156.83) {40};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 36.41, 37.85) --
	( 39.16, 37.85);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 36.41, 68.28) --
	( 39.16, 68.28);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 36.41, 98.72) --
	( 39.16, 98.72);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 36.41,129.15) --
	( 39.16,129.15);

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 36.41,159.59) --
	( 39.16,159.59);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.4pt,line join=round] ( 48.60, 29.01) --
	( 48.60, 31.76);

\path[draw=drawColor,line width= 0.4pt,line join=round] (106.72, 29.01) --
	(106.72, 31.76);

\path[draw=drawColor,line width= 0.4pt,line join=round] (164.85, 29.01) --
	(164.85, 31.76);

\path[draw=drawColor,line width= 0.4pt,line join=round] (222.97, 29.01) --
	(222.97, 31.76);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 48.60, 21.30) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (106.72, 21.30) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (164.85, 21.30) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (222.97, 21.30) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (143.05,  7.77) {Exposure to soluble MWF (mg/m$^3$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 17.51, 98.72) {Density};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (170.95,144.14) rectangle (246.94,167.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (170.95,161.15) {Target exposure limit:};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (170.95,146.64) rectangle (178.95,154.64);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 7pt off 3pt ,line join=round] (174.95,146.64) -- (174.95,154.64);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (184.45,147.89) {0.25 mg/m$^3$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (170.95,110.11) rectangle (246.35,141.64);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (170.95,135.12) {Distribution:};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (170.95,120.61) rectangle (178.95,128.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{fillColor}{RGB}{211,211,211}

\path[fill=fillColor] (170.95,120.61) rectangle (178.95,128.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (170.95,112.61) rectangle (178.95,120.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt] (171.45,113.11) rectangle (178.45,120.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (184.45,121.85) {Post-intervention};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (252.94,180.67);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (184.45,113.85) {Observed};
\end{scope}
\end{tikzpicture}

\end{document}